### 2021-01-11 Manuscript_pirAnalysis_Github/AncestralPir/AnalysisCode/Manuscript_Github_AncestralPir.Rmd
Add files via upload

---
title: "Manuscript_Github_ancestral_pir_210801"
author: "Timothy Little"
date: "08/01/2021"
output: 
  html_document:
      toc: true
      number_sections: true
      toc_float: true
---

Got the 19 individual orthologs of PBANKA_0100500 from PlasmoDB, did orthology transformation twice to get everything. Then added the PccCB ortholog = 20 genes.
With the 20 genes did a MUSCLE (v3.8.31) alignment, PhyML (v3.1)* building with LG model 100 bootstrap replicates, on SEAVIEW v4.7.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/AncestralPir/AnalysisCode")
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r load_packages}
library('treeio') # Download trees from various formats
library('ggtree') # Draw phylogenetic trees with ggplot
library('ggstance')
library('ggnewscale')
library('plotly')
library('ComplexHeatmap')
library('circlize') #for ColorRamp2 function
library('ggplot2')
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('cowplot')
library('kableExtra')
library('viridis') # color palettes
library('RColorBrewer') # color palettes
library('corrplot')
library('reshape2')
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('readxl') # for reading in Excel spreadsheets
library('readr') # for reading in everything else
# library('tidyverse')
library('dplyr')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r function_uniquify}
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
    name <- paste(vector_names[i],
                  df_num$rep[match(vector_names[i], df_num$names)],
                  sep = '_')
    #Increase rep counter
    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
    new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r tpm_function}
tpm = function(count_data, lengths) {

  #Print an error if the id columns do not match
  # first if statement to test that both have a GeneID column
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'Geneid', 
                                                  #This needs to be Geneid when some of the samples have 'id' in them
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      print('Warning:Count data and length gene ids DO NOT MATCH')
    }
  } else {
    print('Count data and/or length table supplied do not have an id column, 
          make sure they are ordered correctly')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = TRUE))){
    gene_names <- count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
    counts_tpm <- count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = TRUE)]
  }else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = TRUE))){
    lengths.table = lengths[,grep(colnames(lengths), pattern = 'length', ignore.case = T)]
  }else{
    lengths.table <- lengths}
  #Making lengths.table a matrix sometimes throws up problems
  lengths.table = as.matrix(lengths.table)
  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  colnames(counts_tpm) <- 'tpm'
  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), 
                                            pattern = 'Geneid', 
                                            ignore.case = T), 
                                      drop = FALSE],
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

```{r function_get_ecdf_percentile}
ecdf_percentile <- function(df_tpm){
  Fn <- ecdf(
    unlist( #To convert into an atomic vector that ecdf will like.
      select_at(
        df_tpm,
        vars(
          matches('med_fpkm|med_tpm|^tpm$') #Default matches behaviour is ignore.case = TRUE
        ) #Could be med_fpkm or med_tpm.
      )
    )
  )
  #Using a lapply to make a list in case there are multiple ancestral genes.
  list_percentile <- lapply(
    df_tpm$Geneid[
      df_tpm$ancestral_ortholog == 1
      ],
    function(ancestor_gene){
      #For debugging
      # ancestor_gene <- "PCHCB_0101200"
      list(
        Fn(
          filter(df_tpm, Geneid %in% ancestor_gene) %>% 
            select_at(
              vars(
                matches('med_fpkm|med_tpm|^tpm$')
              ) #Could be med_fpkm or med_tpm
            ) 
        )
      )
    }
  )
  #Create data.frame of the output
  gene_percentile_df <- data.frame(
    Geneid = df_tpm$Geneid[
      df_tpm$ancestral_ortholog ==1
      ],
    percentile = unlist(
      lapply(
        list_percentile, `[`, 1
      )
    )
  )
  plot(Fn)
  return(gene_percentile_df)
}
# ecdf_percentile(cyno_genes_fpkm)
```

```{r get_phylip_tree}
phylog <- read.tree('PlasmoDBv48_GenesByOrthologs_ancestralpir_withPccCBandPberghei-PhyML_100bootstrap_tree.phylip')
```

```{r removing_unncessary_tips}
phylog <- drop.tip(phylog, 
                   tip = c("PKNOH_S130216000",
                           "YYE_04766",
                           "YYG_04764",
                           # "PY17X_0102100",
                           "PY06119",
                           'PYYM_0101500',
                           'PVL_110036700')
)
```

```{r gene_and_species_names}
gene_names <- phylog$tip.label %>% 
  str_split_fixed(pattern = '_P.', n =2) %>% 
  .[,1]
write(gene_names,
            file = 'gene_names_ancestral_includedintree.txt',
            sep = '\n')
phylog$old.tip.label <- phylog$tip.label

#Download the gene info with species names and add in the PccCB name.
ortho_gene_info_tbl <- rbind(
  read_csv('PlasmoDBv48_ancestralpir_orthologs_GenesByOrthologs_Summary.csv') %>% 
    select(c('Gene ID', Organism)) %>% 
    dplyr::rename(Geneid = 'Gene ID'),
  data.frame(Geneid = 'PCHCB_0101200',
             'Organism' = 'Plasmodium chabaudi chabaudi CB')
)

phylog$tip.label <- ortho_gene_info_tbl$Organism[match(phylog$old.tip.label, 
                                                       ortho_gene_info_tbl$Geneid)] %>% 
  str_replace(pattern = 'Plasmodium ', 
              replacement = 'P. ' ) %>% 
  str_replace(pattern = '(?<!cynomolgi) strain.+| curtisi|(?<=chabaudi) chabaudi|(?<=yoelii) yoelii', 
              replacement = '') %>% 
  str_extract(pattern = 'P. [[:lower:]]+(-like)?( [[:lower:]]+)?( M| B| CB| Sal-1| P01|)') %>% 
  gsub(pattern = 'chabaudi$', 
       replacement = 'chabaudi AS') %>% 
  gsub(pattern = 'strain ',
       replacement = '') %>% 
  uniquify(first_one = TRUE) 
#There shouldn't be any extra ones now
ovale_gene_names <- str_subset(gene_names, 
                               pattern = 'Poc')
```

'A glance of the blood stage transcriptome of a Southeast Asian Plasmodium ovale isolate'

ovale data is in TPM

```{r ovale_transcriptome_data}
ovale_path <- '../other_species_transcriptome_data/Brashear_2019_ovale_transcriptome_journal.pntd.0007850.s008.xlsx'
ovale <- read_xlsx(ovale_path, sheet = 1)
ovale <- ovale %>% 
  mutate(ancestral_ortholog = ifelse(Gene %in% gene_names, 1, 0))
#Sanity test
(ovale$ancestral_ortholog %>% sum) == 9

ovale_anc_v_nonanc <- ovale %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(med_tpm = median(TPM),
                   mean_tpm = mean(TPM),
                   max_tpm = max(TPM),
                   min_tpm = min(TPM))
#Median TPM of non-ancestral pirs
ovale_nonanc_medtpm <- ovale_anc_v_nonanc %>% filter(ancestral_ortholog == 0) %>% .$med_tpm
#Ratio of the TPM of the ancestrals to the median non-ancestrals
ovale_ancestral <- ovale %>% 
  filter(ancestral_ortholog ==1) %>% 
  mutate(Geneid = Gene,
         ratio = TPM/ovale_nonanc_medtpm,
         median_tpm = TPM) %>% 
  select(Geneid, ratio, median_tpm)
```

```{r ovale_percentile}
#Calculate the percentile of the ancestral v other pirs

ovale %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

ovale[ovale$ancestral_ortholog == 1,]

ovale_tpm_collated_pir <- ovale  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo') | ancestral_ortholog == 1) %>% 
  select(c(Gene, Description, ancestral_ortholog, TPM)) %>% 
  dplyr::arrange(desc(TPM))

#ecdf produces a function which can be used to find the percentile that a value is present in in a distribution
Fn_ovale <- ecdf(ovale_tpm_collated_pir$TPM)
plot(Fn_ovale)
list_ovale_percantile <- lapply(ovale_tpm_collated_pir$Gene[ovale_tpm_collated_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_ovale(filter(ovale_tpm_collated_pir, Gene %in% ancestr)$TPM))
                                })
ovale_percentile <- data.frame(Geneid = ovale_tpm_collated_pir$Gene[ovale_tpm_collated_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_ovale_percantile, `[`, 1)))

#Percentile for all genes
# Fn_ovale_all <- ecdf(ovale$TPM)
# plot(Fn_ovale_all)
# list_ovale_percantile_all <- lapply(ovale$Gene[ovale$ancestral_ortholog == 1], 
#                                 function(ancestr){
#                                   list(Fn_ovale_all(filter(ovale, Gene %in% ancestr)$TPM))
#                                 })
# (ovale_percentile_all <- data.frame(Geneid = ovale$Gene[ovale$ancestral_ortholog == 1],
#                                percentile = unlist(lapply(list_ovale_percantile_all, `[`, 1))))

#Wow must genes in the genome are expressed super low (52% expressed under 25 TPM)
# nrow(filter(ovale, TPM < 25)) /72.80

# ecdf_percentile(mutate(ovale, med_tpm = TPM, Geneid = Gene))
ovale_percentile_all <- ecdf_percentile(mutate(ovale, Geneid = Gene))

```

Data from eLife _A comparative transcriptomic analysis of replicating and dormant liver stages of the relapsing malaria parasite Plasmodium cynomolgi_ 2017.

Note: FPKM

```{r cynomolgi_transcriptome_data}
cyno_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'cynomolgi',
                        full.names = TRUE)
cyno <- read_xlsx(cyno_path, sheet = 1, skip = 3)
colnames(cyno) <- make.names(colnames(cyno))
apply(cyno %>% select_at(vars(matches('Sz|Hz'))), 2, sum)
cyno <- cyno %>% 
  mutate(ancestral_ortholog = ifelse(Gene.Id %in% gene_names,1,0))
(cyno$ancestral_ortholog %>% sum)
cyno_anc_v_nonanc <- cyno %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('Sz\\.|Hz\\.')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  #gsub to remove the numbers for easy summarising across the replicates
  mutate(stage = gsub(variable, pattern = '.[[:digit:]]', replacement = '')) %>% 
  group_by(ancestral_ortholog, stage) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
cyno_schiz_ratio <- filter(cyno_anc_v_nonanc, 
                           ancestral_ortholog == 1) %>% 
  filter(stage == 'Sz') %>% 
  .$median / 
  filter(cyno_anc_v_nonanc, ancestral_ortholog == 0) %>% 
  filter(stage == 'Sz') %>% 
  .$median
cyno_hyp_ratio <- filter(cyno_anc_v_nonanc, ancestral_ortholog == 1) %>% 
  filter(stage == 'Hz') %>% 
  .$median / 
  filter(cyno_anc_v_nonanc, ancestral_ortholog == 0) %>% 
  filter(stage == 'Hz') %>% 
  .$median
cyno_data <- data.frame(Geneid = cyno$Gene.Id[which(cyno$ancestral_ortholog == 1)],
                        ratio = cyno_schiz_ratio,
                        median_tpm = filter(cyno_anc_v_nonanc, ancestral_ortholog == 1) %>% 
                          filter(stage == 'Sz') %>% 
                          .$median)
```

```{r cyno_percentile}
#Calculate the percentile of the ancestral v other pirs
cyno_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PlasmoDBv46_PcynoM',
                        full.names = TRUE)
cyno_genes <- read_csv(cyno_gene_path) %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))

cyno_genes %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

cyno_genes %>% 
  filter(str_detect(Description, pattern = 'pseuodgene'))

cyno_genes_fpkm <- cyno %>% 
  select_at(vars(matches('^Gene|Average'))) %>% 
  melt(id.vars = 'Gene.Id') %>% 
  group_by(Gene.Id) %>% 
  dplyr::summarise(med_tpm = median(value),
                   sd_tpm = sd(value)) %>% 
  mutate(Description = cyno_genes$Description[match(Gene.Id, cyno_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Gene.Id %in% gene_names,1,0)) %>% 
  mutate(Geneid = Gene.Id)

cyno_pir_fpkm <- cyno_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_tpm)) %>% 
  dplyr::arrange(desc(med_tpm))
cyno_pir_fpkm[which(cyno_pir_fpkm$Geneid %in% gene_names),]
cyno_pir_fpkm[which(cyno_pir_fpkm$ancestral_ortholog == 1),]

#The percentile among the pir genes
Fn_cyno <- ecdf(cyno_pir_fpkm$med_tpm)
plot(Fn_cyno)
list_cyno_percantile <- lapply(cyno_pir_fpkm$Geneid[cyno_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_cyno(filter(cyno_pir_fpkm, Geneid %in% ancestr)$med_tpm))
                                })
cyno_percentile <- data.frame(Geneid = cyno_pir_fpkm$Geneid[cyno_pir_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_cyno_percantile, `[`, 1)))

#The percentile among all genes
Fn_cyno_all <- ecdf(cyno_genes_fpkm$med_tpm)
plot(Fn_cyno_all)
list_cyno_percentile_all <- lapply(cyno_genes_fpkm$Geneid[cyno_genes_fpkm$ancestral_ortholog == 1], 
                                   function(ancestr){
                                     list(Fn_cyno_all(filter(cyno_genes_fpkm, Geneid %in% ancestr)$med_tpm))
                                   })
cyno_percentile_all <- data.frame(Geneid = cyno_genes_fpkm$Geneid[cyno_genes_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_cyno_percentile_all, `[`, 1)))

ecdf_percentile(cyno_genes_fpkm)

```

Vivax data from Siegal et al 2020 SciRep 'Analysis of Plasmodium vivax schizont transcriptomes from field isolates reveals heterogeneity of expression of genes involved in host-parasite interactions'.

* FPKM data. 
* Clinical isolates.

```{r}
vivax_genes_fpkm <- read_xlsx('../other_species_transcriptome_data/Siegal2020_vivax-schizont-rnaseq_41598_2020_73562_MOESM2_ESM.xlsx',
          sheet = 3,
          skip = 1,
          col_names = TRUE) %>% 
  mutate(Geneid = gene_id,
         rpkm = `mean RPKM`,
         ancestral_ortholog = ifelse(gene_id %in% gene_names,1,0)) %>% 
  select(Geneid, 
         rpkm,
         ancestral_ortholog)

# sum(vivax_genes_fpkm$ancestral_ortholog) == 1

group_by(vivax_genes_fpkm, 
         ancestral_ortholog) %>% 
  dplyr::summarise(median = median(rpkm),
                   standev = sd(rpkm))

#The percentile among all genes
Fn_vivax_all <- ecdf(vivax_genes_fpkm$rpkm)
plot(Fn_vivax_all)
list_vivax_percentile_all <- lapply(vivax_genes_fpkm$Geneid[vivax_genes_fpkm$ancestral_ortholog == 1], 
                                   function(ancestr){
                                     list(Fn_vivax_all(filter(vivax_genes_fpkm, Geneid %in% ancestr)$rpkm))
                                   })
vivax_percentile_all <- data.frame(Geneid = vivax_genes_fpkm$Geneid[vivax_genes_fpkm$ancestral_ortholog == 1],
                              percentile = unlist(lapply(list_vivax_percentile_all, `[`, 1)))

```

P. chabaudi chabaudi CB data from Lim et al, 2019.

```{r cb_transcriptome_data}
cb_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PccCB_rnaseq',
                        full.names = TRUE)
cb <- read_delim(cb_path, delim = '\t')
apply(cb[,-1], 2, sum) #FPKM, do not add up to 1e6, although they are close
cb <- cb %>% 
  mutate(Geneid = id) %>% 
  select(-id) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(cb$ancestral_ortholog %>% sum)
cb_anc_v_nonanc <- cb %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('CB')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
cb_data <- data.frame(Geneid = cb$Geneid[which(cb$ancestral_ortholog == 1)],
                      ratio = cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 1)]/
                        cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 0)],
                      median_tpm = cb_anc_v_nonanc$median[which(cb_anc_v_nonanc$ancestral_ortholog == 1)])
```

Used ftp://ftp.sanger.ac.uk/pub/project/pathogens/Plasmodium/chabaudi/CB_v2/ for the gene names

```{r cb_percentile}
#Calculate the percentile of the ancestral v other pirs
cb_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'PccCB_gene',
                        full.names = TRUE)

read_delim(cb_gene_path, delim = '\t')[!(apply((apply(read_delim(cb_gene_path, delim = '\t'),
             1,
             str_extract,
             pattern = 'pseudo') %>%
         is.na),
      2, all)),] #Are there any pseudo columns? Yes : 'attributes'

cb_genes <- read_delim(cb_gene_path, delim = '\t', 
                       col_names = c('Geneid', 'Description'))

cb_genes %>% 
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

# cb_genes %>% 
#   group_by(Description) %>% 
#   dplyr::summarise(num = n()) %>% 
#   dplyr::arrange(desc(num))

cb_genes_fpkm <- cb  %>% 
  select_at(vars(matches('TCB|Geneid'))) %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_fpkm = median(value),
                   sd_fpkm = sd(value)) %>% 
  mutate(Description = cb_genes$Description[match(Geneid, cb_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

cb_pir_fpkm <- cb_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'CIR|PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_fpkm)) %>% 
  dplyr::arrange(desc(med_fpkm))
cb_pir_fpkm[which(cb_pir_fpkm$Geneid %in% gene_names),]

Fn_cb <- ecdf(cb_pir_fpkm$med_fpkm)
plot(Fn_cb)
list_cb_percantile <- lapply(cb_pir_fpkm$Geneid[cb_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_cb(filter(cb_pir_fpkm, Geneid %in% ancestr)$med_fpkm))
                                })
cb_percentile <- data.frame(Geneid = cb_pir_fpkm$Geneid[cb_pir_fpkm$ancestral_ortholog == 1],
                            percentile = unlist(lapply(list_cb_percantile, `[`, 1)))

ecdf_percentile(cb_pir_fpkm)
cb_percentile_all <- ecdf_percentile(cb_genes_fpkm)

```

vinckei data from Ramaprasad 2019 'A fast and cost-effective microsampling protocol incorporating reduced animal usage for time-series transcriptomics in rodent malaria parasites'

PlasmoDB v48 is not updated enough to have the vinckei assembly that the paper used. By BLASTing the PlasmoDB sequence against the PvvCY assembly on https://www.ncbi.nlm.nih.gov/assembly/GCA_900681995.1 and using the gff file from ftp://ftp.ncbi.nlm.nih.gov/genomes/all/GCA/900/681/995/GCA_900681995.1_PVVCY_v1 my guess is that it is this gene:

LR215057.1	EMBL	gene	71066	72501	.	-	.	ID=gene-PVVCY_0100190;Name=PVVCY_0100190;gbkey=Gene;gene_biotype=protein_coding;locus_tag=PVVCY_0100190

corresponding CDS data:
\>lcl|LR215057.1_cds_VEV54408.1_19 \[locus_tag=PVVCY_0100190\] \[protein=PIR protein CIR protein\] \[protein_id=VEV54408.1\] [location=complement(join(71066..71140,71320..72291,72493..72501))] \[gbkey=CDS\]

Also have downloaded the features table from the same ftp site as above.

Is this TPM or FPKM? FPKM

```{r vinck_data}
vinck_path <- list.files('../other_species_transcriptome_data/', 
                      #Use glob2rx to get the regex we want
                      pattern = glob2rx('*vinckei*xlsx*'),
                      full.names = TRUE)
vinck <- read_xlsx(vinck_path, skip =1)
select_at(vinck, vars(matches('[[:digit:]]h'))) %>% apply(., 2, sum)
#Definitely FPKM
vinck <- vinck %>% 
  mutate(Geneid = gene_id) %>% 
  select(-gene_id) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(vinck$ancestral_ortholog %>% sum)
vinck_anc_v_nonanc <- vinck %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('[[:digit:]]h')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
vinck_data <- data.frame(Geneid = vinck$Geneid[which(vinck$ancestral_ortholog == 1)],
                        ratio = vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 1)]/
                        vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = vinck_anc_v_nonanc$median[which(vinck_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r vinckei_percentile}
#Calculate the percentile of the ancestral v other pirs
vinck_gene_path <- list.files('../other_species_transcriptome_data/', 
                        pattern = 'vinckei_feature',
                        full.names = TRUE)

read_delim(vinck_gene_path, delim = '\t')[!(apply((apply(read_delim(vinck_gene_path, delim = '\t'),
             1,
             str_extract,
             pattern = 'pseudo') %>%
         is.na),
      2, all)),] #Are there any pseudo columns? Yes : 'attributes'

vinck_genes <- read_delim(vinck_gene_path, delim = '\t') %>% 
  mutate(Geneid = locus_tag,
         Description = ifelse(is.na(attributes),
                              name,
                              paste(name, attributes))) %>% 
  select(c(Geneid, Description)) %>% 
  filter(!is.na(Description))

vinck_genes %>% filter(Geneid %in% 'PVVCY_0200040')

vinck_genes %>% 
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

vinck_genes_fpkm <- vinck %>% 
  select_at(vars(matches('^[[:digit:]]+h|Geneid'))) %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_fpkm = median(value),
                   sd_fpkm = sd(value)) %>% 
  mutate(Description = vinck_genes$Description[match(Geneid, vinck_genes$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

vinck_pir_fpkm <- vinck_genes_fpkm  %>% 
  filter(str_detect(Description, pattern = 'PIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  select(c(Geneid, Description, ancestral_ortholog, med_fpkm)) %>% 
  dplyr::arrange(desc(med_fpkm))
vinck_pir_fpkm[which(vinck_pir_fpkm$Geneid %in% gene_names),]

Fn_vinck <- ecdf(vinck_pir_fpkm$med_fpkm)
plot(Fn_vinck)
list_vinck_percantile <- lapply(vinck_pir_fpkm$Geneid[vinck_pir_fpkm$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_vinck(filter(vinck_pir_fpkm, Geneid %in% ancestr)$med_fpkm))
                                })
vinck_percentile <- data.frame(Geneid = vinck_pir_fpkm$Geneid[vinck_pir_fpkm$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_vinck_percantile, `[`, 1)))
ecdf_percentile(vinck_pir_fpkm)
vinck_percentile_all <- ecdf_percentile(vinck_genes_fpkm)

```

This distribution of gene expression is different compared to that of the other samples..

##Coatneyi data

'Distinct amino acid and lipid perturbations characterize acute versus chronic malaria'

from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE103259

Want to convert the counts from this file into TPM with the transcript lengths.

_P. coatneyi_ _ancestral pir_ is not annotated as a _pir_ !

```{r coatneyi_data}
coat_path <- list.files('../other_species_transcriptome_data/', 
                      pattern = '2019_coatneyi_rawcounts',
                      full.names = TRUE)
coat <- read_xlsx(coat_path, skip =3, sheet  =2)
colnames(coat)[1:2] <- c('gene', 'Geneid')
coat <- coat %>% 
  select(-gene) %>% 
  arrange(Geneid) 
#apply(coat[,-1], 2, sum)
plot(apply(coat[,-1], 2, sum), apply(coat[,-1], 2, function(x) sum(x > 0)),
     main = 'coatneyi data features detected vs total counts',
     xlab = 'Total counts of sample',
     ylab = 'Total detected genes (>0 counts) per sample')
#Great ranges in the sum of features counted in each sample, should filter out low ones
coat <- coat[,c(TRUE, apply(coat[,-1], 2, sum) > 1e7)]

coat_transcript_lengths_path <- list.files('../../genomes_genes_annotations/', 
                      pattern = 'PlasmoDBv48_Pcoatneyi',
                      full.names = TRUE)
coat_transcript_lengths <- read_csv(coat_transcript_lengths_path) %>% 
  mutate(Geneid = `Gene ID`) %>% 
  select(c(Geneid, `Transcript Length`))

coat_tpm <- tpm(coat, 
                coat_transcript_lengths)

apply(coat_tpm[,-1], 
      2, 
      sum)

coat_tpm <- coat_tpm %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

(coat_tpm$ancestral_ortholog %>% sum)
coat_anc_v_nonanc <- coat_tpm %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(matches('tpm')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
                   standev = sd(value))
coat_data <- data.frame(Geneid = coat_tpm$Geneid[which(coat_tpm$ancestral_ortholog == 1)],
                        ratio = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)]/
                          coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = coat_anc_v_nonanc$median[which(coat_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r coatneyi_percentile}
#Calculate the percentile of the ancestral v other pirs
coat_gene_info <- read_csv(coat_transcript_lengths_path) %>% 
  mutate(Geneid = `Gene ID`) %>% 
  select(Geneid, `Product Description`)

coat_gene_info %>% 
  filter(str_detect(`Product Description`, pattern = 'KIR')) %>% 
  group_by(`Product Description`) %>% 
  dplyr::summarise(n())

coat_tpm_collated <- coat_tpm %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(value),
                   sd_tpm = sd(value)) %>% 
  mutate(Product = coat_gene_info$`Product Description`[match(Geneid, coat_gene_info$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

coat_tpm_collated_pir <- coat_tpm_collated  %>% 
  filter(str_detect(Product, pattern = 'KIR') | ancestral_ortholog == 1) %>% 
  dplyr::arrange(desc(med_tpm))

Fn_coat <- ecdf(coat_tpm_collated_pir$med_tpm)

plot(Fn_coat)

list_coat_percantile <- lapply(coat_tpm_collated_pir$Geneid[coat_tpm_collated_pir$ancestral_ortholog == 1], 
                               function(ancestr){
                                 list(Fn_coat(filter(coat_tpm_collated_pir, Geneid %in% ancestr)$med_tpm))
                               })
coat_percentile <- data.frame(Geneid = coat_tpm_collated_pir$Geneid[coat_tpm_collated_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_coat_percantile, `[`, 1)))

ecdf_percentile(coat_tpm_collated_pir)
coat_percentile_all <- ecdf_percentile(coat_tpm_collated)
```

berghei data - using Brugat et al., 2017

Highest expressed pir gene of any kind is 'PBANKA_0201250' - a pseudogene.

```{r berghei_data}
berg_path <- list.files('../other_species_transcriptome_data/',
                        pattern = 'Expt02',
                        recursive = TRUE,
                        full.names = TRUE)
berg <- read_csv(berg_path) %>%  
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
(berg$ancestral_ortholog %>% sum)
berg_anc_v_nonanc <- berg %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise_at(vars(-matches('Geneid')), 
                      list(~median(.))) %>% 
  melt(id.vars = 'ancestral_ortholog') %>% 
  group_by(ancestral_ortholog) %>% 
  dplyr::summarise(median = median(value),
            standev = sd(value))
berg_data <- data.frame(Geneid = berg$Geneid[which(berg$ancestral_ortholog == 1)],
                        ratio = berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 1)]/
                          berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 0)],
                        median_tpm = berg_anc_v_nonanc$median[which(berg_anc_v_nonanc$ancestral_ortholog == 1)])
```

```{r berghei_percentile}

berg_descr <- read_csv('../../genomes_genes_annotations/standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv') %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))
berg_pir <- berg %>% 
  melt(id.vars = 'Geneid') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(value),
            sd_tpm = sd(value)) %>% 
  mutate(Description = berg_descr$Description[match(Geneid, berg_descr$Geneid)]) %>% 
  filter(str_detect(Description, pattern = 'PIR')) %>% 
  filter(str_detect(Description, pattern = 'pseudo|fragment', negate = TRUE)) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0)) %>% 
  dplyr::arrange(desc(med_tpm))

Fn_bir <- ecdf(berg_pir$med_tpm)
plot(Fn_bir)
list_bir_percentile <- lapply(berg_pir$Geneid[berg_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_bir(filter(berg_pir, Geneid %in% ancestr)$med_tpm))
                                })
bir_percentile <- data.frame(Geneid = berg_pir$Geneid[berg_pir$ancestral_ortholog == 1],
                             percentile = unlist(lapply(list_bir_percentile, `[`, 1)))

berg_percentile_all <- ecdf_percentile(
  berg %>% 
    melt(id.vars = 'Geneid') %>% 
    group_by(Geneid) %>% 
    dplyr::summarise(med_tpm = median(value),
              sd_tpm = sd(value)) %>% 
    mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))
)
```

Py17XNL data from gametocytes (Hart ... Lindner et al 2019) 'Plasmodium male gametocyte development and transmission are critically regulated by the two putative deadenylases of the CAF1/CCR4/NOT complex'

```{r yoelii_data}
yeol_path <- list.files('..',
                        pattern = 'Hart_2019_Pyeolii',
                        recursive = TRUE,
                        full.names = TRUE)
yeol_counts_full <- read_xlsx(yeol_path, 
                         sheet = 2) %>% 
  mutate(Geneid = GeneID...1,
         Description = `[Product Description]`,
         Counts = `489 Average`) %>% 
  select(c(Geneid, Description, Counts)) %>% 
  arrange(Geneid)

yeol_counts_full %>%
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

yeol_counts <- select(yeol_counts_full, -Description)

yeol_lengths_path <- list.files('..',
                                pattern = 'PlasmoDBv46_Pyoelii17X_',
                                recursive = TRUE,
                                full.names = TRUE)
yeol_lengths <- read_csv(yeol_lengths_path) %>% 
  mutate(Geneid = `Gene ID`,
         length = `Transcript Length`) %>% 
  select(c(Geneid, length))
#One gene 'PY17X_0902700' has two different length transcripts - take median
yeol_lengths %>% group_by(Geneid) %>% dplyr::summarise(n = n()) %>% dplyr::arrange(desc(n))

#This gives 5074.5 so need to make sure that not having an integer count messes up things but it shouldn't
yeol_lengths <- yeol_lengths %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(length = median(length))
nrow(yeol_lengths)
nrow(yeol_counts)

yeol_tpm <- tpm(count_data = yeol_counts, lengths = yeol_lengths)
yeol_tpm <- yeol_tpm %>% 
  mutate(Description = yeol_counts_full$Description[match(Geneid, 
                                                          yeol_counts_full$Geneid)]) %>% 
  mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

yeol_tpm_pir <- yeol_tpm %>% 
  filter(str_detect(Description, pattern = 'PIR|YIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>% 
  dplyr::arrange(desc(tpm))

Fn_yeol <- ecdf(yeol_tpm_pir$tpm)

plot(Fn_yeol)

list_yeol_percantile <- lapply(yeol_tpm_pir$Geneid[yeol_tpm_pir$ancestral_ortholog == 1], 
                               function(ancestr){
                                 list(Fn_yeol(filter(yeol_tpm_pir, Geneid %in% ancestr)$tpm))
                               })
yeol_percentile <- data.frame(Geneid = yeol_tpm_pir$Geneid[yeol_tpm_pir$ancestral_ortholog == 1],
                               percentile = unlist(lapply(list_yeol_percantile, `[`, 1))) 
yeol_percentile

yeol_percentile_all <- ecdf_percentile(yeol_tpm)

```

```{r chabaudi_percentile}

chab_info <- read_csv('../../genomes_genes_annotations/standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv') %>% 
  mutate(Geneid = `Gene ID`,
         Description = `Product Description`) %>% 
  select(c(Geneid, Description))

chab_info %>% 
  filter(str_detect(Description, pattern = 'PIR|CIR|IR')) %>% 
  group_by(Description) %>% 
  dplyr::summarise(n())

chab_tpm <- read_xlsx('../other_species_transcriptome_data/Manuscript_Frontiers_SuppInfo_chabaudi_24h.xlsx',
                      sheet = 'chab_TPM_avg_rMT') %>% 
  melt(variable.name = 'stage', value.name = 'tpm') %>% 
  group_by(Geneid) %>% 
  dplyr::summarise(med_tpm = median(tpm),
                sd_tpm = sd(tpm)) %>% 
  mutate(Description = chab_info$Description[match(Geneid, 
                                                       chab_info$Geneid)]) %>% 
      mutate(ancestral_ortholog = ifelse(Geneid %in% gene_names,1,0))

chab_pir <- chab_tpm %>% 
  filter(str_detect(Description, pattern = 'PIR|YIR') | ancestral_ortholog == 1) %>% 
  filter(!str_detect(Description, pattern = 'pseudo|fragment') | ancestral_ortholog == 1) %>%
  dplyr::arrange(desc(med_tpm))

Fn_chab <- ecdf(chab_pir$med_tpm)
plot(Fn_chab)
list_chab_percentile <- lapply(chab_pir$Geneid[chab_pir$ancestral_ortholog == 1], 
                                function(ancestr){
                                  list(Fn_chab(filter(chab_pir, Geneid %in% ancestr)$med_tpm))
                                })
chab_percentile <- data.frame(Geneid = chab_pir$Geneid[chab_pir$ancestral_ortholog == 1],
                            percentile = unlist(lapply(list_chab_percentile, `[`, 1)))

chab_percentile_all <- ecdf_percentile(chab_tpm)
```

## Phylogenetic Tree / Heatmap of percentile compared to all genes

```{r tree_heatmap_all-genes}

ancestral_df <- bind_rows(cyno_percentile_all,
                          ovale_percentile_all,
                          cb_percentile_all,
                          vinck_percentile_all,
                          coat_percentile_all,
                          yeol_percentile_all,
                          berg_percentile_all,
                          chab_percentile_all,
                          vivax_percentile_all)
ancestral_data <- lapply(phylog$old.tip.label, function(id){
  #If any of the pattern is in the string one will be TRUE
  ratio <- if(any(str_detect(string = paste(id), 
                             pattern = ancestral_df$Geneid))){
    index_id <- str_detect(string = paste(id), 
                           pattern = ancestral_df$Geneid)
    ancestral_df$percentile[index_id]*100
  }else{
    NA
  }
  return(list(ratio = ratio))
}) %>% 
  bind_rows
ancestral_data <- as.data.frame(ancestral_data)

rownames(ancestral_data) <- phylog$tip.label

breaks <- seq(0, 100, by = 10)

(tr <- ggtree(phylog, 
            branch.length = 'none',
            layout = 'circular') +
  geom_tiplab2(aes(angle=angle, x = x + 2.5), 
               cex = 3, 
               color='black',
               font = 3) +
  ggplot2::xlim(0,20) 
)
```

```{r heatmap_with_tree_figure3b}

p <- gheatmap(tr, 
              ancestral_data, 
              offset=.2, 
              width=.15, 
              colnames=FALSE) +
  scale_fill_viridis_c(option='C', 
                       name=substitute(
                         atop(
                           '',
                           atop(
                             '', 
                             atop(
                               atop(
                                 paste(
                                   "Percentile of ",
                                   italic('ancestral pir')
                                 ), 
                                 " expression compared"
                               ), 
                               "to all genes"
                             )
                           )
                         )
                       ), 
                       #Use substitute to ensure that the italics function is passed to the name argument
                       #atop ensures that the first line is above the word 'expression'
                       #Two atops used for every two lines 
                       # as a hacky way to ensure that the lines are closer together
                       na.value = 'white', 
                       breaks = breaks) +
  #Then alter theme as hacky way to make text size back to normal and increase margin so the empty line will fit
  theme(legend.title = element_text(size = 15,
                                    margin = margin(t = -2, 
                                                    r = 0, 
                                                    b = 0, 
                                                    l = 0,
                                                    unit = "mm")))

p
```

### 2021-01-11 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Add files via upload

---
title: "P.berghei pir analysis"
author: "Timothy Little"
date: "05/01/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)
# Set below to your own working directory.
knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/Pberghei")
options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
# getwd()

```

```{r load_packages, include = FALSE, message = FALSE}
library('readxl')
library('readr')
library('dplyr')
library('purrr')
library('ggplot2')
library('DESeq2')
library("RColorBrewer")
library('colorspace')
library('data.table')
library('gridExtra')
library('tibble')
library('scales')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('plotly')
library('ggpattern')
library('ComplexHeatmap')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select

```

```{r function_mround}
mround <- function(x,base){ 
  base*ceiling(x/base) 
} 
```

```{r function_my_max}
my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
```

```{r function_uniquify}
uniquify <- function(vector_names, first_one = TRUE){
  new_vector <- c()
  df_num <- data.frame(names = unique(vector_names), rep = 1)
  if(first_one){
    duplicated_names <- unique(vector_names[duplicated(vector_names)])
  }else{
    duplicated_names <- c()
  }
  for(i in 1:length(vector_names)){
    if(vector_names[i] %in% duplicated_names|!first_one){
    name <- paste(vector_names[i],
                  df_num$rep[match(vector_names[i], df_num$names)],
                  sep = '_')
    #Increase rep counter
    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
    new_vector[i] <- name
    }else{
      new_vector[i] <- vector_names[i]
    }
  }
  new_vector
}
```

```{r function_firstup}
#Capitalize first letter
firstup <- function(x) {
  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
  x
}
```

```{r function_ordering}
ordering = function(to_order, order_vector){
  order = unlist(sapply(order_vector, 
                        function(x){unique(str_subset(to_order, 
                                                      pattern = paste(x)))}))
  order_missing = unique(to_order[!to_order %in% order])
  unique(c(order, order_missing), fromLast = TRUE)
}

order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
                 'Liv-2h','Liv-4h','Liv-12h', 'Liv-18h',
                 'Liv-24h' ,'Liv-36h','48h','54h','60h',
                 'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')
r_order_vector <- str_replace_all(order_vector, 
                                  pattern = '-', 
                                  replacement = '\\.')
```

```{r tpm, results='hide'}

# We need to write a function to calculate the TPM from the count data we have generated. This function uses the _lengths.transcript_ object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.

tpm = function(count_data,lengths) {

  #Print an error if the id columns do not match
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'id', 
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'id', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      print('Warning:Count data and length gene ids DO NOT MATCH')
    }
  } else {
    print('Count data and/or length table supplied do not have an id column, make sure they are ordered correctly')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'id',ignore.case = T)) == T){
    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'id', ignore.case = T)]
  } else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
  }else{
    lengths.table <- lengths}

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j,]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  if(any(grepl(colnames(count_data), pattern = 'id', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'id', ignore.case = T), drop = FALSE], 
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

```{r pir/fam_id, message=FALSE, results='hide'}

# Getting the pir/fam gene names and transcript lengths needed for calculating the transcript-per-million

# First we need to get the _pir gene names. These have been downloaded using a search of PlasmoDB P. berghei ANKA genome v3 using  "\*pir\*" search and manual removal of genes which clearly were not pir.

allgene.info.file <- '../standard_files/PlasmoDb-47_beta_Pberghei_GenesByTaxon_Summary.csv'
allgene.info <- as_tibble(
  read_csv(allgene.info.file)
) %>%
  mutate(Geneid = `Gene ID`, 
         Description = `Product Description`,
         is_pseudo = `Is Pseudo`,
         transcript_length = `Transcript Length`,
         paralog_count = `Paralog count`) %>% 
  select(Geneid, 
         Description, 
         is_pseudo, 
         transcript_length, 
         paralog_count)

pir_info <- allgene.info %>% 
  filter(str_detect(Description, 
                    pattern = 'PIR(?!.+pseud)'), 
         is_pseudo == 'No') %>% 
  filter(!Geneid %in% 'PBANKA_0524600')
#Removing the pir-like gene.

pir_id <- pir_info$Geneid

#pir

pir_genes_list <- lapply(
  unique(pir_info$Description), 
  function(pir_group){
    select(
      filter(
        pir_info, 
        Description %in% pir_group
      ),
      Geneid
    )
  }
)
names(pir_genes_list) <- unique(pir_info$Description)

#How many pirs are present on each chromosome?

pir_tbl = as_tibble(data.frame(Geneid = pir_id))
pir_chrom_tbl = pir_tbl %>% 
  mutate(chrom = unlist( #Getting the chromosomes by a complicated set of lapply and sapplys...
    sapply(
      lapply(
        sapply(
          as.character(pir_tbl$Geneid), 
          strsplit,
          split = ''
        ), 
        '[', 
        8:9
      ), 
      paste, 
      collapse = ''
    )
  )
  ) 
pir_chrom_tbl %>% 
  group_by(chrom) %>% 
  dplyr::summarise(number_of_pirs_in_each_chromosome = n(),
                   percentage_of_all_pirs = round(n()/nrow(pir_tbl), 3)*100) %>% 
  dplyr::arrange(desc(percentage_of_all_pirs))

#Getting the loci information
pir_loci <- data.frame(
  Geneid = pir_id,
  chrom = str_extract(pir_id, 
                      pattern = '(?<=_)[[:digit:]]{2}'),
  number = as.numeric(str_extract(pir_id, 
                                  pattern = '(?<=_[[:digit:]]{2})[[:digit:]]+'))
) %>% 
  filter(!chrom %in% '00') %>% #Remove chrom00 (contigs)
  group_by(chrom) %>% 
  mutate(jump = ifelse(lag(number, 
                           default = 1001) < 1000 & number - lag(number, 
                                                                 default = 0) > 1000,
                       1,
                       0)) %>% 
  mutate(
    jump = ifelse(cumsum(jump) == 1 | jump == 1, 1, 0)
  ) %>% 
  mutate(
    locus = paste0(
      chrom,
      ifelse(jump == 0 & number < 1000, 
             'Left',
             'Right')
    )
  )

# Detect 'jumps' of the gene number over 1000 within each chromosome, setting the lag default as 1001 in the
#   first part of the ifelse statement so that chrom2 doesn't have an erroneous jump as it starts at >1000
# Later cumsum is do that every gene after the 'jump' is included allowing us to define telomere ends.

pir_clade <- mutate(pir_clade, 
                    locus = pir_loci$locus[match(Geneid, pir_loci$Geneid)])

pir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
                        delim = '\t')
pir_id <- pir_clade$Geneid

pir_clade_list <- lapply(unique(pir_clade$clade), function(clade_name){
  pir_clade$Geneid[pir_clade$clade %in% clade_name]
})
names(pir_clade_list) <- as.character(unique(pir_clade$clade))
```

```{r transcript_redundancy, message = FALSE, results='hide'}

#Note that there is some redundancy
two.transcript <- allgene.info %>% 
  dplyr::count(Geneid) %>% 
  filter(n > 1)

#Nine genes have two transcripts - need to give them one average of the transcripts since featurecounts has only
# calculated one count for the locus
two.transcript.info <- allgene.info[allgene.info$Geneid %in% two.transcript$Geneid,]
two.transcript.mean = aggregate(transcript_length~Geneid, data=two.transcript.info, 
                                FUN=function(x) c(mean=mean(x)))

#Remove the first match
allgene.info <- allgene.info[-(match(two.transcript.mean$Geneid, 
                                     allgene.info$Geneid)),]

#For the second match replace the transcript value with the mean calculated with aggregate
allgene.info$transcript_length <- as.double(allgene.info$transcript_length)
allgene.info[match(two.transcript.mean$Geneid, 
                   allgene.info$Geneid),]$transcript_length <- two.transcript.mean$transcript_length
lengths.transcript <- subset(allgene.info, select = c(Geneid, transcript_length), drop = F)
```

## Read in counts and calculate TPM

```{r read_files}
#Reading in the counts from file
berghei_counts <- read_xlsx('Manuscript_Frontiers_berghei_counts.xlsx')

berghei_tpm <- tpm(berghei_counts,
                   lengths = lengths.transcript)

apply(berghei_tpm[-1], 2, sum)

#Get a list of df for each experiment.

expt_nams <- unique(
  na.omit(
    str_extract(colnames(berghei_tpm), 
                pattern = 'E[[:digit:]]{2}')
  )
)

expt_tpms <- lapply(
  expt_nams,
  function(expt_nam){
    expt_df <- select(
      berghei_tpm,
      matches(paste0('Geneid|', expt_nam))
    ) %>% 
      rename_with(~ str_remove(.x, 'E[[:digit:]]{2}_'),
                  matches('E[[:digit:]]{2}'))
    list(tpm = expt_df,
         experiment = expt_nam)
  }
)
names(expt_tpms) <- expt_nams

#Remove the Caldelari 6h samples, as they did in their paper.

expt_tpms$E06$tpm <- select_at(expt_tpms$E06$tpm,
                                          vars(-matches('Liv-6h')))
```

Going to combine certain samples based on their distinctive profiles, e.g. <4h, 6-48h, >48h liver stages.

```{r combining_samples}
names_expt_tpms <- names(expt_tpms)

expt_tpms <- lapply(names(expt_tpms), function(expt){
  col_stage <- colnames(
    getElement(
      getElement(
        expt_tpms, 
        paste(expt)
      ),
      'tpm'
    )
  )

  liver_stage <- col_stage[
    str_detect(
      col_stage,
      pattern = 'Liv-[[:digit:]]{1,2}h'
    )
    ]

  new_liver_stage <- case_when(
    str_detect(liver_stage, pattern = 'Liv-(2|4)h') ~ 'Liv-pre4h',
    # str_detect(liver_stage, pattern = 'Liv-([5-9]|[1-3][0-9]|4[0-7])h') ~ 'Liv-pre47h',
    str_detect(liver_stage, pattern = 'Liv-(4[8-9]|[5-9][0-9])h') ~ 'Liv-post48h',
    TRUE ~ as.character(liver_stage),
  )

  if(length(new_liver_stage) > 0){
    new_liver_stage <-uniquify(new_liver_stage)
  }

  col_stage[
    str_detect(
      col_stage,
      pattern = 'Liv-[[:digit:]]{1,2}h'
    )
    ] <- new_liver_stage

  bloodmeal_stage <- col_stage[
    str_detect(
      col_stage, 
      pattern = 'BldMl-(1h|3h)_[[:digit:]]+'
    )
    ]

  bloodmeal_stage <- str_replace(
    bloodmeal_stage, 
    pattern = '(1h|3h)_[[:digit:]]+', 
    replacement = 'Pre3h'
  )
  if(length(bloodmeal_stage)>0){
    bloodmeal_stage <- uniquify(bloodmeal_stage)
  }

  col_stage[
    str_detect(
      col_stage, 
      pattern = 'BldMl-(1h|3h)_[[:digit:]]+'
    )
    ] <- bloodmeal_stage

  tpm_element <- getElement(
    getElement(
      expt_tpms, 
      paste(expt)
    ),
    'tpm'
  )

  colnames(tpm_element) <- col_stage

  list(tpm = tpm_element,
       experiment =   getElement(
         getElement(
           expt_tpms,
           paste(expt)
         ),
         'experiment'
       )
  )

})
names(expt_tpms) <- names_expt_tpms

#Re-ordering the list
expt_tpms <- (expt_tpms[order(names(expt_tpms))])
```

```{r hclusts, message = FALSE}

hclusts = function(tpm){
  if(ncol(tpm$tpm) > 3){
    dists = dist(t(tpm$tpm[,-1]))
    par(cex=0.7)
    plot(hclust(dists),
         main = tpm$experiment)
  }
}

lapply(expt_tpms, hclusts)
```

```{r remove_badly_clustering_samples}
expt_tpms$E08$tpm$`Liv-pre4h_1` <- NULL
expt_tpms$E03$tpm$`Asex-Schiz_1` <- NULL
expt_tpms$E03$tpm$`Asex-Schiz_3` <- NULL
```

## Supp figure - PCA of all the samples.

```{r supp_figure_PCA_allgenes}

big_tpm <- bind_cols(
  lapply(
    expt_tpms,
    function(x){
      tpm_df <- x$tpm
      colnames(tpm_df)[-1] <- paste0(x$experiment,
                                     '_',
                                     colnames(tpm_df)[-1])
      tpm_df
    }
  )
) %>% 
  dplyr::rename(Geneid = Geneid...1) %>% 
  select_at(vars(-matches('Geneid\\.\\.\\.\\d+$'))) %>% 
  data.frame

rownames(big_tpm) <- big_tpm$Geneid

big_tpm <- big_tpm[,-1,drop=FALSE]

big_tpm_pca <-  prcomp(big_tpm, 
                       center = TRUE, 
                       scale. = TRUE)

big_tpm_pca_var <- get_pca_var(big_tpm_pca)
big_tpm_pca_coords <- data.frame(sample = rownames(big_tpm_pca_var$coord),
                                 big_tpm_pca_var$coord)
big_tpm_pca_coords$stage_experiment <- str_extract(
  big_tpm_pca_coords$sample,
  pattern = '.+(?=_[[:digit:]]+$)'
)
big_tpm_pca_coords$stage <- str_extract(big_tpm_pca_coords$sample, 
                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
big_tpm_pca_coords$experiment <- str_extract(big_tpm_pca_coords$sample, 
                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')

colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
                               length.out = length(unique(big_tpm_pca_coords$stage))),
                       unique(big_tpm_pca_coords$stage))

pca_importance <- as.data.frame(summary(big_tpm_pca)$importance)

plot <- ggplot(big_tpm_pca_coords,
       aes(
         x = Dim.1,
         y = Dim.2,
         col = stage,
         shape = experiment
       )) +
  geom_point() +
  scale_fill_manual(values = colornames) +
  scale_shape_manual(
    values = setNames(
      c(1:10),
      unique(big_tpm_pca_coords$experiment)
    )
  ) +
  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
  guides(shape=guide_legend(ncol=2),
         col = guide_legend(ncol=2)) +
  theme_classic() +
  theme(legend.key.size = unit(5, 'mm'),
        legend.text = element_text(size = 6))

plot

```

```{r supp_figure_PCA_pir}
big_tpm_pir <- big_tpm[rownames(big_tpm) %in% pir_id,,drop = FALSE]

big_tpm_pir_pca <-  prcomp(big_tpm_pir, 
                       center = TRUE, 
                       scale. = TRUE)

big_tpm_pir_pca_var <- get_pca_var(big_tpm_pir_pca)
big_tpm_pir_pca_coords <- data.frame(sample = rownames(big_tpm_pir_pca_var$coord),
                                 big_tpm_pir_pca_var$coord)
big_tpm_pir_pca_coords$stage_experiment <- str_extract(
  big_tpm_pir_pca_coords$sample,
  pattern = '.+(?=_[[:digit:]]+$)'
)
big_tpm_pir_pca_coords$stage <- str_extract(big_tpm_pir_pca_coords$sample, 
                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
big_tpm_pir_pca_coords$experiment <- str_extract(big_tpm_pir_pca_coords$sample, 
                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')

colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
                               length.out = length(unique(big_tpm_pir_pca_coords$stage))),
                       unique(big_tpm_pir_pca_coords$stage))

pca_importance <- as.data.frame(summary(big_tpm_pir_pca)$importance)

plot_pir <- ggplot(big_tpm_pir_pca_coords,
       aes(
         x = Dim.1,
         y = Dim.2,
         col = stage,
         shape = experiment
       )) +
  geom_point() +
  scale_fill_manual(values = colornames) +
  scale_shape_manual(
    values = setNames(
      c(1:10),
      unique(big_tpm_pca_coords$experiment)
    )
  ) +
  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
  guides(shape=guide_legend(ncol=2),
         col = guide_legend(ncol=2)) +
  theme_classic() +
  theme(legend.key.size = unit(5, 'mm'),
        legend.text = element_text(size = 6))

plot_pir
```

How many experiments cover each life stage?

```{r experiment_per_life_stage}

big_tpm %>% 
  rownames_to_column(var = 'Geneid') %>% 
  melt(id.vars = 'Geneid') %>% 
  mutate(
    stage = str_extract(variable, pattern = '(?<=E[:digit:]{2}_).+(?=_[:digit:])'),
    expt = str_extract(variable, pattern = 'E[:digit:]{2}')
  ) %>% 
  group_by(stage) %>% 
    summarise(num_expt = n_distinct(expt)) %>% 
    arrange(desc(num_expt))

```

## The highest expressed _pir_ genes across the samples

```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
top_n_genes <- function(tpm_input, 
                       n = 5, 
                       threshold = 0){
  tpm = tpm_input$tpm %>% 
    #Melt the data frame for using dplyr
    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
    mutate(stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1))) %>% 
    group_by(Geneid, stage) %>%  
    dplyr::summarise(median_tpm = median(tpm),
                     mean_tpm = mean(tpm), 
                     min_tpm = min(tpm), 
                     max_tpm = max(tpm), 
                     sd_tpm = sd(tpm)) %>% 
    group_by(stage) %>% 
    mutate(experiment = tpm_input$experiment)

  #Calculating the index used to make the kable table

  kable_index = rep(n, times = length(unique(tpm$stage)))
  names(kable_index) = unique(tpm$stage)

  all_genes = tpm %>% 
    top_n(wt = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue")))
  all_genes$short_descr = allgene.info$handy_descriptions[match(all_genes$Geneid, 
                                                                allgene.info$Geneid)]
  all_genes$description = allgene.info$Description[match(all_genes$Geneid, 
                                                         allgene.info$Geneid)]
  all_genes = all_genes %>% 
    kable(format = 'html', 
          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed total genes'), 
          escape = FALSE) %>%
    kable_styling() %>% 
    pack_rows(index = kable_index)

  pir_genes = tpm %>% 
    filter(Geneid %in% pir_id) %>% 
    top_n(wt = mean_tpm, n = n) %>% 
    dplyr::arrange(stage, desc(mean_tpm)) %>% 
    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
                                color = ifelse(mean_tpm > threshold, "red", "blue"))) %>% 
    mutate(Geneid = as.character(Geneid)) %>% 
    mutate(Description = allgene.info$handy_descriptions[match(Geneid, 
                                                               allgene.info$Geneid)]) %>% 
    mutate(Geneid = ifelse(grepl(Geneid,
                                 pattern = 'PBANKA_0100500'),
                           cell_spec(Geneid, bold = TRUE),
                           Geneid)) %>%
    kable(format = 'html', 
          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed pir genes'), 
          escape = FALSE) %>% 
    pack_rows(index = kable_index) %>% 
    kable_styling()

  # print(paste0('Completed analysis of ', tpm_input$experiment))
  list = list(tpm, all_genes, pir_genes, tpm_input$experiment)
  names(list) = c('tpm', 'all_genes', 'pir_genes', 'experiment')
  return(list)
}

### Generate and save top genes - core dataset all_top

all_top <- lapply(expt_tpms, 
                  top_n_genes)

###

### Read in top genes

###

lapply(all_top, `[`, c('all_genes', 'pir_genes'))

```

## Pearson correlation plot of collated samples.

```{r all_samples_collated_heatmap}
## Collated across the stages for a more easily understood heatmap

big_top = bind_rows(lapply(all_top,function(x){
  data = x$tpm
  data$stage = paste(x$experiment, data$stage, sep = '_')
  data
}))

big_top_recast = dcast(big_top, Geneid ~ stage, value.var = 'mean_tpm')
big_top_recast <- column_to_rownames(big_top_recast, var = 'Geneid')
# big_top_recast = big_top_recast[,-1,drop=FALSE]

# apply(big_top_recast, 2, sum)

corrplot(cor(big_top_recast), method = 'color', 
         order = 'hclust', hclust.method = 'ward.D2',
         cl.lim = c(-0.036,1), tl.cex = 0.50,
         col = c(viridis(100), magma(115)), addrect = 6)
```

```{r combining_stages_data}
combined_data <- big_tpm %>% 
  rownames_to_column(var = 'Geneid') %>% 
  melt(value.name = 'tpm') %>% 
  mutate(stage = str_extract(variable, 
                             pattern = '(?<=_).+(?=_)')) %>% 
  group_by(Geneid, stage) %>% 
  dplyr::summarise(tpm_mean = mean(tpm),
                   tpm_med = median(tpm),
                   tpm_sd = sd(tpm),
                   tpm_max = max(tpm),
                   tpm_min = min(tpm))

combined_data <- dcast(combined_data, 
                       Geneid ~ stage, 
                       value.var = 'tpm_mean') %>% 
  column_to_rownames(var = 'Geneid')

#Ordering by life cycle progression
combined_data <- combined_data[,
                               ordering(colnames(combined_data), 
                                        order_vector = gsub(order_vector,
                                                            pattern = '-',
                                                            replacement = '\\.')
                               )]

#Check that they still sum to 1e6

apply(combined_data, 2, sum)

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9',
        '#E69F00'),
  sort(
    unique(pir_clade %>% 
             mutate(
               SorL = str_extract(clade, 
                                  pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent')
             ) %>% 
             .$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(pir_clade$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9',
    '#009E73',
    '#0072B2',
    '#E69F00',
    '#D55E00',
    '#CC79A7',
    '#999999',
    'black',
    'red'),
  sort(
   unique(pir_clade$clade)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(pir_clade$clade == clade_name),
                                   ')'
                                 )
                               })
```

# pir expression, S or L-form, stacked bar chart, stack for each gene

```{r SorL_stacked_eachGene_fig1b}
  tpm_data <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = Geneid) %>% 
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

tpm_data <- mutate(tpm_data,
                   order = row_number())

tpm_data_SorL_stackgenes <- mutate(tpm_data,
                        SorL = str_extract(clade, 
                                           pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent'))
tpm_data_SorL_stackgenes <- group_by(tpm_data_SorL_stackgenes,
                                     stage, 
                                     SorL,
                                     stack_group) %>% 
  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
  #Include the number of genes in each fill category for the labels
  mutate(SorL = paste0(SorL, 
                       ' (n = ', 
                       n_distinct(stack_group),
                       ')')) %>% 
#Order by SorL and then by total_tpm, then convert stack_group to a factor so ggplot orders correctly
  arrange(SorL, 
          total_tpm) %>% 
  mutate(stage_num = as.integer(stage))
tpm_data_SorL_stackgenes$stage %>% as.integer %>% unique
tpm_data_SorL_stackgenes$stage %>% unique

tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))
```

```{r stacked_bar_chart_figure2b}
#Calculate the range for the y value to make it look nicer
stacked_max <- tpm_data %>% 
  ungroup %>% 
  group_by(stage) %>% 
  dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
  .$max_tpm %>% 
  max

max_tpm <- c(
  seq(
    0,
    mround(
      stacked_max,
      500),
    length.out = 11)
)

stack_label <- 'sub-family'

write_csv(x = tpm_data_SorL_stackgenes,
          path = 'tl20-05_SorL-subfamily_pir.csv')

stage_vec <- unique(
  levels(tpm_data$stage)
)

stage_vec <- tpm_data_SorL_stackgenes$stage
col_group <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'zzAsex' ,
                   str_detect(stage_vec, pattern = 'Liv') ~ 'zzLiv',
                   str_detect(stage_vec, pattern = 'Gam') ~ 'zzGam',
                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'zzMos')
tpm_data_SorL_stackgenes$stage_col <- col_group
fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') #Note that this order is influenced by sort(col_group)

y <- ggplot(data = tpm_data_SorL_stackgenes, 
            aes(
              x = stage_num,
              y = total_tpm, 
              fill = SorL,
              group = stage_num
            )) +
  geom_rect(
    aes(
      xmin = stage_num - .7,
      xmax = stage_num + .7,
      ymin = -Inf,
      ymax = Inf,
      fill = stage_col
    ),
    alpha = .05
  ) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) +
  scale_fill_manual(
    breaks = unique(tpm_data_SorL_stackgenes$SorL),
    values = c(
      SorL_col_num,
      fills
    )
  )+
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  # scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) +
  scale_x_continuous(
    breaks = tpm_data_SorL_stackgenes$stage_num %>% unique(),
    labels = tpm_data_SorL_stackgenes$stage %>% unique(),
    expand = c(0, 0)
  ) +
    theme(
      axis.text.x = element_text(angle = 90, 
                                 vjust = .5,
                                 size = 10),
      axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    ) +
    xlab('stage') + 
    facet_grid(.~stage, 
               scales = "free_x",
               space = "free_x") +
    scale_x_discrete(expand=c(0,0)) +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_blank(),
      axis.ticks.x = element_blank(),
      # axis.line.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.title.y = element_text(size = 10),
      legend.title = element_text(size = 12),
      legend.text = element_text(size = 10),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=8, 
                                  face = 'bold', 
                                  angle = 90),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      ),
      panel.spacing.x = unit(0, "null")
    )

g <- ggplot_gtable(ggplot_build(y))
strip_t <- which(grepl('strip-t', g$layout$name))
stage_vec <- unique(
  levels(tpm_data$stage)
)
fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
k <- 1
for (i in strip_t) {
  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
  k <- k+1
}

grid.newpage()
grid.draw(g)

```

# figure2binset

```{r stack_bar_graph_SorL_lowexpression-figure2binset}
#Include all stages expressed lower than a certain amount

low_express <- tpm_data_SorL_stackgenes %>% 
  group_by(stage) %>% 
  summarise(sum_tpm = sum(total_tpm)) %>% 
  filter(sum_tpm < 350)

#Calculate the range for the y value to make it look nicer

stacked_max <- max(low_express$sum_tpm)

max_tpm <- c(
  seq(
    0,
    mround(
      stacked_max,
      100),
    length.out = 11)
)

tpm_data_SorL_stackgenes_low <- filter(tpm_data_SorL_stackgenes,
                                       stage %in% low_express$stage)

y <- ggplot(data = tpm_data_SorL_stackgenes_low, 
            aes(
              x = stage,
              y = total_tpm, 
              fill = SorL,
              group = stack_group_order
            )) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) +
  scale_fill_manual(
    breaks = unique(tpm_data_SorL_stackgenes_low$SorL),
    values = SorL_col_num
  ) +
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 15),
    axis.title.y = element_text(size = 8),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 5),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=6, 
                                face = 'bold', 
                                angle = 90),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )

  #This bit is to change the colour of the facet boxes

  g <- ggplot_gtable(ggplot_build(y))
  strip_t <- which(grepl('strip-t', g$layout$name))
  stage_vec <- unique(
  levels(tpm_data_SorL_stackgenes_low$stage)[levels(tpm_data_SorL_stackgenes_low$stage) %in% as.character(tpm_data_SorL_stackgenes_low$stage)]
  )
  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
  k <- 1
  for (i in strip_t) {
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }
  grid.newpage()
  grid.draw(g)
```

```{r table_of_SorL_highest_supptable}
#For a neat table
SorL_kable_index <- rep.int(2, 
                            times = length(unique(tpm_data_SorL_stackgenes$stage)))
names(SorL_kable_index) <- levels(tpm_data_SorL_stackgenes$stage)

tpm_data_SorL_stackgenes_top_table <- tpm_data_SorL_stackgenes %>% 
  mutate(Geneid = stack_group) %>% 
  filter(str_detect(SorL, 
                    pattern = 'ancestral', 
                    negate = TRUE)) %>% 
  group_by(stage) %>% 
  mutate(sum_tpm = sum(total_tpm)) %>% 
  group_by(stage, 
           SorL) %>% 
  slice_max(order_by = total_tpm,
            n = 1) %>% 
  mutate(total_tpm = round(total_tpm, digits = 2),
         prop_tpm = round(total_tpm/sum_tpm, digits = 2)) %>% 
  select(stage, Geneid, SorL, total_tpm, prop_tpm)

tpm_data_SorL_stackgenes_top_table %>% 
  kable() %>% 
  kable_styling() %>% 
  pack_rows(index = SorL_kable_index) %>% 
  scroll_box(width = "100%", height = "500px")
```

## Figure 1C

```{r combined_tpm_stacked_df}

  tpm_data_subfam <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    filter(Geneid %in% pir_id) %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(clade = pir_clade$clade[match(Geneid, 
                                       pir_clade$Geneid)]
         ) %>% 
  group_by(clade) %>% 
  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
    mutate(stack_group = clade) %>% 
    group_by(stage, 
             stack_group,
             clade) %>% 
    dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(clade),
          total_tpm
)

#Separate out the pan-rodent pir
# tpm_data_subfam$clade[tpm_data_subfam$stack_group == 'PBANKA_0524600'] <- 'pan-rodent'

tpm_data_subfam <- mutate(tpm_data_subfam,
                   order = row_number())

  # #Apply ordering
  # tpm_data_subfam$stage = factor(tpm_data_subfam$stage, 
  #                         levels = ordering(tpm_data_subfam$stage,
  #                                           order_vector))

  #Calculate the range for the y value to make it look nicer
  stacked_max <- tpm_data_subfam %>% 
    ungroup %>% 
    group_by(stage) %>% 
    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
    .$max_tpm %>% 
    max

  max_tpm <- c(
    seq(
      0,
      mround(
        stacked_max,
        500),
      length.out = 11)
  )
```

```{r combined_tpm_stacked_plot}

  stack_label <- 'sub-family'

y <- ggplot(data = tpm_data_subfam, 
            aes(
              x = stage,
              y = total_tpm, 
              fill = clade,
              group = order
            )) +
  geom_bar(
    # position = position_stack(),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1) +
  scale_fill_manual(
    values = clade_col_num
  ) +
  scale_color_manual('black') + 
  labs(fill = stack_label) +
  scale_y_continuous(breaks = seq(0,1,0.1)) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('Proportion of '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~stage, 
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=8, 
                                face = 'bold', 
                                angle = 90),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )

  #This bit is to change the colour of the facet boxes

  g <- ggplot_gtable(ggplot_build(y))
  strip_t <- which(grepl('strip-t', g$layout$name))
  stage_vec <- unique(

    levels(tpm_data_subfam$stage)
  )
  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
  k <- 1
  for (i in strip_t) {
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }

  grid.newpage()
  grid.draw(g)
```

## Chromosome proportional bar chart supp fig

```{r chrom_prop_bar_chart}
tpm_data_chrom <- combined_data %>% 
  rownames_to_column(var = 'Geneid') %>% 
  filter(Geneid %in% pir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'stage', 
       value.name = 'tpm') %>% 
  mutate(
    chrom = str_extract(Geneid, pattern = '(?<=PBANKA_)[[:digit:]]{2}')
  ) %>% 
  group_by(chrom) %>% 
  mutate(chrom = paste0(chrom, ' (n = ', n_distinct(Geneid) ,')')) %>% 
  group_by(stage, 
           chrom) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(chrom),
          total_tpm
) %>% 
  mutate(chrom = factor(chrom, 
                        levels = unique(chrom)))

  stacked_max <- tpm_data_chrom %>% 
    ungroup %>% 
    group_by(stage) %>% 
    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
    .$max_tpm %>% 
    max

  max_tpm <- c(
    seq(
      0,
      mround(
        stacked_max,
        500),
      length.out = 11)
  )

y <- ggplot(data = tpm_data_chrom) +
  geom_bar_pattern(
    aes(
      x = stage,
      y = total_tpm, 
      fill = chrom,
      group = stage,
      pattern = chrom
    ),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1,
    pattern_density = 0.2,
    pattern_spacing = 0.05,
    pattern_key_scale_factor = 0.25
  ) +
  scale_pattern_manual(
    values = rep(c('none', 
                   'circle'), 
                 each = length(c(cbPalette, 
                                 'red', 
                                 'white')), 
                 length.out = length(unique(tpm_data_chrom$chrom))),
    labels = unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_fill_manual(
    breaks = unique(tpm_data_chrom$chrom),
    values = rep_len(c(cbPalette, 
                       'red', 
                       'white'),
                     length.out = length(unique(tpm_data_chrom$chrom))),
    labels =  unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_color_manual('black') +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) +
  facet_grid(.~stage,
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme_classic() +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=8,
                                face = 'bold',
                                angle = 90),
    strip.background = element_rect(
      color="black",
      size=0.1,
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  )

  #This bit is to change the colour of the facet boxes

  g <- ggplot_gtable(ggplot_build(y))
  strip_t <- which(grepl('strip-t', g$layout$name))
  stage_vec <- unique(

    levels(tpm_data$stage)
  )
  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
  k <- 1
  for (i in strip_t) {
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }

  grid.newpage()
  grid.draw(g)

```

## Expression by locus. Proportion. Supp figure.

Loci are defined by L or R of each chromosome number, will define this by the jump in the gene numbers.

```{r defining_loci}

tpm_data_locus <- combined_data %>% 
  rownames_to_column(var = 'Geneid') %>% 
  filter(Geneid %in% pir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'stage', 
       value.name = 'tpm') %>% 
  mutate(
    locus = pir_loci$locus[match(Geneid, pir_loci$Geneid)]
  ) %>% 
  filter(!is.na(locus)) %>% 
  group_by(locus) %>% 
  mutate(locus = paste0(locus, ' (n = ', n_distinct(Geneid) ,')')) %>% 
  group_by(stage, 
           locus) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>%
  arrange(
          desc(locus),
          total_tpm
) %>% 
  mutate(locus = factor(locus, levels = unique(locus)))

  stacked_max <- tpm_data_locus %>% 
    ungroup %>% 
    group_by(stage) %>% 
    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
    .$max_tpm %>% 
    max

  max_tpm <- c(
    seq(
      0,
      mround(
        stacked_max,
        500),
      length.out = 11)
  )

(y <- ggplot(data = tpm_data_locus) +
  geom_bar_pattern(
    aes(
      x = stage,
      y = total_tpm, 
      fill = locus,
      group = stage,
      pattern = locus
    ),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1,
        pattern_density = 0.05,
    pattern_spacing = 0.1,
    pattern_key_scale_factor = 0.1
    #  pattern_density = 0.1,
    # pattern_spacing = 0.1,
    # pattern_key_scale_factor = 0.5
  ) +
  scale_pattern_manual(
    values = rep(c('none', 
                   'circle',
                   'stripe'), 
                 each = length(c(cbPalette, 
                                 'red', 
                                 'white')), 
                 length.out = length(unique(tpm_data_locus$locus))),
    labels = unique(tpm_data_locus$locus),
    guide = guide_legend(reverse = TRUE),
    name = 'locus'
  ) +
  scale_fill_manual(
    breaks = unique(tpm_data_locus$locus),
    values = rep_len(c(cbPalette, 
                       'red', 
                       'white'),
                     length.out = length(unique(tpm_data_locus$locus))),
    labels =  unique(tpm_data_locus$locus),
    guide = guide_legend(reverse = TRUE),
    name = 'locus'
  ) +
  scale_color_manual('black') +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) +
  facet_grid(.~stage,
             scales = "free_x",
             space = "free_x") +
  scale_x_discrete(expand=expand_scale(add=1)) +
  theme_classic() +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x = element_blank(),
    # axis.line.x = element_blank(),
    axis.text.y = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    legend.title = element_blank(),
    legend.text = element_text(size = 8),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=8,
                                face = 'bold',
                                angle = 90),
    strip.background = element_rect(
      color="black",
      size=0.1,
      linetype="solid"
    ),
    panel.spacing.x = unit(0, "null")
  ))

  #This bit is to change the colour of the facet boxes

  g <- ggplot_gtable(ggplot_build(y))
  strip_t <- which(grepl('strip-t', g$layout$name))
  stage_vec <- unique(

    levels(tpm_data$stage)
  )
  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
  k <- 1
  for (i in strip_t) {
    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
    k <- k+1
  }

  grid.newpage()
  grid.draw(g)
```
## Gametocyte marker gene expression - supp figure

```{r}
list_genes <- c(
  list('P28' = 'PBANKA_0514900'),
  list('nek4' = 'PBANKA_0616700'),
  list('map2' = 'PBANKA_0933700'),
  list('hap2' = 'PBANKA_1212600')
)
gam_markers <- unlist(list_genes)

list_tpms_line <- lapply(expt_tpms, 
                         function(input){
                           tpm_data <- input$tpm %>% 
                             filter(Geneid %in% gam_markers) %>% 
                             mutate(Geneid = names(gam_markers)[match(Geneid, gam_markers)]) %>% 
                             melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
                             mutate(stage = unlist(lapply(strsplit(as.character(sample), 
                                                                   split = '_'),
                                                          `[`,
                                                          1))) %>% 
                             group_by(stage, 
                                      sample,
                                      Geneid) %>% 
                             dplyr::summarise(total_tpm = sum(tpm))

                           #Make stages with no cirs in the sample give a zero 
                           if(!all(colnames(input$tpm) %in% tpm_data$sample)){
                             missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
                             missing_stages_data_zero = data.frame(lapply(missing_stages, function(missing_st){
                               data.frame(stage = unlist(lapply(strsplit(as.character(missing_st), split = '_'),`[`,2)),
                                          sample = missing_st,
                                          total_tpm = 0)
                             }))
                             tpm_data <- bind_rows(tpm_data, missing_stages_data_zero)
                           }

                           #Add experiment column
                           tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
                           max_tpm = if(max(tpm_data$total_tpm) < 10){
                             c(seq(0,mround(max(tpm_data$total_tpm),1),1))
                           }else{
                             c(seq(0,mround(max(tpm_data$total_tpm),5),5))
                           }

                           #Get the average data for the line graph
                           tpm_data_line <- tpm_data %>% 
                             ungroup %>% 
                             group_by(stage, 
                                      experiment,
                                      Geneid) %>% 
                             dplyr::summarise(total_tpm_median = median(total_tpm),
                                              Number_of_replicates = n(),
                                              total_tpm_max = max(total_tpm),
                                              total_tpm_min = min(total_tpm)) %>% 
                             arrange(stage)

                           return(tpm_data_line)
                         })

tpm_data_line <- bind_rows(list_tpms_line)

tpm_data_line$stage <- factor(tpm_data_line$stage, 
                             levels = ordering(tpm_data_line$stage,
                                               order_vector))

#Calculate the range for the y value to make it look nicer
if(max(
  tpm_data_line$total_tpm_max
) >= 500){
  max_tpm = c(
    seq(
      0,
      mround(
        max(
          tpm_data_line$total_tpm_max
        ),
        500),
      length.out = 11)
  )
}else{
  max_tpm = c(
    seq(
      0,
      mround(
        max(
          tpm_data_line$total_tpm_max
        ),
        10),
      length.out = 11)
  )
}
```

super hacky way to change the scales in each facet from:
https://fishandwhistle.net/post/2018/modifying-facet-scales-in-ggplot2/

```{r}
scale_override <- function(which, scale) {
  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
    stop("which must be an integer of length 1")
  }

  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
    stop("scale must be an x or y position scale")
  }

  structure(list(which = which, scale = scale), class = "scale_override")
}

CustomFacetWrap <- ggproto(
  "CustomFacetWrap", FacetWrap,
  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
    # make the initial x, y scales list
    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)

    if(is.null(params$scale_overrides)) return(scales)

    max_scale_x <- length(scales$x)
    max_scale_y <- length(scales$y)

    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
    for(scale_override in params$scale_overrides) {
      which <- scale_override$which
      scale <- scale_override$scale

      if("x" %in% scale$aesthetics) {
        if(!is.null(scales$x)) {
          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
          scales$x[[which]] <- scale$clone()
        }
      } else if("y" %in% scale$aesthetics) {
        if(!is.null(scales$y)) {
          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
          scales$y[[which]] <- scale$clone()
        }
      } else {
        stop("Invalid scale")
      }
    }

    # return scales
    scales
  }
)

facet_wrap_custom <- function(..., scale_overrides = NULL) {
  # take advantage of the sanitizing that happens in facet_wrap
  facet_super <- facet_wrap(...)

  # sanitize scale overrides
  if(inherits(scale_overrides, "scale_override")) {
    scale_overrides <- list(scale_overrides)
  } else if(!is.list(scale_overrides) || 
            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) {
    stop("scale_overrides must be a scale_override object or a list of scale_override objects")
  }

  facet_super$params$scale_overrides <- scale_overrides

  ggproto(NULL, CustomFacetWrap,
    shrink = facet_super$shrink,
    params = facet_super$params
  )
}

(
  y <- ggplot(data = tpm_data_line, 
              aes(x = stage, 
                  y = total_tpm_median, 
                  fill = experiment, 
                  group = experiment)) +
    geom_bar(position = position_dodge2(preserve = 'single'),
             stat = 'identity') +
    scale_fill_manual(values = c(cbPalette, 'red', 'purple')[1:(length(unique(tpm_data_line$experiment)))]) +
    geom_errorbar(aes(x = stage,
                      ymax = total_tpm_max, 
                      ymin = total_tpm_min),
                  position = position_dodge2(preserve = 'single'),
                  size = 0.25) +
    scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    theme(axis.title.x=element_blank(),
          axis.text.x=element_text(angle = 90, 
                                   size = 6, 
                                   vjust = 0.5),
          axis.text.y = element_text(size = 5),
          axis.title.y = element_text(size = 8),
          legend.title = element_text(size = 7),
          legend.text = element_text(size = 5),
          legend.key.size = unit(3,'mm'),
          legend.key.width = unit(2,'mm')) +
    scale_size(name = 'Number of genes detected over threshold', 
               range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) +
    facet_wrap_custom(~Geneid, 
                      scales = 'free_y',
                      scale_overrides = list(
                        scale_override(3, scale_y_continuous(breaks = seq(0,650,length.out = 11))),
                        scale_override(2, scale_y_continuous(breaks = seq(0,900,length.out = 11))),
                        scale_override(1, scale_y_continuous(breaks = seq(0,310,length.out = 11)))
                      )) +
    annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)
)
#Could also use functions from a package called 'lemon' to add the lines to the top two facets.
```
```

## Lower expression of _pir_ genes versus the rest of the genome

```{r low_pir_expression}
set.seed(4908)

gene_expression_df <- combined_data %>% 
    rownames_to_column(var = 'Geneid') %>% 
    melt(id.vars = 'Geneid', 
         variable.name = 'stage', 
         value.name = 'tpm') %>% 
  mutate(pir_gene = ifelse(Geneid %in% pir_id,
                           'pir',
                           'non-pir')) %>% 
  group_by(stage, pir_gene) %>% 
  slice_sample(n = length(pir_id)) %>%
  summarise(mean_tpm = mean(tpm),
            med_tpm = round(median(tpm),
                            digits = 6))

g <- ggplot(gene_expression_df,
       aes(x = pir_gene,
           y = log2(1+med_tpm))) +
  xlab('Gene group')+
  ylab('l0g(Median TPM + 1)')+
  # geom_violin()+
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               # binwidth = 1,
               # dotsize = 2,
               method = 'dotdensity') +
  geom_path(aes(group = stage,
                col = stage)) +
  theme_classic() +
  theme(
    legend.text = element_text(size = 9),
    legend.key.size = unit(4, 'mm')
  )

g
```

_pir_ expression on average ranges from minimum mean `r min(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$mean_tpm)` TPM and median `r min(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$med_tpm)` TPM to maximum mean `r max(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$mean_tpm)` TPM and median `r max(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$med_tpm)` TPM. Non-\ _pir_ expression ranges from minimum mean `r min(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$mean_tpm)` TPM and median `r min(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$med_tpm)` TPM to maximum mean `r max(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$mean_tpm)` TPM and median `r max(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$med_tpm)` TPM.

### 2021-01-11 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Add files via upload

---
title: "Manuscript_Github_Pchabaudi_pir_210105"
author: "Timothy Little"
date: "05/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      results = TRUE, 
                      message = FALSE, 
                      warning = FALSE)

knitr::opts_knit$set(root.dir = 
                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/Pchabaudi24h")
#Note that the above doesn't give an error if the dir doesn't exist

options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
```

```{r loading_packages, include = FALSE, message = FALSE}
# Load the required packages

library('readr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('DESeq2')
library("pheatmap")
library("RColorBrewer")
library('data.table')
library('gridExtra')
library('tibble')
library('stringr')
library('kableExtra')
library('viridis')
library('corrplot')
library('reshape2')
library('circlize') # for circular plots and colours
library('cluster')    # clustering algorithms
library('factoextra') # clustering algorithms & visualization
library('ggfortify') # so ggplot understands prcomp objects
library('ape') # for phylogeny plotting
library('dendextend') # for dendrogram plotting
library('readxl') # for reading in Excel spreadsheets
library('plotly')
library('scales')
library('processx') # for downloading the plotly plots
library('cowplot')
library('ComplexHeatmap')
library('ggbiplot')
library('ggpattern')

#Colour blind friendly palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
show_col(cbPalette)

#Want the defaults for arrange, mutate and summarise to be from dplyr
arrange <- dplyr::arrange
mutate <- dplyr::mutate
summarise <- dplyr::summarise
select <- dplyr::select
```

```{r tpm, results = 'hide', message = FALSE}
# # Calculating the transcript-per-million 
# 
# We need to write a function to calculate the TPM from the count data we have generated. This function uses the 'chabaudi_transcript_lengths' object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.

tpm = function(count_data, lengths = chabaudi_transcript_lengths) {

  #Print an error if the id columns do not match
  # first if statement to test that both have a GeneID column
  if(any(grepl(colnames(lengths), 
               pattern = 'id', 
               ignore.case = T)) == T & any(grepl(colnames(count_data), 
                                                  pattern = 'Geneid', 
                                                  #This needs to be Geneid when some of the samples have 'id' in them
                                                  ignore.case = T)) == T){
    if(all(count_data[grep(colnames(count_data), 
                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
                                                                             pattern = 'id', ignore.case = T)])) {
      print('Count data and length gene ids both match')
    } else {
      stop('Error: Count data and length gene ids DO NOT MATCH, please make sure they are ordered identically.')
    }
  } else {
    stop('Error: Count data and/or length table supplied do not have an id column.')
  }

  #Remove the id column from supplied count_data
  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = T)) == T){
    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = T)]
  } else{
    counts_tpm <- count_data}

  #Remove the id column from supplied lengths
  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
  }else{
    lengths.table <- lengths}
  lengths.table = matrix(lengths.table)
  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop

  #Now for the actual calculation
  rpm <- counts_tpm
  for (i in 1:dim(counts_tpm)[2]) {
    for (j in 1:dim(counts_tpm)[1]) {
      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
    }
    scale.factor <- sum(rpm[,i])/(1E6)
    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
  }
  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = T), drop = FALSE], 
                           counts_tpm)
  } 
  return(counts_tpm)
}
```

```{r function_mround}
mround <- function(x,base){ 
  base*ceiling(x/base) 
} 
```

```{r function_my_max}
my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
```

The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome.

```{r cir_id, message=FALSE, results='hide'}

# Getting the cir gene names and transcript lengths needed for calculating the transcript-per-million

chabaudi_gene_info <- read.csv('../standard_files/PlasmoDB-47_beta_Pchabaudichabaudi_GenesByTaxon_Summary.csv',
                              header = TRUE,
                              stringsAsFactors = FALSE)

chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
  arrange(Gene.ID)

chabaudi_gene_info <- dplyr::rename(chabaudi_gene_info, 
                             Geneid = Gene.ID,
                             Description = Product.Description)

```

```{r cir_info}

cir_info <- read_delim('../standard_files/Pchabaudi_pir_info.txt',
           delim = '\t')
cir_id <- cir_info$Geneid

```

```{r times_stages_labels}
#Use the rMT stage differential count in order to name the samples, but keep time to retain order
times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
                                 stage_diff_count = c('Late-Rings',
                                                      'Rings-Troph-conversion',
                                                      'Early-Trophs',
                                                      'Mid-Trophs',
                                                      'Late-Trophs',
                                                      'Early-Rings',
                                                      'Early-Mid-Rings',
                                                      'Late-Mid-Rings'))
# R will convert the dashes to '.' at some point so let's have this notation here also
times_stages_table$r_stage_diff_count <- str_replace_all(times_stages_table$stage_diff_count, 
                                                         pattern = '-', 
                                                         replacement = '\\.')
```

```{r read_tpm_rescaled_data}

chab_count_table <- read_xlsx('Manuscript_Frontiers_chabaudi_24h_counts.xlsx')

twentyfour_ct_comb <- tpm(chab_count_table, 
                          lengths = chabaudi_transcript_lengths)

#Setting twentyfour_tpm_ct core dataset - remove pseudogene and re-scaled

twentyfour_tpm_nopseudo <- twentyfour_ct_comb %>% 
  filter(!Geneid %in% 'PCHAS_0420800')

twentyfour_tpm_rescale_nopseudo <- lapply(colnames(twentyfour_tpm_nopseudo), 
                                          function(column_name){
                                            column <- getElement(twentyfour_tpm_nopseudo, 
                                                                 paste(column_name))
                                            if(is.numeric(column)){
                                              #is.numeric to avoid trying to scale the Geneid column
                                              tpm_scale_factor <- 1e6 / sum(column)
                                              tpm_scaled <- as.numeric(column*tpm_scale_factor)
                                            }else{
                                              tpm_scaled <- column
                                            }
                                            return(tpm_scaled)
                                          }) %>% 
  bind_cols

colnames(twentyfour_tpm_rescale_nopseudo) <- colnames(twentyfour_tpm_nopseudo)
# apply(twentyfour_tpm_rescale_nopseudo[,-1], 2, sum)

#Setting the re-scaled version as the default from now on
#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo

twentyfour_rMT <- select(twentyfour_ct_comb, 
                         matches('Geneid|rMT')) %>% 
  rename_with(~ str_extract(.x,pattern = '(?<=rMT_).+'),
              matches('rMT'))

twentyfour_rMT_avg <- melt(twentyfour_rMT, 
                           variable.name = 'sample',
                           value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
  group_by(Geneid, stage) %>% 
  summarise(mean_tpm = mean(tpm)) %>% 
  dcast(Geneid ~ stage)

twentyfour_SBP <- select(twentyfour_ct_comb, 
                         matches('Geneid|SBP')) %>% 
  rename_with(~ str_extract(.x,pattern = '(?<=SBP_).+'),
              matches('SBP'))

twentyfour_SBP_avg <- melt(twentyfour_SBP, 
                           variable.name = 'sample',
                           value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
  group_by(Geneid, stage) %>% 
  summarise(mean_tpm = mean(tpm)) %>% 
  dcast(Geneid ~ stage)

twentyfour_avgtpm <- list(
  'rMT' = list(
    tpm = twentyfour_rMT_avg,
    experiment = 'rMT'
  ),
  'SBP' = list(
    tpm = twentyfour_SBP_avg,
    experiment = 'SBP'
  )
)

```

```{r colours_clades_phase}
SorL_col <- setNames(
  c(    '#F0E442',
        '#56B4E9', #L
        '#E69F00', #S
        '#999999'),
  sort(
    unique(cir_info$SorL)
  )
)

SorL_col_num <- SorL_col
names(SorL_col_num) <- sapply(names(SorL_col), 
                              function(clade_name) {
                                paste0(
                                  clade_name, 
                                  ' (n = ',
                                  sum(cir_info$SorL == clade_name),
                                  ')'
                                )
                              })

clade_col <- setNames(
  c('#F0E442',
    '#56B4E9', #L1
    '#009E73',
    '#0072B2',
    '#E69F00', #S1
    '#D55E00',
    '#CC79A7',
    '#999999'),
  sort(
    unique(cir_info$subfam)
  )
)

clade_col_num <- clade_col
names(clade_col_num) <- sapply(names(clade_col), 
                               function(clade_name) {
                                 paste0(
                                   clade_name, 
                                   ' (n = ',
                                   sum(cir_info$subfam == clade_name),
                                   ')'
                                 )
                               })

```

## Figure 4 - Expression of major expressed cirs rMT/SBP side-by-side.

_pir_ genes selected by being expressed above 1 TPM in all samples.

Adding in housekeeping genes for comparison:
HSP70 putative  PCHAS_0721000
MSP1 PCHAS_0831300
AMA1 PCHAS_0931000

```{r pirs_side-by-side_figure4}

housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
                             'MSP1' = 'PCHAS_0831300',
                             'AMA1' = 'PCHAS_0931000')

cir_housekeeping_id <- c(cir_id, 
                         'PCHAS_0721000',
                         'PCHAS_0831300',
                         'PCHAS_0931000')

cir_plot <- twentyfour_ct_comb %>% 
  filter(Geneid %in% cir_housekeeping_id) %>%
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(
    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
    subfam = cir_info$subfam[match(Geneid, cir_info$Geneid)]
  ) %>% 
  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
  filter(!(time %in% c('sch.culture', 
                       'Unknown'))) %>% 
  mutate(subfam = str_replace(subfam, 
                              'ancestral',
                              'a'))

cir_plot$Geneid <- ifelse(
  cir_plot$Geneid %in% housekeeping_genes[1,],
  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
  cir_plot$Geneid
)

limits_cir <- c(0, 
                mround(max(cir_plot$tpm), 
                       base = 100))

cir_plot_order <- cir_plot %>% 
  filter(time == '14h',
         transmission == 'rMT') %>% 
  arrange(desc(tpm))

cir_plot <- mutate(cir_plot, 
                   Geneid = factor(Geneid, 
                                   levels = unique(cir_plot_order$Geneid)))

side_by_side_cir_plot <- function(time_vec = '14h',
                                  log_it = TRUE,
                                  limits = limits_cir,
                                  point_size = 1,
                                  text_size = 10,
                                  file_name_end = '',
                                  threshold_min = 100,
                                  threshold_max = Inf){

  cir_plot <- group_by(cir_plot, 
                       Geneid) %>% 
    filter(
      time == time_vec,
      !Geneid %in% cir_id | any(tpm >= threshold_min),
      !Geneid %in% cir_id | all(tpm <= threshold_max)
    )

  # cir_plot %>% group_by(Geneid) %>% summarise(n())
  if(log_it){
    cir_plot$tpm <- log2(1+cir_plot$tpm)
      limits = log2(1+limits)
      breaks <- round(seq.int(0, 
                              max(limits), 
                              length.out = round(max(limits)) + 1, 
                              digits = 0), 
                      digits = 0)
    }else{
      breaks <- seq.int(0, max(limits), length.out = 11)
    }

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    cir_plot$subfam <- factor(
      cir_plot$subfam,
      levels = c(
        sort(unique(cir_plot$subfam)[!(unique(cir_plot$subfam) %in% 'ref')]),
        'ref'
      )
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
      stat_summary(fun.y = median, 
                   fun.ymin = median, 
                   fun.ymax = median,
                   geom = "bar", 
                   width = 0.75, 
                   size = 0.25, 
                   position = 'dodge') +
      geom_point(stat = 'identity', 
                 position = position_dodge(width = 0.75),
                 size = point_size) +
      scale_y_continuous(breaks = breaks,
                         limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      facet_grid(.~subfam, 
                 scales = "free_x",
                 space = "free_x") +
      theme_classic() +
      theme(
        axis.text.x = element_text(size = text_size,
                                   angle = 90)
      ) +
      ylab('log(TPM + 1)')+
      ggtitle(paste(unique(cir_plot$time)))

    #This bit is to change the colour, and text colour, of the facet boxes

    g <- ggplot_gtable(ggplot_build(y))
    strip_t <- which(grepl('strip-t', g$layout$name))
    subfam_vec <- sort(unique(cir_plot$subfam))
    fills <- case_when(str_detect(subfam_vec, 
                                  pattern = 'a') ~ clade_col['ancestral'],
                       subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
                       str_detect(subfam_vec, 
                                  pattern = 'ref') ~ 'white')
    k <- 1
    for (i in strip_t) {
      j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
      g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
      k <- k+1
    }

    grid.newpage()
    grid.draw(g)

}

 lapply(
    sort(unique(cir_plot$time)),
    side_by_side_cir_plot,
    threshold_min = 100,
    threshold_max = Inf,
    point_size = 1,
    text_size = 10,
    file_name_end = '_Over100TPM'
  )

```

# Supp figure - Pir expression of AAPLs and ChAPLs.

```{r Supp_fig_AapL-ChaAPl_expression}

  cir_plot <- twentyfour_ct_comb%>% 
    filter(Geneid %in% cir_id) %>%
    melt(id.vars = 'Geneid', 
         variable.name = 'sample', 
         value.name = 'tpm') %>% 
    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
           subfam = cir_info$subfam[match(Geneid, cir_info$Geneid)],
           chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
                                                                    cir_info$Geneid)],
                               levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
                                                                                         cir_info$Geneid)])))) %>% 
    filter(!(time %in% c('sch', 
                         'Unknown')),
           !is.na(chapl_aapl) ) %>% 
    group_by(Geneid) %>% 
    # filter(
    #   # all(tpm < 100),
    #   #      any(tpm > 50)
    #   subfam == 'S7'
    #   ) %>% 
    arrange(desc(tpm))

  cir_plot_order <- cir_plot %>% 
    filter(time == '14h',
           transmission == 'SBP') %>% 
    arrange(desc(tpm))

  cir_plot <- mutate(cir_plot, 
                     Geneid = factor(Geneid, 
                                     levels = unique(cir_plot_order$Geneid)))
  cir_plot_chapl_aapl_supp_info <- cir_plot

  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))

  breaks = seq(0, max_aapl_chapl, by = 1)
    limits = c(0, max_aapl_chapl)
    point_size = 0.25

  lapply(
    sort(unique(cir_plot$time)),
    function(time_vec){

    cir_plot <- filter(cir_plot, 
                       time == time_vec)

      cir_plot$tpm <- log2(1+cir_plot$tpm)

    cir_plot$Geneid <- factor(
      str_remove(cir_plot$Geneid, 
                 pattern = 'PCHAS_'),
      levels = str_remove(levels(cir_plot$Geneid), 
                          pattern = 'PCHAS_')
    )

    y <- ggplot(
      data = cir_plot,
      mapping = aes(
        x = Geneid,
        fill = transmission,
        y = tpm
      )
    ) +
    stat_summary(fun.y = median, 
                 fun.ymin = median, 
                 fun.ymax = median,
                 geom = "bar", 
                 width = 0.75, 
                 size = 0.25, 
                 position = 'dodge') +
    geom_point(stat = 'identity', 
               position = position_dodge(width = 0.75),
               size = point_size) +
    scale_y_continuous(breaks = breaks,
                       limits = limits) +
      scale_fill_manual(values = c('skyblue','red')) +
      ylab('log(TPM +1)') +
    facet_grid(.~chapl_aapl, 
             scales = "free_x",
             space = "free_x") +
    theme_classic() +
    theme(
      axis.text.x = element_text(angle = 90,
                                 size = 5)
    ) +
    ggtitle(paste(unique(cir_plot$time)))

  return(y)
})
```

## Bar chart of S and L clade expression across time.

```{r SorL_barchart}
tpm_data_SorL <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm))

y <- ggplot(data = tpm_data_SorL, 
              aes(
                x = transmission,
                y = total_tpm, 
                fill = SorL,
                group = transmission
              )) +
    geom_bar(position = position_stack(),
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
    scale_fill_manual(
      values = SorL_col,
      guide = guide_legend(reverse = TRUE)
    ) +
    scale_color_manual('black') + 
    # labs(fill = stack_label) +
    # scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('total '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~stage, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

```

## Clades proportion bar chart

```{r clades_prop_barchart}
tpm_data_clade <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  # mutate(stage =  factor(stage, 
  #                        levels = ordering(stage, 
  #                                          order_vector))) %>% 
  group_by(transmission,
           stage, 
           clade) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  arrange(stage)

y <- ggplot(data = tpm_data_clade, 
              aes(
                x = stage,
                y = total_tpm, 
                fill = clade,
                group = transmission
              )) +
    geom_bar(position = 'fill',
             stat = 'identity',
             colour = 'black',
             size = 0.1) +
    scale_fill_manual(
      breaks = sort(unique(tpm_data_clade$clade)),
      values = clade_col,
      guide = guide_legend(reverse = TRUE) 
    ) +
    scale_color_manual('black') + 
    # labs(fill = stack_label) +
    # scale_y_continuous(breaks = max_tpm) +
    theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('proportion of total '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~transmission, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_text(size = 7, 
                                 angle = 90),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

```

```{r add_berghei_SorL_prop}
#Add berghei data
berghei_SorL_file <- list.files('..', 
                                recursive = TRUE, 
                                pattern = 'tl20-05_SorL-subfamily_pir.csv', 
                                full.names = TRUE)
berghei_SorL <- read_csv(berghei_SorL_file) %>% 
  filter(stage %in% c('Asex.SchizA',
                      'Asex.SchizB',
                      'Asex.Ring',
                      'Asex.Troph',
                      'Asex.Mixed')) 
berghei_SorL_avg <- berghei_SorL %>% 
  group_by(stage,
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
  mutate(SorL = str_extract(SorL,
                            pattern = 'S|L|ancestral'),
         transmission = 'berghei')

```

```{r}
y <- ggplot(data = rbind(
  tpm_data_SorL,
  berghei_SorL_avg
), 
aes(
  x = stage,
  y = total_tpm, 
  fill = SorL,
  # group = transmission
)) +
  geom_bar(position = 'fill',
           stat = 'identity',
           colour = 'black',
           size = 0.1) +
  scale_fill_manual(
    values = SorL_col,
    guide = guide_legend(reverse = FALSE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('proportion of total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~transmission, 
             scales = "free_x",
             space = "free_x") +
  labs(fill = 'sub-family') +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_text(angle = 90),
    axis.text.y = element_text(size = 5),
    axis.title.y = element_text(size = 8),
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 5),
    legend.key.size = unit(3,'mm'),
    legend.key.width = unit(2,'mm'),
    strip.text.x = element_text(size=7, 
                                face = 'bold', 
                                angle = 0),
    strip.background = element_rect(
      color="black", 
      size=0.1, 
      linetype="solid"
    )
  ) 
y

```

```{r SorL_barchart_transmission-facet}
tpm_data_SorL <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)]) %>% 
  mutate(SorL = str_extract(clade, 
                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
         # stage =  factor(stage, levels = ordering(stage, order_vector))
         ) %>% 
  group_by(transmission,
           stage, 
           SorL) %>% 
  dplyr::summarise(total_tpm = sum(tpm)) %>% 
  mutate(SorL = factor(SorL, 
                       levels = sort(unique(SorL))))

y <- ggplot(data = tpm_data_SorL, 
              aes(
                x = stage,
                y = total_tpm, 
                fill = SorL,
                # group = transmission
              )) +
  geom_bar(position = position_stack(),
           stat = 'identity',
           colour = 'black',
           size = 0.1) +
  scale_fill_manual(
    values = SorL_col,
    guide = guide_legend(reverse = FALSE)
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
  scale_size(range = c(0,3)) +
  ylab(
    bquote('total '~italic(.('pir'))~' TPM')
  ) + 
  facet_grid(.~transmission, 
             scales = "free_x",
             space = "free_x") +
  labs(fill = 'sub-family') +
    theme(
      axis.title.x=element_blank(),
      # axis.text.x=element_blank(),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=7, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

```

```{r proportion_data}

tpm_data_twentyfour <- twentyfour_ct_comb%>% 
  filter(Geneid %in% cir_id) %>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(clade = cir_info$subfam[match(Geneid, 
                                            cir_info$Geneid)],
         chrom = str_extract(Geneid, 
                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
         ) %>% 
  mutate(
    SorL = str_extract(clade, 
                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
    # stage =  factor(stage, 
    #                 levels = ordering(stage, order_vector))
  ) %>% 
  group_by(transmission,
           stage, 
           Geneid,
           clade,
           chrom,
           SorL) %>% 
  dplyr::summarise(tpm = mean(tpm))

```

```{r chrom_prop_bar_chart}
tpm_data_chrom <- tpm_data_twentyfour %>%
  mutate(stage = times_stages_table$time[match(stage,
                                               times_stages_table$r_stage_diff_count)]) %>%
  group_by(transmission,
           stage,
           chrom) %>%
  dplyr::summarise(total_tpm = sum(tpm))

y <- ggplot(data = tpm_data_chrom) +
  geom_bar_pattern(
    aes(
      x = stage,
      y = total_tpm, 
      fill = chrom,
      group = transmission,
      pattern = chrom
    ),
    position = 'fill',
    stat = 'identity',
    colour = 'black',
    size = 0.1,
    pattern_density = 0.05,
    pattern_spacing = 0.02,
    pattern_key_scale_factor = 0.5
  ) +
  scale_pattern_manual(
    values = rep(c('none', 
                   'circle'), 
                 each = length(c(cbPalette, 
                                 'red', 
                                 'white')), 
                 length.out = length(unique(tpm_data_chrom$chrom))),
    labels = unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_fill_manual(
    breaks = unique(tpm_data_chrom$chrom),
    values = rep_len(c(cbPalette, 
                       'red', 
                       'white'),
                     length.out = length(unique(tpm_data_chrom$chrom))),
    labels =  unique(tpm_data_chrom$chrom),
    guide = guide_legend(reverse = TRUE),
    name = 'chromosome'
  ) +
  scale_color_manual('black') + 
  # labs(fill = stack_label) +
  # scale_y_continuous(breaks = max_tpm) +
  theme_classic() +
    scale_size(range = c(0,3)) +
    ylab(
      bquote('proportion of '~italic(.('pir'))~' TPM')
    ) + 
    facet_grid(.~transmission, 
               scales = "free_x",
               space = "free_x") +
    theme(
      axis.title.x=element_blank(),
      axis.text.x=element_text(size = 10, 
                                 angle = 0),
      axis.text.y = element_text(size = 5),
      axis.title.y = element_text(size = 8),
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 5),
      legend.key.size = unit(3,'mm'),
      legend.key.width = unit(2,'mm'),
      strip.text.x = element_text(size=10, 
                                  face = 'bold', 
                                  angle = 0),
      strip.background = element_rect(
        color="black", 
        size=0.1, 
        linetype="solid"
      )
    ) 
y

```

### 2D plot of PC1 v PC2 - Supp figure

```{r pca_calculation, results = TRUE}
twentyfour_tpm <- twentyfour_ct_comb
rownames_geneid <- twentyfour_tpm$Geneid
twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
rownames(twentyfour_tpm_forpca) <- rownames_geneid
twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
#Need to look into what exactly scale and center are doing
```

```{r 2d_pca, results = TRUE}
twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)

twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])

colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
colornames <- setNames(colornames, 
                       sort(unique(twentyfour_tpm_pca_coords$time)))

pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)

(
  plot <- ggplot(twentyfour_tpm_pca_coords,
                 aes(
                   x = PC1,
                   y = PC2,
                   col = time,
                   shape = transmission
                 )) +
    geom_point(size = 3,
               alpha = 0.8) +
    scale_color_manual(values = colornames) +
    scale_shape_manual(
      values = setNames(
        c(17,19),
        unique(twentyfour_tpm_pca_coords$transmission)
      )
    ) +
    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
    guides(shape=guide_legend(ncol=2),
           col = guide_legend(ncol=2)) +
    theme_classic() +
    theme(legend.key.size = unit(4, 'mm'),
          legend.text = element_text(size = 7))
)

```

# Calculate and cluster with z-scores {.tabset}

z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:

1. Convert the count/RPKM values of each gene into log values.
2. Calculate the mean and standard deviation of X gene log values in each sample
3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.

* Need to make sure that the z-score is calculated within each transmission method!

```{r zscores}
tpm_to_zscore = function(input){

  #If the input does not have a TPM element then assume that the input itself is the TPM object
  if(any(names(input) %in% 'tpm')){
    tpm_object <- input$tpm
  }else{
    tpm_object <- input
    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
  }

  #Want all of them so will just ignore Geneid
  input_logged <- tpm_object %>%
    mutate_at(vars(-matches('Geneid')), 
              log1p)

  gene_mean_scores = input_logged %>% 
    melt(id.vars = 'Geneid') %>% 
    group_by(Geneid) %>% 
    dplyr::summarise(mean = mean(value),
                     std_dev = sd(value))

  gene_vector = as.character(input_logged$Geneid)

  input_z = lapply(gene_vector, function(gene){
    mean_gene = filter(gene_mean_scores, 
                       Geneid == paste(gene)) %>% 
      select(mean)
    std_gene = filter(gene_mean_scores, 
                      Geneid == paste(gene)) %>% 
      select(std_dev)
    gene_scores = input_logged %>% 
      filter(Geneid == paste(gene)) %>% 
      select(-Geneid)
    if(std_gene != 0){
      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
    }else{
      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
    }
    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
  })

  #Use reduce and rbind to turn the list into one df
  input_z_df = purrr::reduce(input_z, 
                             rbind)
  return(list(zscores = input_z_df,
              tpm = tpm_object,
              experiment = input$experiment))
}

#Make and save z-scores
twentyfour_z <- lapply(twentyfour_avgtpm,
                       tpm_to_zscore)

###
```

# Phaseograms of the data.

```{r row_order_phaseograms}

#Getting the row order for each phaseogram for each experiment
# Want the data ordered by the highest zscore for each gene in each stage.
phaseogram_row_order <- lapply(twentyfour_z, 
                               function(list_elements){
                                 zscore_df <- list_elements$zscores
                                 zscore_ordered <- melt(zscore_df, 
                                                        id = 'Geneid',
                                                        variable.name = 'stage',
                                                        value.name = 'zscore') %>% 
                                   group_by(Geneid) %>% 
                                   top_n(zscore, n = 1) %>% 
                                   arrange(stage, 
                                           desc(zscore)) %>%
                                   distinct(Geneid, 
                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
                                   filter(zscore > 0) #For some reason some remain with zero zscore
                                 # zscore_ordered <- zscore_ordered[order(
                                 #   match(
                                 #     zscore_ordered$stage, 
                                 #     order_vector
                                 #   )
                                 # ),]
                                 return(c(list_elements, 
                                          list(zscore_order = zscore_ordered)))
                               })

phaseogram_time_order <- lapply(names(phaseogram_row_order), 
                                function(list_element){
                                  transmission_list <- getElement(phaseogram_row_order,
                                                                  list_element)
                                  order_by_time <- transmission_list$zscore_order %>% 
                                    mutate(stage = times_stages_table$time[match(stage,
                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
                                    arrange(stage)
                                  return(list(zscore_order = order_by_time))
                                }
)
names(phaseogram_time_order) <- names(phaseogram_row_order)

```

```{r function_draw_heatmap_zscore}

draw_heatmap_zscore <- function(zscore_list,
                                title = '',
                                gene_ordering_vector = NULL,
                                filter_genes_vector = NULL,
                                extra_title = NULL,
                                annotation_clade = NULL,
                                annotation_col,
                                threshold_tpm = NULL,
                                use_times = FALSE,
                                use_order_vector = TRUE,
                                tpm_mat = TRUE,
                                tpm_breaks = NULL,
                                ...){
  if(is.null(tpm_breaks) & tpm_mat){
    #Need the generate the breaks in the data for the heatmap legend if not provided.
    max_tpm <- lapply(zscore_list, 
                      function(list_element){
                        #list_element <- zscore_list[[1]]
                        heatmap_data <- as.data.frame(list_element$zscores)
                        title <- list_element$experiment

                        if(!'Geneid' %in% colnames(heatmap_data)){
                          stop('No Geneid column in the heatmap data.frame')
                        }

                        if(!is.null(filter_genes_vector)){
                          heatmap_data <- as.data.frame(
                            filter(heatmap_data, 
                                   Geneid %in% filter_genes_vector)
                          )
                        }

                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                        if(!is.null(gene_ordering_vector)){
                          #Remove genes not in the main data frame but in the ordering vector, 
                          # and then order based on the vector
                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          heatmap_data <- heatmap_data[na.omit(
                            match(gene_ordering_vector, 
                                  rownames(heatmap_data))
                          ),,
                          drop=FALSE]
                        }

                        geneids <- rownames(heatmap_data)

                        #Get max/min TPM scores for the side annotation heatmap                       
                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                          list_element$tpm$Geneid),] %>% 
                          select(-Geneid) %>% 
                          as.matrix %>% 
                          log1p
                        rownames(tpm_mat) <- geneids

                        tpm_min <- rowMins(tpm_mat)
                        names(tpm_min) <- geneids
                        tpm_max <- rowMaxs(tpm_mat)
                        names(tpm_max) <- geneids
                        return(list(tpm_max = tpm_max,
                                    tpm_min = tpm_min))
                      })

    tpm_max_all <- mround(
      max(
        unlist(
          lapply(max_tpm, 
                 getElement,
                 'tpm_max')
        )
      ), 
      base = 1)

    tpm_breaks <- round(
      c(0,
        tpm_max_all/4,
        tpm_max_all/2,
        tpm_max_all*(3/4),
        tpm_max_all),
      digits = 2
    )
  }

  #Generate the colour function from the TPM legend breaks for the TPM annotation
  if(tpm_mat){
    tpm_col_fun <- colorRamp2(
      breaks = tpm_breaks,
      colors = viridis(length(tpm_breaks))
    )
  }

  if(is.numeric(threshold_tpm)){
    #Apply a threshold to the TPM in order to remove low expressed genes.
    # Remember to add log1p here! 
    # Threshold uses the max value, so it ony needs to be above thresh 
    #   in one stage avg in one transmission method.
    #Will alter the gene_filter vector in order to only select the genes above threshold

    gene_maxs <- bind_rows(
      lapply(max_tpm, 
             getElement,
             'tpm_max')
    )

    remove_genes <- names(gene_maxs)[apply(
      gene_maxs, 
      2, 
      function(maxs){
        all(maxs < log2(1+threshold_tpm))
      }
    )]

    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]

  }

  if(is.character(annotation_clade)){
    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
    clade_info <- getElement(cir_info,
                             annotation_clade)[match(order_cir, 
                                                     cir_info$Geneid)]
    names(clade_info) <- str_split_fixed(order_cir, 
                                         pattern = '_', 
                                         n = 2)[,2]
    annotation_col <- annotation_col
  }

  heatmaps <- lapply(zscore_list, 
                     function(list_element){
                       #list_element <- zscore_list[[1]]
                       heatmap_data <- as.data.frame(list_element$zscores)
                       title <- list_element$experiment

                       if(!'Geneid' %in% colnames(heatmap_data)){
                         stop('No Geneid column in the heatmap data.frame')
                       }

                       if(!is.null(filter_genes_vector)){
                         heatmap_data <- as.data.frame(
                           filter(heatmap_data, 
                                  Geneid %in% filter_genes_vector)
                         )
                       }

                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!

                       if(!is.null(gene_ordering_vector)){
                         #Remove genes not in the main data frame but in the ordering vector,
                         #  and then order based on the vector
                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                         heatmap_data <- heatmap_data[na.omit(
                           match(gene_ordering_vector, 
                                 rownames(heatmap_data))
                         ),,
                         drop=FALSE]
                       }

                       if(tpm_mat){
                         #Get max/min TPM scores for the side annotation heatmap                       
                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                           list_element$tpm$Geneid),] %>%
                           select(-Geneid) %>% 
                           as.matrix %>% 
                           log1p

                         tpm_min <- rowMins(tpm_mat)
                         tpm_max <- rowMaxs(tpm_mat)

                         names(tpm_min) = rownames(heatmap_data)
                         names(tpm_max) = rownames(heatmap_data)

                         tpm_ha <- rowAnnotation(
                           min.max = anno_simple(cbind(tpm_min, 
                                                       tpm_max), 
                                                 col = tpm_col_fun,
                                                 gp = gpar(col = 'black',
                                                           lwd = 0.1),
                                                 border = TRUE,
                                                 simple_anno_size = unit(0.2, "cm")),
                           show_annotation_name = FALSE)
                       }

                       #Draw the left annotation is needed, only for the first element.

                       if(
                         identical(list_element,
                                   first(zscore_list)) & is.character(annotation_clade)
                       ){
                         clade_ha <- rowAnnotation(
                           'subfamily' = anno_simple(
                             clade_info,
                             col = annotation_col
                           )
                         )
                       }

                       #Change rownames to remove PCHAS_
                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
                                                                pattern = '_', 
                                                                n = 2)[,2]

                       if(use_times){
                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
                                                                                 times_stages_table$r_stage_diff_count)]
                       }

                       if(use_order_vector){
                         #Order the columns by ordering_vector
                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
                                                                   order_vector)),
                                                      drop = FALSE]
                       }else{
                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
                                                      drop = FALSE] 
                       }

                       h = Heatmap(heatmap_data,
                                   name = paste0('z-score',
                                                 extra_title),
                                   col = c(cbPalette[6],
                                           'white', 
                                           cbPalette[7]),
                                   row_names_gp = gpar(fontsize = ifelse(
                                     nrow(heatmap_data) > 13, 
                                     12*(13/nrow(heatmap_data)), 
                                     12)
                                   ),
                                   column_names_gp = gpar(fontsize = 10),
                                   row_dend_width = unit(3, "cm"),
                                   right_annotation = if(tpm_mat){
                                     tpm_ha
                                   }else{
                                     NULL
                                   },
                                   left_annotation = 
                                     if(identical(list_element,
                                                  first(zscore_list)) & is.character(annotation_clade)){
                                       clade_ha
                                     }else{
                                       NULL
                                     }, 
                                   #Only want the clade annotation when it is first heatmap
                                   width = unit(4.5, 'cm'),
                                   cluster_columns = FALSE,
                                   height = unit(7,'cm'),
                                   row_title = NULL,
                                   cluster_rows = FALSE,
                                   column_title = paste(title),
                                   heatmap_legend_param = list(
                                     at = c(3,2,1,0,-1,-2,-3),
                                     labels_gp = gpar(fontsize = 10),
                                     title_gp = gpar(fontsize = 10)
                                   ),
                                   ...
                       )
                       return(h)
                     })

  ht_list <- Reduce('+', heatmaps) #To add the list elements together

  if(tpm_mat){
    #Draw the legend
    tpm_lgd <- Legend(
      title = "Min/Max log(TPM + 1)", 
      col = tpm_col_fun, 
      at = tpm_breaks, 
      labels = tpm_breaks,
      title_gp = gpar(fontsize = 7)
    )
  }

  if(is.character(annotation_clade)){
    clade_lgd <- Legend(
      title = 'sub-family',
      at = names(annotation_col),
      legend_gp = gpar(fill = annotation_col)
    )
  }

  draw(ht_list, 
       row_title_side = 'left',
       column_title_side = 'bottom',
       heatmap_legend_side = 'left',
       #Only include the annotation legends for the annotations we actually have.
       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
         list(tpm_lgd,
              clade_lgd)
       }else if(tpm_mat & !is.character(annotation_clade) ){
         list(tpm_lgd)
       } else if(!tpm_mat & is.character(annotation_clade) ){
         list(clade_lgd)
       })
}
```

## Phaseograms of the entire genome.

```{r phaseograms_whole_genome}

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

```

## Phaseograms of the _cir_ genes with clade annotations.

```{r pir_gene_phaseogram_breaks}
  max_tpm <- lapply(twentyfour_z, 
                    function(list_element){
                      #list_element <- zscore_list[[1]]
                      heatmap_data <- as.data.frame(list_element$zscores)
                      title <- list_element$experiment

                      if(!'Geneid' %in% colnames(heatmap_data)){
                        stop('No Geneid column in the heatmap data.frame')
                      }

                            heatmap_data <- as.data.frame(
                              filter(heatmap_data, 
                                     Geneid %in% cir_id)
                            )

                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)

                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
                          # 
                          # if(!is.null(gene_ordering_vector)){
                          #   #Remove genes not in the main data frame but in the ordering vector, 
                          #   # and then order based on the vector
                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
                          #   heatmap_data <- heatmap_data[na.omit(
                          #     match(gene_ordering_vector, 
                          #           rownames(heatmap_data))
                          #   ),,
                          #   drop=FALSE]
                          # }

                          geneids <- rownames(heatmap_data)

                          #Get max/min TPM scores for the side annotation heatmap                       
                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
                                                            list_element$tpm$Geneid),] %>% 
                            select(-Geneid) %>% 
                            as.matrix %>% 
                            log1p
                          rownames(tpm_mat) <- geneids

                          tpm_min <- rowMins(tpm_mat)
                          names(tpm_min) <- geneids
                          tpm_max <- rowMaxs(tpm_mat)
                          names(tpm_max) <- geneids
                          return(list(tpm_max = tpm_max,
                                      tpm_min = tpm_min))
                        })

  tpm_max_all <- mround(
    max(
      unlist(
        lapply(max_tpm, 
               getElement,
               'tpm_max')
      )
    ), 
    base = 1)

  tpm_breaks <- round(
    c(0,
      tpm_max_all/4,
      tpm_max_all/2,
      tpm_max_all*(3/4),
      tpm_max_all),
    digits = 2
  )
```

```{r phaseograms_cir}

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = FALSE)

draw_heatmap_zscore(zscore_list = twentyfour_z,
                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
                    filter_genes_vector = cir_id,
                    annotation_clade = 'SorL',
                    annotation_col = SorL_col,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_breaks = tpm_breaks,
                    tpm_mat = FALSE)

```
## Figure 3b - Highest expressed all _pir_ heatmap

```{r highest_expressed_stage_heatmap_fig3b}

# Get order of pir genes by highest expressed genes in each sample.

combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm), 
       function(list_name){

         combined_data_pir <- getElement(twentyfour_avgtpm, 
                                     list_name)$tpm %>% 
           filter(Geneid %in% cir_id) %>% 
           column_to_rownames(var = 'Geneid')

         #Ordering that a more understandable order

         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
                                                            order_vector)),
                                                drop = FALSE]

         # combined_data_ordered <- melt(
         #   (
         #     combined_data_pir %>% 
         #       rownames_to_column(var = 'Geneid')
         #   ),
         #   id = 'Geneid',
         #   variable.name = 'stage',
         #   value.name = 'tpm') %>% 
         #   group_by(Geneid) %>% 
         #   top_n(tpm, 
         #         n = 1) %>% 
         #   arrange(stage,
         #           desc(tpm)) %>%
         #   distinct(Geneid, 
         #            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
         #   filter(tpm > 0)

         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
                                         means = apply(combined_data_pir, 1, mean),
                                         subfam = cir_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
                                                                            cir_info$Geneid)]) %>% 
           filter(means > 1) %>% 
           arrange(subfam, 
                   desc(means)) %>% 
           .$Geneid

         #Order the heatmap data by the order defined above (SorL, then mean expression).
         combined_data_pir_ht <- combined_data_pir[
           order(
             match(rownames(combined_data_pir),
                   combined_data_ordered),
             na.last = NA #Remove NAs
           ),
         ]

         #Change the colnames to the times
         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                                                           pattern = '[[:digit:]]{2}h')

         #Order by time point
         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                      drop = FALSE] 
         #Add a column for the SorL
         combined_data_pir_ht$SorL <- cir_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
                                                                cir_info$Geneid)]

         return(combined_data_pir_ht)
       })
names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)

maxcol <- ceiling(
  max(
    unlist(
      lapply(combined_data_pir_ht_list,
             function(list_element){
               max(
                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
               )
             }
      )
    )
  )
)

lapply(names(combined_data_pir_ht_list),
       function(list_name_to_order){

         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
                                            list_name_to_order)

         pir_highest_tpm_order <- rownames(combined_data_pir_ht)

         #Sub-family annotation
         clade_info <- combined_data_pir_ht$SorL
         annotation_col <- SorL_col

         #Subfamily annotation

         clade_ha <- rowAnnotation(
           'sub-family' = anno_simple(
             clade_info,
             col = annotation_col,
           ),
           annotation_name_gp = gpar(fontsize = 10)
         )

         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
                                                 labels = 'ancestral',
                                                 labels_gp = gpar(fontsize = 7),
                                                 link_width = unit(3, "mm")))

         #lapply across the datasets now, for each order
         ht_list <- lapply(names(combined_data_pir_ht_list),
                           function(list_name){

                             combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
                                                                list_name)$tpm %>% 
                               column_to_rownames(var = 'Geneid')

                             #Get the times instead of the stage names.
                             colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                                                           pattern = '[[:digit:]]{2}h')

                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                                          drop = FALSE]

                             Heatmap(
                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
                               cluster_rows = FALSE,
                               cluster_columns = FALSE,
                               show_row_names = FALSE,
                               name = 'log(TPM+1)', 
                               col = colorRamp2(c(0,
                                                  maxcol/4,
                                                  maxcol/2,
                                                  3*maxcol/4,
                                                  maxcol),
                                                viridis(5)),
                               na_col = 'white', 
                               border = TRUE,
                               border_gp = gpar(lwd = 0.1),
                               column_names_gp = gpar(fontsize = 10),
                               row_dend_width = unit(3, "cm"),
                               width = unit(4.5, 'cm'),
                               height = unit(7,'cm'),
                               row_title = NULL,
                               left_annotation = if(
                                 list_name %in% first(names(combined_data_pir_ht_list))
                               ){
                                 clade_ha
                               }else{
                                 NULL
                               },
                               right_annotation = if(
                                 list_name %in% last(names(combined_data_pir_ht_list))
                               ){
                                 anc_ha
                               }else{
                                 NULL
                               },
                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
                               #                                        cir_info$Geneid)],
                               column_title = paste(list_name),
                               heatmap_legend_param = list(
                                 labels_gp = gpar(fontsize = 10),
                                 title_gp = gpar(fontsize = 10,
                                                 fontface = 'bold'),
                                 at = c(0,
                                        maxcol/4,
                                        maxcol/2,
                                        3*maxcol/4,
                                        maxcol)
                               )
                             )
                           }
         )
         ht_list <- Reduce(`+`, ht_list)
         clade_lgd <- Legend(
           title = 'sub-family',
           at = names(SorL_col),
           legend_gp = gpar(fill = SorL_col)
         )

         return(
           draw(ht_list, 
                row_title_side = 'left',
                column_title_side = 'bottom',
                heatmap_legend_side = 'right',
                annotation_legend_side = 'right',
                annotation_legend_list = list(clade_lgd),
                ht_gap = unit(.35,'cm'),
                merge_legend = TRUE
           ))
       })

```

```{r figure_3b_suppinfo}

combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
                                            'rMT')

         pir_highest_tpm_order <- rownames(combined_data_pir_ht)

supp_info_fig_3b <- lapply(names(combined_data_pir_ht_list),
       function(list_name){

         combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
                                            list_name)$tpm %>% 
           column_to_rownames(var = 'Geneid')

         #Get the times instead of the stage names.
         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
                                                       pattern = '[[:digit:]]{2}h')

         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
                                                      drop = FALSE]

         colnames(combined_data_pir_ht) <- paste0(list_name, 
                                                  '_', 
                                                  colnames(combined_data_pir_ht))

         log2(1+combined_data_pir_ht)[pir_highest_tpm_order,]
       }) %>% 
  bind_cols

supp_info_fig_3b$SorL <- cir_info$SorL[match(names(apply(supp_info_fig_3b, 1, mean)), 
                                                   cir_info$Geneid)]

head(supp_info_fig_3b)

```

# Lower pir expression vs rest of genome

```{r low_pir_expression}
set.seed(4534)

gene_expression_df <- twentyfour_ct_comb%>% 
  melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'tpm') %>% 
  mutate(stage = str_extract(sample, 
                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
  filter(stage != 'sch.culture') %>% 
  mutate(pir_gene = ifelse(Geneid %in% cir_id,
                           'pir',
                           'non-pir')) %>% 
  group_by(stage, 
           pir_gene,
           transmission) %>% 
  slice_sample(n = length(cir_id)) %>%
  summarise(mean_tpm = mean(tpm),
            med_tpm = round(median(tpm),
                            digits = 6))

gene_expression_df %>% 
  group_by(transmission, 
           pir_gene) %>% 
  slice_min(mean_tpm, n =1)

gene_expression_df %>% 
  group_by(transmission, 
           pir_gene) %>% 
  slice_max(mean_tpm, n =1)

y <- ggplot(gene_expression_df,
       aes(x = pir_gene,
           y = log2(1+mean_tpm))) +
  facet_grid(~transmission) +
  ggtitle(expression(paste(italic('P. chabaudi'), ' 24h cycle'))) +
  xlab('Gene group')+
  ylab('log(Median TPM + 1)')+
  # geom_violin()+
  geom_dotplot(binaxis = 'y',
               stackdir = 'center',
               # binwidth = 1,
               # dotsize = 2,
               method = 'dotdensity') +
  geom_path(aes(group = stage,
                col = stage)) +
  theme_classic()
y

```

### 2021-03-12 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd

@@
     mutate(stack_group = Geneid) %>% 
+    mutate(tpm = ifelse(tpm > 1, tpm, 0)) %>% #Filtering step for low value pir genes.
     group_by(stage, 
@@
 ```
-```
 

### 2021-03-12 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd

@@
 my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
+```
+
+```{r function_ordering}
+ordering <- function(to_order, order_vector){
+  order = unlist(sapply(order_vector, 
+                        function(x){unique(str_subset(to_order, 
+                                                      pattern = paste(x)))}))
+  order_missing = unique(to_order[!to_order %in% order])
+  unique(c(order, order_missing), fromLast = TRUE)
+}
+
+order_vector <- c("Early.Rings",
+                  "Early.Mid.Rings",
+                  "Late.Mid.Rings",
+                  "Late.Rings",
+                  "Rings.Troph.conversion",
+                  "Early.Trophs",
+                  "Mid.Trophs",
+                  "Late.Trophs")
 ```
@@
                          identical(list_element,
-                                   first(zscore_list)) & is.character(annotation_clade)
+                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
                        ){
@@
                                      if(identical(list_element,
-                                                  first(zscore_list)) & is.character(annotation_clade)){
+                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
                                        clade_ha
@@
 ```
+
 ## Figure 3b - Highest expressed all _pir_ heatmap
@@
 ```{r highest_expressed_stage_heatmap_fig3b}
+
+num_SorL <- cir_info %>% 
+  group_by(SorL) %>% 
+  summarise(num = n_distinct(Gene.ID))
+
+labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
+                        num_SorL$SorL)
 
@@
                                                 drop = FALSE]
-
-         # combined_data_ordered <- melt(
-         #   (
-         #     combined_data_pir %>% 
-         #       rownames_to_column(var = 'Geneid')
-         #   ),
-         #   id = 'Geneid',
-         #   variable.name = 'stage',
-         #   value.name = 'tpm') %>% 
-         #   group_by(Geneid) %>% 
-         #   top_n(tpm, 
-         #         n = 1) %>% 
-         #   arrange(stage,
-         #           desc(tpm)) %>%
-         #   distinct(Geneid, 
-         #            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-         #   filter(tpm > 0)
 
@@
                                left_annotation = if(
-                                 list_name %in% first(names(combined_data_pir_ht_list))
+                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
                                ){
@@
                                right_annotation = if(
-                                 list_name %in% last(names(combined_data_pir_ht_list))
+                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
                                ){
@@
            title = 'sub-family',
-           at = names(SorL_col),
+           at = labels_SorL,
            legend_gp = gpar(fill = SorL_col)

### 2021-03-12 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Change directory

@@
 author: "Timothy Little"
-date: "05/01/2021"
+date: "12/03/2021"
 output: 
@@
 knitr::opts_knit$set(root.dir = 
-                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/Pberghei")
+                       ".")
 options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)

### 2021-03-12 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Some better descriptions

@@
 options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
-# getwd()
-
 ```
@@
 ```{r function_mround}
+#Make nicer numbers rounded up, useful for making scales for a plot.
 mround <- function(x,base){ 
@@
 ```{r function_my_max}
-my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
+#max robust to NAs
+my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) 
 ```
@@
 ```{r function_uniquify}
+#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
 uniquify <- function(vector_names, first_one = TRUE){
@@
 ```{r function_ordering}
+#Use an order_vector of partial string matches to order.
 ordering = function(to_order, order_vector){
@@
 # We need to write a function to calculate the TPM from the count data we have generated.
-This function uses the _lengths.transcript_ object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.
+This function uses the _lengths.transcript_ object and should generate errors if it detects discrepancies between the Geneids so that the wrong transcript length is not used.
 
@@
 
-```{r pir/fam_id, message=FALSE, results='hide'}
-
-# Getting the pir/fam gene names and transcript lengths needed for calculating the transcript-per-million
+```{r pir_id, message=FALSE, results='hide'}
+
+# Getting the pir gene names and transcript lengths needed for calculating the transcript-per-million
 
@@
 
-#Remove the Caldelari 6h samples, as they did in their paper.
+#Remove the Caldelari (E06) 6h samples, as they did in their paper.
 

### 2021-03-12 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Updated names, orders, to match the manuscript. Removed figures.

@@
 
+pir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
+                        delim = '\t')
 pir_clade <- mutate(pir_clade, 
                     locus = pir_loci$locus[match(Geneid, pir_loci$Geneid)])
-
-pir_clade <- read_delim('../standard_files/Pberghei_pir_clades.txt', 
-                        delim = '\t')
 pir_id <- pir_clade$Geneid
@@
 expt_tpms$E03$tpm$`Asex-Schiz_3` <- NULL
-```
-
-## Supp figure - PCA of all the samples.
-
-```{r supp_figure_PCA_allgenes}
-
-big_tpm <- bind_cols(
-  lapply(
-    expt_tpms,
-    function(x){
-      tpm_df <- x$tpm
-      colnames(tpm_df)[-1] <- paste0(x$experiment,
-                                     '_',
-                                     colnames(tpm_df)[-1])
-      tpm_df
-    }
-  )
-) %>% 
-  dplyr::rename(Geneid = Geneid...1) %>% 
-  select_at(vars(-matches('Geneid\\.\\.\\.\\d+$'))) %>% 
-  data.frame
-
-rownames(big_tpm) <- big_tpm$Geneid
-
-big_tpm <- big_tpm[,-1,drop=FALSE]
-
-big_tpm_pca <-  prcomp(big_tpm, 
-                       center = TRUE, 
-                       scale. = TRUE)
-
-big_tpm_pca_var <- get_pca_var(big_tpm_pca)
-big_tpm_pca_coords <- data.frame(sample = rownames(big_tpm_pca_var$coord),
-                                 big_tpm_pca_var$coord)
-big_tpm_pca_coords$stage_experiment <- str_extract(
-  big_tpm_pca_coords$sample,
-  pattern = '.+(?=_[[:digit:]]+$)'
-)
-big_tpm_pca_coords$stage <- str_extract(big_tpm_pca_coords$sample, 
-                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
-big_tpm_pca_coords$experiment <- str_extract(big_tpm_pca_coords$sample, 
-                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')
-
-colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
-                               length.out = length(unique(big_tpm_pca_coords$stage))),
-                       unique(big_tpm_pca_coords$stage))
-
-pca_importance <- as.data.frame(summary(big_tpm_pca)$importance)
-
-plot <- ggplot(big_tpm_pca_coords,
-       aes(
-         x = Dim.1,
-         y = Dim.2,
-         col = stage,
-         shape = experiment
-       )) +
-  geom_point() +
-  scale_fill_manual(values = colornames) +
-  scale_shape_manual(
-    values = setNames(
-      c(1:10),
-      unique(big_tpm_pca_coords$experiment)
-    )
-  ) +
-  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-  guides(shape=guide_legend(ncol=2),
-         col = guide_legend(ncol=2)) +
-  theme_classic() +
-  theme(legend.key.size = unit(5, 'mm'),
-        legend.text = element_text(size = 6))
-
-plot
-
-```
-
-```{r supp_figure_PCA_pir}
-big_tpm_pir <- big_tpm[rownames(big_tpm) %in% pir_id,,drop = FALSE]
-
-big_tpm_pir_pca <-  prcomp(big_tpm_pir, 
-                       center = TRUE, 
-                       scale. = TRUE)
-
-big_tpm_pir_pca_var <- get_pca_var(big_tpm_pir_pca)
-big_tpm_pir_pca_coords <- data.frame(sample = rownames(big_tpm_pir_pca_var$coord),
-                                 big_tpm_pir_pca_var$coord)
-big_tpm_pir_pca_coords$stage_experiment <- str_extract(
-  big_tpm_pir_pca_coords$sample,
-  pattern = '.+(?=_[[:digit:]]+$)'
-)
-big_tpm_pir_pca_coords$stage <- str_extract(big_tpm_pir_pca_coords$sample, 
-                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
-big_tpm_pir_pca_coords$experiment <- str_extract(big_tpm_pir_pca_coords$sample, 
-                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')
-
-colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
-                               length.out = length(unique(big_tpm_pir_pca_coords$stage))),
-                       unique(big_tpm_pir_pca_coords$stage))
-
-pca_importance <- as.data.frame(summary(big_tpm_pir_pca)$importance)
-
-plot_pir <- ggplot(big_tpm_pir_pca_coords,
-       aes(
-         x = Dim.1,
-         y = Dim.2,
-         col = stage,
-         shape = experiment
-       )) +
-  geom_point() +
-  scale_fill_manual(values = colornames) +
-  scale_shape_manual(
-    values = setNames(
-      c(1:10),
-      unique(big_tpm_pca_coords$experiment)
-    )
-  ) +
-  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-  guides(shape=guide_legend(ncol=2),
-         col = guide_legend(ncol=2)) +
-  theme_classic() +
-  theme(legend.key.size = unit(5, 'mm'),
-        legend.text = element_text(size = 6))
-
-plot_pir
-```
-
-How many experiments cover each life stage?
-
-```{r experiment_per_life_stage}
-
-big_tpm %>% 
-  rownames_to_column(var = 'Geneid') %>% 
-  melt(id.vars = 'Geneid') %>% 
-  mutate(
-    stage = str_extract(variable, pattern = '(?<=E[:digit:]{2}_).+(?=_[:digit:])'),
-    expt = str_extract(variable, pattern = 'E[:digit:]{2}')
-  ) %>% 
-  group_by(stage) %>% 
-    summarise(num_expt = n_distinct(expt)) %>% 
-    arrange(desc(num_expt))
-
 ```
@@
 
-###
-
-### Read in top genes
-
-###
-
 lapply(all_top, `[`, c('all_genes', 'pir_genes'))
@@
 
-## Pearson correlation plot of collated samples.
+##Supp figure 1A - Pearson correlation plot of collated samples.
 
 ```{r all_samples_collated_heatmap}
+
+big_tpm <- bind_cols(
+  lapply(
+    expt_tpms,
+    function(x){
+      tpm_df <- x$tpm
+      colnames(tpm_df)[-1] <- paste0(x$experiment,
+                                     '_',
+                                     colnames(tpm_df)[-1])
+      tpm_df
+    }
+  )
+) %>% 
+  dplyr::rename(Geneid = Geneid...1) %>% 
+  select_at(vars(-matches('Geneid\\.\\.\\.\\d+$'))) %>% 
+  data.frame
+
+rownames(big_tpm) <- big_tpm$Geneid
+
+big_tpm <- big_tpm[,-1,drop=FALSE]
+
 ## Collated across the stages for a more easily understood heatmap
@@
          col = c(viridis(100), magma(115)), addrect = 6)
+```
+
+## Supp figure 1B - PCA of all the samples.
+
+```{r supp_figure_PCA_allgenes}
+
+big_tpm_pca <-  prcomp(big_tpm, 
+                       center = TRUE, 
+                       scale. = TRUE)
+
+big_tpm_pca_var <- get_pca_var(big_tpm_pca)
+big_tpm_pca_coords <- data.frame(sample = rownames(big_tpm_pca_var$coord),
+                                 big_tpm_pca_var$coord)
+big_tpm_pca_coords$stage_experiment <- str_extract(
+  big_tpm_pca_coords$sample,
+  pattern = '.+(?=_[[:digit:]]+$)'
+)
+big_tpm_pca_coords$stage <- str_extract(big_tpm_pca_coords$sample, 
+                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
+big_tpm_pca_coords$experiment <- str_extract(big_tpm_pca_coords$sample, 
+                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')
+
+colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
+                               length.out = length(unique(big_tpm_pca_coords$stage))),
+                       unique(big_tpm_pca_coords$stage))
+
+pca_importance <- as.data.frame(summary(big_tpm_pca)$importance)
+
+plot <- ggplot(big_tpm_pca_coords,
+       aes(
+         x = Dim.1,
+         y = Dim.2,
+         col = stage,
+         shape = experiment
+       )) +
+  geom_point() +
+  scale_fill_manual(values = colornames) +
+  scale_shape_manual(
+    values = setNames(
+      c(1:10),
+      unique(big_tpm_pca_coords$experiment)
+    )
+  ) +
+  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
+  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
+  guides(shape=guide_legend(ncol=2),
+         col = guide_legend(ncol=2)) +
+  theme_classic() +
+  theme(legend.key.size = unit(5, 'mm'),
+        legend.text = element_text(size = 6))
+
+plot
+
+```
+
+#Supp figure 1C - PCA using just the pir genes.
+
+```{r supp_figure_PCA_pir}
+big_tpm_pir <- big_tpm[rownames(big_tpm) %in% pir_id,,drop = FALSE]
+
+big_tpm_pir_pca <-  prcomp(big_tpm_pir, 
+                       center = TRUE, 
+                       scale. = TRUE)
+
+big_tpm_pir_pca_var <- get_pca_var(big_tpm_pir_pca)
+big_tpm_pir_pca_coords <- data.frame(sample = rownames(big_tpm_pir_pca_var$coord),
+                                 big_tpm_pir_pca_var$coord)
+big_tpm_pir_pca_coords$stage_experiment <- str_extract(
+  big_tpm_pir_pca_coords$sample,
+  pattern = '.+(?=_[[:digit:]]+$)'
+)
+big_tpm_pir_pca_coords$stage <- str_extract(big_tpm_pir_pca_coords$sample, 
+                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
+big_tpm_pir_pca_coords$experiment <- str_extract(big_tpm_pir_pca_coords$sample, 
+                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')
+
+colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
+                               length.out = length(unique(big_tpm_pir_pca_coords$stage))),
+                       unique(big_tpm_pir_pca_coords$stage))
+
+pca_importance <- as.data.frame(summary(big_tpm_pir_pca)$importance)
+
+plot_pir <- ggplot(big_tpm_pir_pca_coords,
+       aes(
+         x = Dim.1,
+         y = Dim.2,
+         col = stage,
+         shape = experiment
+       )) +
+  geom_point() +
+  scale_fill_manual(values = colornames) +
+  scale_shape_manual(
+    values = setNames(
+      c(1:10),
+      unique(big_tpm_pca_coords$experiment)
+    )
+  ) +
+  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
+  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
+  guides(shape=guide_legend(ncol=2),
+         col = guide_legend(ncol=2)) +
+  theme_classic() +
+  theme(legend.key.size = unit(5, 'mm'),
+        legend.text = element_text(size = 6))
+
+plot_pir
 ```
@@
 
-```{r SorL_stacked_eachGene_fig1b}
+```{r SorL_stacked_eachGene_fig2a}
   tpm_data <- combined_data %>% 
@@
 
-## Figure 1C
+## Figure 2B - sub-family proportions
 
@@
 )
-
-#Separate out the pan-rodent pir
-# tpm_data_subfam$clade[tpm_data_subfam$stack_group == 'PBANKA_0524600'] <- 'pan-rodent'
 
@@
 
-## Chromosome proportional bar chart supp fig
-
-```{r chrom_prop_bar_chart}
-tpm_data_chrom <- combined_data %>% 
-  rownames_to_column(var = 'Geneid') %>% 
-  filter(Geneid %in% pir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'stage', 
-       value.name = 'tpm') %>% 
-  mutate(
-    chrom = str_extract(Geneid, pattern = '(?<=PBANKA_)[[:digit:]]{2}')
-  ) %>% 
-  group_by(chrom) %>% 
-  mutate(chrom = paste0(chrom, ' (n = ', n_distinct(Geneid) ,')')) %>% 
-  group_by(stage, 
-           chrom) %>% 
-  dplyr::summarise(total_tpm = sum(tpm)) %>%
-  arrange(
-          desc(chrom),
-          total_tpm
-) %>% 
-  mutate(chrom = factor(chrom, 
-                        levels = unique(chrom)))
-
-  stacked_max <- tpm_data_chrom %>% 
-    ungroup %>% 
-    group_by(stage) %>% 
-    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
-    .$max_tpm %>% 
-    max
-
-  max_tpm <- c(
-    seq(
-      0,
-      mround(
-        stacked_max,
-        500),
-      length.out = 11)
-  )
-
-y <- ggplot(data = tpm_data_chrom) +
-  geom_bar_pattern(
-    aes(
-      x = stage,
-      y = total_tpm, 
-      fill = chrom,
-      group = stage,
-      pattern = chrom
-    ),
-    position = 'fill',
-    stat = 'identity',
-    colour = 'black',
-    size = 0.1,
-    pattern_density = 0.2,
-    pattern_spacing = 0.05,
-    pattern_key_scale_factor = 0.25
-  ) +
-  scale_pattern_manual(
-    values = rep(c('none', 
-                   'circle'), 
-                 each = length(c(cbPalette, 
-                                 'red', 
-                                 'white')), 
-                 length.out = length(unique(tpm_data_chrom$chrom))),
-    labels = unique(tpm_data_chrom$chrom),
-    guide = guide_legend(reverse = TRUE),
-    name = 'chromosome'
-  ) +
-  scale_fill_manual(
-    breaks = unique(tpm_data_chrom$chrom),
-    values = rep_len(c(cbPalette, 
-                       'red', 
-                       'white'),
-                     length.out = length(unique(tpm_data_chrom$chrom))),
-    labels =  unique(tpm_data_chrom$chrom),
-    guide = guide_legend(reverse = TRUE),
-    name = 'chromosome'
-  ) +
-  scale_color_manual('black') +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) +
-  facet_grid(.~stage,
-             scales = "free_x",
-             space = "free_x") +
-  scale_x_discrete(expand=expand_scale(add=1)) +
-  theme_classic() +
-  theme(
-    axis.title.x=element_blank(),
-    axis.text.x=element_blank(),
-    axis.ticks.x = element_blank(),
-    # axis.line.x = element_blank(),
-    axis.text.y = element_text(size = 10),
-    axis.title.y = element_text(size = 10),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 8),
-    legend.key.size = unit(3,'mm'),
-    legend.key.width = unit(2,'mm'),
-    strip.text.x = element_text(size=8,
-                                face = 'bold',
-                                angle = 90),
-    strip.background = element_rect(
-      color="black",
-      size=0.1,
-      linetype="solid"
-    ),
-    panel.spacing.x = unit(0, "null")
-  )
-
-  #This bit is to change the colour of the facet boxes
-
-  g <- ggplot_gtable(ggplot_build(y))
-  strip_t <- which(grepl('strip-t', g$layout$name))
-  stage_vec <- unique(
-
-    levels(tpm_data$stage)
-  )
-  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
-                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
-                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
-                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
-  k <- 1
-  for (i in strip_t) {
-    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-    k <- k+1
-  }
-
-  grid.newpage()
-  grid.draw(g)
-
-```
-
-## Expression by locus. Proportion. Supp figure.
-
-Loci are defined by L or R of each chromosome number, will define this by the jump in the gene numbers.
-
-```{r defining_loci}
-
-tpm_data_locus <- combined_data %>% 
-  rownames_to_column(var = 'Geneid') %>% 
-  filter(Geneid %in% pir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'stage', 
-       value.name = 'tpm') %>% 
-  mutate(
-    locus = pir_loci$locus[match(Geneid, pir_loci$Geneid)]
-  ) %>% 
-  filter(!is.na(locus)) %>% 
-  group_by(locus) %>% 
-  mutate(locus = paste0(locus, ' (n = ', n_distinct(Geneid) ,')')) %>% 
-  group_by(stage, 
-           locus) %>% 
-  dplyr::summarise(total_tpm = sum(tpm)) %>%
-  arrange(
-          desc(locus),
-          total_tpm
-) %>% 
-  mutate(locus = factor(locus, levels = unique(locus)))
-
-  stacked_max <- tpm_data_locus %>% 
-    ungroup %>% 
-    group_by(stage) %>% 
-    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
-    .$max_tpm %>% 
-    max
-
-  max_tpm <- c(
-    seq(
-      0,
-      mround(
-        stacked_max,
-        500),
-      length.out = 11)
-  )
-
-(y <- ggplot(data = tpm_data_locus) +
-  geom_bar_pattern(
-    aes(
-      x = stage,
-      y = total_tpm, 
-      fill = locus,
-      group = stage,
-      pattern = locus
-    ),
-    position = 'fill',
-    stat = 'identity',
-    colour = 'black',
-    size = 0.1,
-        pattern_density = 0.05,
-    pattern_spacing = 0.1,
-    pattern_key_scale_factor = 0.1
-    #  pattern_density = 0.1,
-    # pattern_spacing = 0.1,
-    # pattern_key_scale_factor = 0.5
-  ) +
-  scale_pattern_manual(
-    values = rep(c('none', 
-                   'circle',
-                   'stripe'), 
-                 each = length(c(cbPalette, 
-                                 'red', 
-                                 'white')), 
-                 length.out = length(unique(tpm_data_locus$locus))),
-    labels = unique(tpm_data_locus$locus),
-    guide = guide_legend(reverse = TRUE),
-    name = 'locus'
-  ) +
-  scale_fill_manual(
-    breaks = unique(tpm_data_locus$locus),
-    values = rep_len(c(cbPalette, 
-                       'red', 
-                       'white'),
-                     length.out = length(unique(tpm_data_locus$locus))),
-    labels =  unique(tpm_data_locus$locus),
-    guide = guide_legend(reverse = TRUE),
-    name = 'locus'
-  ) +
-  scale_color_manual('black') +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) +
-  facet_grid(.~stage,
-             scales = "free_x",
-             space = "free_x") +
-  scale_x_discrete(expand=expand_scale(add=1)) +
-  theme_classic() +
-  theme(
-    axis.title.x=element_blank(),
-    axis.text.x=element_blank(),
-    axis.ticks.x = element_blank(),
-    # axis.line.x = element_blank(),
-    axis.text.y = element_text(size = 10),
-    axis.title.y = element_text(size = 10),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 8),
-    legend.key.size = unit(3,'mm'),
-    legend.key.width = unit(2,'mm'),
-    strip.text.x = element_text(size=8,
-                                face = 'bold',
-                                angle = 90),
-    strip.background = element_rect(
-      color="black",
-      size=0.1,
-      linetype="solid"
-    ),
-    panel.spacing.x = unit(0, "null")
-  ))
-
-  #This bit is to change the colour of the facet boxes
-
-  g <- ggplot_gtable(ggplot_build(y))
-  strip_t <- which(grepl('strip-t', g$layout$name))
-  stage_vec <- unique(
-
-    levels(tpm_data$stage)
-  )
-  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
-                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
-                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
-                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
-  k <- 1
-  for (i in strip_t) {
-    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-    k <- k+1
-  }
-
-  grid.newpage()
-  grid.draw(g)
-```
-## Gametocyte marker gene expression - supp figure
+## Gametocyte marker gene expression - supp figure 2
 

### 2021-03-15 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Reorder bits

@@
 ---
-title: "Manuscript_Github_Pchabaudi_pir_210105"
+title: "Manuscript_Github_Pchabaudi_pir"
 author: "Timothy Little"
-date: "05/01/2021"
+date: "12/03/2021"
 output: html_document
@@
 
-The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome.
+The _cir_ gene names were downloaded using a search of PlasmoDB _P. chabaudi_ genome. _cir_ genes are the names of _pir_ genes in _P chabaudi_.
 
@@
 
+cir_clade_info <- read_delim('../standard_files/Pchabaudi_pir_clades.txt',
+            delim = '\t')
 ```
@@
 
-## Figure 4 - Expression of major expressed cirs rMT/SBP side-by-side.
+### 2D plot of PC1 v PC2 - Supp figure 4A
+
+```{r pca_calculation, results = TRUE}
+twentyfour_tpm <- twentyfour_ct_comb
+rownames_geneid <- twentyfour_tpm$Geneid
+twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
+rownames(twentyfour_tpm_forpca) <- rownames_geneid
+twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
+#Need to look into what exactly scale and center are doing
+```
+
+```{r 2d_pca, results = TRUE}
+twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
+twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
+
+twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
+                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
+                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
+                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
+  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
+  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
+  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
+
+colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
+colornames <- setNames(colornames, 
+                       sort(unique(twentyfour_tpm_pca_coords$time)))
+
+pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
+pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
+pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
+pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
+
+(
+  plot <- ggplot(twentyfour_tpm_pca_coords,
+                 aes(
+                   x = PC1,
+                   y = PC2,
+                   col = time,
+                   shape = transmission
+                 )) +
+    geom_point(size = 3,
+               alpha = 0.8) +
+    scale_color_manual(values = colornames) +
+    scale_shape_manual(
+      values = setNames(
+        c(17,19),
+        unique(twentyfour_tpm_pca_coords$transmission)
+      )
+    ) +
+    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
+    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
+    guides(shape=guide_legend(ncol=2),
+           col = guide_legend(ncol=2)) +
+    theme_classic() +
+    theme(legend.key.size = unit(4, 'mm'),
+          legend.text = element_text(size = 7))
+)
+
+```
+
+# Calculate and cluster with z-scores {.tabset}
+
+z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:
+
+1. Convert the count/RPKM values of each gene into log values.
+2. Calculate the mean and standard deviation of X gene log values in each sample
+3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
+4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.
+
+* Need to make sure that the z-score is calculated within each transmission method!
+
+```{r zscores}
+tpm_to_zscore = function(input){
+
+  #If the input does not have a TPM element then assume that the input itself is the TPM object
+  if(any(names(input) %in% 'tpm')){
+    tpm_object <- input$tpm
+  }else{
+    tpm_object <- input
+    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
+  }
+
+  #Want all of them so will just ignore Geneid
+  input_logged <- tpm_object %>%
+    mutate_at(vars(-matches('Geneid')), 
+              log1p)
+
+  gene_mean_scores = input_logged %>% 
+    melt(id.vars = 'Geneid') %>% 
+    group_by(Geneid) %>% 
+    dplyr::summarise(mean = mean(value),
+                     std_dev = sd(value))
+
+  gene_vector = as.character(input_logged$Geneid)
+
+  input_z = lapply(gene_vector, function(gene){
+    mean_gene = filter(gene_mean_scores, 
+                       Geneid == paste(gene)) %>% 
+      select(mean)
+    std_gene = filter(gene_mean_scores, 
+                      Geneid == paste(gene)) %>% 
+      select(std_dev)
+    gene_scores = input_logged %>% 
+      filter(Geneid == paste(gene)) %>% 
+      select(-Geneid)
+    if(std_gene != 0){
+      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
+    }else{
+      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
+    }
+    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
+    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
+  })
+
+  #Use reduce and rbind to turn the list into one df
+  input_z_df = purrr::reduce(input_z, 
+                             rbind)
+  return(list(zscores = input_z_df,
+              tpm = tpm_object,
+              experiment = input$experiment))
+}
+
+#Make and save z-scores
+twentyfour_z <- lapply(twentyfour_avgtpm,
+                       tpm_to_zscore)
+
+###
+```
+
+# Phaseograms of the data.
+
+```{r row_order_phaseograms}
+
+#Getting the row order for each phaseogram for each experiment
+# Want the data ordered by the highest zscore for each gene in each stage.
+phaseogram_row_order <- lapply(twentyfour_z, 
+                               function(list_elements){
+                                 zscore_df <- list_elements$zscores
+                                 zscore_ordered <- melt(zscore_df, 
+                                                        id = 'Geneid',
+                                                        variable.name = 'stage',
+                                                        value.name = 'zscore') %>% 
+                                   group_by(Geneid) %>% 
+                                   top_n(zscore, n = 1) %>% 
+                                   arrange(stage, 
+                                           desc(zscore)) %>%
+                                   distinct(Geneid, 
+                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
+                                   filter(zscore > 0) #For some reason some remain with zero zscore
+                                 # zscore_ordered <- zscore_ordered[order(
+                                 #   match(
+                                 #     zscore_ordered$stage, 
+                                 #     order_vector
+                                 #   )
+                                 # ),]
+                                 return(c(list_elements, 
+                                          list(zscore_order = zscore_ordered)))
+                               })
+
+phaseogram_time_order <- lapply(names(phaseogram_row_order), 
+                                function(list_element){
+                                  transmission_list <- getElement(phaseogram_row_order,
+                                                                  list_element)
+                                  order_by_time <- transmission_list$zscore_order %>% 
+                                    mutate(stage = times_stages_table$time[match(stage,
+                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
+                                    arrange(stage)
+                                  return(list(zscore_order = order_by_time))
+                                }
+)
+names(phaseogram_time_order) <- names(phaseogram_row_order)
+
+```
+
+```{r function_draw_heatmap_zscore}
+
+draw_heatmap_zscore <- function(zscore_list,
+                                title = '',
+                                gene_ordering_vector = NULL,
+                                filter_genes_vector = NULL,
+                                extra_title = NULL,
+                                annotation_clade = NULL,
+                                annotation_col,
+                                threshold_tpm = NULL,
+                                use_times = FALSE,
+                                use_order_vector = TRUE,
+                                tpm_mat = TRUE,
+                                tpm_breaks = NULL,
+                                ...){
+  if(is.null(tpm_breaks) & tpm_mat){
+    #Need the generate the breaks in the data for the heatmap legend if not provided.
+    max_tpm <- lapply(zscore_list, 
+                      function(list_element){
+                        #list_element <- zscore_list[[1]]
+                        heatmap_data <- as.data.frame(list_element$zscores)
+                        title <- list_element$experiment
+
+                        if(!'Geneid' %in% colnames(heatmap_data)){
+                          stop('No Geneid column in the heatmap data.frame')
+                        }
+
+                        if(!is.null(filter_genes_vector)){
+                          heatmap_data <- as.data.frame(
+                            filter(heatmap_data, 
+                                   Geneid %in% filter_genes_vector)
+                          )
+                        }
+
+                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+
+                        if(!is.null(gene_ordering_vector)){
+                          #Remove genes not in the main data frame but in the ordering vector, 
+                          # and then order based on the vector
+                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                          heatmap_data <- heatmap_data[na.omit(
+                            match(gene_ordering_vector, 
+                                  rownames(heatmap_data))
+                          ),,
+                          drop=FALSE]
+                        }
+
+                        geneids <- rownames(heatmap_data)
+
+                        #Get max/min TPM scores for the side annotation heatmap                       
+                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                          list_element$tpm$Geneid),] %>% 
+                          select(-Geneid) %>% 
+                          as.matrix %>% 
+                          log1p
+                        rownames(tpm_mat) <- geneids
+
+                        tpm_min <- rowMins(tpm_mat)
+                        names(tpm_min) <- geneids
+                        tpm_max <- rowMaxs(tpm_mat)
+                        names(tpm_max) <- geneids
+                        return(list(tpm_max = tpm_max,
+                                    tpm_min = tpm_min))
+                      })
+
+    tpm_max_all <- mround(
+      max(
+        unlist(
+          lapply(max_tpm, 
+                 getElement,
+                 'tpm_max')
+        )
+      ), 
+      base = 1)
+
+    tpm_breaks <- round(
+      c(0,
+        tpm_max_all/4,
+        tpm_max_all/2,
+        tpm_max_all*(3/4),
+        tpm_max_all),
+      digits = 2
+    )
+  }
+
+  #Generate the colour function from the TPM legend breaks for the TPM annotation
+  if(tpm_mat){
+    tpm_col_fun <- colorRamp2(
+      breaks = tpm_breaks,
+      colors = viridis(length(tpm_breaks))
+    )
+  }
+
+  if(is.numeric(threshold_tpm)){
+    #Apply a threshold to the TPM in order to remove low expressed genes.
+    # Remember to add log1p here! 
+    # Threshold uses the max value, so it ony needs to be above thresh 
+    #   in one stage avg in one transmission method.
+    #Will alter the gene_filter vector in order to only select the genes above threshold
+
+    gene_maxs <- bind_rows(
+      lapply(max_tpm, 
+             getElement,
+             'tpm_max')
+    )
+
+    remove_genes <- names(gene_maxs)[apply(
+      gene_maxs, 
+      2, 
+      function(maxs){
+        all(maxs < log2(1+threshold_tpm))
+      }
+    )]
+
+    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
+
+  }
+
+  if(is.character(annotation_clade)){
+    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
+    clade_info <- getElement(cir_info,
+                             annotation_clade)[match(order_cir, 
+                                                     cir_info$Geneid)]
+    names(clade_info) <- str_split_fixed(order_cir, 
+                                         pattern = '_', 
+                                         n = 2)[,2]
+    annotation_col <- annotation_col
+  }
+
+  heatmaps <- lapply(zscore_list, 
+                     function(list_element){
+                       #list_element <- zscore_list[[1]]
+                       heatmap_data <- as.data.frame(list_element$zscores)
+                       title <- list_element$experiment
+
+                       if(!'Geneid' %in% colnames(heatmap_data)){
+                         stop('No Geneid column in the heatmap data.frame')
+                       }
+
+                       if(!is.null(filter_genes_vector)){
+                         heatmap_data <- as.data.frame(
+                           filter(heatmap_data, 
+                                  Geneid %in% filter_genes_vector)
+                         )
+                       }
+
+                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+
+                       if(!is.null(gene_ordering_vector)){
+                         #Remove genes not in the main data frame but in the ordering vector,
+                         #  and then order based on the vector
+                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                         heatmap_data <- heatmap_data[na.omit(
+                           match(gene_ordering_vector, 
+                                 rownames(heatmap_data))
+                         ),,
+                         drop=FALSE]
+                       }
+
+                       if(tpm_mat){
+                         #Get max/min TPM scores for the side annotation heatmap                       
+                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                           list_element$tpm$Geneid),] %>%
+                           select(-Geneid) %>% 
+                           as.matrix %>% 
+                           log1p
+
+                         tpm_min <- rowMins(tpm_mat)
+                         tpm_max <- rowMaxs(tpm_mat)
+
+                         names(tpm_min) = rownames(heatmap_data)
+                         names(tpm_max) = rownames(heatmap_data)
+
+                         tpm_ha <- rowAnnotation(
+                           min.max = anno_simple(cbind(tpm_min, 
+                                                       tpm_max), 
+                                                 col = tpm_col_fun,
+                                                 gp = gpar(col = 'black',
+                                                           lwd = 0.1),
+                                                 border = TRUE,
+                                                 simple_anno_size = unit(0.2, "cm")),
+                           show_annotation_name = FALSE)
+                       }
+
+                       #Draw the left annotation is needed, only for the first element.
+
+                       if(
+                         identical(list_element,
+                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
+                       ){
+                         clade_ha <- rowAnnotation(
+                           'subfamily' = anno_simple(
+                             clade_info,
+                             col = annotation_col
+                           )
+                         )
+                       }
+
+                       #Change rownames to remove PCHAS_
+                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
+                                                                pattern = '_', 
+                                                                n = 2)[,2]
+
+                       if(use_times){
+                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
+                                                                                 times_stages_table$r_stage_diff_count)]
+                       }
+
+                       if(use_order_vector){
+                         #Order the columns by ordering_vector
+                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
+                                                                   order_vector)),
+                                                      drop = FALSE]
+                       }else{
+                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
+                                                      drop = FALSE] 
+                       }
+
+                       h = Heatmap(heatmap_data,
+                                   name = paste0('z-score',
+                                                 extra_title),
+                                   col = c(cbPalette[6],
+                                           'white', 
+                                           cbPalette[7]),
+                                   row_names_gp = gpar(fontsize = ifelse(
+                                     nrow(heatmap_data) > 13, 
+                                     12*(13/nrow(heatmap_data)), 
+                                     12)
+                                   ),
+                                   column_names_gp = gpar(fontsize = 10),
+                                   row_dend_width = unit(3, "cm"),
+                                   right_annotation = if(tpm_mat){
+                                     tpm_ha
+                                   }else{
+                                     NULL
+                                   },
+                                   left_annotation = 
+                                     if(identical(list_element,
+                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
+                                       clade_ha
+                                     }else{
+                                       NULL
+                                     }, 
+                                   #Only want the clade annotation when it is first heatmap
+                                   width = unit(4.5, 'cm'),
+                                   cluster_columns = FALSE,
+                                   height = unit(7,'cm'),
+                                   row_title = NULL,
+                                   cluster_rows = FALSE,
+                                   column_title = paste(title),
+                                   heatmap_legend_param = list(
+                                     at = c(3,2,1,0,-1,-2,-3),
+                                     labels_gp = gpar(fontsize = 10),
+                                     title_gp = gpar(fontsize = 10)
+                                   ),
+                                   ...
+                       )
+                       return(h)
+                     })
+
+  ht_list <- Reduce('+', heatmaps) #To add the list elements together
+
+  if(tpm_mat){
+    #Draw the legend
+    tpm_lgd <- Legend(
+      title = "Min/Max log(TPM + 1)", 
+      col = tpm_col_fun, 
+      at = tpm_breaks, 
+      labels = tpm_breaks,
+      title_gp = gpar(fontsize = 7)
+    )
+  }
+
+  if(is.character(annotation_clade)){
+    clade_lgd <- Legend(
+      title = 'sub-family',
+      at = names(annotation_col),
+      legend_gp = gpar(fill = annotation_col)
+    )
+  }
+
+  draw(ht_list, 
+       row_title_side = 'left',
+       column_title_side = 'bottom',
+       heatmap_legend_side = 'left',
+       #Only include the annotation legends for the annotations we actually have.
+       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
+         list(tpm_lgd,
+              clade_lgd)
+       }else if(tpm_mat & !is.character(annotation_clade) ){
+         list(tpm_lgd)
+       } else if(!tpm_mat & is.character(annotation_clade) ){
+         list(clade_lgd)
+       })
+}
+```
+
+## Phaseograms of the entire genome.
+
+```{r phaseograms_whole_genome}
+
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE)
+
+```
+
+## Phaseograms of the _cir_ genes with clade annotations.
+
+```{r pir_gene_phaseogram_breaks}
+  max_tpm <- lapply(twentyfour_z, 
+                    function(list_element){
+                      #list_element <- zscore_list[[1]]
+                      heatmap_data <- as.data.frame(list_element$zscores)
+                      title <- list_element$experiment
+
+                      if(!'Geneid' %in% colnames(heatmap_data)){
+                        stop('No Geneid column in the heatmap data.frame')
+                      }
+
+                            heatmap_data <- as.data.frame(
+                              filter(heatmap_data, 
+                                     Geneid %in% cir_id)
+                            )
+
+                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+                          # 
+                          # if(!is.null(gene_ordering_vector)){
+                          #   #Remove genes not in the main data frame but in the ordering vector, 
+                          #   # and then order based on the vector
+                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                          #   heatmap_data <- heatmap_data[na.omit(
+                          #     match(gene_ordering_vector, 
+                          #           rownames(heatmap_data))
+                          #   ),,
+                          #   drop=FALSE]
+                          # }
+
+                          geneids <- rownames(heatmap_data)
+
+                          #Get max/min TPM scores for the side annotation heatmap                       
+                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                            list_element$tpm$Geneid),] %>% 
+                            select(-Geneid) %>% 
+                            as.matrix %>% 
+                            log1p
+                          rownames(tpm_mat) <- geneids
+
+                          tpm_min <- rowMins(tpm_mat)
+                          names(tpm_min) <- geneids
+                          tpm_max <- rowMaxs(tpm_mat)
+                          names(tpm_max) <- geneids
+                          return(list(tpm_max = tpm_max,
+                                      tpm_min = tpm_min))
+                        })
+
+  tpm_max_all <- mround(
+    max(
+      unlist(
+        lapply(max_tpm, 
+               getElement,
+               'tpm_max')
+      )
+    ), 
+    base = 1)
+
+  tpm_breaks <- round(
+    c(0,
+      tpm_max_all/4,
+      tpm_max_all/2,
+      tpm_max_all*(3/4),
+      tpm_max_all),
+    digits = 2
+  )
+```
+
+```{r phaseograms_cir}
+
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE)
+
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_breaks = tpm_breaks,
+                    tpm_mat = FALSE)
+
+```
+
+## Figure 3b - Highest expressed all _pir_ heatmap
+
+```{r highest_expressed_stage_heatmap_fig3b}
+
+num_SorL <- cir_info %>% 
+  group_by(SorL) %>% 
+  summarise(num = n_distinct(Gene.ID))
+
+labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
+                        num_SorL$SorL)
+
+# Get order of pir genes by highest expressed genes in each sample.
+
+combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm), 
+       function(list_name){
+
+         combined_data_pir <- getElement(twentyfour_avgtpm, 
+                                     list_name)$tpm %>% 
+           filter(Geneid %in% cir_id) %>% 
+           column_to_rownames(var = 'Geneid')
+
+         #Ordering that a more understandable order
+
+         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
+                                                            order_vector)),
+                                                drop = FALSE]
+
+         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
+                                         means = apply(combined_data_pir, 1, mean),
+                                         subfam = cir_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
+                                                                            cir_info$Geneid)]) %>% 
+           filter(means > 1) %>% 
+           arrange(subfam, 
+                   desc(means)) %>% 
+           .$Geneid
+
+         #Order the heatmap data by the order defined above (SorL, then mean expression).
+         combined_data_pir_ht <- combined_data_pir[
+           order(
+             match(rownames(combined_data_pir),
+                   combined_data_ordered),
+             na.last = NA #Remove NAs
+           ),
+         ]
+
+         #Change the colnames to the times
+         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
+                                                                           pattern = '[[:digit:]]{2}h')
+
+         #Order by time point
+         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+                                                      drop = FALSE] 
+         #Add a column for the SorL
+         combined_data_pir_ht$SorL <- cir_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
+                                                                cir_info$Geneid)]
+
+         return(combined_data_pir_ht)
+       })
+names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)
+
+maxcol <- ceiling(
+  max(
+    unlist(
+      lapply(combined_data_pir_ht_list,
+             function(list_element){
+               max(
+                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
+               )
+             }
+      )
+    )
+  )
+)
+
+lapply(names(combined_data_pir_ht_list),
+       function(list_name_to_order){
+
+         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
+                                            list_name_to_order)
+
+         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
+
+         #Sub-family annotation
+         clade_info <- combined_data_pir_ht$SorL
+         annotation_col <- SorL_col
+
+         #Subfamily annotation
+
+         clade_ha <- rowAnnotation(
+           'sub-family' = anno_simple(
+             clade_info,
+             col = annotation_col,
+           ),
+           annotation_name_gp = gpar(fontsize = 10)
+         )
+
+         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
+                                                 labels = 'ancestral',
+                                                 labels_gp = gpar(fontsize = 7),
+                                                 link_width = unit(3, "mm")))
+
+         #lapply across the datasets now, for each order
+         ht_list <- lapply(names(combined_data_pir_ht_list),
+                           function(list_name){
+
+                             combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
+                                                                list_name)$tpm %>% 
+                               column_to_rownames(var = 'Geneid')
+
+                             #Get the times instead of the stage names.
+                             colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
+                                                                           pattern = '[[:digit:]]{2}h')
+
+                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+                                                                          drop = FALSE]
+
+                             Heatmap(
+                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
+                               cluster_rows = FALSE,
+                               cluster_columns = FALSE,
+                               show_row_names = FALSE,
+                               name = 'log(TPM+1)', 
+                               col = colorRamp2(c(0,
+                                                  maxcol/4,
+                                                  maxcol/2,
+                                                  3*maxcol/4,
+                                                  maxcol),
+                                                viridis(5)),
+                               na_col = 'white', 
+                               border = TRUE,
+                               border_gp = gpar(lwd = 0.1),
+                               column_names_gp = gpar(fontsize = 10),
+                               row_dend_width = unit(3, "cm"),
+                               width = unit(4.5, 'cm'),
+                               height = unit(7,'cm'),
+                               row_title = NULL,
+                               left_annotation = if(
+                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
+                               ){
+                                 clade_ha
+                               }else{
+                                 NULL
+                               },
+                               right_annotation = if(
+                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
+                               ){
+                                 anc_ha
+                               }else{
+                                 NULL
+                               },
+                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
+                               #                                        cir_info$Geneid)],
+                               column_title = paste(list_name),
+                               heatmap_legend_param = list(
+                                 labels_gp = gpar(fontsize = 10),
+                                 title_gp = gpar(fontsize = 10,
+                                                 fontface = 'bold'),
+                                 at = c(0,
+                                        maxcol/4,
+                                        maxcol/2,
+                                        3*maxcol/4,
+                                        maxcol)
+                               )
+                             )
+                           }
+         )
+         ht_list <- Reduce(`+`, ht_list)
+         clade_lgd <- Legend(
+           title = 'sub-family',
+           at = labels_SorL,
+           legend_gp = gpar(fill = SorL_col)
+         )
+
+         return(
+           draw(ht_list, 
+                row_title_side = 'left',
+                column_title_side = 'bottom',
+                heatmap_legend_side = 'right',
+                annotation_legend_side = 'right',
+                annotation_legend_list = list(clade_lgd),
+                ht_gap = unit(.35,'cm'),
+                merge_legend = TRUE
+           ))
+       })
+
+```
+
+```{r figure_3b_suppinfo}
+
+combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
+                                            'rMT')
+
+         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
+
+supp_info_fig_3b <- lapply(names(combined_data_pir_ht_list),
+       function(list_name){
+
+         combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
+                                            list_name)$tpm %>% 
+           column_to_rownames(var = 'Geneid')
+
+         #Get the times instead of the stage names.
+         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
+                                                       pattern = '[[:digit:]]{2}h')
+
+         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+                                                      drop = FALSE]
+
+         colnames(combined_data_pir_ht) <- paste0(list_name, 
+                                                  '_', 
+                                                  colnames(combined_data_pir_ht))
+
+         log2(1+combined_data_pir_ht)[pir_highest_tpm_order,]
+       }) %>% 
+  bind_cols
+
+supp_info_fig_3b$SorL <- cir_info$SorL[match(names(apply(supp_info_fig_3b, 1, mean)), 
+                                                   cir_info$Geneid)]
+
+head(supp_info_fig_3b)
+
+```
+
+## Figure 4 and supplementary figure 7/8
 
@@
 
-```{r pirs_side-by-side_figure4}
+```{r pirs_side-by-side_figure4_RMT-SBP-completelySeparate}
 
@@
     time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
-    subfam = cir_info$subfam[match(Geneid, cir_info$Geneid)]
+    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
   ) %>% 
@@
                               'ancestral',
-                              'a'))
+                              'a')) %>% 
+  mutate(
+    subfam = interaction(
+      transmission,
+      forcats::fct_shift(
+        relevel(
+          as.factor(subfam),
+          ref = 'ref')
+      ),
+      sep = '_',
+      lex.order = TRUE
+    )
+  )
+#Need to add the transmission name to the subfam so that they are ordered correctly.
+# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
+# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.
 
@@
 
-side_by_side_cir_plot <- function(time_vec = '14h',
+transmissionSep_cir_plot <- function(time_vec = '14h',
                                      log_it = TRUE,
@@
                                      text_size = 10,
-                                  file_name_end = '',
                                      threshold_min = 100,
@@
 
-  cir_plot <- group_by(cir_plot, 
+  cir_plot_filter <- group_by(cir_plot, 
                        Geneid) %>% 
@@
     )
-
-  # cir_plot %>% group_by(Geneid) %>% summarise(n())
+  #Filter for pirs expressed above or below the thresholds, have the OR so that the ref genes aren't removed.
+  # e.g. Either 'Not a cir' OR 'Any TPM is above the minimum'.
+
+  # cir_plot_filter %>% group_by(Geneid) %>% summarise(n())
   if(log_it){
-    cir_plot$tpm <- log2(1+cir_plot$tpm)
+    cir_plot_filter$tpm <- log2(1+cir_plot_filter$tpm)
     limits = log2(1+limits)
@@
 
-    cir_plot$Geneid <- factor(
-      str_remove(cir_plot$Geneid, 
+  cir_plot_filter$Geneid <- factor(
+    str_remove(cir_plot_filter$Geneid, 
                pattern = 'PCHAS_'),
-      levels = str_remove(levels(cir_plot$Geneid), 
+    levels = str_remove(levels(cir_plot_filter$Geneid), 
                         pattern = 'PCHAS_')
@@
 
-    cir_plot$subfam <- factor(
-      cir_plot$subfam,
-      levels = c(
-        sort(unique(cir_plot$subfam)[!(unique(cir_plot$subfam) %in% 'ref')]),
-        'ref'
-      )
-    )
+  lapply(unique(cir_plot_filter$transmission), 
+         function(transmission_method){
+
+           cir_plot_filter <- filter(cir_plot_filter, 
+                              transmission %in% transmission_method)
 
            y <- ggplot(
-      data = cir_plot,
+             data = cir_plot_filter,
              mapping = aes(
                x = Geneid,
-        fill = transmission,
                y = tpm
@@
                           geom = "bar",
+                          fill = 'lightblue',
                           width = 0.75, 
@@
                                 limits = limits) +
-      scale_fill_manual(values = c('skyblue','red')) +
              facet_grid(.~subfam, 
                         scales = "free_x",
-                 space = "free_x") +
+                        space = "free_x",
+                        labeller = labeller(subfam = function(string) {
+                          str_remove_all(string, 
+                                         pattern = '(rMT|SBP)_')
+                        })) +
              theme_classic() +
-      theme(
+             ylab('log(TPM + 1)')+
+             ggtitle(paste(unique(cir_plot_filter$time)))
+
+           if(grepl(time_vec, pattern = '20|23')) {
+             y <- y + theme(
                axis.text.x = element_text(size = text_size,
-                                   angle = 90)
-      ) +
-      ylab('log(TPM + 1)')+
-      ggtitle(paste(unique(cir_plot$time)))
+                                          angle = 90,
+                                          vjust = 0.5)
+             )
+           } else {
+             y <- y + theme(
+               axis.text.x = element_blank(),
+               axis.ticks.x = element_blank()
+             ) 
+           }
 
@@
            g <- ggplot_gtable(ggplot_build(y))
+
+           # g$widths[5] = 2*g$widths[5]
+
            strip_t <- which(grepl('strip-t', g$layout$name))
-    subfam_vec <- sort(unique(cir_plot$subfam))
+           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
            fills <- case_when(str_detect(subfam_vec, 
@@
 
+           #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
+           # manually make them a bit wider.
+
+           short_panels <- which(str_detect(subfam_vec, pattern = 'a|S1'))
+
+           for(i in short_panels){
+             num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
+             g$widths[num] = 1.75*g$widths[num]
+             #1.75 times the width of the right panels.
+             # print(num)
+           }
+
            grid.newpage()
            grid.draw(g)
+         })
 
@@
   sort(unique(cir_plot$time)),
-    side_by_side_cir_plot,
-    threshold_min = 100,
+  transmissionSep_cir_plot,
+  threshold_min = 50,
   threshold_max = Inf,
   point_size = 1,
-    text_size = 10,
-    file_name_end = '_Over100TPM'
-  )
-
-```
-
-# Supp figure - Pir expression of AAPLs and ChAPLs.
+  text_size = 7.5
+)
+```
+
+# Supp figure 9 - Pir expression of AAPLs and ChAPLs.
 
@@
 
-```{r chrom_prop_bar_chart}
-tpm_data_chrom <- tpm_data_twentyfour %>%
-  mutate(stage = times_stages_table$time[match(stage,
-                                               times_stages_table$r_stage_diff_count)]) %>%
-  group_by(transmission,
-           stage,
-           chrom) %>%
-  dplyr::summarise(total_tpm = sum(tpm))
-
-y <- ggplot(data = tpm_data_chrom) +
-  geom_bar_pattern(
-    aes(
-      x = stage,
-      y = total_tpm, 
-      fill = chrom,
-      group = transmission,
-      pattern = chrom
-    ),
-    position = 'fill',
-    stat = 'identity',
-    colour = 'black',
-    size = 0.1,
-    pattern_density = 0.05,
-    pattern_spacing = 0.02,
-    pattern_key_scale_factor = 0.5
-  ) +
-  scale_pattern_manual(
-    values = rep(c('none', 
-                   'circle'), 
-                 each = length(c(cbPalette, 
-                                 'red', 
-                                 'white')), 
-                 length.out = length(unique(tpm_data_chrom$chrom))),
-    labels = unique(tpm_data_chrom$chrom),
-    guide = guide_legend(reverse = TRUE),
-    name = 'chromosome'
-  ) +
-  scale_fill_manual(
-    breaks = unique(tpm_data_chrom$chrom),
-    values = rep_len(c(cbPalette, 
-                       'red', 
-                       'white'),
-                     length.out = length(unique(tpm_data_chrom$chrom))),
-    labels =  unique(tpm_data_chrom$chrom),
-    guide = guide_legend(reverse = TRUE),
-    name = 'chromosome'
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  # scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-    scale_size(range = c(0,3)) +
-    ylab(
-      bquote('proportion of '~italic(.('pir'))~' TPM')
-    ) + 
-    facet_grid(.~transmission, 
-               scales = "free_x",
-               space = "free_x") +
-    theme(
-      axis.title.x=element_blank(),
-      axis.text.x=element_text(size = 10, 
-                                 angle = 0),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=10, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-```
-
-### 2D plot of PC1 v PC2 - Supp figure
-
-```{r pca_calculation, results = TRUE}
-twentyfour_tpm <- twentyfour_ct_comb
-rownames_geneid <- twentyfour_tpm$Geneid
-twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
-rownames(twentyfour_tpm_forpca) <- rownames_geneid
-twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
-#Need to look into what exactly scale and center are doing
-```
-
-```{r 2d_pca, results = TRUE}
-twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
-twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
-
-twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
-                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
-                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
-                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
-  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
-  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
-
-colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(twentyfour_tpm_pca_coords$time)))
-
-pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
-pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
-pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
-pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
-
-(
-  plot <- ggplot(twentyfour_tpm_pca_coords,
-                 aes(
-                   x = PC1,
-                   y = PC2,
-                   col = time,
-                   shape = transmission
-                 )) +
-    geom_point(size = 3,
-               alpha = 0.8) +
-    scale_color_manual(values = colornames) +
-    scale_shape_manual(
-      values = setNames(
-        c(17,19),
-        unique(twentyfour_tpm_pca_coords$transmission)
-      )
-    ) +
-    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-    guides(shape=guide_legend(ncol=2),
-           col = guide_legend(ncol=2)) +
-    theme_classic() +
-    theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7))
-)
-
-```
-
-# Calculate and cluster with z-scores {.tabset}
-
-z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:
-
-1. Convert the count/RPKM values of each gene into log values.
-2. Calculate the mean and standard deviation of X gene log values in each sample
-3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
-4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.
-
-* Need to make sure that the z-score is calculated within each transmission method!
-
-```{r zscores}
-tpm_to_zscore = function(input){
-
-  #If the input does not have a TPM element then assume that the input itself is the TPM object
-  if(any(names(input) %in% 'tpm')){
-    tpm_object <- input$tpm
-  }else{
-    tpm_object <- input
-    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
-  }
-
-  #Want all of them so will just ignore Geneid
-  input_logged <- tpm_object %>%
-    mutate_at(vars(-matches('Geneid')), 
-              log1p)
-
-  gene_mean_scores = input_logged %>% 
-    melt(id.vars = 'Geneid') %>% 
-    group_by(Geneid) %>% 
-    dplyr::summarise(mean = mean(value),
-                     std_dev = sd(value))
-
-  gene_vector = as.character(input_logged$Geneid)
-
-  input_z = lapply(gene_vector, function(gene){
-    mean_gene = filter(gene_mean_scores, 
-                       Geneid == paste(gene)) %>% 
-      select(mean)
-    std_gene = filter(gene_mean_scores, 
-                      Geneid == paste(gene)) %>% 
-      select(std_dev)
-    gene_scores = input_logged %>% 
-      filter(Geneid == paste(gene)) %>% 
-      select(-Geneid)
-    if(std_gene != 0){
-      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
-    }else{
-      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
-    }
-    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
-    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
-  })
-
-  #Use reduce and rbind to turn the list into one df
-  input_z_df = purrr::reduce(input_z, 
-                             rbind)
-  return(list(zscores = input_z_df,
-              tpm = tpm_object,
-              experiment = input$experiment))
-}
-
-#Make and save z-scores
-twentyfour_z <- lapply(twentyfour_avgtpm,
-                       tpm_to_zscore)
-
-###
-```
-
-# Phaseograms of the data.
-
-```{r row_order_phaseograms}
-
-#Getting the row order for each phaseogram for each experiment
-# Want the data ordered by the highest zscore for each gene in each stage.
-phaseogram_row_order <- lapply(twentyfour_z, 
-                               function(list_elements){
-                                 zscore_df <- list_elements$zscores
-                                 zscore_ordered <- melt(zscore_df, 
-                                                        id = 'Geneid',
-                                                        variable.name = 'stage',
-                                                        value.name = 'zscore') %>% 
-                                   group_by(Geneid) %>% 
-                                   top_n(zscore, n = 1) %>% 
-                                   arrange(stage, 
-                                           desc(zscore)) %>%
-                                   distinct(Geneid, 
-                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-                                   filter(zscore > 0) #For some reason some remain with zero zscore
-                                 # zscore_ordered <- zscore_ordered[order(
-                                 #   match(
-                                 #     zscore_ordered$stage, 
-                                 #     order_vector
-                                 #   )
-                                 # ),]
-                                 return(c(list_elements, 
-                                          list(zscore_order = zscore_ordered)))
-                               })
-
-phaseogram_time_order <- lapply(names(phaseogram_row_order), 
-                                function(list_element){
-                                  transmission_list <- getElement(phaseogram_row_order,
-                                                                  list_element)
-                                  order_by_time <- transmission_list$zscore_order %>% 
-                                    mutate(stage = times_stages_table$time[match(stage,
-                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
-                                    arrange(stage)
-                                  return(list(zscore_order = order_by_time))
-                                }
-)
-names(phaseogram_time_order) <- names(phaseogram_row_order)
-
-```
-
-```{r function_draw_heatmap_zscore}
-
-draw_heatmap_zscore <- function(zscore_list,
-                                title = '',
-                                gene_ordering_vector = NULL,
-                                filter_genes_vector = NULL,
-                                extra_title = NULL,
-                                annotation_clade = NULL,
-                                annotation_col,
-                                threshold_tpm = NULL,
-                                use_times = FALSE,
-                                use_order_vector = TRUE,
-                                tpm_mat = TRUE,
-                                tpm_breaks = NULL,
-                                ...){
-  if(is.null(tpm_breaks) & tpm_mat){
-    #Need the generate the breaks in the data for the heatmap legend if not provided.
-    max_tpm <- lapply(zscore_list, 
-                      function(list_element){
-                        #list_element <- zscore_list[[1]]
-                        heatmap_data <- as.data.frame(list_element$zscores)
-                        title <- list_element$experiment
-
-                        if(!'Geneid' %in% colnames(heatmap_data)){
-                          stop('No Geneid column in the heatmap data.frame')
-                        }
-
-                        if(!is.null(filter_genes_vector)){
-                          heatmap_data <- as.data.frame(
-                            filter(heatmap_data, 
-                                   Geneid %in% filter_genes_vector)
-                          )
-                        }
-
-                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                        if(!is.null(gene_ordering_vector)){
-                          #Remove genes not in the main data frame but in the ordering vector, 
-                          # and then order based on the vector
-                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          heatmap_data <- heatmap_data[na.omit(
-                            match(gene_ordering_vector, 
-                                  rownames(heatmap_data))
-                          ),,
-                          drop=FALSE]
-                        }
-
-                        geneids <- rownames(heatmap_data)
-
-                        #Get max/min TPM scores for the side annotation heatmap                       
-                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                          list_element$tpm$Geneid),] %>% 
-                          select(-Geneid) %>% 
-                          as.matrix %>% 
-                          log1p
-                        rownames(tpm_mat) <- geneids
-
-                        tpm_min <- rowMins(tpm_mat)
-                        names(tpm_min) <- geneids
-                        tpm_max <- rowMaxs(tpm_mat)
-                        names(tpm_max) <- geneids
-                        return(list(tpm_max = tpm_max,
-                                    tpm_min = tpm_min))
-                      })
-
-    tpm_max_all <- mround(
-      max(
-        unlist(
-          lapply(max_tpm, 
-                 getElement,
-                 'tpm_max')
-        )
-      ), 
-      base = 1)
-
-    tpm_breaks <- round(
-      c(0,
-        tpm_max_all/4,
-        tpm_max_all/2,
-        tpm_max_all*(3/4),
-        tpm_max_all),
-      digits = 2
-    )
-  }
-
-  #Generate the colour function from the TPM legend breaks for the TPM annotation
-  if(tpm_mat){
-    tpm_col_fun <- colorRamp2(
-      breaks = tpm_breaks,
-      colors = viridis(length(tpm_breaks))
-    )
-  }
-
-  if(is.numeric(threshold_tpm)){
-    #Apply a threshold to the TPM in order to remove low expressed genes.
-    # Remember to add log1p here! 
-    # Threshold uses the max value, so it ony needs to be above thresh 
-    #   in one stage avg in one transmission method.
-    #Will alter the gene_filter vector in order to only select the genes above threshold
-
-    gene_maxs <- bind_rows(
-      lapply(max_tpm, 
-             getElement,
-             'tpm_max')
-    )
-
-    remove_genes <- names(gene_maxs)[apply(
-      gene_maxs, 
-      2, 
-      function(maxs){
-        all(maxs < log2(1+threshold_tpm))
-      }
-    )]
-
-    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
-
-  }
-
-  if(is.character(annotation_clade)){
-    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
-    clade_info <- getElement(cir_info,
-                             annotation_clade)[match(order_cir, 
-                                                     cir_info$Geneid)]
-    names(clade_info) <- str_split_fixed(order_cir, 
-                                         pattern = '_', 
-                                         n = 2)[,2]
-    annotation_col <- annotation_col
-  }
-
-  heatmaps <- lapply(zscore_list, 
-                     function(list_element){
-                       #list_element <- zscore_list[[1]]
-                       heatmap_data <- as.data.frame(list_element$zscores)
-                       title <- list_element$experiment
-
-                       if(!'Geneid' %in% colnames(heatmap_data)){
-                         stop('No Geneid column in the heatmap data.frame')
-                       }
-
-                       if(!is.null(filter_genes_vector)){
-                         heatmap_data <- as.data.frame(
-                           filter(heatmap_data, 
-                                  Geneid %in% filter_genes_vector)
-                         )
-                       }
-
-                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                       if(!is.null(gene_ordering_vector)){
-                         #Remove genes not in the main data frame but in the ordering vector,
-                         #  and then order based on the vector
-                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                         heatmap_data <- heatmap_data[na.omit(
-                           match(gene_ordering_vector, 
-                                 rownames(heatmap_data))
-                         ),,
-                         drop=FALSE]
-                       }
-
-                       if(tpm_mat){
-                         #Get max/min TPM scores for the side annotation heatmap                       
-                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                           list_element$tpm$Geneid),] %>%
-                           select(-Geneid) %>% 
-                           as.matrix %>% 
-                           log1p
-
-                         tpm_min <- rowMins(tpm_mat)
-                         tpm_max <- rowMaxs(tpm_mat)
-
-                         names(tpm_min) = rownames(heatmap_data)
-                         names(tpm_max) = rownames(heatmap_data)
-
-                         tpm_ha <- rowAnnotation(
-                           min.max = anno_simple(cbind(tpm_min, 
-                                                       tpm_max), 
-                                                 col = tpm_col_fun,
-                                                 gp = gpar(col = 'black',
-                                                           lwd = 0.1),
-                                                 border = TRUE,
-                                                 simple_anno_size = unit(0.2, "cm")),
-                           show_annotation_name = FALSE)
-                       }
-
-                       #Draw the left annotation is needed, only for the first element.
-
-                       if(
-                         identical(list_element,
-                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
-                       ){
-                         clade_ha <- rowAnnotation(
-                           'subfamily' = anno_simple(
-                             clade_info,
-                             col = annotation_col
-                           )
-                         )
-                       }
-
-                       #Change rownames to remove PCHAS_
-                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
-                                                                pattern = '_', 
-                                                                n = 2)[,2]
-
-                       if(use_times){
-                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
-                                                                                 times_stages_table$r_stage_diff_count)]
-                       }
-
-                       if(use_order_vector){
-                         #Order the columns by ordering_vector
-                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
-                                                                   order_vector)),
-                                                      drop = FALSE]
-                       }else{
-                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
-                                                      drop = FALSE] 
-                       }
-
-                       h = Heatmap(heatmap_data,
-                                   name = paste0('z-score',
-                                                 extra_title),
-                                   col = c(cbPalette[6],
-                                           'white', 
-                                           cbPalette[7]),
-                                   row_names_gp = gpar(fontsize = ifelse(
-                                     nrow(heatmap_data) > 13, 
-                                     12*(13/nrow(heatmap_data)), 
-                                     12)
-                                   ),
-                                   column_names_gp = gpar(fontsize = 10),
-                                   row_dend_width = unit(3, "cm"),
-                                   right_annotation = if(tpm_mat){
-                                     tpm_ha
-                                   }else{
-                                     NULL
-                                   },
-                                   left_annotation = 
-                                     if(identical(list_element,
-                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
-                                       clade_ha
-                                     }else{
-                                       NULL
-                                     }, 
-                                   #Only want the clade annotation when it is first heatmap
-                                   width = unit(4.5, 'cm'),
-                                   cluster_columns = FALSE,
-                                   height = unit(7,'cm'),
-                                   row_title = NULL,
-                                   cluster_rows = FALSE,
-                                   column_title = paste(title),
-                                   heatmap_legend_param = list(
-                                     at = c(3,2,1,0,-1,-2,-3),
-                                     labels_gp = gpar(fontsize = 10),
-                                     title_gp = gpar(fontsize = 10)
-                                   ),
-                                   ...
-                       )
-                       return(h)
-                     })
-
-  ht_list <- Reduce('+', heatmaps) #To add the list elements together
-
-  if(tpm_mat){
-    #Draw the legend
-    tpm_lgd <- Legend(
-      title = "Min/Max log(TPM + 1)", 
-      col = tpm_col_fun, 
-      at = tpm_breaks, 
-      labels = tpm_breaks,
-      title_gp = gpar(fontsize = 7)
-    )
-  }
-
-  if(is.character(annotation_clade)){
-    clade_lgd <- Legend(
-      title = 'sub-family',
-      at = names(annotation_col),
-      legend_gp = gpar(fill = annotation_col)
-    )
-  }
-
-  draw(ht_list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'left',
-       #Only include the annotation legends for the annotations we actually have.
-       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
-         list(tpm_lgd,
-              clade_lgd)
-       }else if(tpm_mat & !is.character(annotation_clade) ){
-         list(tpm_lgd)
-       } else if(!tpm_mat & is.character(annotation_clade) ){
-         list(clade_lgd)
-       })
-}
-```
-
-## Phaseograms of the entire genome.
-
-```{r phaseograms_whole_genome}
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
-```
-
-## Phaseograms of the _cir_ genes with clade annotations.
-
-```{r pir_gene_phaseogram_breaks}
-  max_tpm <- lapply(twentyfour_z, 
-                    function(list_element){
-                      #list_element <- zscore_list[[1]]
-                      heatmap_data <- as.data.frame(list_element$zscores)
-                      title <- list_element$experiment
-
-                      if(!'Geneid' %in% colnames(heatmap_data)){
-                        stop('No Geneid column in the heatmap data.frame')
-                      }
-
-                            heatmap_data <- as.data.frame(
-                              filter(heatmap_data, 
-                                     Geneid %in% cir_id)
-                            )
-
-                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-                          # 
-                          # if(!is.null(gene_ordering_vector)){
-                          #   #Remove genes not in the main data frame but in the ordering vector, 
-                          #   # and then order based on the vector
-                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          #   heatmap_data <- heatmap_data[na.omit(
-                          #     match(gene_ordering_vector, 
-                          #           rownames(heatmap_data))
-                          #   ),,
-                          #   drop=FALSE]
-                          # }
-
-                          geneids <- rownames(heatmap_data)
-
-                          #Get max/min TPM scores for the side annotation heatmap                       
-                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                            list_element$tpm$Geneid),] %>% 
-                            select(-Geneid) %>% 
-                            as.matrix %>% 
-                            log1p
-                          rownames(tpm_mat) <- geneids
-
-                          tpm_min <- rowMins(tpm_mat)
-                          names(tpm_min) <- geneids
-                          tpm_max <- rowMaxs(tpm_mat)
-                          names(tpm_max) <- geneids
-                          return(list(tpm_max = tpm_max,
-                                      tpm_min = tpm_min))
-                        })
-
-  tpm_max_all <- mround(
-    max(
-      unlist(
-        lapply(max_tpm, 
-               getElement,
-               'tpm_max')
-      )
-    ), 
-    base = 1)
-
-  tpm_breaks <- round(
-    c(0,
-      tpm_max_all/4,
-      tpm_max_all/2,
-      tpm_max_all*(3/4),
-      tpm_max_all),
-    digits = 2
-  )
-```
-
-```{r phaseograms_cir}
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_breaks = tpm_breaks,
-                    tpm_mat = FALSE)
-
-```
-
-## Figure 3b - Highest expressed all _pir_ heatmap
-
-```{r highest_expressed_stage_heatmap_fig3b}
-
-num_SorL <- cir_info %>% 
-  group_by(SorL) %>% 
-  summarise(num = n_distinct(Gene.ID))
-
-labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
-                        num_SorL$SorL)
-
-# Get order of pir genes by highest expressed genes in each sample.
-
-combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm), 
-       function(list_name){
-
-         combined_data_pir <- getElement(twentyfour_avgtpm, 
-                                     list_name)$tpm %>% 
-           filter(Geneid %in% cir_id) %>% 
-           column_to_rownames(var = 'Geneid')
-
-         #Ordering that a more understandable order
-
-         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
-                                                            order_vector)),
-                                                drop = FALSE]
-
-         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
-                                         means = apply(combined_data_pir, 1, mean),
-                                         subfam = cir_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
-                                                                            cir_info$Geneid)]) %>% 
-           filter(means > 1) %>% 
-           arrange(subfam, 
-                   desc(means)) %>% 
-           .$Geneid
-
-         #Order the heatmap data by the order defined above (SorL, then mean expression).
-         combined_data_pir_ht <- combined_data_pir[
-           order(
-             match(rownames(combined_data_pir),
-                   combined_data_ordered),
-             na.last = NA #Remove NAs
-           ),
-         ]
-
-         #Change the colnames to the times
-         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h')
-
-         #Order by time point
-         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                      drop = FALSE] 
-         #Add a column for the SorL
-         combined_data_pir_ht$SorL <- cir_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
-                                                                cir_info$Geneid)]
-
-         return(combined_data_pir_ht)
-       })
-names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)
-
-maxcol <- ceiling(
-  max(
-    unlist(
-      lapply(combined_data_pir_ht_list,
-             function(list_element){
-               max(
-                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
-               )
-             }
-      )
-    )
-  )
-)
-
-lapply(names(combined_data_pir_ht_list),
-       function(list_name_to_order){
-
-         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
-                                            list_name_to_order)
-
-         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
-         #Sub-family annotation
-         clade_info <- combined_data_pir_ht$SorL
-         annotation_col <- SorL_col
-
-         #Subfamily annotation
-
-         clade_ha <- rowAnnotation(
-           'sub-family' = anno_simple(
-             clade_info,
-             col = annotation_col,
-           ),
-           annotation_name_gp = gpar(fontsize = 10)
-         )
-
-         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
-                                                 labels = 'ancestral',
-                                                 labels_gp = gpar(fontsize = 7),
-                                                 link_width = unit(3, "mm")))
-
-         #lapply across the datasets now, for each order
-         ht_list <- lapply(names(combined_data_pir_ht_list),
-                           function(list_name){
-
-                             combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
-                                                                list_name)$tpm %>% 
-                               column_to_rownames(var = 'Geneid')
-
-                             #Get the times instead of the stage names.
-                             colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h')
-
-                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                                          drop = FALSE]
-
-                             Heatmap(
-                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
-                               cluster_rows = FALSE,
-                               cluster_columns = FALSE,
-                               show_row_names = FALSE,
-                               name = 'log(TPM+1)', 
-                               col = colorRamp2(c(0,
-                                                  maxcol/4,
-                                                  maxcol/2,
-                                                  3*maxcol/4,
-                                                  maxcol),
-                                                viridis(5)),
-                               na_col = 'white', 
-                               border = TRUE,
-                               border_gp = gpar(lwd = 0.1),
-                               column_names_gp = gpar(fontsize = 10),
-                               row_dend_width = unit(3, "cm"),
-                               width = unit(4.5, 'cm'),
-                               height = unit(7,'cm'),
-                               row_title = NULL,
-                               left_annotation = if(
-                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
-                               ){
-                                 clade_ha
-                               }else{
-                                 NULL
-                               },
-                               right_annotation = if(
-                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
-                               ){
-                                 anc_ha
-                               }else{
-                                 NULL
-                               },
-                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
-                               #                                        cir_info$Geneid)],
-                               column_title = paste(list_name),
-                               heatmap_legend_param = list(
-                                 labels_gp = gpar(fontsize = 10),
-                                 title_gp = gpar(fontsize = 10,
-                                                 fontface = 'bold'),
-                                 at = c(0,
-                                        maxcol/4,
-                                        maxcol/2,
-                                        3*maxcol/4,
-                                        maxcol)
-                               )
-                             )
-                           }
-         )
-         ht_list <- Reduce(`+`, ht_list)
-         clade_lgd <- Legend(
-           title = 'sub-family',
-           at = labels_SorL,
-           legend_gp = gpar(fill = SorL_col)
-         )
-
-         return(
-           draw(ht_list, 
-                row_title_side = 'left',
-                column_title_side = 'bottom',
-                heatmap_legend_side = 'right',
-                annotation_legend_side = 'right',
-                annotation_legend_list = list(clade_lgd),
-                ht_gap = unit(.35,'cm'),
-                merge_legend = TRUE
-           ))
-       })
-
-```
-
-```{r figure_3b_suppinfo}
-
-combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
-                                            'rMT')
-
-         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
-supp_info_fig_3b <- lapply(names(combined_data_pir_ht_list),
-       function(list_name){
-
-         combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
-                                            list_name)$tpm %>% 
-           column_to_rownames(var = 'Geneid')
-
-         #Get the times instead of the stage names.
-         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                       pattern = '[[:digit:]]{2}h')
-
-         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                      drop = FALSE]
-
-         colnames(combined_data_pir_ht) <- paste0(list_name, 
-                                                  '_', 
-                                                  colnames(combined_data_pir_ht))
-
-         log2(1+combined_data_pir_ht)[pir_highest_tpm_order,]
-       }) %>% 
-  bind_cols
-
-supp_info_fig_3b$SorL <- cir_info$SorL[match(names(apply(supp_info_fig_3b, 1, mean)), 
-                                                   cir_info$Geneid)]
-
-head(supp_info_fig_3b)
-
-```
-
 # Lower pir expression vs rest of genome

### 2021-03-15 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Added supp fig names and reorder figures

@@
 
-### 2D plot of PC1 v PC2 - Supp figure 4A
-
-```{r pca_calculation, results = TRUE}
-twentyfour_tpm <- twentyfour_ct_comb
-rownames_geneid <- twentyfour_tpm$Geneid
-twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
-rownames(twentyfour_tpm_forpca) <- rownames_geneid
-twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
-#Need to look into what exactly scale and center are doing
-```
-
-```{r 2d_pca, results = TRUE}
-twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
-twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
-
-twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
-                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
-                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
-                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
-  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
-  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
-
-colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(twentyfour_tpm_pca_coords$time)))
-
-pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
-pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
-pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
-pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
-
-(
-  plot <- ggplot(twentyfour_tpm_pca_coords,
-                 aes(
-                   x = PC1,
-                   y = PC2,
-                   col = time,
-                   shape = transmission
-                 )) +
-    geom_point(size = 3,
-               alpha = 0.8) +
-    scale_color_manual(values = colornames) +
-    scale_shape_manual(
-      values = setNames(
-        c(17,19),
-        unique(twentyfour_tpm_pca_coords$transmission)
-      )
-    ) +
-    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-    guides(shape=guide_legend(ncol=2),
-           col = guide_legend(ncol=2)) +
-    theme_classic() +
-    theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7))
-)
-
-```
-
-# Calculate and cluster with z-scores {.tabset}
-
-z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:
-
-1. Convert the count/RPKM values of each gene into log values.
-2. Calculate the mean and standard deviation of X gene log values in each sample
-3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
-4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.
-
-* Need to make sure that the z-score is calculated within each transmission method!
-
-```{r zscores}
-tpm_to_zscore = function(input){
-
-  #If the input does not have a TPM element then assume that the input itself is the TPM object
-  if(any(names(input) %in% 'tpm')){
-    tpm_object <- input$tpm
-  }else{
-    tpm_object <- input
-    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
-  }
-
-  #Want all of them so will just ignore Geneid
-  input_logged <- tpm_object %>%
-    mutate_at(vars(-matches('Geneid')), 
-              log1p)
-
-  gene_mean_scores = input_logged %>% 
-    melt(id.vars = 'Geneid') %>% 
-    group_by(Geneid) %>% 
-    dplyr::summarise(mean = mean(value),
-                     std_dev = sd(value))
-
-  gene_vector = as.character(input_logged$Geneid)
-
-  input_z = lapply(gene_vector, function(gene){
-    mean_gene = filter(gene_mean_scores, 
-                       Geneid == paste(gene)) %>% 
-      select(mean)
-    std_gene = filter(gene_mean_scores, 
-                      Geneid == paste(gene)) %>% 
-      select(std_dev)
-    gene_scores = input_logged %>% 
-      filter(Geneid == paste(gene)) %>% 
-      select(-Geneid)
-    if(std_gene != 0){
-      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
-    }else{
-      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
-    }
-    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
-    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
-  })
-
-  #Use reduce and rbind to turn the list into one df
-  input_z_df = purrr::reduce(input_z, 
-                             rbind)
-  return(list(zscores = input_z_df,
-              tpm = tpm_object,
-              experiment = input$experiment))
-}
-
-#Make and save z-scores
-twentyfour_z <- lapply(twentyfour_avgtpm,
-                       tpm_to_zscore)
-
-###
-```
-
-# Phaseograms of the data.
-
-```{r row_order_phaseograms}
-
-#Getting the row order for each phaseogram for each experiment
-# Want the data ordered by the highest zscore for each gene in each stage.
-phaseogram_row_order <- lapply(twentyfour_z, 
-                               function(list_elements){
-                                 zscore_df <- list_elements$zscores
-                                 zscore_ordered <- melt(zscore_df, 
-                                                        id = 'Geneid',
-                                                        variable.name = 'stage',
-                                                        value.name = 'zscore') %>% 
-                                   group_by(Geneid) %>% 
-                                   top_n(zscore, n = 1) %>% 
-                                   arrange(stage, 
-                                           desc(zscore)) %>%
-                                   distinct(Geneid, 
-                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-                                   filter(zscore > 0) #For some reason some remain with zero zscore
-                                 # zscore_ordered <- zscore_ordered[order(
-                                 #   match(
-                                 #     zscore_ordered$stage, 
-                                 #     order_vector
-                                 #   )
-                                 # ),]
-                                 return(c(list_elements, 
-                                          list(zscore_order = zscore_ordered)))
-                               })
-
-phaseogram_time_order <- lapply(names(phaseogram_row_order), 
-                                function(list_element){
-                                  transmission_list <- getElement(phaseogram_row_order,
-                                                                  list_element)
-                                  order_by_time <- transmission_list$zscore_order %>% 
-                                    mutate(stage = times_stages_table$time[match(stage,
-                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
-                                    arrange(stage)
-                                  return(list(zscore_order = order_by_time))
-                                }
-)
-names(phaseogram_time_order) <- names(phaseogram_row_order)
-
-```
-
-```{r function_draw_heatmap_zscore}
-
-draw_heatmap_zscore <- function(zscore_list,
-                                title = '',
-                                gene_ordering_vector = NULL,
-                                filter_genes_vector = NULL,
-                                extra_title = NULL,
-                                annotation_clade = NULL,
-                                annotation_col,
-                                threshold_tpm = NULL,
-                                use_times = FALSE,
-                                use_order_vector = TRUE,
-                                tpm_mat = TRUE,
-                                tpm_breaks = NULL,
-                                ...){
-  if(is.null(tpm_breaks) & tpm_mat){
-    #Need the generate the breaks in the data for the heatmap legend if not provided.
-    max_tpm <- lapply(zscore_list, 
-                      function(list_element){
-                        #list_element <- zscore_list[[1]]
-                        heatmap_data <- as.data.frame(list_element$zscores)
-                        title <- list_element$experiment
-
-                        if(!'Geneid' %in% colnames(heatmap_data)){
-                          stop('No Geneid column in the heatmap data.frame')
-                        }
-
-                        if(!is.null(filter_genes_vector)){
-                          heatmap_data <- as.data.frame(
-                            filter(heatmap_data, 
-                                   Geneid %in% filter_genes_vector)
-                          )
-                        }
-
-                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                        if(!is.null(gene_ordering_vector)){
-                          #Remove genes not in the main data frame but in the ordering vector, 
-                          # and then order based on the vector
-                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          heatmap_data <- heatmap_data[na.omit(
-                            match(gene_ordering_vector, 
-                                  rownames(heatmap_data))
-                          ),,
-                          drop=FALSE]
-                        }
-
-                        geneids <- rownames(heatmap_data)
-
-                        #Get max/min TPM scores for the side annotation heatmap                       
-                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                          list_element$tpm$Geneid),] %>% 
-                          select(-Geneid) %>% 
-                          as.matrix %>% 
-                          log1p
-                        rownames(tpm_mat) <- geneids
-
-                        tpm_min <- rowMins(tpm_mat)
-                        names(tpm_min) <- geneids
-                        tpm_max <- rowMaxs(tpm_mat)
-                        names(tpm_max) <- geneids
-                        return(list(tpm_max = tpm_max,
-                                    tpm_min = tpm_min))
-                      })
-
-    tpm_max_all <- mround(
-      max(
-        unlist(
-          lapply(max_tpm, 
-                 getElement,
-                 'tpm_max')
-        )
-      ), 
-      base = 1)
-
-    tpm_breaks <- round(
-      c(0,
-        tpm_max_all/4,
-        tpm_max_all/2,
-        tpm_max_all*(3/4),
-        tpm_max_all),
-      digits = 2
-    )
-  }
-
-  #Generate the colour function from the TPM legend breaks for the TPM annotation
-  if(tpm_mat){
-    tpm_col_fun <- colorRamp2(
-      breaks = tpm_breaks,
-      colors = viridis(length(tpm_breaks))
-    )
-  }
-
-  if(is.numeric(threshold_tpm)){
-    #Apply a threshold to the TPM in order to remove low expressed genes.
-    # Remember to add log1p here! 
-    # Threshold uses the max value, so it ony needs to be above thresh 
-    #   in one stage avg in one transmission method.
-    #Will alter the gene_filter vector in order to only select the genes above threshold
-
-    gene_maxs <- bind_rows(
-      lapply(max_tpm, 
-             getElement,
-             'tpm_max')
-    )
-
-    remove_genes <- names(gene_maxs)[apply(
-      gene_maxs, 
-      2, 
-      function(maxs){
-        all(maxs < log2(1+threshold_tpm))
-      }
-    )]
-
-    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
-
-  }
-
-  if(is.character(annotation_clade)){
-    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
-    clade_info <- getElement(cir_info,
-                             annotation_clade)[match(order_cir, 
-                                                     cir_info$Geneid)]
-    names(clade_info) <- str_split_fixed(order_cir, 
-                                         pattern = '_', 
-                                         n = 2)[,2]
-    annotation_col <- annotation_col
-  }
-
-  heatmaps <- lapply(zscore_list, 
-                     function(list_element){
-                       #list_element <- zscore_list[[1]]
-                       heatmap_data <- as.data.frame(list_element$zscores)
-                       title <- list_element$experiment
-
-                       if(!'Geneid' %in% colnames(heatmap_data)){
-                         stop('No Geneid column in the heatmap data.frame')
-                       }
-
-                       if(!is.null(filter_genes_vector)){
-                         heatmap_data <- as.data.frame(
-                           filter(heatmap_data, 
-                                  Geneid %in% filter_genes_vector)
-                         )
-                       }
-
-                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                       if(!is.null(gene_ordering_vector)){
-                         #Remove genes not in the main data frame but in the ordering vector,
-                         #  and then order based on the vector
-                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                         heatmap_data <- heatmap_data[na.omit(
-                           match(gene_ordering_vector, 
-                                 rownames(heatmap_data))
-                         ),,
-                         drop=FALSE]
-                       }
-
-                       if(tpm_mat){
-                         #Get max/min TPM scores for the side annotation heatmap                       
-                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                           list_element$tpm$Geneid),] %>%
-                           select(-Geneid) %>% 
-                           as.matrix %>% 
-                           log1p
-
-                         tpm_min <- rowMins(tpm_mat)
-                         tpm_max <- rowMaxs(tpm_mat)
-
-                         names(tpm_min) = rownames(heatmap_data)
-                         names(tpm_max) = rownames(heatmap_data)
-
-                         tpm_ha <- rowAnnotation(
-                           min.max = anno_simple(cbind(tpm_min, 
-                                                       tpm_max), 
-                                                 col = tpm_col_fun,
-                                                 gp = gpar(col = 'black',
-                                                           lwd = 0.1),
-                                                 border = TRUE,
-                                                 simple_anno_size = unit(0.2, "cm")),
-                           show_annotation_name = FALSE)
-                       }
-
-                       #Draw the left annotation is needed, only for the first element.
-
-                       if(
-                         identical(list_element,
-                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
-                       ){
-                         clade_ha <- rowAnnotation(
-                           'subfamily' = anno_simple(
-                             clade_info,
-                             col = annotation_col
-                           )
-                         )
-                       }
-
-                       #Change rownames to remove PCHAS_
-                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
-                                                                pattern = '_', 
-                                                                n = 2)[,2]
-
-                       if(use_times){
-                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
-                                                                                 times_stages_table$r_stage_diff_count)]
-                       }
-
-                       if(use_order_vector){
-                         #Order the columns by ordering_vector
-                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
-                                                                   order_vector)),
-                                                      drop = FALSE]
-                       }else{
-                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
-                                                      drop = FALSE] 
-                       }
-
-                       h = Heatmap(heatmap_data,
-                                   name = paste0('z-score',
-                                                 extra_title),
-                                   col = c(cbPalette[6],
-                                           'white', 
-                                           cbPalette[7]),
-                                   row_names_gp = gpar(fontsize = ifelse(
-                                     nrow(heatmap_data) > 13, 
-                                     12*(13/nrow(heatmap_data)), 
-                                     12)
-                                   ),
-                                   column_names_gp = gpar(fontsize = 10),
-                                   row_dend_width = unit(3, "cm"),
-                                   right_annotation = if(tpm_mat){
-                                     tpm_ha
-                                   }else{
-                                     NULL
-                                   },
-                                   left_annotation = 
-                                     if(identical(list_element,
-                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
-                                       clade_ha
-                                     }else{
-                                       NULL
-                                     }, 
-                                   #Only want the clade annotation when it is first heatmap
-                                   width = unit(4.5, 'cm'),
-                                   cluster_columns = FALSE,
-                                   height = unit(7,'cm'),
-                                   row_title = NULL,
-                                   cluster_rows = FALSE,
-                                   column_title = paste(title),
-                                   heatmap_legend_param = list(
-                                     at = c(3,2,1,0,-1,-2,-3),
-                                     labels_gp = gpar(fontsize = 10),
-                                     title_gp = gpar(fontsize = 10)
-                                   ),
-                                   ...
-                       )
-                       return(h)
-                     })
-
-  ht_list <- Reduce('+', heatmaps) #To add the list elements together
-
-  if(tpm_mat){
-    #Draw the legend
-    tpm_lgd <- Legend(
-      title = "Min/Max log(TPM + 1)", 
-      col = tpm_col_fun, 
-      at = tpm_breaks, 
-      labels = tpm_breaks,
-      title_gp = gpar(fontsize = 7)
-    )
-  }
-
-  if(is.character(annotation_clade)){
-    clade_lgd <- Legend(
-      title = 'sub-family',
-      at = names(annotation_col),
-      legend_gp = gpar(fill = annotation_col)
-    )
-  }
-
-  draw(ht_list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'left',
-       #Only include the annotation legends for the annotations we actually have.
-       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
-         list(tpm_lgd,
-              clade_lgd)
-       }else if(tpm_mat & !is.character(annotation_clade) ){
-         list(tpm_lgd)
-       } else if(!tpm_mat & is.character(annotation_clade) ){
-         list(clade_lgd)
-       })
-}
-```
-
-## Phaseograms of the entire genome.
-
-```{r phaseograms_whole_genome}
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
-```
-
-## Phaseograms of the _cir_ genes with clade annotations.
-
-```{r pir_gene_phaseogram_breaks}
-  max_tpm <- lapply(twentyfour_z, 
-                    function(list_element){
-                      #list_element <- zscore_list[[1]]
-                      heatmap_data <- as.data.frame(list_element$zscores)
-                      title <- list_element$experiment
-
-                      if(!'Geneid' %in% colnames(heatmap_data)){
-                        stop('No Geneid column in the heatmap data.frame')
-                      }
-
-                            heatmap_data <- as.data.frame(
-                              filter(heatmap_data, 
-                                     Geneid %in% cir_id)
-                            )
-
-                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-                          # 
-                          # if(!is.null(gene_ordering_vector)){
-                          #   #Remove genes not in the main data frame but in the ordering vector, 
-                          #   # and then order based on the vector
-                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          #   heatmap_data <- heatmap_data[na.omit(
-                          #     match(gene_ordering_vector, 
-                          #           rownames(heatmap_data))
-                          #   ),,
-                          #   drop=FALSE]
-                          # }
-
-                          geneids <- rownames(heatmap_data)
-
-                          #Get max/min TPM scores for the side annotation heatmap                       
-                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                            list_element$tpm$Geneid),] %>% 
-                            select(-Geneid) %>% 
-                            as.matrix %>% 
-                            log1p
-                          rownames(tpm_mat) <- geneids
-
-                          tpm_min <- rowMins(tpm_mat)
-                          names(tpm_min) <- geneids
-                          tpm_max <- rowMaxs(tpm_mat)
-                          names(tpm_max) <- geneids
-                          return(list(tpm_max = tpm_max,
-                                      tpm_min = tpm_min))
-                        })
-
-  tpm_max_all <- mround(
-    max(
-      unlist(
-        lapply(max_tpm, 
-               getElement,
-               'tpm_max')
-      )
-    ), 
-    base = 1)
-
-  tpm_breaks <- round(
-    c(0,
-      tpm_max_all/4,
-      tpm_max_all/2,
-      tpm_max_all*(3/4),
-      tpm_max_all),
-    digits = 2
-  )
-```
-
-```{r phaseograms_cir}
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_breaks = tpm_breaks,
-                    tpm_mat = FALSE)
-
-```
-
-## Figure 3b - Highest expressed all _pir_ heatmap
+## Figure 3c and Supp figure 3 - Highest expressed all _pir_ heatmap
 
@@
 
-```{r figure_3b_suppinfo}
-
-combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
-                                            'rMT')
-
-         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
-supp_info_fig_3b <- lapply(names(combined_data_pir_ht_list),
-       function(list_name){
-
-         combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
-                                            list_name)$tpm %>% 
-           column_to_rownames(var = 'Geneid')
-
-         #Get the times instead of the stage names.
-         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                       pattern = '[[:digit:]]{2}h')
-
-         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+## 2D plot of PC1 v PC2 - Supp figure 4A
+
+```{r pca_calculation, results = TRUE}
+twentyfour_tpm <- twentyfour_ct_comb
+rownames_geneid <- twentyfour_tpm$Geneid
+twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
+rownames(twentyfour_tpm_forpca) <- rownames_geneid
+twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
+#Need to look into what exactly scale and center are doing
+```
+
+```{r 2d_pca, results = TRUE}
+twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
+twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
+
+twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
+                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
+                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
+                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
+  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
+  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
+  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
+
+colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
+colornames <- setNames(colornames, 
+                       sort(unique(twentyfour_tpm_pca_coords$time)))
+
+pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
+pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
+pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
+pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
+
+(
+  plot <- ggplot(twentyfour_tpm_pca_coords,
+                 aes(
+                   x = PC1,
+                   y = PC2,
+                   col = time,
+                   shape = transmission
+                 )) +
+    geom_point(size = 3,
+               alpha = 0.8) +
+    scale_color_manual(values = colornames) +
+    scale_shape_manual(
+      values = setNames(
+        c(17,19),
+        unique(twentyfour_tpm_pca_coords$transmission)
+      )
+    ) +
+    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
+    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
+    guides(shape=guide_legend(ncol=2),
+           col = guide_legend(ncol=2)) +
+    theme_classic() +
+    theme(legend.key.size = unit(4, 'mm'),
+          legend.text = element_text(size = 7))
+)
+
+```
+
+# Calculate and cluster with z-scores {.tabset}
+
+z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:
+
+1. Convert the count/RPKM values of each gene into log values.
+2. Calculate the mean and standard deviation of X gene log values in each sample
+3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
+4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.
+
+* Need to make sure that the z-score is calculated within each transmission method!
+
+```{r zscores}
+tpm_to_zscore = function(input){
+
+  #If the input does not have a TPM element then assume that the input itself is the TPM object
+  if(any(names(input) %in% 'tpm')){
+    tpm_object <- input$tpm
+  }else{
+    tpm_object <- input
+    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
+  }
+
+  #Want all of them so will just ignore Geneid
+  input_logged <- tpm_object %>%
+    mutate_at(vars(-matches('Geneid')), 
+              log1p)
+
+  gene_mean_scores = input_logged %>% 
+    melt(id.vars = 'Geneid') %>% 
+    group_by(Geneid) %>% 
+    dplyr::summarise(mean = mean(value),
+                     std_dev = sd(value))
+
+  gene_vector = as.character(input_logged$Geneid)
+
+  input_z = lapply(gene_vector, function(gene){
+    mean_gene = filter(gene_mean_scores, 
+                       Geneid == paste(gene)) %>% 
+      select(mean)
+    std_gene = filter(gene_mean_scores, 
+                      Geneid == paste(gene)) %>% 
+      select(std_dev)
+    gene_scores = input_logged %>% 
+      filter(Geneid == paste(gene)) %>% 
+      select(-Geneid)
+    if(std_gene != 0){
+      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
+    }else{
+      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
+    }
+    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
+    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
+  })
+
+  #Use reduce and rbind to turn the list into one df
+  input_z_df = purrr::reduce(input_z, 
+                             rbind)
+  return(list(zscores = input_z_df,
+              tpm = tpm_object,
+              experiment = input$experiment))
+}
+
+#Make and save z-scores
+twentyfour_z <- lapply(twentyfour_avgtpm,
+                       tpm_to_zscore)
+
+###
+```
+
+# Phaseograms of all the genes in both transmission methods - supplementary figure 5.
+
+```{r row_order_phaseograms}
+
+#Getting the row order for each phaseogram for each experiment
+# Want the data ordered by the highest zscore for each gene in each stage.
+phaseogram_row_order <- lapply(twentyfour_z, 
+                               function(list_elements){
+                                 zscore_df <- list_elements$zscores
+                                 zscore_ordered <- melt(zscore_df, 
+                                                        id = 'Geneid',
+                                                        variable.name = 'stage',
+                                                        value.name = 'zscore') %>% 
+                                   group_by(Geneid) %>% 
+                                   top_n(zscore, n = 1) %>% 
+                                   arrange(stage, 
+                                           desc(zscore)) %>%
+                                   distinct(Geneid, 
+                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
+                                   filter(zscore > 0) #For some reason some remain with zero zscore
+                                 # zscore_ordered <- zscore_ordered[order(
+                                 #   match(
+                                 #     zscore_ordered$stage, 
+                                 #     order_vector
+                                 #   )
+                                 # ),]
+                                 return(c(list_elements, 
+                                          list(zscore_order = zscore_ordered)))
+                               })
+
+phaseogram_time_order <- lapply(names(phaseogram_row_order), 
+                                function(list_element){
+                                  transmission_list <- getElement(phaseogram_row_order,
+                                                                  list_element)
+                                  order_by_time <- transmission_list$zscore_order %>% 
+                                    mutate(stage = times_stages_table$time[match(stage,
+                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
+                                    arrange(stage)
+                                  return(list(zscore_order = order_by_time))
+                                }
+)
+names(phaseogram_time_order) <- names(phaseogram_row_order)
+
+```
+
+```{r function_draw_heatmap_zscore}
+
+draw_heatmap_zscore <- function(zscore_list,
+                                title = '',
+                                gene_ordering_vector = NULL,
+                                filter_genes_vector = NULL,
+                                extra_title = NULL,
+                                annotation_clade = NULL,
+                                annotation_col,
+                                threshold_tpm = NULL,
+                                use_times = FALSE,
+                                use_order_vector = TRUE,
+                                tpm_mat = TRUE,
+                                tpm_breaks = NULL,
+                                ...){
+  if(is.null(tpm_breaks) & tpm_mat){
+    #Need the generate the breaks in the data for the heatmap legend if not provided.
+    max_tpm <- lapply(zscore_list, 
+                      function(list_element){
+                        #list_element <- zscore_list[[1]]
+                        heatmap_data <- as.data.frame(list_element$zscores)
+                        title <- list_element$experiment
+
+                        if(!'Geneid' %in% colnames(heatmap_data)){
+                          stop('No Geneid column in the heatmap data.frame')
+                        }
+
+                        if(!is.null(filter_genes_vector)){
+                          heatmap_data <- as.data.frame(
+                            filter(heatmap_data, 
+                                   Geneid %in% filter_genes_vector)
+                          )
+                        }
+
+                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+
+                        if(!is.null(gene_ordering_vector)){
+                          #Remove genes not in the main data frame but in the ordering vector, 
+                          # and then order based on the vector
+                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                          heatmap_data <- heatmap_data[na.omit(
+                            match(gene_ordering_vector, 
+                                  rownames(heatmap_data))
+                          ),,
+                          drop=FALSE]
+                        }
+
+                        geneids <- rownames(heatmap_data)
+
+                        #Get max/min TPM scores for the side annotation heatmap                       
+                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                          list_element$tpm$Geneid),] %>% 
+                          select(-Geneid) %>% 
+                          as.matrix %>% 
+                          log1p
+                        rownames(tpm_mat) <- geneids
+
+                        tpm_min <- rowMins(tpm_mat)
+                        names(tpm_min) <- geneids
+                        tpm_max <- rowMaxs(tpm_mat)
+                        names(tpm_max) <- geneids
+                        return(list(tpm_max = tpm_max,
+                                    tpm_min = tpm_min))
+                      })
+
+    tpm_max_all <- mround(
+      max(
+        unlist(
+          lapply(max_tpm, 
+                 getElement,
+                 'tpm_max')
+        )
+      ), 
+      base = 1)
+
+    tpm_breaks <- round(
+      c(0,
+        tpm_max_all/4,
+        tpm_max_all/2,
+        tpm_max_all*(3/4),
+        tpm_max_all),
+      digits = 2
+    )
+  }
+
+  #Generate the colour function from the TPM legend breaks for the TPM annotation
+  if(tpm_mat){
+    tpm_col_fun <- colorRamp2(
+      breaks = tpm_breaks,
+      colors = viridis(length(tpm_breaks))
+    )
+  }
+
+  if(is.numeric(threshold_tpm)){
+    #Apply a threshold to the TPM in order to remove low expressed genes.
+    # Remember to add log1p here! 
+    # Threshold uses the max value, so it ony needs to be above thresh 
+    #   in one stage avg in one transmission method.
+    #Will alter the gene_filter vector in order to only select the genes above threshold
+
+    gene_maxs <- bind_rows(
+      lapply(max_tpm, 
+             getElement,
+             'tpm_max')
+    )
+
+    remove_genes <- names(gene_maxs)[apply(
+      gene_maxs, 
+      2, 
+      function(maxs){
+        all(maxs < log2(1+threshold_tpm))
+      }
+    )]
+
+    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
+
+  }
+
+  if(is.character(annotation_clade)){
+    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
+    clade_info <- getElement(cir_info,
+                             annotation_clade)[match(order_cir, 
+                                                     cir_info$Geneid)]
+    names(clade_info) <- str_split_fixed(order_cir, 
+                                         pattern = '_', 
+                                         n = 2)[,2]
+    annotation_col <- annotation_col
+  }
+
+  heatmaps <- lapply(zscore_list, 
+                     function(list_element){
+                       #list_element <- zscore_list[[1]]
+                       heatmap_data <- as.data.frame(list_element$zscores)
+                       title <- list_element$experiment
+
+                       if(!'Geneid' %in% colnames(heatmap_data)){
+                         stop('No Geneid column in the heatmap data.frame')
+                       }
+
+                       if(!is.null(filter_genes_vector)){
+                         heatmap_data <- as.data.frame(
+                           filter(heatmap_data, 
+                                  Geneid %in% filter_genes_vector)
+                         )
+                       }
+
+                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+
+                       if(!is.null(gene_ordering_vector)){
+                         #Remove genes not in the main data frame but in the ordering vector,
+                         #  and then order based on the vector
+                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                         heatmap_data <- heatmap_data[na.omit(
+                           match(gene_ordering_vector, 
+                                 rownames(heatmap_data))
+                         ),,
+                         drop=FALSE]
+                       }
+
+                       if(tpm_mat){
+                         #Get max/min TPM scores for the side annotation heatmap                       
+                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                           list_element$tpm$Geneid),] %>%
+                           select(-Geneid) %>% 
+                           as.matrix %>% 
+                           log1p
+
+                         tpm_min <- rowMins(tpm_mat)
+                         tpm_max <- rowMaxs(tpm_mat)
+
+                         names(tpm_min) = rownames(heatmap_data)
+                         names(tpm_max) = rownames(heatmap_data)
+
+                         tpm_ha <- rowAnnotation(
+                           min.max = anno_simple(cbind(tpm_min, 
+                                                       tpm_max), 
+                                                 col = tpm_col_fun,
+                                                 gp = gpar(col = 'black',
+                                                           lwd = 0.1),
+                                                 border = TRUE,
+                                                 simple_anno_size = unit(0.2, "cm")),
+                           show_annotation_name = FALSE)
+                       }
+
+                       #Draw the left annotation is needed, only for the first element.
+
+                       if(
+                         identical(list_element,
+                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
+                       ){
+                         clade_ha <- rowAnnotation(
+                           'subfamily' = anno_simple(
+                             clade_info,
+                             col = annotation_col
+                           )
+                         )
+                       }
+
+                       #Change rownames to remove PCHAS_
+                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
+                                                                pattern = '_', 
+                                                                n = 2)[,2]
+
+                       if(use_times){
+                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
+                                                                                 times_stages_table$r_stage_diff_count)]
+                       }
+
+                       if(use_order_vector){
+                         #Order the columns by ordering_vector
+                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
+                                                                   order_vector)),
                                                       drop = FALSE]
-
-         colnames(combined_data_pir_ht) <- paste0(list_name, 
-                                                  '_', 
-                                                  colnames(combined_data_pir_ht))
-
-         log2(1+combined_data_pir_ht)[pir_highest_tpm_order,]
-       }) %>% 
-  bind_cols
-
-supp_info_fig_3b$SorL <- cir_info$SorL[match(names(apply(supp_info_fig_3b, 1, mean)), 
-                                                   cir_info$Geneid)]
-
-head(supp_info_fig_3b)
+                       }else{
+                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
+                                                      drop = FALSE] 
+                       }
+
+                       h = Heatmap(heatmap_data,
+                                   name = paste0('z-score',
+                                                 extra_title),
+                                   col = c(cbPalette[6],
+                                           'white', 
+                                           cbPalette[7]),
+                                   row_names_gp = gpar(fontsize = ifelse(
+                                     nrow(heatmap_data) > 13, 
+                                     12*(13/nrow(heatmap_data)), 
+                                     12)
+                                   ),
+                                   column_names_gp = gpar(fontsize = 10),
+                                   row_dend_width = unit(3, "cm"),
+                                   right_annotation = if(tpm_mat){
+                                     tpm_ha
+                                   }else{
+                                     NULL
+                                   },
+                                   left_annotation = 
+                                     if(identical(list_element,
+                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
+                                       clade_ha
+                                     }else{
+                                       NULL
+                                     }, 
+                                   #Only want the clade annotation when it is first heatmap
+                                   width = unit(4.5, 'cm'),
+                                   cluster_columns = FALSE,
+                                   height = unit(7,'cm'),
+                                   row_title = NULL,
+                                   cluster_rows = FALSE,
+                                   column_title = paste(title),
+                                   heatmap_legend_param = list(
+                                     at = c(3,2,1,0,-1,-2,-3),
+                                     labels_gp = gpar(fontsize = 10),
+                                     title_gp = gpar(fontsize = 10)
+                                   ),
+                                   ...
+                       )
+                       return(h)
+                     })
+
+  ht_list <- Reduce('+', heatmaps) #To add the list elements together
+
+  if(tpm_mat){
+    #Draw the legend
+    tpm_lgd <- Legend(
+      title = "Min/Max log(TPM + 1)", 
+      col = tpm_col_fun, 
+      at = tpm_breaks, 
+      labels = tpm_breaks,
+      title_gp = gpar(fontsize = 7)
+    )
+  }
+
+  if(is.character(annotation_clade)){
+    clade_lgd <- Legend(
+      title = 'sub-family',
+      at = names(annotation_col),
+      legend_gp = gpar(fill = annotation_col)
+    )
+  }
+
+  draw(ht_list, 
+       row_title_side = 'left',
+       column_title_side = 'bottom',
+       heatmap_legend_side = 'left',
+       #Only include the annotation legends for the annotations we actually have.
+       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
+         list(tpm_lgd,
+              clade_lgd)
+       }else if(tpm_mat & !is.character(annotation_clade) ){
+         list(tpm_lgd)
+       } else if(!tpm_mat & is.character(annotation_clade) ){
+         list(clade_lgd)
+       })
+}
+```
+
+## Phaseograms of the entire genome.
+
+```{r phaseograms_whole_genome}
+
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE)
+
+```
+
+## Phaseograms of the _cir_ genes with clade annotations. - Supplementary figure 6
+
+```{r pir_gene_phaseogram_breaks}
+  max_tpm <- lapply(twentyfour_z, 
+                    function(list_element){
+                      #list_element <- zscore_list[[1]]
+                      heatmap_data <- as.data.frame(list_element$zscores)
+                      title <- list_element$experiment
+
+                      if(!'Geneid' %in% colnames(heatmap_data)){
+                        stop('No Geneid column in the heatmap data.frame')
+                      }
+
+                            heatmap_data <- as.data.frame(
+                              filter(heatmap_data, 
+                                     Geneid %in% cir_id)
+                            )
+
+                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
+
+                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
+                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
+                          # 
+                          # if(!is.null(gene_ordering_vector)){
+                          #   #Remove genes not in the main data frame but in the ordering vector, 
+                          #   # and then order based on the vector
+                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
+                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
+                          #   heatmap_data <- heatmap_data[na.omit(
+                          #     match(gene_ordering_vector, 
+                          #           rownames(heatmap_data))
+                          #   ),,
+                          #   drop=FALSE]
+                          # }
+
+                          geneids <- rownames(heatmap_data)
+
+                          #Get max/min TPM scores for the side annotation heatmap                       
+                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
+                                                            list_element$tpm$Geneid),] %>% 
+                            select(-Geneid) %>% 
+                            as.matrix %>% 
+                            log1p
+                          rownames(tpm_mat) <- geneids
+
+                          tpm_min <- rowMins(tpm_mat)
+                          names(tpm_min) <- geneids
+                          tpm_max <- rowMaxs(tpm_mat)
+                          names(tpm_max) <- geneids
+                          return(list(tpm_max = tpm_max,
+                                      tpm_min = tpm_min))
+                        })
+
+  tpm_max_all <- mround(
+    max(
+      unlist(
+        lapply(max_tpm, 
+               getElement,
+               'tpm_max')
+      )
+    ), 
+    base = 1)
+
+  tpm_breaks <- round(
+    c(0,
+      tpm_max_all/4,
+      tpm_max_all/2,
+      tpm_max_all*(3/4),
+      tpm_max_all),
+    digits = 2
+  )
+```
+
+```{r phaseograms_cir}
+
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE)
+
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_breaks = tpm_breaks,
+                    tpm_mat = FALSE)
 

### 2021-03-17 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_berghei_pir_DESEQ.Rmd
Adding files for stats

---
title: "tl20-05_berghei_pir_DESEQ"
author: "Timothy Little"
date: "25/01/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

Using the P. berghei systematic analysis of pir gene expression data we want to perform statistical tests for enrichment of individual pirs in certain stages of the malaria life cycle, using the package DESEQ2.

Notes:
* Would be worth looking at http://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#levels-without-samples to see if we can incorporate the experiment values into this too.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)

# Set the working directory for the document where the data is located.
knitr::opts_knit$set(root.dir = 
                       ".")

# Change options so that numbers are displayed with all the zeros (e,g, as 1000000 not 1e6)
options(scipen = 999) 
```

```{r load_packages}
library('dplyr')
library('readr')
library('readxl')
library('data.table')
library('tibble')
library('ggplot2')
library('stringr')
library('RColorBrewer')
library('DESeq2')
```

Download the data, and remove heterogeneous life cycle stages like 'Gam' and 'BldMl'. Will keep Asex.Mixed since we want to be able to compare that to other stages, although for each Asex stage we should remove the individual stages (like Rings) when looking at mixed, and the Mixed stages when looking at the inidvidual stages.

Note that I'm using the count data with names updated by how I combined similar stages e.g. Liver.post48h.

```{r read_count_data}
berg_counts <- read_xlsx('Manuscript_Frontiers_Counts_UpdatedNames.xlsx')

#Filter low expressed genes (less than 1 count in all samples), convert to a matrix, add 1 for the log transformation, then transform to log2
berg_counts <- column_to_rownames(berg_counts, var = 'Geneid') %>% 
  select(!matches('BldMl|Gam_'))
berg_counts.mat <- berg_counts[!rowSums(berg_counts) <= 10,] %>% 
  as.matrix

berg_counts.mat[1:5,1:5]
```

```{r generate_genesets}
pir_info <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
                      sheet = 1)

pir_geneset <- list(pirs = pir_info$Geneid)

```

```{r colData}
( coldata <- data.frame(
  row.names = colnames(berg_counts.mat),
  experiment = str_extract(colnames(berg_counts.mat), pattern = '^E[[:digit:]]{2}'),
  stage = str_extract(colnames(berg_counts.mat), pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]{1,2})')
) )
```

```{r dds}
dds <- DESeqDataSetFromMatrix(countData = berg_counts.mat,
                              colData = coldata,
                              design = ~ stage)
dds
```

```{r deseq-ing}
dds.deseq <- DESeq(dds)
resultsNames(dds.deseq)
```

Individual _pir_ genes which we use as examples of being highly upregulated in individual stages.

```{r gamMale_v_asexMix}
coef <- "stage_Gam.Male_vs_Asex.Mixed"
dds.deseq.gamMale_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  type = "apeglm")
gamMale_v_asexMixed_df <- as.data.frame(
  dds.deseq.gamMale_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0317181'),]
) %>% 
  mutate(contrast = coef)
```

```{r}
coef <- "stage_Liv.Meroz_vs_Asex.Mixed"
dds.deseq.livMeroz_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  type = "apeglm")

livMeroz_v_asexMixed_df <- as.data.frame(
  dds.deseq.livMeroz_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0100500','PBANKA_0600031'),]) %>%  
  mutate(contrast = coef)
```

```{r asexRing_v_asexMixed}
coef <- 'stage_Asex.Ring_vs_Asex.Mixed'

dds.deseq.asexRing_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                  coef = coef,
                                                  type = "apeglm")
asexRing_v_asexMixed_df <- as.data.frame(
  dds.deseq.asexRing_v_asexMixed.lfcShr[rownames(dds.deseq.asexRing_v_asexMixed.lfcShr) %in% c('PBANKA_1300200'),]
) %>% 
  mutate(contrast = coef)
```

```{r}
dds$stage <- relevel(dds$stage, ref = 'Gam-Male')
dds.maleGam.deseq <- DESeq(dds)
resultsNames(dds.maleGam.deseq)
```

```{r}
coef <- "stage_Liv.Meroz_vs_Gam.Male"
dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr <- lfcShrink(dds.maleGam.deseq, 
                                                         coef = coef,
                                                         type = "apeglm")

livMeroz_v_gamMale_df <- as.data.frame(
  dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr[rownames(dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr) %in% c('PBANKA_0317181',
                                                                                                           'PBANKA_0100500',
                                                                                                           'PBANKA_0600031'),]) %>%  
  mutate(contrast = coef)
```

```{r}
bind_rows(livMeroz_v_gamMale_df, 
          as.data.frame(gamMale_v_asexMixed_df),
          as.data.frame(livMeroz_v_asexMixed_df)) %>% 
  mutate(FDR = p.adjust(pvalue, 
                        method = 'fdr')) %>% 
  select(-c(lfcSE, 
            padj, 
            pvalue))
```

### 2021-03-17 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Adding files for stats

---
title: "tl20-05_berghei_pir_GeneSetEnrichmentQusage"
author: "Timothy Little"
date: "15/01/2021"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

Using the P. berghei systematic analysis of pir gene expression data we want to perform statistical tests for enrichment of sets of pirs in certain stages of the malaria life cycle, using the package Qusage.

Notes:

* Do I want to remove the other Asexual samples, or the other Gam samples, when looking into these stages? Probably.
* Probably want to remove samples which express almost no pirs. E.g. Sporoz is signif. downregulated for everything which is no surprise since it basically doesn't express pir genes.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)

# Set the working directory for the document where the data is located.
knitr::opts_knit$set(root.dir = 
                       ".")

# Change options so that numbers are displayed with all the zeros (e,g, as 1000000 not 1e6)
options(scipen = 999) 
```

```{r load_packages}
library('dplyr')
library('readr')
library('readxl')
library('data.table')
library('tibble')
library('qusage')
library('ggplot2')
library('stringr')
library('RColorBrewer')
```

Download the data, and remove heterogeneous life cycle stages like 'Gam' and 'BldMl'. Will keep Asex.Mixed since we want to be able to compare that to other stages, although for each Asex stage we should remove the individual stages (like Rings) when looking at mixed, and the Mixed stages when looking at the inidvidual stages.

Initially I used counts here but I don't know if I made that up or not... Going to try with log(TPM+1).

```{r read_count_data}
# berg_counts <- read_xlsx('Manuscript_Frontiers_Counts_UpdatedNames.xlsx')

berg_tpm <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
                      sheet = 'berg_TPM_allExpts_bioReps') %>% 
  dplyr::rename_with(.fn = ~ str_replace(.x, 
                                         pattern = 'Expt', 
                                         replacement = 'E'))

#Filter low expressed genes (greater than 10 tpm in at least one sample), convert to a matrix, add 1 for the log transformation, then transform to log2

berg_tpm <- column_to_rownames(berg_tpm, var = 'Geneid') %>% 
  select(!matches('BldMl|Gam_'))

berg_tpm <- berg_tpm[apply(berg_tpm, 1, function(x) {any(x >= 10)}),]

berg_tpm_log2plus1.mat <- berg_tpm %>% 
  as.matrix %>% 
  `+`(.+1) %>% 
  log(base = 2)

berg_tpm_log2plus1.mat[1:5,1:5]

```

```{r generate_labels}
#Use make.names to remove the '-' from the labels
colnames(berg_tpm_log2plus1.mat) <- make.names(colnames(berg_tpm_log2plus1.mat))

labels <- colnames(berg_tpm_log2plus1.mat)

#We want the life cycle stages to be the different groups for now.

lifeCycle_labels <- str_extract(labels, 
                                pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]{1,2})')
unique(lifeCycle_labels) %>% sort

```

```{r generate_genesets}
pir_info <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
                      sheet = 1)

pir_geneset <- list(pirs = pir_info$Geneid)

chrom_list <- lapply(unique(pir_info$chrom), 
                     function(chrom_nam) filter(pir_info, 
                                                chrom == chrom_nam)$Geneid)

names(chrom_list) <- paste0('chr', 
                            unique(pir_info$chrom))

#Remove groups with only three members or fewer.
chrom_list <- chrom_list[unlist(lapply(chrom_list, function(x) length(x) > 3))]

subfam_list <- lapply(unique(pir_info$clade), 
                     function(clade_nam) filter(pir_info, 
                                                clade == clade_nam)$Geneid)
names(subfam_list) <- unique(pir_info$clade)
subfam_list <- subfam_list[unlist(lapply(subfam_list, function(x) length(x) > 3))]

ShortLong_list <- 
  lapply(unique(pir_info$SorL), 
                     function(SorL_nam) filter(pir_info, 
                                                SorL == SorL_nam)$Geneid)
names(ShortLong_list) <- unique(pir_info$SorL)
ShortLong_list <- ShortLong_list[unlist(lapply(ShortLong_list, function(x) length(x) > 3))]

```

What is the question that I actually want to answer? I have a lot of life stages and I want to know if each life stage contains an enrichment for transcription of certain groups of pirs compared to the others. So does this mean that I want to do a pairwise comparison, for each life cycle stage, versus every other identified stage? i.e. lapply for each stage with the 'labels' set as that stage and the rest all set to 'other'?

```{r running_qusage_malevfemale}
maleVfemale <- lapply(
  list(pir_geneset), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Gam.Female',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)

```

```{r lateliverstages}

post48livvearlyliv <- lapply(
  list(pir_geneset,
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Liv.post48h-Liv.24h',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)

```

```{r}
livmero_livpost48.results <- lapply(
  list(pir_geneset,
       ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Liv.Meroz-Liv.post48h',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)

```

```{r}
asexRing_asexMix.results <- lapply(
  list(pir_geneset,
       subfam_list,
       ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Asex.Ring-Asex.Mixed',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
names(asexRing_asexMix.results) <- c('all', 'subfam', 'SorL')
asexRing_asexMix.results$subfam$qstable
asexRing_asexMix.results$SorL$qstable
```

```{r}
maleGam_v_Asex.Mix.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list, 
    chrom_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Asex.Mixed',
                                gene_list,
                                n.points = 2^18)

    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
    #                                 FDR < 0.05)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
names(maleGam_v_Asex.Mix.results) <- c('all', 'subfam', 'SorL', 'chrom')

plotGeneSetDistributions(maleGam_v_Asex.Mix.results$SorL$results, path.index =  'S')
plotGeneSetDistributions(maleGam_v_Asex.Mix.results$SorL$results, path.index = 'L')
plotGeneSetDistributions(maleGam_v_Asex.Mix.results$subfam$results, path.index = 'S4')
mutate(maleGam_v_Asex.Mix.results$subfam$qstable , adjust = p.adjust(p.Value, method = 'fdr'))
plotGeneSetDistributions(maleGam_v_Asex.Mix.results$chrom$results, path.index = 'chr14')
```

```{r}
maleGam_v_Liv.Meroz.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list, 
    chrom_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Liv.Meroz',
                                gene_list,
                                n.points = 2^18)

    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
    #                                 FDR < 0.05)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
```

```{r}
Asex.Mix_v_Liv.Meroz.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list, 
    chrom_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Asex.Mixed-Liv.Meroz',
                                gene_list,
                                n.points = 2^18)

    # qs.results.df.fdr0.05 <- filter(qsTable(qs.results.subfam), 
    #                                 FDR < 0.05)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
names(Asex.Mix_v_Liv.Meroz.results) <- c('all', 'subfam', 'SorL', 'chrom')

plotGeneSetDistributions(Asex.Mix_v_Liv.Meroz.results$SorL$results, 
                         path.index = 'L')
plotGeneSetDistributions(Asex.Mix_v_Liv.Meroz.results$SorL$results, 
                         path.index = 'S')

bind_rows(lapply(Asex.Mix_v_Liv.Meroz.results, 
                 getElement, 
                 'qstable')) %>% 
  mutate(adjust = p.adjust(p.Value)) %>% 
  arrange(adjust)
```

```{r}
berghei_comparison_stats_df <- bind_rows(
  lapply(c(maleGam_v_Asex.Mix.results, 
           maleGam_v_Liv.Meroz.results, 
           Asex.Mix_v_Liv.Meroz.results,
           maleVfemale, 
           post48livvearlyliv, 
           livmero_livpost48.results),
         getElement,
         'qstable')
) %>% 
  select(contrast, everything()) %>% 
  mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
  filter(FDR < 0.05) %>% 
  arrange(contrast)

max_logfold <- ceiling(max(max(berghei_comparison_stats_df$log.fold.change), 
                           abs(min(berghei_comparison_stats_df$log.fold.change))))

(
x <- ggplot(berghei_comparison_stats_df, 
            aes(x = contrast, 
                y = reorder(pathway.name, dplyr::desc(pathway.name)))) +
  xlab('Comparison') + ylab('Geneset') + 
  # scale_x_discrete(labels= c(1,2,3,4,6)) +
  geom_count(aes(
                 colour = log.fold.change), 
             na.rm = TRUE) +
  labs(colour = 'log2 Fold Change') +
  scale_color_gradientn(
    colours = rev(brewer.pal(11, "RdBu")),
    na.value="transparent",
    limits = c(-max_logfold,max_logfold),
    # limits = c(-0.5,1.25)
  ) +
  scale_size_area(max_size = 6) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = 'white'),
        plot.background = element_blank(),
        legend.background = element_rect(fill = 'transparent'),
        axis.text.x = element_text(angle = 90))
  )
```

```{r}
berghei_comparison_stats_df
```

```{r binom_glm}
berghei_tpm <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
                      sheet = 2) %>% 
  melt(value.name = 'tpm', 
       variable.name = 'stage') %>% 
  mutate(experiment = as.factor(
    str_extract(stage, 
                pattern = '^Expt[[:digit:]]{2}')
    ),
stage = as.factor(
  str_extract(stage, 
              pattern = '(?<=Expt[[:digit:]]{2}_).+(?=_[[:digit:]]{1,2})')
  ),
Geneid = as.factor(Geneid),
tpm = as.factor(tpm)
)

```

### 2021-03-17 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_berghei_pir_DESEQ.Rmd
Updated stats Rmd files & removed Cooks distance cut off for PBANKA_0600031 analysis

@@
 Notes:
+* Have to remove Cook's distance cut off when looking at gene PBANKA_0600031 because otherwise DESEQ says that it is an outlier, from the data itself I'm not sure if I agree.
+
+Future:
 * Would be worth looking at http://bioconductor.org/packages/release/bioc/vignettes/DESeq2/inst/doc/DESeq2.html#levels-without-samples to see if we can incorporate the experiment values into this too.
@@
 options(scipen = 999)
+
+set.seed(8)
 ```
@@
 ```{r read_count_data}
-berg_counts <- read_xlsx('Manuscript_Frontiers_Counts_UpdatedNames.xlsx')
+berg_counts <- read_xlsx('Manuscript_Frontiers_suppinfo_bergheiCounts_UpdatedNames.xlsx')
 
@@
 coef <- "stage_Liv.Meroz_vs_Asex.Mixed"
+res <- results(dds.deseq,
+               name = 'stage_Liv.Meroz_vs_Asex.Mixed',
+               cooksCutoff = FALSE)
 dds.deseq.livMeroz_v_asexMixed.lfcShr <- lfcShrink(dds.deseq, 
                                                   coef = coef,
+                                                  res = res,
                                                   type = "apeglm")
@@
 livMeroz_v_asexMixed_df <- as.data.frame(
-  dds.deseq.livMeroz_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0100500','PBANKA_0600031'),]) %>%  
+  dds.deseq.livMeroz_v_asexMixed.lfcShr[rownames(dds.deseq.gamMale_v_asexMixed.lfcShr) %in% c('PBANKA_0100500',
+                                                                                              'PBANKA_0600031'),]) %>%  
   mutate(contrast = coef)
@@
 coef <- "stage_Liv.Meroz_vs_Gam.Male"
+res <- results(dds.maleGam.deseq,
+               name = 'stage_Liv.Meroz_vs_Gam.Male',
+               cooksCutoff = FALSE)
 dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr <- lfcShrink(dds.maleGam.deseq, 
                                                          coef = coef,
+                                                         res= res,
                                                          type = "apeglm")
 
-livMeroz_v_gamMale_df <- as.data.frame(
+(livMeroz_v_gamMale_df <- as.data.frame(
   dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr[rownames(dds.maleGam.deseq.livMeroz_v_gamMale.lfcShr) %in% c('PBANKA_0317181',
                                                                                                            'PBANKA_0100500',
-                                                                                                           'PBANKA_0600031'),]) %>%  
-  mutate(contrast = coef)
+                                                                                                           'PBANKA_0600031'),]
+  ) %>%  
+  mutate(contrast = coef))
 ```

### 2021-03-17 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Updated stats Rmd files & removed Cooks distance cut off for PBANKA_0600031 analysis

@@
 options(scipen = 999)
+
+set.seed(89)
 ```
@@
 Download the data, and remove heterogeneous life cycle stages like 'Gam' and 'BldMl'.
-Will keep Asex.Mixed since we want to be able to compare that to other stages, although for each Asex stage we should remove the individual stages (like Rings) when looking at mixed, and the Mixed stages when looking at the inidvidual stages.
-
-Initially I used counts here but I don't know if I made that up or not... Going to try with log(TPM+1).
+Will keep Asex.Mixed since we want to be able to compare that to other stages, although for each Asex stage we should remove the individual stages (like Rings) when looking at mixed, and the Mixed stages when looking at the individual stages.
 
 ```{r read_count_data}
-# berg_counts <- read_xlsx('Manuscript_Frontiers_Counts_UpdatedNames.xlsx')
 
@@
 ```
-
-```{r binom_glm}
-berghei_tpm <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
-                      sheet = 2) %>% 
-  melt(value.name = 'tpm', 
-       variable.name = 'stage') %>% 
-  mutate(experiment = as.factor(
-    str_extract(stage, 
-                pattern = '^Expt[[:digit:]]{2}')
-    ),
-stage = as.factor(
-  str_extract(stage, 
-              pattern = '(?<=Expt[[:digit:]]{2}_).+(?=_[[:digit:]]{1,2})')
-  ),
-Geneid = as.factor(Geneid),
-tpm = as.factor(tpm)
-)
-
-```

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Split chabaudi TPM calculation into new file and started removing SBP

@@
 author: "Timothy Little"
-date: "12/03/2021"
+date: "22/03/2021"
 output: html_document
@@
 
-```{r read_tpm_rescaled_data}
-
-chab_count_table <- read_xlsx('Manuscript_Frontiers_chabaudi_24h_counts.xlsx')
-
-twentyfour_ct_comb <- tpm(chab_count_table, 
-                          lengths = chabaudi_transcript_lengths)
-
-#Setting twentyfour_tpm_ct core dataset - remove pseudogene and re-scaled
-
-twentyfour_tpm_nopseudo <- twentyfour_ct_comb %>% 
-  filter(!Geneid %in% 'PCHAS_0420800')
-
-twentyfour_tpm_rescale_nopseudo <- lapply(colnames(twentyfour_tpm_nopseudo), 
-                                          function(column_name){
-                                            column <- getElement(twentyfour_tpm_nopseudo, 
-                                                                 paste(column_name))
-                                            if(is.numeric(column)){
-                                              #is.numeric to avoid trying to scale the Geneid column
-                                              tpm_scale_factor <- 1e6 / sum(column)
-                                              tpm_scaled <- as.numeric(column*tpm_scale_factor)
-                                            }else{
-                                              tpm_scaled <- column
-                                            }
-                                            return(tpm_scaled)
-                                          }) %>% 
-  bind_cols
-
-colnames(twentyfour_tpm_rescale_nopseudo) <- colnames(twentyfour_tpm_nopseudo)
-# apply(twentyfour_tpm_rescale_nopseudo[,-1], 2, sum)
-
-#Setting the re-scaled version as the default from now on
-#Setting twentyfour_tpm_ct core dataset - removed pseudogene and re-scaled
-twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo
-
-twentyfour_rMT <- select(twentyfour_ct_comb, 
-                         matches('Geneid|rMT')) %>% 
-  rename_with(~ str_extract(.x,pattern = '(?<=rMT_).+'),
-              matches('rMT'))
-
-twentyfour_rMT_avg <- melt(twentyfour_rMT, 
-                           variable.name = 'sample',
-                           value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
-  group_by(Geneid, stage) %>% 
-  summarise(mean_tpm = mean(tpm)) %>% 
-  dcast(Geneid ~ stage)
-
-twentyfour_SBP <- select(twentyfour_ct_comb, 
-                         matches('Geneid|SBP')) %>% 
-  rename_with(~ str_extract(.x,pattern = '(?<=SBP_).+'),
-              matches('SBP'))
-
-twentyfour_SBP_avg <- melt(twentyfour_SBP, 
-                           variable.name = 'sample',
-                           value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
-  group_by(Geneid, stage) %>% 
-  summarise(mean_tpm = mean(tpm)) %>% 
-  dcast(Geneid ~ stage)
-
-twentyfour_avgtpm <- list(
-  'rMT' = list(
-    tpm = twentyfour_rMT_avg,
-    experiment = 'rMT'
-  ),
-  'SBP' = list(
-    tpm = twentyfour_SBP_avg,
-    experiment = 'SBP'
-  )
-)
-
-```
-
 ```{r colours_clades_phase}
@@
 
-## Figure 3c and Supp figure 3 - Highest expressed all _pir_ heatmap
+## Figure 3c - Highest expressed all _pir_ heatmap
 

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_calculateTPM.Rmd
Pchabaudi TPM file

--- 2021-03-15 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
@@
 ---
-title: "Manuscript_Github_Pchabaudi_pir"
+title: "Manuscript_Github_Pchabaudi_calculateTPM"
 author: "Timothy Little"
-date: "12/03/2021"
+date: "22/03/2021"
 output: html_document
@@
 
-twentyfour_SBP <- select(twentyfour_ct_comb, 
-                         matches('Geneid|SBP')) %>% 
-  rename_with(~ str_extract(.x,pattern = '(?<=SBP_).+'),
-              matches('SBP'))
-
-twentyfour_SBP_avg <- melt(twentyfour_SBP, 
-                           variable.name = 'sample',
-                           value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
-  group_by(Geneid, stage) %>% 
-  summarise(mean_tpm = mean(tpm)) %>% 
-  dcast(Geneid ~ stage)
+# twentyfour_SBP <- select(twentyfour_ct_comb, 
+#                          matches('Geneid|SBP')) %>% 
+#   rename_with(~ str_extract(.x,pattern = '(?<=SBP_).+'),
+#               matches('SBP'))
+# 
+# twentyfour_SBP_avg <- melt(twentyfour_SBP, 
+#                            variable.name = 'sample',
+#                            value.name = 'tpm') %>% 
+#   mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
+#   group_by(Geneid, stage) %>% 
+#   summarise(mean_tpm = mean(tpm)) %>% 
+#   dcast(Geneid ~ stage)
 
@@
     experiment = 'rMT'
-  ),
-  'SBP' = list(
-    tpm = twentyfour_SBP_avg,
-    experiment = 'SBP'
+  # ),
+  # 'SBP' = list(
+  #   tpm = twentyfour_SBP_avg,
+  #   experiment = 'SBP'
   )
@@
 ```
-
-```{r colours_clades_phase}
-SorL_col <- setNames(
-  c(    '#F0E442',
-        '#56B4E9', #L
-        '#E69F00', #S
-        '#999999'),
-  sort(
-    unique(cir_info$SorL)
-  )
-)
-
-SorL_col_num <- SorL_col
-names(SorL_col_num) <- sapply(names(SorL_col), 
-                              function(clade_name) {
-                                paste0(
-                                  clade_name, 
-                                  ' (n = ',
-                                  sum(cir_info$SorL == clade_name),
-                                  ')'
-                                )
-                              })
-
-clade_col <- setNames(
-  c('#F0E442',
-    '#56B4E9', #L1
-    '#009E73',
-    '#0072B2',
-    '#E69F00', #S1
-    '#D55E00',
-    '#CC79A7',
-    '#999999'),
-  sort(
-    unique(cir_info$subfam)
-  )
-)
-
-clade_col_num <- clade_col
-names(clade_col_num) <- sapply(names(clade_col), 
-                               function(clade_name) {
-                                 paste0(
-                                   clade_name, 
-                                   ' (n = ',
-                                   sum(cir_info$subfam == clade_name),
-                                   ')'
-                                 )
-                               })
-
-```
-
-## Figure 3c and Supp figure 3 - Highest expressed all _pir_ heatmap
-
-```{r highest_expressed_stage_heatmap_fig3b}
-
-num_SorL <- cir_info %>% 
-  group_by(SorL) %>% 
-  summarise(num = n_distinct(Gene.ID))
-
-labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
-                        num_SorL$SorL)
-
-# Get order of pir genes by highest expressed genes in each sample.
-
-combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm), 
-       function(list_name){
-
-         combined_data_pir <- getElement(twentyfour_avgtpm, 
-                                     list_name)$tpm %>% 
-           filter(Geneid %in% cir_id) %>% 
-           column_to_rownames(var = 'Geneid')
-
-         #Ordering that a more understandable order
-
-         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
-                                                            order_vector)),
-                                                drop = FALSE]
-
-         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
-                                         means = apply(combined_data_pir, 1, mean),
-                                         subfam = cir_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
-                                                                            cir_info$Geneid)]) %>% 
-           filter(means > 1) %>% 
-           arrange(subfam, 
-                   desc(means)) %>% 
-           .$Geneid
-
-         #Order the heatmap data by the order defined above (SorL, then mean expression).
-         combined_data_pir_ht <- combined_data_pir[
-           order(
-             match(rownames(combined_data_pir),
-                   combined_data_ordered),
-             na.last = NA #Remove NAs
-           ),
-         ]
-
-         #Change the colnames to the times
-         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h')
-
-         #Order by time point
-         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                      drop = FALSE] 
-         #Add a column for the SorL
-         combined_data_pir_ht$SorL <- cir_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
-                                                                cir_info$Geneid)]
-
-         return(combined_data_pir_ht)
-       })
-names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)
-
-maxcol <- ceiling(
-  max(
-    unlist(
-      lapply(combined_data_pir_ht_list,
-             function(list_element){
-               max(
-                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
-               )
-             }
-      )
-    )
-  )
-)
-
-lapply(names(combined_data_pir_ht_list),
-       function(list_name_to_order){
-
-         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
-                                            list_name_to_order)
-
-         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
-         #Sub-family annotation
-         clade_info <- combined_data_pir_ht$SorL
-         annotation_col <- SorL_col
-
-         #Subfamily annotation
-
-         clade_ha <- rowAnnotation(
-           'sub-family' = anno_simple(
-             clade_info,
-             col = annotation_col,
-           ),
-           annotation_name_gp = gpar(fontsize = 10)
-         )
-
-         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
-                                                 labels = 'ancestral',
-                                                 labels_gp = gpar(fontsize = 7),
-                                                 link_width = unit(3, "mm")))
-
-         #lapply across the datasets now, for each order
-         ht_list <- lapply(names(combined_data_pir_ht_list),
-                           function(list_name){
-
-                             combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
-                                                                list_name)$tpm %>% 
-                               column_to_rownames(var = 'Geneid')
-
-                             #Get the times instead of the stage names.
-                             colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h')
-
-                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                                          drop = FALSE]
-
-                             Heatmap(
-                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
-                               cluster_rows = FALSE,
-                               cluster_columns = FALSE,
-                               show_row_names = FALSE,
-                               name = 'log(TPM+1)', 
-                               col = colorRamp2(c(0,
-                                                  maxcol/4,
-                                                  maxcol/2,
-                                                  3*maxcol/4,
-                                                  maxcol),
-                                                viridis(5)),
-                               na_col = 'white', 
-                               border = TRUE,
-                               border_gp = gpar(lwd = 0.1),
-                               column_names_gp = gpar(fontsize = 10),
-                               row_dend_width = unit(3, "cm"),
-                               width = unit(4.5, 'cm'),
-                               height = unit(7,'cm'),
-                               row_title = NULL,
-                               left_annotation = if(
-                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
-                               ){
-                                 clade_ha
-                               }else{
-                                 NULL
-                               },
-                               right_annotation = if(
-                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
-                               ){
-                                 anc_ha
-                               }else{
-                                 NULL
-                               },
-                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
-                               #                                        cir_info$Geneid)],
-                               column_title = paste(list_name),
-                               heatmap_legend_param = list(
-                                 labels_gp = gpar(fontsize = 10),
-                                 title_gp = gpar(fontsize = 10,
-                                                 fontface = 'bold'),
-                                 at = c(0,
-                                        maxcol/4,
-                                        maxcol/2,
-                                        3*maxcol/4,
-                                        maxcol)
-                               )
-                             )
-                           }
-         )
-         ht_list <- Reduce(`+`, ht_list)
-         clade_lgd <- Legend(
-           title = 'sub-family',
-           at = labels_SorL,
-           legend_gp = gpar(fill = SorL_col)
-         )
-
-         return(
-           draw(ht_list, 
-                row_title_side = 'left',
-                column_title_side = 'bottom',
-                heatmap_legend_side = 'right',
-                annotation_legend_side = 'right',
-                annotation_legend_list = list(clade_lgd),
-                ht_gap = unit(.35,'cm'),
-                merge_legend = TRUE
-           ))
-       })
-
-```
-
-## 2D plot of PC1 v PC2 - Supp figure 4A
-
-```{r pca_calculation, results = TRUE}
-twentyfour_tpm <- twentyfour_ct_comb
-rownames_geneid <- twentyfour_tpm$Geneid
-twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
-rownames(twentyfour_tpm_forpca) <- rownames_geneid
-twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
-#Need to look into what exactly scale and center are doing
-```
-
-```{r 2d_pca, results = TRUE}
-twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
-twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
-
-twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
-                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
-                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
-                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
-  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
-  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
-
-colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(twentyfour_tpm_pca_coords$time)))
-
-pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
-pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
-pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
-pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
-
-(
-  plot <- ggplot(twentyfour_tpm_pca_coords,
-                 aes(
-                   x = PC1,
-                   y = PC2,
-                   col = time,
-                   shape = transmission
-                 )) +
-    geom_point(size = 3,
-               alpha = 0.8) +
-    scale_color_manual(values = colornames) +
-    scale_shape_manual(
-      values = setNames(
-        c(17,19),
-        unique(twentyfour_tpm_pca_coords$transmission)
-      )
-    ) +
-    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-    guides(shape=guide_legend(ncol=2),
-           col = guide_legend(ncol=2)) +
-    theme_classic() +
-    theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7))
-)
-
-```
-
-# Calculate and cluster with z-scores {.tabset}
-
-z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:
-
-1. Convert the count/RPKM values of each gene into log values.
-2. Calculate the mean and standard deviation of X gene log values in each sample
-3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
-4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.
-
-* Need to make sure that the z-score is calculated within each transmission method!
-
-```{r zscores}
-tpm_to_zscore = function(input){
-
-  #If the input does not have a TPM element then assume that the input itself is the TPM object
-  if(any(names(input) %in% 'tpm')){
-    tpm_object <- input$tpm
-  }else{
-    tpm_object <- input
-    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
-  }
-
-  #Want all of them so will just ignore Geneid
-  input_logged <- tpm_object %>%
-    mutate_at(vars(-matches('Geneid')), 
-              log1p)
-
-  gene_mean_scores = input_logged %>% 
-    melt(id.vars = 'Geneid') %>% 
-    group_by(Geneid) %>% 
-    dplyr::summarise(mean = mean(value),
-                     std_dev = sd(value))
-
-  gene_vector = as.character(input_logged$Geneid)
-
-  input_z = lapply(gene_vector, function(gene){
-    mean_gene = filter(gene_mean_scores, 
-                       Geneid == paste(gene)) %>% 
-      select(mean)
-    std_gene = filter(gene_mean_scores, 
-                      Geneid == paste(gene)) %>% 
-      select(std_dev)
-    gene_scores = input_logged %>% 
-      filter(Geneid == paste(gene)) %>% 
-      select(-Geneid)
-    if(std_gene != 0){
-      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
-    }else{
-      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
-    }
-    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
-    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
-  })
-
-  #Use reduce and rbind to turn the list into one df
-  input_z_df = purrr::reduce(input_z, 
-                             rbind)
-  return(list(zscores = input_z_df,
-              tpm = tpm_object,
-              experiment = input$experiment))
-}
-
-#Make and save z-scores
-twentyfour_z <- lapply(twentyfour_avgtpm,
-                       tpm_to_zscore)
-
-###
-```
-
-# Phaseograms of all the genes in both transmission methods - supplementary figure 5.
-
-```{r row_order_phaseograms}
-
-#Getting the row order for each phaseogram for each experiment
-# Want the data ordered by the highest zscore for each gene in each stage.
-phaseogram_row_order <- lapply(twentyfour_z, 
-                               function(list_elements){
-                                 zscore_df <- list_elements$zscores
-                                 zscore_ordered <- melt(zscore_df, 
-                                                        id = 'Geneid',
-                                                        variable.name = 'stage',
-                                                        value.name = 'zscore') %>% 
-                                   group_by(Geneid) %>% 
-                                   top_n(zscore, n = 1) %>% 
-                                   arrange(stage, 
-                                           desc(zscore)) %>%
-                                   distinct(Geneid, 
-                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-                                   filter(zscore > 0) #For some reason some remain with zero zscore
-                                 # zscore_ordered <- zscore_ordered[order(
-                                 #   match(
-                                 #     zscore_ordered$stage, 
-                                 #     order_vector
-                                 #   )
-                                 # ),]
-                                 return(c(list_elements, 
-                                          list(zscore_order = zscore_ordered)))
-                               })
-
-phaseogram_time_order <- lapply(names(phaseogram_row_order), 
-                                function(list_element){
-                                  transmission_list <- getElement(phaseogram_row_order,
-                                                                  list_element)
-                                  order_by_time <- transmission_list$zscore_order %>% 
-                                    mutate(stage = times_stages_table$time[match(stage,
-                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
-                                    arrange(stage)
-                                  return(list(zscore_order = order_by_time))
-                                }
-)
-names(phaseogram_time_order) <- names(phaseogram_row_order)
-
-```
-
-```{r function_draw_heatmap_zscore}
-
-draw_heatmap_zscore <- function(zscore_list,
-                                title = '',
-                                gene_ordering_vector = NULL,
-                                filter_genes_vector = NULL,
-                                extra_title = NULL,
-                                annotation_clade = NULL,
-                                annotation_col,
-                                threshold_tpm = NULL,
-                                use_times = FALSE,
-                                use_order_vector = TRUE,
-                                tpm_mat = TRUE,
-                                tpm_breaks = NULL,
-                                ...){
-  if(is.null(tpm_breaks) & tpm_mat){
-    #Need the generate the breaks in the data for the heatmap legend if not provided.
-    max_tpm <- lapply(zscore_list, 
-                      function(list_element){
-                        #list_element <- zscore_list[[1]]
-                        heatmap_data <- as.data.frame(list_element$zscores)
-                        title <- list_element$experiment
-
-                        if(!'Geneid' %in% colnames(heatmap_data)){
-                          stop('No Geneid column in the heatmap data.frame')
-                        }
-
-                        if(!is.null(filter_genes_vector)){
-                          heatmap_data <- as.data.frame(
-                            filter(heatmap_data, 
-                                   Geneid %in% filter_genes_vector)
-                          )
-                        }
-
-                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                        if(!is.null(gene_ordering_vector)){
-                          #Remove genes not in the main data frame but in the ordering vector, 
-                          # and then order based on the vector
-                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          heatmap_data <- heatmap_data[na.omit(
-                            match(gene_ordering_vector, 
-                                  rownames(heatmap_data))
-                          ),,
-                          drop=FALSE]
-                        }
-
-                        geneids <- rownames(heatmap_data)
-
-                        #Get max/min TPM scores for the side annotation heatmap                       
-                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                          list_element$tpm$Geneid),] %>% 
-                          select(-Geneid) %>% 
-                          as.matrix %>% 
-                          log1p
-                        rownames(tpm_mat) <- geneids
-
-                        tpm_min <- rowMins(tpm_mat)
-                        names(tpm_min) <- geneids
-                        tpm_max <- rowMaxs(tpm_mat)
-                        names(tpm_max) <- geneids
-                        return(list(tpm_max = tpm_max,
-                                    tpm_min = tpm_min))
-                      })
-
-    tpm_max_all <- mround(
-      max(
-        unlist(
-          lapply(max_tpm, 
-                 getElement,
-                 'tpm_max')
-        )
-      ), 
-      base = 1)
-
-    tpm_breaks <- round(
-      c(0,
-        tpm_max_all/4,
-        tpm_max_all/2,
-        tpm_max_all*(3/4),
-        tpm_max_all),
-      digits = 2
-    )
-  }
-
-  #Generate the colour function from the TPM legend breaks for the TPM annotation
-  if(tpm_mat){
-    tpm_col_fun <- colorRamp2(
-      breaks = tpm_breaks,
-      colors = viridis(length(tpm_breaks))
-    )
-  }
-
-  if(is.numeric(threshold_tpm)){
-    #Apply a threshold to the TPM in order to remove low expressed genes.
-    # Remember to add log1p here! 
-    # Threshold uses the max value, so it ony needs to be above thresh 
-    #   in one stage avg in one transmission method.
-    #Will alter the gene_filter vector in order to only select the genes above threshold
-
-    gene_maxs <- bind_rows(
-      lapply(max_tpm, 
-             getElement,
-             'tpm_max')
-    )
-
-    remove_genes <- names(gene_maxs)[apply(
-      gene_maxs, 
-      2, 
-      function(maxs){
-        all(maxs < log2(1+threshold_tpm))
-      }
-    )]
-
-    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
-
-  }
-
-  if(is.character(annotation_clade)){
-    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
-    clade_info <- getElement(cir_info,
-                             annotation_clade)[match(order_cir, 
-                                                     cir_info$Geneid)]
-    names(clade_info) <- str_split_fixed(order_cir, 
-                                         pattern = '_', 
-                                         n = 2)[,2]
-    annotation_col <- annotation_col
-  }
-
-  heatmaps <- lapply(zscore_list, 
-                     function(list_element){
-                       #list_element <- zscore_list[[1]]
-                       heatmap_data <- as.data.frame(list_element$zscores)
-                       title <- list_element$experiment
-
-                       if(!'Geneid' %in% colnames(heatmap_data)){
-                         stop('No Geneid column in the heatmap data.frame')
-                       }
-
-                       if(!is.null(filter_genes_vector)){
-                         heatmap_data <- as.data.frame(
-                           filter(heatmap_data, 
-                                  Geneid %in% filter_genes_vector)
-                         )
-                       }
-
-                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                       if(!is.null(gene_ordering_vector)){
-                         #Remove genes not in the main data frame but in the ordering vector,
-                         #  and then order based on the vector
-                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                         heatmap_data <- heatmap_data[na.omit(
-                           match(gene_ordering_vector, 
-                                 rownames(heatmap_data))
-                         ),,
-                         drop=FALSE]
-                       }
-
-                       if(tpm_mat){
-                         #Get max/min TPM scores for the side annotation heatmap                       
-                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                           list_element$tpm$Geneid),] %>%
-                           select(-Geneid) %>% 
-                           as.matrix %>% 
-                           log1p
-
-                         tpm_min <- rowMins(tpm_mat)
-                         tpm_max <- rowMaxs(tpm_mat)
-
-                         names(tpm_min) = rownames(heatmap_data)
-                         names(tpm_max) = rownames(heatmap_data)
-
-                         tpm_ha <- rowAnnotation(
-                           min.max = anno_simple(cbind(tpm_min, 
-                                                       tpm_max), 
-                                                 col = tpm_col_fun,
-                                                 gp = gpar(col = 'black',
-                                                           lwd = 0.1),
-                                                 border = TRUE,
-                                                 simple_anno_size = unit(0.2, "cm")),
-                           show_annotation_name = FALSE)
-                       }
-
-                       #Draw the left annotation is needed, only for the first element.
-
-                       if(
-                         identical(list_element,
-                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
-                       ){
-                         clade_ha <- rowAnnotation(
-                           'subfamily' = anno_simple(
-                             clade_info,
-                             col = annotation_col
-                           )
-                         )
-                       }
-
-                       #Change rownames to remove PCHAS_
-                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
-                                                                pattern = '_', 
-                                                                n = 2)[,2]
-
-                       if(use_times){
-                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
-                                                                                 times_stages_table$r_stage_diff_count)]
-                       }
-
-                       if(use_order_vector){
-                         #Order the columns by ordering_vector
-                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
-                                                                   order_vector)),
-                                                      drop = FALSE]
-                       }else{
-                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
-                                                      drop = FALSE] 
-                       }
-
-                       h = Heatmap(heatmap_data,
-                                   name = paste0('z-score',
-                                                 extra_title),
-                                   col = c(cbPalette[6],
-                                           'white', 
-                                           cbPalette[7]),
-                                   row_names_gp = gpar(fontsize = ifelse(
-                                     nrow(heatmap_data) > 13, 
-                                     12*(13/nrow(heatmap_data)), 
-                                     12)
-                                   ),
-                                   column_names_gp = gpar(fontsize = 10),
-                                   row_dend_width = unit(3, "cm"),
-                                   right_annotation = if(tpm_mat){
-                                     tpm_ha
-                                   }else{
-                                     NULL
-                                   },
-                                   left_annotation = 
-                                     if(identical(list_element,
-                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
-                                       clade_ha
-                                     }else{
-                                       NULL
-                                     }, 
-                                   #Only want the clade annotation when it is first heatmap
-                                   width = unit(4.5, 'cm'),
-                                   cluster_columns = FALSE,
-                                   height = unit(7,'cm'),
-                                   row_title = NULL,
-                                   cluster_rows = FALSE,
-                                   column_title = paste(title),
-                                   heatmap_legend_param = list(
-                                     at = c(3,2,1,0,-1,-2,-3),
-                                     labels_gp = gpar(fontsize = 10),
-                                     title_gp = gpar(fontsize = 10)
-                                   ),
-                                   ...
-                       )
-                       return(h)
-                     })
-
-  ht_list <- Reduce('+', heatmaps) #To add the list elements together
-
-  if(tpm_mat){
-    #Draw the legend
-    tpm_lgd <- Legend(
-      title = "Min/Max log(TPM + 1)", 
-      col = tpm_col_fun, 
-      at = tpm_breaks, 
-      labels = tpm_breaks,
-      title_gp = gpar(fontsize = 7)
-    )
-  }
-
-  if(is.character(annotation_clade)){
-    clade_lgd <- Legend(
-      title = 'sub-family',
-      at = names(annotation_col),
-      legend_gp = gpar(fill = annotation_col)
-    )
-  }
-
-  draw(ht_list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'left',
-       #Only include the annotation legends for the annotations we actually have.
-       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
-         list(tpm_lgd,
-              clade_lgd)
-       }else if(tpm_mat & !is.character(annotation_clade) ){
-         list(tpm_lgd)
-       } else if(!tpm_mat & is.character(annotation_clade) ){
-         list(clade_lgd)
-       })
-}
-```
-
-## Phaseograms of the entire genome.
-
-```{r phaseograms_whole_genome}
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
-```
-
-## Phaseograms of the _cir_ genes with clade annotations. - Supplementary figure 6
-
-```{r pir_gene_phaseogram_breaks}
-  max_tpm <- lapply(twentyfour_z, 
-                    function(list_element){
-                      #list_element <- zscore_list[[1]]
-                      heatmap_data <- as.data.frame(list_element$zscores)
-                      title <- list_element$experiment
-
-                      if(!'Geneid' %in% colnames(heatmap_data)){
-                        stop('No Geneid column in the heatmap data.frame')
-                      }
-
-                            heatmap_data <- as.data.frame(
-                              filter(heatmap_data, 
-                                     Geneid %in% cir_id)
-                            )
-
-                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-                          # 
-                          # if(!is.null(gene_ordering_vector)){
-                          #   #Remove genes not in the main data frame but in the ordering vector, 
-                          #   # and then order based on the vector
-                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          #   heatmap_data <- heatmap_data[na.omit(
-                          #     match(gene_ordering_vector, 
-                          #           rownames(heatmap_data))
-                          #   ),,
-                          #   drop=FALSE]
-                          # }
-
-                          geneids <- rownames(heatmap_data)
-
-                          #Get max/min TPM scores for the side annotation heatmap                       
-                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                            list_element$tpm$Geneid),] %>% 
-                            select(-Geneid) %>% 
-                            as.matrix %>% 
-                            log1p
-                          rownames(tpm_mat) <- geneids
-
-                          tpm_min <- rowMins(tpm_mat)
-                          names(tpm_min) <- geneids
-                          tpm_max <- rowMaxs(tpm_mat)
-                          names(tpm_max) <- geneids
-                          return(list(tpm_max = tpm_max,
-                                      tpm_min = tpm_min))
-                        })
-
-  tpm_max_all <- mround(
-    max(
-      unlist(
-        lapply(max_tpm, 
-               getElement,
-               'tpm_max')
-      )
-    ), 
-    base = 1)
-
-  tpm_breaks <- round(
-    c(0,
-      tpm_max_all/4,
-      tpm_max_all/2,
-      tpm_max_all*(3/4),
-      tpm_max_all),
-    digits = 2
-  )
-```
-
-```{r phaseograms_cir}
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_breaks = tpm_breaks,
-                    tpm_mat = FALSE)
-
-```
-
-## Figure 4 and supplementary figure 7/8
-
-_pir_ genes selected by being expressed above 1 TPM in all samples.
-
-Adding in housekeeping genes for comparison:
-HSP70 putative  PCHAS_0721000
-MSP1 PCHAS_0831300
-AMA1 PCHAS_0931000
-
-```{r pirs_side-by-side_figure4_RMT-SBP-completelySeparate}
-
-housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
-                             'MSP1' = 'PCHAS_0831300',
-                             'AMA1' = 'PCHAS_0931000')
-
-cir_housekeeping_id <- c(cir_id, 
-                         'PCHAS_0721000',
-                         'PCHAS_0831300',
-                         'PCHAS_0931000')
-
-cir_plot <- twentyfour_ct_comb %>% 
-  filter(Geneid %in% cir_housekeeping_id) %>%
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(
-    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
-    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
-    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
-  ) %>% 
-  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
-  filter(!(time %in% c('sch.culture', 
-                       'Unknown'))) %>% 
-  mutate(subfam = str_replace(subfam, 
-                              'ancestral',
-                              'a')) %>% 
-  mutate(
-    subfam = interaction(
-      transmission,
-      forcats::fct_shift(
-        relevel(
-          as.factor(subfam),
-          ref = 'ref')
-      ),
-      sep = '_',
-      lex.order = TRUE
-    )
-  )
-#Need to add the transmission name to the subfam so that they are ordered correctly.
-# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
-# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.
-
-cir_plot$Geneid <- ifelse(
-  cir_plot$Geneid %in% housekeeping_genes[1,],
-  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
-  cir_plot$Geneid
-)
-
-limits_cir <- c(0, 
-                mround(max(cir_plot$tpm), 
-                       base = 100))
-
-cir_plot_order <- cir_plot %>% 
-  filter(time == '14h',
-         transmission == 'rMT') %>% 
-  arrange(desc(tpm))
-
-cir_plot <- mutate(cir_plot, 
-                   Geneid = factor(Geneid, 
-                                   levels = unique(cir_plot_order$Geneid)))
-
-transmissionSep_cir_plot <- function(time_vec = '14h',
-                                     log_it = TRUE,
-                                     limits = limits_cir,
-                                     point_size = 1,
-                                     text_size = 10,
-                                     threshold_min = 100,
-                                     threshold_max = Inf){
-
-  cir_plot_filter <- group_by(cir_plot, 
-                       Geneid) %>% 
-    filter(
-      time == time_vec,
-      !Geneid %in% cir_id | any(tpm >= threshold_min),
-      !Geneid %in% cir_id | all(tpm <= threshold_max)
-    )
-  #Filter for pirs expressed above or below the thresholds, have the OR so that the ref genes aren't removed.
-  # e.g. Either 'Not a cir' OR 'Any TPM is above the minimum'.
-
-  # cir_plot_filter %>% group_by(Geneid) %>% summarise(n())
-  if(log_it){
-    cir_plot_filter$tpm <- log2(1+cir_plot_filter$tpm)
-    limits = log2(1+limits)
-    breaks <- round(seq.int(0, 
-                            max(limits), 
-                            length.out = round(max(limits)) + 1, 
-                            digits = 0), 
-                    digits = 0)
-  }else{
-    breaks <- seq.int(0, max(limits), length.out = 11)
-  }
-
-  cir_plot_filter$Geneid <- factor(
-    str_remove(cir_plot_filter$Geneid, 
-               pattern = 'PCHAS_'),
-    levels = str_remove(levels(cir_plot_filter$Geneid), 
-                        pattern = 'PCHAS_')
-  )
-
-  lapply(unique(cir_plot_filter$transmission), 
-         function(transmission_method){
-
-           cir_plot_filter <- filter(cir_plot_filter, 
-                              transmission %in% transmission_method)
-
-           y <- ggplot(
-             data = cir_plot_filter,
-             mapping = aes(
-               x = Geneid,
-               y = tpm
-             )
-           ) +
-             stat_summary(fun.y = median, 
-                          fun.ymin = median, 
-                          fun.ymax = median,
-                          geom = "bar",
-                          fill = 'lightblue',
-                          width = 0.75, 
-                          size = 0.25, 
-                          position = 'dodge') +
-             geom_point(stat = 'identity', 
-                        position = position_dodge(width = 0.75),
-                        size = point_size) +
-             scale_y_continuous(breaks = breaks,
-                                limits = limits) +
-             facet_grid(.~subfam, 
-                        scales = "free_x",
-                        space = "free_x",
-                        labeller = labeller(subfam = function(string) {
-                          str_remove_all(string, 
-                                         pattern = '(rMT|SBP)_')
-                        })) +
-             theme_classic() +
-             ylab('log(TPM + 1)')+
-             ggtitle(paste(unique(cir_plot_filter$time)))
-
-           if(grepl(time_vec, pattern = '20|23')) {
-             y <- y + theme(
-               axis.text.x = element_text(size = text_size,
-                                          angle = 90,
-                                          vjust = 0.5)
-             )
-           } else {
-             y <- y + theme(
-               axis.text.x = element_blank(),
-               axis.ticks.x = element_blank()
-             ) 
-           }
-
-           #This bit is to change the colour, and text colour, of the facet boxes
-
-           g <- ggplot_gtable(ggplot_build(y))
-
-           # g$widths[5] = 2*g$widths[5]
-
-           strip_t <- which(grepl('strip-t', g$layout$name))
-           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
-           fills <- case_when(str_detect(subfam_vec, 
-                                         pattern = 'a') ~ clade_col['ancestral'],
-                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
-                              str_detect(subfam_vec, 
-                                         pattern = 'ref') ~ 'white')
-           k <- 1
-           for (i in strip_t) {
-             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-             k <- k+1
-           }
-
-           #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
-           # manually make them a bit wider.
-
-           short_panels <- which(str_detect(subfam_vec, pattern = 'a|S1'))
-
-           for(i in short_panels){
-             num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
-             g$widths[num] = 1.75*g$widths[num]
-             #1.75 times the width of the right panels.
-             # print(num)
-           }
-
-           grid.newpage()
-           grid.draw(g)
-         })
-
-}
-
-lapply(
-  sort(unique(cir_plot$time)),
-  transmissionSep_cir_plot,
-  threshold_min = 50,
-  threshold_max = Inf,
-  point_size = 1,
-  text_size = 7.5
-)
-```
-
-# Supp figure 9 - Pir expression of AAPLs and ChAPLs.
-
-```{r Supp_fig_AapL-ChaAPl_expression}
-
-  cir_plot <- twentyfour_ct_comb%>% 
-    filter(Geneid %in% cir_id) %>%
-    melt(id.vars = 'Geneid', 
-         variable.name = 'sample', 
-         value.name = 'tpm') %>% 
-    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
-           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
-           subfam = cir_info$subfam[match(Geneid, cir_info$Geneid)],
-           chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
-                                                                    cir_info$Geneid)],
-                               levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
-                                                                                         cir_info$Geneid)])))) %>% 
-    filter(!(time %in% c('sch', 
-                         'Unknown')),
-           !is.na(chapl_aapl) ) %>% 
-    group_by(Geneid) %>% 
-    # filter(
-    #   # all(tpm < 100),
-    #   #      any(tpm > 50)
-    #   subfam == 'S7'
-    #   ) %>% 
-    arrange(desc(tpm))
-
-  cir_plot_order <- cir_plot %>% 
-    filter(time == '14h',
-           transmission == 'SBP') %>% 
-    arrange(desc(tpm))
-
-  cir_plot <- mutate(cir_plot, 
-                     Geneid = factor(Geneid, 
-                                     levels = unique(cir_plot_order$Geneid)))
-  cir_plot_chapl_aapl_supp_info <- cir_plot
-
-  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))
-
-  breaks = seq(0, max_aapl_chapl, by = 1)
-    limits = c(0, max_aapl_chapl)
-    point_size = 0.25
-
-  lapply(
-    sort(unique(cir_plot$time)),
-    function(time_vec){
-
-    cir_plot <- filter(cir_plot, 
-                       time == time_vec)
-
-      cir_plot$tpm <- log2(1+cir_plot$tpm)
-
-    cir_plot$Geneid <- factor(
-      str_remove(cir_plot$Geneid, 
-                 pattern = 'PCHAS_'),
-      levels = str_remove(levels(cir_plot$Geneid), 
-                          pattern = 'PCHAS_')
-    )
-
-    y <- ggplot(
-      data = cir_plot,
-      mapping = aes(
-        x = Geneid,
-        fill = transmission,
-        y = tpm
-      )
-    ) +
-    stat_summary(fun.y = median, 
-                 fun.ymin = median, 
-                 fun.ymax = median,
-                 geom = "bar", 
-                 width = 0.75, 
-                 size = 0.25, 
-                 position = 'dodge') +
-    geom_point(stat = 'identity', 
-               position = position_dodge(width = 0.75),
-               size = point_size) +
-    scale_y_continuous(breaks = breaks,
-                       limits = limits) +
-      scale_fill_manual(values = c('skyblue','red')) +
-      ylab('log(TPM +1)') +
-    facet_grid(.~chapl_aapl, 
-             scales = "free_x",
-             space = "free_x") +
-    theme_classic() +
-    theme(
-      axis.text.x = element_text(angle = 90,
-                                 size = 5)
-    ) +
-    ggtitle(paste(unique(cir_plot$time)))
-
-  return(y)
-})
-```
-
-## Bar chart of S and L clade expression across time.
-
-```{r SorL_barchart}
-tpm_data_SorL <- twentyfour_ct_comb%>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_info$subfam[match(Geneid, 
-                                            cir_info$Geneid)]) %>% 
-  mutate(SorL = str_extract(clade, 
-                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-         # stage =  factor(stage, levels = ordering(stage, order_vector))
-         ) %>% 
-  group_by(transmission,
-           stage, 
-           SorL) %>% 
-  dplyr::summarise(total_tpm = sum(tpm))
-
-y <- ggplot(data = tpm_data_SorL, 
-              aes(
-                x = transmission,
-                y = total_tpm, 
-                fill = SorL,
-                group = transmission
-              )) +
-    geom_bar(position = position_stack(),
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-    scale_fill_manual(
-      values = SorL_col,
-      guide = guide_legend(reverse = TRUE)
-    ) +
-    scale_color_manual('black') + 
-    # labs(fill = stack_label) +
-    # scale_y_continuous(breaks = max_tpm) +
-    theme_classic() +
-    scale_size(range = c(0,3)) +
-    ylab(
-      bquote('total '~italic(.('pir'))~' TPM')
-    ) + 
-    facet_grid(.~stage, 
-               scales = "free_x",
-               space = "free_x") +
-    theme(
-      axis.title.x=element_blank(),
-      # axis.text.x=element_blank(),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-```
-
-## Clades proportion bar chart
-
-```{r clades_prop_barchart}
-tpm_data_clade <- twentyfour_ct_comb%>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-  mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_info$subfam[match(Geneid, 
-                                            cir_info$Geneid)]) %>% 
-  # mutate(stage =  factor(stage, 
-  #                        levels = ordering(stage, 
-  #                                          order_vector))) %>% 
-  group_by(transmission,
-           stage, 
-           clade) %>% 
-  dplyr::summarise(total_tpm = sum(tpm)) %>% 
-  arrange(stage)
-
-y <- ggplot(data = tpm_data_clade, 
-              aes(
-                x = stage,
-                y = total_tpm, 
-                fill = clade,
-                group = transmission
-              )) +
-    geom_bar(position = 'fill',
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-    scale_fill_manual(
-      breaks = sort(unique(tpm_data_clade$clade)),
-      values = clade_col,
-      guide = guide_legend(reverse = TRUE) 
-    ) +
-    scale_color_manual('black') + 
-    # labs(fill = stack_label) +
-    # scale_y_continuous(breaks = max_tpm) +
-    theme_classic() +
-    scale_size(range = c(0,3)) +
-    ylab(
-      bquote('proportion of total '~italic(.('pir'))~' TPM')
-    ) + 
-    facet_grid(.~transmission, 
-               scales = "free_x",
-               space = "free_x") +
-    theme(
-      axis.title.x=element_blank(),
-      axis.text.x=element_text(size = 7, 
-                                 angle = 90),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-```
-
-```{r add_berghei_SorL_prop}
-#Add berghei data
-berghei_SorL_file <- list.files('..', 
-                                recursive = TRUE, 
-                                pattern = 'tl20-05_SorL-subfamily_pir.csv', 
-                                full.names = TRUE)
-berghei_SorL <- read_csv(berghei_SorL_file) %>% 
-  filter(stage %in% c('Asex.SchizA',
-                      'Asex.SchizB',
-                      'Asex.Ring',
-                      'Asex.Troph',
-                      'Asex.Mixed')) 
-berghei_SorL_avg <- berghei_SorL %>% 
-  group_by(stage,
-           SorL) %>% 
-  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
-  mutate(SorL = str_extract(SorL,
-                            pattern = 'S|L|ancestral'),
-         transmission = 'berghei')
-
-```
-
-```{r}
-y <- ggplot(data = rbind(
-  tpm_data_SorL,
-  berghei_SorL_avg
-), 
-aes(
-  x = stage,
-  y = total_tpm, 
-  fill = SorL,
-  # group = transmission
-)) +
-  geom_bar(position = 'fill',
-           stat = 'identity',
-           colour = 'black',
-           size = 0.1) +
-  scale_fill_manual(
-    values = SorL_col,
-    guide = guide_legend(reverse = FALSE)
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  # scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('proportion of total '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~transmission, 
-             scales = "free_x",
-             space = "free_x") +
-  labs(fill = 'sub-family') +
-  theme(
-    axis.title.x=element_blank(),
-    axis.text.x=element_text(angle = 90),
-    axis.text.y = element_text(size = 5),
-    axis.title.y = element_text(size = 8),
-    legend.title = element_text(size = 7),
-    legend.text = element_text(size = 5),
-    legend.key.size = unit(3,'mm'),
-    legend.key.width = unit(2,'mm'),
-    strip.text.x = element_text(size=7, 
-                                face = 'bold', 
-                                angle = 0),
-    strip.background = element_rect(
-      color="black", 
-      size=0.1, 
-      linetype="solid"
-    )
-  ) 
-y
-
-```
-
-```{r SorL_barchart_transmission-facet}
-tpm_data_SorL <- twentyfour_ct_comb%>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_info$subfam[match(Geneid, 
-                                            cir_info$Geneid)]) %>% 
-  mutate(SorL = str_extract(clade, 
-                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-         # stage =  factor(stage, levels = ordering(stage, order_vector))
-         ) %>% 
-  group_by(transmission,
-           stage, 
-           SorL) %>% 
-  dplyr::summarise(total_tpm = sum(tpm)) %>% 
-  mutate(SorL = factor(SorL, 
-                       levels = sort(unique(SorL))))
-
-y <- ggplot(data = tpm_data_SorL, 
-              aes(
-                x = stage,
-                y = total_tpm, 
-                fill = SorL,
-                # group = transmission
-              )) +
-  geom_bar(position = position_stack(),
-           stat = 'identity',
-           colour = 'black',
-           size = 0.1) +
-  scale_fill_manual(
-    values = SorL_col,
-    guide = guide_legend(reverse = FALSE)
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  # scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~transmission, 
-             scales = "free_x",
-             space = "free_x") +
-  labs(fill = 'sub-family') +
-    theme(
-      axis.title.x=element_blank(),
-      # axis.text.x=element_blank(),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-```
-
-```{r proportion_data}
-
-tpm_data_twentyfour <- twentyfour_ct_comb%>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_info$subfam[match(Geneid, 
-                                            cir_info$Geneid)],
-         chrom = str_extract(Geneid, 
-                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
-         ) %>% 
-  mutate(
-    SorL = str_extract(clade, 
-                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-    # stage =  factor(stage, 
-    #                 levels = ordering(stage, order_vector))
-  ) %>% 
-  group_by(transmission,
-           stage, 
-           Geneid,
-           clade,
-           chrom,
-           SorL) %>% 
-  dplyr::summarise(tpm = mean(tpm))
-
-```
-
-# Lower pir expression vs rest of genome
-
-```{r low_pir_expression}
-set.seed(4534)
-
-gene_expression_df <- twentyfour_ct_comb%>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(pir_gene = ifelse(Geneid %in% cir_id,
-                           'pir',
-                           'non-pir')) %>% 
-  group_by(stage, 
-           pir_gene,
-           transmission) %>% 
-  slice_sample(n = length(cir_id)) %>%
-  summarise(mean_tpm = mean(tpm),
-            med_tpm = round(median(tpm),
-                            digits = 6))
-
-gene_expression_df %>% 
-  group_by(transmission, 
-           pir_gene) %>% 
-  slice_min(mean_tpm, n =1)
-
-gene_expression_df %>% 
-  group_by(transmission, 
-           pir_gene) %>% 
-  slice_max(mean_tpm, n =1)
-
-y <- ggplot(gene_expression_df,
-       aes(x = pir_gene,
-           y = log2(1+mean_tpm))) +
-  facet_grid(~transmission) +
-  ggtitle(expression(paste(italic('P. chabaudi'), ' 24h cycle'))) +
-  xlab('Gene group')+
-  ylab('log(Median TPM + 1)')+
-  # geom_violin()+
-  geom_dotplot(binaxis = 'y',
-               stackdir = 'center',
-               # binwidth = 1,
-               # dotsize = 2,
-               method = 'dotdensity') +
-  geom_path(aes(group = stage,
-                col = stage)) +
-  theme_classic()
-y
-
-```

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_calculateTPM.Rmd
Saving TPM files
Reading in the TPM to the figure file

@@
 knitr::opts_knit$set(root.dir = 
-                       "/Users/littlet/Dropbox (The Francis Crick)/Data/Manuscript_pirAnalysis_Github/Pchabaudi24h")
+                       ".")
 #Note that the above doesn't give an error if the dir doesn't exist
@@
 
+write_csv(twentyfour_rMT,
+          file = 'Pchabaudi_TPM_allBioReps.csv')
+write_csv(twentyfour_rMT_avg,
+          file = 'Pchabaudi_TPM_average.csv')
+
 ```

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Reading in the TPM to the figure file
PCA just RMT

@@
 ```{r times_stages_labels}
-#Use the rMT stage differential count in order to name the samples, but keep time to retain order
+#Use the RMT stage differential count in order to name the samples, but keep time to retain order
 times_stages_table <- data.frame(time = c('02h', '05h', '08h', '11h', '14h', '17h','20h','23h'),
@@
                                                          replacement = '\\.')
+```
+
+```{r read_TPM}
+#Need to use the calculateTPM script first.
+
+twentyfour_rMT <- read_csv('Pchabaudi_TPM_allBioReps.csv')
+
+twentyfour_avgtpm <- list(
+  'RMT' = list(
+    tpm = read_csv('Pchabaudi_TPM_average.csv'),
+    experiment = 'RMT'
+  )
+)
+
 ```
@@
 ```{r pca_calculation, results = TRUE}
-twentyfour_tpm <- twentyfour_ct_comb
+twentyfour_tpm <- twentyfour_rMT
 rownames_geneid <- twentyfour_tpm$Geneid
@@
   dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,2]) %>% 
-  dplyr::mutate(transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1])
+  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,1])
 
@@
                    y = PC2,
-                   col = time,
-                   shape = transmission
+                   col = time
                  )) +
@@
     scale_color_manual(values = colornames) +
-    scale_shape_manual(
-      values = setNames(
-        c(17,19),
-        unique(twentyfour_tpm_pca_coords$transmission)
-      )
-    ) +
     xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
@@
 draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                     use_times = TRUE,
@@
 draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
+                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
                     filter_genes_vector = cir_id,
@@
   filter(time == '14h',
-         transmission == 'rMT') %>% 
+         transmission == 'RMT') %>% 
   arrange(desc(tpm))
@@
                           str_remove_all(string, 
-                                         pattern = '(rMT|SBP)_')
+                                         pattern = '(RMT|SBP)_')
                         })) +
@@
            strip_t <- which(grepl('strip-t', g$layout$name))
-           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(rMT|SBP)_')
+           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(RMT|SBP)_')
            fills <- case_when(str_detect(subfam_vec, 
@@
     mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
+  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
   filter(stage != 'sch.culture') %>% 
@@
   mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
+  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
   filter(stage != 'sch.culture') %>% 
@@
                              pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-  mutate(transmission = str_extract(sample, pattern = 'rMT|SBP')) %>% 
+  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
   filter(stage != 'sch.culture') %>% 

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Phaseograms only RMT

@@
                        if(use_times){
-                         colnames(heatmap_data) <- times_stages_table$time[match(colnames(heatmap_data),
-                                                                                 times_stages_table$r_stage_diff_count)]
+                         colnames(heatmap_data) <- str_extract(colnames(heatmap_data), 
+                                                               pattern = '[[:digit:]]{2}h')
                        }
@@
                     show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$SBP$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_breaks = tpm_breaks,
                     tpm_mat = FALSE)

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
figure 4 only RMT

@@
 
-cir_plot <- twentyfour_ct_comb %>% 
+cir_plot <- twentyfour_rMT %>% 
   filter(Geneid %in% cir_housekeeping_id) %>%
@@
     stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-    transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
-    time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
+    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
+    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
     subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
@@
                               'ancestral',
-                              'a')) %>% 
-  mutate(
-    subfam = interaction(
-      transmission,
-      forcats::fct_shift(
-        relevel(
-          as.factor(subfam),
-          ref = 'ref')
-      ),
-      sep = '_',
-      lex.order = TRUE
-    )
-  )
+                              'a'))
+
 #Need to add the transmission name to the subfam so that they are ordered correctly.
@@
 cir_plot_order <- cir_plot %>% 
-  filter(time == '14h',
-         transmission == 'RMT') %>% 
+  filter(time == '14h') %>% 
   arrange(desc(tpm))
@@
   )
-
-  lapply(unique(cir_plot_filter$transmission), 
-         function(transmission_method){
-
-           cir_plot_filter <- filter(cir_plot_filter, 
-                              transmission %in% transmission_method)
 
@@
            grid.draw(g)
-         })
 

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Relevelling 'ref' to end with forcats

@@
                               'ancestral',
-                              'a'))
+                              'a')) %>% 
+  mutate(subfam =
+           forcats::fct_shift(
+             relevel(
+               as.factor(subfam),
+               ref = 'ref')
+           )
+  )
 

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
supp fig 9 only RMT

@@
 
-  cir_plot <- twentyfour_ct_comb%>% 
+cir_plot <- twentyfour_rMT %>% 
   filter(Geneid %in% cir_id) %>%
@@
        value.name = 'tpm') %>% 
-    mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-           transmission = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-           diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,3],
-           time = str_split_fixed(sample, pattern = '_', n = 3)[,2],
-           subfam = cir_info$subfam[match(Geneid, cir_info$Geneid)],
+  mutate(
+    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
+    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
+    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
+    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
     chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
@@
   group_by(Geneid) %>% 
-    # filter(
-    #   # all(tpm < 100),
-    #   #      any(tpm > 50)
-    #   subfam == 'S7'
-    #   ) %>% 
     arrange(desc(tpm))
@@
   cir_plot_order <- cir_plot %>% 
-    filter(time == '14h',
-           transmission == 'SBP') %>% 
+    filter(time == '14h') %>% 
     arrange(desc(tpm))
@@
         x = Geneid,
-        fill = transmission,
         y = tpm
@@
                  size = 0.25, 
-                 position = 'dodge') +
+                 position = 'dodge',
+                 fill = 'lightblue') +
     geom_point(stat = 'identity', 

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Removed figures not in the manuscript

@@
 ```
-
-## Bar chart of S and L clade expression across time.
-
-```{r SorL_barchart}
-tpm_data_SorL <- twentyfour_ct_comb%>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_info$subfam[match(Geneid, 
-                                            cir_info$Geneid)]) %>% 
-  mutate(SorL = str_extract(clade, 
-                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-         # stage =  factor(stage, levels = ordering(stage, order_vector))
-         ) %>% 
-  group_by(transmission,
-           stage, 
-           SorL) %>% 
-  dplyr::summarise(total_tpm = sum(tpm))
-
-y <- ggplot(data = tpm_data_SorL, 
-              aes(
-                x = transmission,
-                y = total_tpm, 
-                fill = SorL,
-                group = transmission
-              )) +
-    geom_bar(position = position_stack(),
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-    scale_fill_manual(
-      values = SorL_col,
-      guide = guide_legend(reverse = TRUE)
-    ) +
-    scale_color_manual('black') + 
-    # labs(fill = stack_label) +
-    # scale_y_continuous(breaks = max_tpm) +
-    theme_classic() +
-    scale_size(range = c(0,3)) +
-    ylab(
-      bquote('total '~italic(.('pir'))~' TPM')
-    ) + 
-    facet_grid(.~stage, 
-               scales = "free_x",
-               space = "free_x") +
-    theme(
-      axis.title.x=element_blank(),
-      # axis.text.x=element_blank(),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-```
-
-## Clades proportion bar chart
-
-```{r clades_prop_barchart}
-tpm_data_clade <- twentyfour_ct_comb%>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-  mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_info$subfam[match(Geneid, 
-                                            cir_info$Geneid)]) %>% 
-  # mutate(stage =  factor(stage, 
-  #                        levels = ordering(stage, 
-  #                                          order_vector))) %>% 
-  group_by(transmission,
-           stage, 
-           clade) %>% 
-  dplyr::summarise(total_tpm = sum(tpm)) %>% 
-  arrange(stage)
-
-y <- ggplot(data = tpm_data_clade, 
-              aes(
-                x = stage,
-                y = total_tpm, 
-                fill = clade,
-                group = transmission
-              )) +
-    geom_bar(position = 'fill',
-             stat = 'identity',
-             colour = 'black',
-             size = 0.1) +
-    scale_fill_manual(
-      breaks = sort(unique(tpm_data_clade$clade)),
-      values = clade_col,
-      guide = guide_legend(reverse = TRUE) 
-    ) +
-    scale_color_manual('black') + 
-    # labs(fill = stack_label) +
-    # scale_y_continuous(breaks = max_tpm) +
-    theme_classic() +
-    scale_size(range = c(0,3)) +
-    ylab(
-      bquote('proportion of total '~italic(.('pir'))~' TPM')
-    ) + 
-    facet_grid(.~transmission, 
-               scales = "free_x",
-               space = "free_x") +
-    theme(
-      axis.title.x=element_blank(),
-      axis.text.x=element_text(size = 7, 
-                                 angle = 90),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-```
-
-```{r add_berghei_SorL_prop}
-#Add berghei data
-berghei_SorL_file <- list.files('..', 
-                                recursive = TRUE, 
-                                pattern = 'tl20-05_SorL-subfamily_pir.csv', 
-                                full.names = TRUE)
-berghei_SorL <- read_csv(berghei_SorL_file) %>% 
-  filter(stage %in% c('Asex.SchizA',
-                      'Asex.SchizB',
-                      'Asex.Ring',
-                      'Asex.Troph',
-                      'Asex.Mixed')) 
-berghei_SorL_avg <- berghei_SorL %>% 
-  group_by(stage,
-           SorL) %>% 
-  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
-  mutate(SorL = str_extract(SorL,
-                            pattern = 'S|L|ancestral'),
-         transmission = 'berghei')
-
-```
-
-```{r}
-y <- ggplot(data = rbind(
-  tpm_data_SorL,
-  berghei_SorL_avg
-), 
-aes(
-  x = stage,
-  y = total_tpm, 
-  fill = SorL,
-  # group = transmission
-)) +
-  geom_bar(position = 'fill',
-           stat = 'identity',
-           colour = 'black',
-           size = 0.1) +
-  scale_fill_manual(
-    values = SorL_col,
-    guide = guide_legend(reverse = FALSE)
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  # scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('proportion of total '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~transmission, 
-             scales = "free_x",
-             space = "free_x") +
-  labs(fill = 'sub-family') +
-  theme(
-    axis.title.x=element_blank(),
-    axis.text.x=element_text(angle = 90),
-    axis.text.y = element_text(size = 5),
-    axis.title.y = element_text(size = 8),
-    legend.title = element_text(size = 7),
-    legend.text = element_text(size = 5),
-    legend.key.size = unit(3,'mm'),
-    legend.key.width = unit(2,'mm'),
-    strip.text.x = element_text(size=7, 
-                                face = 'bold', 
-                                angle = 0),
-    strip.background = element_rect(
-      color="black", 
-      size=0.1, 
-      linetype="solid"
-    )
-  ) 
-y
-
-```
-
-```{r SorL_barchart_transmission-facet}
-tpm_data_SorL <- twentyfour_ct_comb%>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-    mutate(stage = times_stages_table$time[match(stage, times_stages_table$r_stage_diff_count)]) %>% 
-  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_info$subfam[match(Geneid, 
-                                            cir_info$Geneid)]) %>% 
-  mutate(SorL = str_extract(clade, 
-                            pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-         # stage =  factor(stage, levels = ordering(stage, order_vector))
-         ) %>% 
-  group_by(transmission,
-           stage, 
-           SorL) %>% 
-  dplyr::summarise(total_tpm = sum(tpm)) %>% 
-  mutate(SorL = factor(SorL, 
-                       levels = sort(unique(SorL))))
-
-y <- ggplot(data = tpm_data_SorL, 
-              aes(
-                x = stage,
-                y = total_tpm, 
-                fill = SorL,
-                # group = transmission
-              )) +
-  geom_bar(position = position_stack(),
-           stat = 'identity',
-           colour = 'black',
-           size = 0.1) +
-  scale_fill_manual(
-    values = SorL_col,
-    guide = guide_legend(reverse = FALSE)
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  # scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~transmission, 
-             scales = "free_x",
-             space = "free_x") +
-  labs(fill = 'sub-family') +
-    theme(
-      axis.title.x=element_blank(),
-      # axis.text.x=element_blank(),
-      axis.text.y = element_text(size = 5),
-      axis.title.y = element_text(size = 8),
-      legend.title = element_text(size = 7),
-      legend.text = element_text(size = 5),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=7, 
-                                  face = 'bold', 
-                                  angle = 0),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      )
-    ) 
-y
-
-```
-
-```{r proportion_data}
-
-tpm_data_twentyfour <- twentyfour_ct_comb%>% 
-  filter(Geneid %in% cir_id) %>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(clade = cir_info$subfam[match(Geneid, 
-                                            cir_info$Geneid)],
-         chrom = str_extract(Geneid, 
-                             pattern = '(?<=PCHAS_)[[:digit:]]{2}')
-         ) %>% 
-  mutate(
-    SorL = str_extract(clade, 
-                       pattern = '(L|S(?=[[:digit:]]))|ancestral|unk|pir-like'),
-    # stage =  factor(stage, 
-    #                 levels = ordering(stage, order_vector))
-  ) %>% 
-  group_by(transmission,
-           stage, 
-           Geneid,
-           clade,
-           chrom,
-           SorL) %>% 
-  dplyr::summarise(tpm = mean(tpm))
-
-```
-
-# Lower pir expression vs rest of genome
-
-```{r low_pir_expression}
-set.seed(4534)
-
-gene_expression_df <- twentyfour_ct_comb%>% 
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(stage = str_extract(sample, 
-                             pattern = '(?<=\\w{3}_\\w{3}_).+(?=_[[:digit:]]$)|sch.culture')) %>%
-  mutate(transmission = str_extract(sample, pattern = 'RMT|SBP')) %>% 
-  filter(stage != 'sch.culture') %>% 
-  mutate(pir_gene = ifelse(Geneid %in% cir_id,
-                           'pir',
-                           'non-pir')) %>% 
-  group_by(stage, 
-           pir_gene,
-           transmission) %>% 
-  slice_sample(n = length(cir_id)) %>%
-  summarise(mean_tpm = mean(tpm),
-            med_tpm = round(median(tpm),
-                            digits = 6))
-
-gene_expression_df %>% 
-  group_by(transmission, 
-           pir_gene) %>% 
-  slice_min(mean_tpm, n =1)
-
-gene_expression_df %>% 
-  group_by(transmission, 
-           pir_gene) %>% 
-  slice_max(mean_tpm, n =1)
-
-y <- ggplot(gene_expression_df,
-       aes(x = pir_gene,
-           y = log2(1+mean_tpm))) +
-  facet_grid(~transmission) +
-  ggtitle(expression(paste(italic('P. chabaudi'), ' 24h cycle'))) +
-  xlab('Gene group')+
-  ylab('log(Median TPM + 1)')+
-  # geom_violin()+
-  geom_dotplot(binaxis = 'y',
-               stackdir = 'center',
-               # binwidth = 1,
-               # dotsize = 2,
-               method = 'dotdensity') +
-  geom_path(aes(group = stage,
-                col = stage)) +
-  theme_classic()
-y
-
-```

### 2021-03-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Adding commands to save the figures

@@
 select <- dplyr::select
-```
-
-```{r tpm, results = 'hide', message = FALSE}
-# # Calculating the transcript-per-million 
-# 
-# We need to write a function to calculate the TPM from the count data we have generated.
-This function uses the 'chabaudi_transcript_lengths' object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.
-
-tpm = function(count_data, lengths = chabaudi_transcript_lengths) {
-
-  #Print an error if the id columns do not match
-  # first if statement to test that both have a GeneID column
-  if(any(grepl(colnames(lengths), 
-               pattern = 'id', 
-               ignore.case = T)) == T & any(grepl(colnames(count_data), 
-                                                  pattern = 'Geneid', 
-                                                  #This needs to be Geneid when some of the samples have 'id' in them
-                                                  ignore.case = T)) == T){
-    if(all(count_data[grep(colnames(count_data), 
-                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
-                                                                             pattern = 'id', ignore.case = T)])) {
-      print('Count data and length gene ids both match')
-    } else {
-      stop('Error: Count data and length gene ids DO NOT MATCH, please make sure they are ordered identically.')
-    }
-  } else {
-    stop('Error: Count data and/or length table supplied do not have an id column.')
-  }
-
-  #Remove the id column from supplied count_data
-  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = T)) == T){
-    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = T)]
-  } else{
-    counts_tpm <- count_data}
-
-  #Remove the id column from supplied lengths
-  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
-    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
-  }else{
-    lengths.table <- lengths}
-  lengths.table = matrix(lengths.table)
-  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop
-
-  #Now for the actual calculation
-  rpm <- counts_tpm
-  for (i in 1:dim(counts_tpm)[2]) {
-    for (j in 1:dim(counts_tpm)[1]) {
-      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
-    }
-    scale.factor <- sum(rpm[,i])/(1E6)
-    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
-  }
-  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
-    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = T), drop = FALSE], 
-                           counts_tpm)
-  } 
-  return(counts_tpm)
-}
+
+#Global save var - set to TRUE if you want to save figures.
+global_save_var <- TRUE
 ```
@@
 
+         if(global_save_var){        
+           pdf('Manuscript_figure3c_heatmap.pdf'
+               , width = 10
+               , height = 4
+           )
+           draw(ht_list, 
+                row_title_side = 'left',
+                column_title_side = 'bottom',
+                heatmap_legend_side = 'right',
+                annotation_legend_side = 'right',
+                annotation_legend_list = list(clade_lgd),
+                ht_gap = unit(.35,'cm'),
+                merge_legend = TRUE)
+           dev.off()
+         }
+
          return(
@@
 
+if(global_save_var){
+  ggsave2(filename = 'Manuscript_SuppFigure3_chabaudi24h_pca_nopseudo.png',
+          plot = plot,
+          dpi = 300,
+          width = 15,
+          height = 10,
+          units = 'cm')
+}
+
 ```
@@
 
-# Phaseograms of all the genes in both transmission methods - supplementary figure 5.
+# Phaseograms of all the genes.
 
@@
 
-## Phaseograms of the entire genome.
+## Supplementary figure 5 - Phaseograms of the entire genome.
 
 ```{r phaseograms_whole_genome}
-
+if(global_save_var){
+  png(filename = 'Manuscript_SuppFig5_phaseogram_allgenes.png',
+      units = 'cm',
+      height = 20,
+      width = 22,
+      res = 330)
   draw_heatmap_zscore(zscore_list = twentyfour_z,
@@
                       tpm_mat = FALSE)
-
-```
-
-## Phaseograms of the _cir_ genes with clade annotations. - Supplementary figure 6
+  dev.off()
+}
+
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE)
+
+```
+
+##  Supplementary figure 6 - Phaseograms of the _cir_ genes with clade annotations
 
@@
 
+if(global_save_var){
+  png(filename = 'Manuscript_SuppFig6_phaseogram_cirgenes.png',
+      units = 'cm',
+      height = 20,
+      width = 22,
+      res = 200)
   draw_heatmap_zscore(zscore_list = twentyfour_z,
@@
                       tpm_mat = FALSE)
-
-```
-
-## Figure 4 and supplementary figure 7/8
+  dev.off()
+}
+
+draw_heatmap_zscore(zscore_list = twentyfour_z,
+                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
+                    filter_genes_vector = cir_id,
+                    annotation_clade = 'SorL',
+                    annotation_col = SorL_col,
+                    threshold_tpm = 1,
+                    use_times = TRUE,
+                    use_order_vector = FALSE,
+                    show_row_names = FALSE,
+                    tpm_mat = FALSE)
+
+```
+
+## Figure 4 and supplementary figure 7
 
@@
              # print(num)
+           }
+
+           if(global_save_var){
+             ggsave2(
+               paste0('Pchabaudi_side-by-side_barchart_facetbySubfam_',
+                      time_vec,
+                      '.png'),
+               plot = g,
+               dpi = 300,
+               width = 28,
+               height = 8,
+               units = 'cm'
+             )
            }
@@
 
+    if(global_save_var){
+      ggsave2(
+        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
+               time_vec,
+               '.png'),
+        plot = y,
+        dpi = 300,
+        width = 25,
+        height = 10,
+        units = 'cm'
+      )
+    }
+
     return(y)

### 2021-03-23 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
Saving png files as PDFs

@@
 if(global_save_var){
-  png(filename = 'Manuscript_SuppFig5_phaseogram_allgenes.png',
-      units = 'cm',
-      height = 20,
-      width = 22,
-      res = 330)
+  pdf(file = 'Manuscript_SuppFig5_phaseogram_allgenes.pdf',
+      height = 5,
+      width = 5)
   draw_heatmap_zscore(zscore_list = twentyfour_z,
@@
 if(global_save_var){
-  png(filename = 'Manuscript_SuppFig6_phaseogram_cirgenes.png',
-      units = 'cm',
-      height = 20,
-      width = 22,
-      res = 200)
+ pdf(file = 'Manuscript_SuppFig6_phaseogram_cirgenes.pdf',
+      height = 5,
+      width = 5)
   draw_heatmap_zscore(zscore_list = twentyfour_z,

### 2021-04-20 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
New chab heatmaps file

--- 2021-03-23 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
@@
 ---
-title: "Manuscript_Github_Pchabaudi_pir"
+title: "Manuscript_Github_Pchabaudi_pir_TPMHeatmaps"
 author: "Timothy Little"
-date: "22/03/2021"
+date: "20/04/2021"
 output: html_document
@@
 ```
-
-## Figure 3c - Highest expressed all _pir_ heatmap
-
-```{r highest_expressed_stage_heatmap_fig3b}
-
-num_SorL <- cir_info %>% 
-  group_by(SorL) %>% 
-  summarise(num = n_distinct(Gene.ID))
-
-labels_SorL <- setNames(paste0(num_SorL$SorL, ' (n = ', num_SorL$num, ')'),
-                        num_SorL$SorL)
-
-# Get order of pir genes by highest expressed genes in each sample.
-
-combined_data_pir_ht_list <- lapply(names(twentyfour_avgtpm), 
-       function(list_name){
-
-         combined_data_pir <- getElement(twentyfour_avgtpm, 
-                                     list_name)$tpm %>% 
-           filter(Geneid %in% cir_id) %>% 
-           column_to_rownames(var = 'Geneid')
-
-         #Ordering that a more understandable order
-
-         combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
-                                                            order_vector)),
-                                                drop = FALSE]
-
-         combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
-                                         means = apply(combined_data_pir, 1, mean),
-                                         subfam = cir_info$SorL[match(names(apply(combined_data_pir, 1, mean)), 
-                                                                            cir_info$Geneid)]) %>% 
-           filter(means > 1) %>% 
-           arrange(subfam, 
-                   desc(means)) %>% 
-           .$Geneid
-
-         #Order the heatmap data by the order defined above (SorL, then mean expression).
-         combined_data_pir_ht <- combined_data_pir[
-           order(
-             match(rownames(combined_data_pir),
-                   combined_data_ordered),
-             na.last = NA #Remove NAs
-           ),
-         ]
-
-         #Change the colnames to the times
-         colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h')
-
-         #Order by time point
-         combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                      drop = FALSE] 
-         #Add a column for the SorL
-         combined_data_pir_ht$SorL <- cir_info$SorL[match(names(apply(combined_data_pir_ht, 1, mean)), 
-                                                                cir_info$Geneid)]
-
-         return(combined_data_pir_ht)
-       })
-names(combined_data_pir_ht_list) <- names(twentyfour_avgtpm)
-
-maxcol <- ceiling(
-  max(
-    unlist(
-      lapply(combined_data_pir_ht_list,
-             function(list_element){
-               max(
-                 log2(1+list_element[!colnames(list_element) %in% 'SorL'])
-               )
-             }
-      )
-    )
-  )
-)
-
-lapply(names(combined_data_pir_ht_list),
-       function(list_name_to_order){
-
-         combined_data_pir_ht <- getElement(combined_data_pir_ht_list,
-                                            list_name_to_order)
-
-         pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
-         #Sub-family annotation
-         clade_info <- combined_data_pir_ht$SorL
-         annotation_col <- SorL_col
-
-         #Subfamily annotation
-
-         clade_ha <- rowAnnotation(
-           'sub-family' = anno_simple(
-             clade_info,
-             col = annotation_col,
-           ),
-           annotation_name_gp = gpar(fontsize = 10)
-         )
-
-         anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
-                                                 labels = 'ancestral',
-                                                 labels_gp = gpar(fontsize = 7),
-                                                 link_width = unit(3, "mm")))
-
-         #lapply across the datasets now, for each order
-         ht_list <- lapply(names(combined_data_pir_ht_list),
-                           function(list_name){
-
-                             combined_data_pir_ht <- getElement(twentyfour_avgtpm, 
-                                                                list_name)$tpm %>% 
-                               column_to_rownames(var = 'Geneid')
-
-                             #Get the times instead of the stage names.
-                             colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-                                                                           pattern = '[[:digit:]]{2}h')
-
-                             combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-                                                                          drop = FALSE]
-
-                             Heatmap(
-                               log2(1+combined_data_pir_ht)[pir_highest_tpm_order,],
-                               cluster_rows = FALSE,
-                               cluster_columns = FALSE,
-                               show_row_names = FALSE,
-                               name = 'log(TPM+1)', 
-                               col = colorRamp2(c(0,
-                                                  maxcol/4,
-                                                  maxcol/2,
-                                                  3*maxcol/4,
-                                                  maxcol),
-                                                viridis(5)),
-                               na_col = 'white', 
-                               border = TRUE,
-                               border_gp = gpar(lwd = 0.1),
-                               column_names_gp = gpar(fontsize = 10),
-                               row_dend_width = unit(3, "cm"),
-                               width = unit(4.5, 'cm'),
-                               height = unit(7,'cm'),
-                               row_title = NULL,
-                               left_annotation = if(
-                                 list_name %in% dplyr::first(names(combined_data_pir_ht_list))
-                               ){
-                                 clade_ha
-                               }else{
-                                 NULL
-                               },
-                               right_annotation = if(
-                                 list_name %in% dplyr::last(names(combined_data_pir_ht_list))
-                               ){
-                                 anc_ha
-                               }else{
-                                 NULL
-                               },
-                               # row_split =  cir_info$SorL[match(pir_highest_tpm_order, 
-                               #                                        cir_info$Geneid)],
-                               column_title = paste(list_name),
-                               heatmap_legend_param = list(
-                                 labels_gp = gpar(fontsize = 10),
-                                 title_gp = gpar(fontsize = 10,
-                                                 fontface = 'bold'),
-                                 at = c(0,
-                                        maxcol/4,
-                                        maxcol/2,
-                                        3*maxcol/4,
-                                        maxcol)
-                               )
-                             )
-                           }
-         )
-         ht_list <- Reduce(`+`, ht_list)
-         clade_lgd <- Legend(
-           title = 'sub-family',
-           at = labels_SorL,
-           legend_gp = gpar(fill = SorL_col)
-         )
-
-         if(global_save_var){        
-           pdf('Manuscript_figure3c_heatmap.pdf'
-               , width = 10
-               , height = 4
-           )
-           draw(ht_list, 
-                row_title_side = 'left',
-                column_title_side = 'bottom',
-                heatmap_legend_side = 'right',
-                annotation_legend_side = 'right',
-                annotation_legend_list = list(clade_lgd),
-                ht_gap = unit(.35,'cm'),
-                merge_legend = TRUE)
-           dev.off()
-         }
-
-         return(
-           draw(ht_list, 
-                row_title_side = 'left',
-                column_title_side = 'bottom',
-                heatmap_legend_side = 'right',
-                annotation_legend_side = 'right',
-                annotation_legend_list = list(clade_lgd),
-                ht_gap = unit(.35,'cm'),
-                merge_legend = TRUE
-           ))
-       })
-
-```
-
-## 2D plot of PC1 v PC2 - Supp figure 4A
-
-```{r pca_calculation, results = TRUE}
-twentyfour_tpm <- twentyfour_rMT
-rownames_geneid <- twentyfour_tpm$Geneid
-twentyfour_tpm_forpca <- select(twentyfour_tpm, -Geneid)
-rownames(twentyfour_tpm_forpca) <- rownames_geneid
-twentyfour_tpm_pca <- prcomp(twentyfour_tpm_forpca, center = TRUE, scale. = TRUE)
-#Need to look into what exactly scale and center are doing
-```
-
-```{r 2d_pca, results = TRUE}
-twentyfour_tpm_pca_var <- get_pca_var(twentyfour_tpm_pca)
-twentyfour_tpm_pca_coords <- data.frame(twentyfour_tpm_pca_var$coord)
-
-twentyfour_tpm_pca_coords <- data.frame(sample = rownames(twentyfour_tpm_pca_var$coord),
-                                        PC1 = twentyfour_tpm_pca_coords$Dim.1,
-                                        PC3 = twentyfour_tpm_pca_coords$Dim.3,
-                                        PC2 = twentyfour_tpm_pca_coords$Dim.2) %>% 
-  dplyr::mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>% 
-  dplyr::mutate(time = str_split_fixed(stage, pattern = '_', n = 2)[,1])
-
-colornames <- viridis(length(unique(twentyfour_tpm_pca_coords$time)))
-colornames <- setNames(colornames, 
-                       sort(unique(twentyfour_tpm_pca_coords$time)))
-
-pca_importance <- as.data.frame(summary(twentyfour_tpm_pca)$importance)
-pc1_importance <- round(pca_importance$PC1[2]*100, digits = 1)
-pc2_importance <- round(pca_importance$PC2[2]*100, digits = 1)
-pc3_importance <- round(pca_importance$PC3[2]*100, digits = 1)
-
-(
-  plot <- ggplot(twentyfour_tpm_pca_coords,
-                 aes(
-                   x = PC1,
-                   y = PC2,
-                   col = time
-                 )) +
-    geom_point(size = 3,
-               alpha = 0.8) +
-    scale_color_manual(values = colornames) +
-    xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-    ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-    guides(shape=guide_legend(ncol=2),
-           col = guide_legend(ncol=2)) +
-    theme_classic() +
-    theme(legend.key.size = unit(4, 'mm'),
-          legend.text = element_text(size = 7))
-)
-
-if(global_save_var){
-  ggsave2(filename = 'Manuscript_SuppFigure3_chabaudi24h_pca_nopseudo.png',
-          plot = plot,
-          dpi = 300,
-          width = 15,
-          height = 10,
-          units = 'cm')
-}
-
-```
-
-# Calculate and cluster with z-scores {.tabset}
-
-z-score represents the number of standard deviations from the mean of the data. To calculate from TPM or counts for a given gene X:
-
-1. Convert the count/RPKM values of each gene into log values.
-2. Calculate the mean and standard deviation of X gene log values in each sample
-3. For first sample: (gene X log value - mean of log values of samples)/ standard deviation of log values of all samples
-4. Now. i have the z-score for gene x in first sample. Using the above protocol, I can convert all genes log values into z-score.
-
-* Need to make sure that the z-score is calculated within each transmission method!
-
-```{r zscores}
-tpm_to_zscore = function(input){
-
-  #If the input does not have a TPM element then assume that the input itself is the TPM object
-  if(any(names(input) %in% 'tpm')){
-    tpm_object <- input$tpm
-  }else{
-    tpm_object <- input
-    warning('No \'tpm\' element in the supplied input so function will assume that the input is the tpm data')
-  }
-
-  #Want all of them so will just ignore Geneid
-  input_logged <- tpm_object %>%
-    mutate_at(vars(-matches('Geneid')), 
-              log1p)
-
-  gene_mean_scores = input_logged %>% 
-    melt(id.vars = 'Geneid') %>% 
-    group_by(Geneid) %>% 
-    dplyr::summarise(mean = mean(value),
-                     std_dev = sd(value))
-
-  gene_vector = as.character(input_logged$Geneid)
-
-  input_z = lapply(gene_vector, function(gene){
-    mean_gene = filter(gene_mean_scores, 
-                       Geneid == paste(gene)) %>% 
-      select(mean)
-    std_gene = filter(gene_mean_scores, 
-                      Geneid == paste(gene)) %>% 
-      select(std_dev)
-    gene_scores = input_logged %>% 
-      filter(Geneid == paste(gene)) %>% 
-      select(-Geneid)
-    if(std_gene != 0){
-      gene_z = sapply(gene_scores, function(score){(score-mean_gene)/std_gene}) %>% data.frame
-    }else{
-      gene_z = (sapply(gene_scores, function(score){0})) %>% data.frame %>% t
-    }
-    colnames(gene_z) = colnames(gene_scores) #For some reason apply makes the column names weird
-    gene_z = cbind(data.frame(Geneid = paste(gene)), gene_z)
-  })
-
-  #Use reduce and rbind to turn the list into one df
-  input_z_df = purrr::reduce(input_z, 
-                             rbind)
-  return(list(zscores = input_z_df,
-              tpm = tpm_object,
-              experiment = input$experiment))
-}
-
-#Make and save z-scores
-twentyfour_z <- lapply(twentyfour_avgtpm,
-                       tpm_to_zscore)
-
-###
-```
-
-# Phaseograms of all the genes.
-
-```{r row_order_phaseograms}
-
-#Getting the row order for each phaseogram for each experiment
-# Want the data ordered by the highest zscore for each gene in each stage.
-phaseogram_row_order <- lapply(twentyfour_z, 
-                               function(list_elements){
-                                 zscore_df <- list_elements$zscores
-                                 zscore_ordered <- melt(zscore_df, 
-                                                        id = 'Geneid',
-                                                        variable.name = 'stage',
-                                                        value.name = 'zscore') %>% 
-                                   group_by(Geneid) %>% 
-                                   top_n(zscore, n = 1) %>% 
-                                   arrange(stage, 
-                                           desc(zscore)) %>%
-                                   distinct(Geneid, 
-                                            .keep_all = TRUE) %>%  #Keep only the ones that are uniquely highest expressed in one sample
-                                   filter(zscore > 0) #For some reason some remain with zero zscore
-                                 # zscore_ordered <- zscore_ordered[order(
-                                 #   match(
-                                 #     zscore_ordered$stage, 
-                                 #     order_vector
-                                 #   )
-                                 # ),]
-                                 return(c(list_elements, 
-                                          list(zscore_order = zscore_ordered)))
-                               })
-
-phaseogram_time_order <- lapply(names(phaseogram_row_order), 
-                                function(list_element){
-                                  transmission_list <- getElement(phaseogram_row_order,
-                                                                  list_element)
-                                  order_by_time <- transmission_list$zscore_order %>% 
-                                    mutate(stage = times_stages_table$time[match(stage,
-                                                                                 times_stages_table$r_stage_diff_count)]) %>% 
-                                    arrange(stage)
-                                  return(list(zscore_order = order_by_time))
-                                }
-)
-names(phaseogram_time_order) <- names(phaseogram_row_order)
-
-```
-
-```{r function_draw_heatmap_zscore}
-
-draw_heatmap_zscore <- function(zscore_list,
-                                title = '',
-                                gene_ordering_vector = NULL,
-                                filter_genes_vector = NULL,
-                                extra_title = NULL,
-                                annotation_clade = NULL,
-                                annotation_col,
-                                threshold_tpm = NULL,
-                                use_times = FALSE,
-                                use_order_vector = TRUE,
-                                tpm_mat = TRUE,
-                                tpm_breaks = NULL,
-                                ...){
-  if(is.null(tpm_breaks) & tpm_mat){
-    #Need the generate the breaks in the data for the heatmap legend if not provided.
-    max_tpm <- lapply(zscore_list, 
-                      function(list_element){
-                        #list_element <- zscore_list[[1]]
-                        heatmap_data <- as.data.frame(list_element$zscores)
-                        title <- list_element$experiment
-
-                        if(!'Geneid' %in% colnames(heatmap_data)){
-                          stop('No Geneid column in the heatmap data.frame')
-                        }
-
-                        if(!is.null(filter_genes_vector)){
-                          heatmap_data <- as.data.frame(
-                            filter(heatmap_data, 
-                                   Geneid %in% filter_genes_vector)
-                          )
-                        }
-
-                        rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                        heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                        #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                        if(!is.null(gene_ordering_vector)){
-                          #Remove genes not in the main data frame but in the ordering vector, 
-                          # and then order based on the vector
-                          remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          heatmap_data <- heatmap_data[na.omit(
-                            match(gene_ordering_vector, 
-                                  rownames(heatmap_data))
-                          ),,
-                          drop=FALSE]
-                        }
-
-                        geneids <- rownames(heatmap_data)
-
-                        #Get max/min TPM scores for the side annotation heatmap                       
-                        tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                          list_element$tpm$Geneid),] %>% 
-                          select(-Geneid) %>% 
-                          as.matrix %>% 
-                          log1p
-                        rownames(tpm_mat) <- geneids
-
-                        tpm_min <- rowMins(tpm_mat)
-                        names(tpm_min) <- geneids
-                        tpm_max <- rowMaxs(tpm_mat)
-                        names(tpm_max) <- geneids
-                        return(list(tpm_max = tpm_max,
-                                    tpm_min = tpm_min))
-                      })
-
-    tpm_max_all <- mround(
-      max(
-        unlist(
-          lapply(max_tpm, 
-                 getElement,
-                 'tpm_max')
-        )
-      ), 
-      base = 1)
-
-    tpm_breaks <- round(
-      c(0,
-        tpm_max_all/4,
-        tpm_max_all/2,
-        tpm_max_all*(3/4),
-        tpm_max_all),
-      digits = 2
-    )
-  }
-
-  #Generate the colour function from the TPM legend breaks for the TPM annotation
-  if(tpm_mat){
-    tpm_col_fun <- colorRamp2(
-      breaks = tpm_breaks,
-      colors = viridis(length(tpm_breaks))
-    )
-  }
-
-  if(is.numeric(threshold_tpm)){
-    #Apply a threshold to the TPM in order to remove low expressed genes.
-    # Remember to add log1p here! 
-    # Threshold uses the max value, so it ony needs to be above thresh 
-    #   in one stage avg in one transmission method.
-    #Will alter the gene_filter vector in order to only select the genes above threshold
-
-    gene_maxs <- bind_rows(
-      lapply(max_tpm, 
-             getElement,
-             'tpm_max')
-    )
-
-    remove_genes <- names(gene_maxs)[apply(
-      gene_maxs, 
-      2, 
-      function(maxs){
-        all(maxs < log2(1+threshold_tpm))
-      }
-    )]
-
-    filter_genes_vector <- filter_genes_vector[!filter_genes_vector %in% remove_genes]
-
-  }
-
-  if(is.character(annotation_clade)){
-    order_cir <- gene_ordering_vector[gene_ordering_vector %in% filter_genes_vector]
-    clade_info <- getElement(cir_info,
-                             annotation_clade)[match(order_cir, 
-                                                     cir_info$Geneid)]
-    names(clade_info) <- str_split_fixed(order_cir, 
-                                         pattern = '_', 
-                                         n = 2)[,2]
-    annotation_col <- annotation_col
-  }
-
-  heatmaps <- lapply(zscore_list, 
-                     function(list_element){
-                       #list_element <- zscore_list[[1]]
-                       heatmap_data <- as.data.frame(list_element$zscores)
-                       title <- list_element$experiment
-
-                       if(!'Geneid' %in% colnames(heatmap_data)){
-                         stop('No Geneid column in the heatmap data.frame')
-                       }
-
-                       if(!is.null(filter_genes_vector)){
-                         heatmap_data <- as.data.frame(
-                           filter(heatmap_data, 
-                                  Geneid %in% filter_genes_vector)
-                         )
-                       }
-
-                       rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                       heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                       #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-
-                       if(!is.null(gene_ordering_vector)){
-                         #Remove genes not in the main data frame but in the ordering vector,
-                         #  and then order based on the vector
-                         remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                         heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                         heatmap_data <- heatmap_data[na.omit(
-                           match(gene_ordering_vector, 
-                                 rownames(heatmap_data))
-                         ),,
-                         drop=FALSE]
-                       }
-
-                       if(tpm_mat){
-                         #Get max/min TPM scores for the side annotation heatmap                       
-                         tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                           list_element$tpm$Geneid),] %>%
-                           select(-Geneid) %>% 
-                           as.matrix %>% 
-                           log1p
-
-                         tpm_min <- rowMins(tpm_mat)
-                         tpm_max <- rowMaxs(tpm_mat)
-
-                         names(tpm_min) = rownames(heatmap_data)
-                         names(tpm_max) = rownames(heatmap_data)
-
-                         tpm_ha <- rowAnnotation(
-                           min.max = anno_simple(cbind(tpm_min, 
-                                                       tpm_max), 
-                                                 col = tpm_col_fun,
-                                                 gp = gpar(col = 'black',
-                                                           lwd = 0.1),
-                                                 border = TRUE,
-                                                 simple_anno_size = unit(0.2, "cm")),
-                           show_annotation_name = FALSE)
-                       }
-
-                       #Draw the left annotation is needed, only for the first element.
-
-                       if(
-                         identical(list_element,
-                                   dplyr::first(zscore_list)) & is.character(annotation_clade)
-                       ){
-                         clade_ha <- rowAnnotation(
-                           'subfamily' = anno_simple(
-                             clade_info,
-                             col = annotation_col
-                           )
-                         )
-                       }
-
-                       #Change rownames to remove PCHAS_
-                       rownames(heatmap_data) = str_split_fixed(rownames(heatmap_data), 
-                                                                pattern = '_', 
-                                                                n = 2)[,2]
-
-                       if(use_times){
-                         colnames(heatmap_data) <- str_extract(colnames(heatmap_data), 
-                                                               pattern = '[[:digit:]]{2}h')
-                       }
-
-                       if(use_order_vector){
-                         #Order the columns by ordering_vector
-                         heatmap_data <- heatmap_data[,order(match(colnames(heatmap_data), 
-                                                                   order_vector)),
-                                                      drop = FALSE]
-                       }else{
-                         heatmap_data <- heatmap_data[,order(colnames(heatmap_data)),
-                                                      drop = FALSE] 
-                       }
-
-                       h = Heatmap(heatmap_data,
-                                   name = paste0('z-score',
-                                                 extra_title),
-                                   col = c(cbPalette[6],
-                                           'white', 
-                                           cbPalette[7]),
-                                   row_names_gp = gpar(fontsize = ifelse(
-                                     nrow(heatmap_data) > 13, 
-                                     12*(13/nrow(heatmap_data)), 
-                                     12)
-                                   ),
-                                   column_names_gp = gpar(fontsize = 10),
-                                   row_dend_width = unit(3, "cm"),
-                                   right_annotation = if(tpm_mat){
-                                     tpm_ha
-                                   }else{
-                                     NULL
-                                   },
-                                   left_annotation = 
-                                     if(identical(list_element,
-                                                  dplyr::first(zscore_list)) & is.character(annotation_clade)){
-                                       clade_ha
-                                     }else{
-                                       NULL
-                                     }, 
-                                   #Only want the clade annotation when it is first heatmap
-                                   width = unit(4.5, 'cm'),
-                                   cluster_columns = FALSE,
-                                   height = unit(7,'cm'),
-                                   row_title = NULL,
-                                   cluster_rows = FALSE,
-                                   column_title = paste(title),
-                                   heatmap_legend_param = list(
-                                     at = c(3,2,1,0,-1,-2,-3),
-                                     labels_gp = gpar(fontsize = 10),
-                                     title_gp = gpar(fontsize = 10)
-                                   ),
-                                   ...
-                       )
-                       return(h)
-                     })
-
-  ht_list <- Reduce('+', heatmaps) #To add the list elements together
-
-  if(tpm_mat){
-    #Draw the legend
-    tpm_lgd <- Legend(
-      title = "Min/Max log(TPM + 1)", 
-      col = tpm_col_fun, 
-      at = tpm_breaks, 
-      labels = tpm_breaks,
-      title_gp = gpar(fontsize = 7)
-    )
-  }
-
-  if(is.character(annotation_clade)){
-    clade_lgd <- Legend(
-      title = 'sub-family',
-      at = names(annotation_col),
-      legend_gp = gpar(fill = annotation_col)
-    )
-  }
-
-  draw(ht_list, 
-       row_title_side = 'left',
-       column_title_side = 'bottom',
-       heatmap_legend_side = 'left',
-       #Only include the annotation legends for the annotations we actually have.
-       annotation_legend_list = if(is.character(annotation_clade) & tpm_mat){
-         list(tpm_lgd,
-              clade_lgd)
-       }else if(tpm_mat & !is.character(annotation_clade) ){
-         list(tpm_lgd)
-       } else if(!tpm_mat & is.character(annotation_clade) ){
-         list(clade_lgd)
-       })
-}
-```
-
-## Supplementary figure 5 - Phaseograms of the entire genome.
-
-```{r phaseograms_whole_genome}
-if(global_save_var){
-  pdf(file = 'Manuscript_SuppFig5_phaseogram_allgenes.pdf',
-      height = 5,
-      width = 5)
-  draw_heatmap_zscore(zscore_list = twentyfour_z,
-                      gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
-                      use_times = TRUE,
-                      use_order_vector = FALSE,
-                      show_row_names = FALSE,
-                      tpm_mat = FALSE)
-  dev.off()
-}
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
-```
-
-##  Supplementary figure 6 - Phaseograms of the _cir_ genes with clade annotations
-
-```{r pir_gene_phaseogram_breaks}
-  max_tpm <- lapply(twentyfour_z, 
-                    function(list_element){
-                      #list_element <- zscore_list[[1]]
-                      heatmap_data <- as.data.frame(list_element$zscores)
-                      title <- list_element$experiment
-
-                      if(!'Geneid' %in% colnames(heatmap_data)){
-                        stop('No Geneid column in the heatmap data.frame')
-                      }
-
-                            heatmap_data <- as.data.frame(
-                              filter(heatmap_data, 
-                                     Geneid %in% cir_id)
-                            )
-
-                          rownames(heatmap_data) <- as.character(heatmap_data$Geneid)
-
-                          heatmap_data <- heatmap_data[, !colnames(heatmap_data) %in% 'Geneid', drop = TRUE]
-                          #This won't work if heatmap_data is a tibble, must be a data.frame, so annoying!!
-                          # 
-                          # if(!is.null(gene_ordering_vector)){
-                          #   #Remove genes not in the main data frame but in the ordering vector, 
-                          #   # and then order based on the vector
-                          #   remove_index <- rownames(heatmap_data) %in% gene_ordering_vector
-                          #   heatmap_data <- heatmap_data[remove_index,,drop = FALSE]
-                          #   heatmap_data <- heatmap_data[na.omit(
-                          #     match(gene_ordering_vector, 
-                          #           rownames(heatmap_data))
-                          #   ),,
-                          #   drop=FALSE]
-                          # }
-
-                          geneids <- rownames(heatmap_data)
-
-                          #Get max/min TPM scores for the side annotation heatmap                       
-                          tpm_mat <- list_element$tpm[match(rownames(heatmap_data),
-                                                            list_element$tpm$Geneid),] %>% 
-                            select(-Geneid) %>% 
-                            as.matrix %>% 
-                            log1p
-                          rownames(tpm_mat) <- geneids
-
-                          tpm_min <- rowMins(tpm_mat)
-                          names(tpm_min) <- geneids
-                          tpm_max <- rowMaxs(tpm_mat)
-                          names(tpm_max) <- geneids
-                          return(list(tpm_max = tpm_max,
-                                      tpm_min = tpm_min))
-                        })
-
-  tpm_max_all <- mround(
-    max(
-      unlist(
-        lapply(max_tpm, 
-               getElement,
-               'tpm_max')
-      )
-    ), 
-    base = 1)
-
-  tpm_breaks <- round(
-    c(0,
-      tpm_max_all/4,
-      tpm_max_all/2,
-      tpm_max_all*(3/4),
-      tpm_max_all),
-    digits = 2
-  )
-```
-
-```{r phaseograms_cir}
-
-if(global_save_var){
- pdf(file = 'Manuscript_SuppFig6_phaseogram_cirgenes.pdf',
-      height = 5,
-      width = 5)
-  draw_heatmap_zscore(zscore_list = twentyfour_z,
-                      gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
-                      filter_genes_vector = cir_id,
-                      annotation_clade = 'SorL',
-                      annotation_col = SorL_col,
-                      threshold_tpm = 1,
-                      use_times = TRUE,
-                      use_order_vector = FALSE,
-                      show_row_names = FALSE,
-                      tpm_mat = FALSE)
-  dev.off()
-}
-
-draw_heatmap_zscore(zscore_list = twentyfour_z,
-                    gene_ordering_vector = phaseogram_time_order$RMT$zscore_order$Geneid,
-                    filter_genes_vector = cir_id,
-                    annotation_clade = 'SorL',
-                    annotation_col = SorL_col,
-                    threshold_tpm = 1,
-                    use_times = TRUE,
-                    use_order_vector = FALSE,
-                    show_row_names = FALSE,
-                    tpm_mat = FALSE)
-
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
 ```
@@
 
-```{r pirs_side-by-side_figure4_RMT-SBP-completelySeparate}
+```{r pirs_side-by-side_figure4}
 
@@
 ```
-
-# Supp figure 9 - Pir expression of AAPLs and ChAPLs.
-
-```{r Supp_fig_AapL-ChaAPl_expression}
-
-cir_plot <- twentyfour_rMT %>% 
-  filter(Geneid %in% cir_id) %>%
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(
-    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
-    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
-    chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
-                                                       cir_info$Geneid)],
-                        levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
-                                                                            cir_info$Geneid)])))) %>% 
-  filter(!(time %in% c('sch', 
-                       'Unknown')),
-         !is.na(chapl_aapl) ) %>% 
-  group_by(Geneid) %>% 
-    arrange(desc(tpm))
-
-  cir_plot_order <- cir_plot %>% 
-    filter(time == '14h') %>% 
-    arrange(desc(tpm))
-
-  cir_plot <- mutate(cir_plot, 
-                     Geneid = factor(Geneid, 
-                                     levels = unique(cir_plot_order$Geneid)))
-  cir_plot_chapl_aapl_supp_info <- cir_plot
-
-  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))
-
-  breaks = seq(0, max_aapl_chapl, by = 1)
-    limits = c(0, max_aapl_chapl)
-    point_size = 0.25
-
-  lapply(
-    sort(unique(cir_plot$time)),
-    function(time_vec){
-
-    cir_plot <- filter(cir_plot, 
-                       time == time_vec)
-
-      cir_plot$tpm <- log2(1+cir_plot$tpm)
-
-    cir_plot$Geneid <- factor(
-      str_remove(cir_plot$Geneid, 
-                 pattern = 'PCHAS_'),
-      levels = str_remove(levels(cir_plot$Geneid), 
-                          pattern = 'PCHAS_')
-    )
-
-    y <- ggplot(
-      data = cir_plot,
-      mapping = aes(
-        x = Geneid,
-        y = tpm
-      )
-    ) +
-      stat_summary(fun.y = median, 
-                   fun.ymin = median, 
-                   fun.ymax = median,
-                   geom = "bar", 
-                   width = 0.75, 
-                   size = 0.25, 
-                   position = 'dodge',
-                   fill = 'lightblue') +
-      geom_point(stat = 'identity', 
-                 position = position_dodge(width = 0.75),
-                 size = point_size) +
-      scale_y_continuous(breaks = breaks,
-                         limits = limits) +
-      scale_fill_manual(values = c('skyblue','red')) +
-      ylab('log(TPM +1)') +
-      facet_grid(.~chapl_aapl, 
-                 scales = "free_x",
-                 space = "free_x") +
-      theme_classic() +
-      theme(
-        axis.text.x = element_text(angle = 90,
-                                   size = 5)
-      ) +
-      ggtitle(paste(unique(cir_plot$time)))
-
-    if(global_save_var){
-      ggsave2(
-        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
-               time_vec,
-               '.png'),
-        plot = y,
-        dpi = 300,
-        width = 25,
-        height = 10,
-        units = 'cm'
-      )
-    }
-
-    return(y)
-    })
-```

### 2021-04-20 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Heatmap separated by sub-family

@@
 ```
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+
+```{r highest_expressed_stage_heatmap_fig3b}
+
+housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
+                                 'MSP1' = 'PCHAS_0831300',
+                                 'AMA1' = 'PCHAS_0931000')
+
+cir_housekeeping_id <- c(cir_id, 
+                         'PCHAS_0721000',
+                         'PCHAS_0831300',
+                         'PCHAS_0931000')
+```
+
+```{r}
+# Get order of pir genes by highest expressed genes in each sample.
+
+combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
+  filter(Geneid %in% cir_housekeeping_id) %>% 
+  column_to_rownames(var = 'Geneid') %>% 
+  filter_all(
+      all_vars(. > 1)
+  )
+
+#Ordering that a more understandable order
+
+combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
+                                                   order_vector)),
+                                       drop = FALSE]
+
+combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
+                                means = apply(combined_data_pir, 1, mean),
+                                subfam = factor(
+                                  ifelse(rownames(combined_data_pir) %in% cir_id,
+                                         cir_info$subfam[match(rownames(combined_data_pir), 
+                                                               cir_info$Geneid)],
+                                         'ref')
+                                )) %>% 
+  mutate(subfam =
+           forcats::fct_shift(
+             relevel(
+               as.factor(subfam),
+               ref = 'ref')
+           )
+  ) %>% 
+  # filter(means > 1) %>% 
+  arrange(subfam, 
+          desc(means)) %>% 
+  .$Geneid
+
+#Order the heatmap data by the order defined above (Subfam, then mean expression).
+combined_data_pir_ht <- combined_data_pir[
+  order(
+    match(rownames(combined_data_pir),
+          combined_data_ordered),
+    na.last = NA #Remove NAs
+  ),
+]
+
+#Change the colnames to the times
+colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
+                                              pattern = '[[:digit:]]{2}h')
+
+#Order by time point
+combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+                                             drop = FALSE]
+
+#Add a column for the subfam (ref for the reference genes)
+combined_data_pir_ht$subfam <- ifelse(rownames(combined_data_pir_ht) %in% cir_id,
+                                      cir_info$subfam[match(rownames(combined_data_pir_ht), 
+                                                            cir_info$Geneid)],
+                                      'ref')
+
+#Get the ref gene actual names instead of the Geneid
+rownames(combined_data_pir_ht) <- ifelse(
+  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
+  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
+  rownames(combined_data_pir_ht)
+)
+
+#Get the colour range
+maxcol <- ceiling(
+  max(
+    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'subfam'])
+  ))
+
+# pir_highest_tpm_order <- rownames(combined_data_pir_ht)
+
+#Sub-family annotation
+clade_info <- combined_data_pir_ht$subfam
+annotation_col <- c(clade_col, 'ref' = 'black')
+
+#Subfamily annotation
+
+clade_ha <- rowAnnotation(
+  'sub-family' = anno_simple(
+    clade_info,
+    col = annotation_col,
+  ),
+  annotation_name_gp = gpar(fontsize = 10)
+)
+
+#Subfamily name labels
+num_subfam <- cir_info %>% 
+  group_by(subfam) %>% 
+  summarise(num = n_distinct(Gene.ID))
+
+labels_subfam <- c(setNames(paste0(num_subfam$subfam, ' (n = ', num_subfam$num, ')'),
+                            num_subfam$subfam),
+                   'ref' = 'ref')
+
+anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
+                                        labels = 'ancestral',
+                                        labels_gp = gpar(fontsize = 7),
+                                        link_width = unit(3, "mm")))
+
+# combined_data_pir_ht <- twentyfour_avgtpm$RMT$tpm %>% 
+#   column_to_rownames(var = 'Geneid')
+# 
+# #Get the times instead of the stage names.
+# colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
+#                                               pattern = '[[:digit:]]{2}h')
+# 
+# combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+#                                              drop = FALSE]
+
+ht <- Heatmap(
+  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
+  cluster_rows = FALSE,
+  cluster_columns = FALSE,
+  show_row_names = FALSE,
+  name = 'log(TPM+1)', 
+  col = colorRamp2(c(0,
+                     maxcol/4,
+                     maxcol/2,
+                     3*maxcol/4,
+                     maxcol),
+                   viridis(5)),
+  na_col = 'white', 
+  border = TRUE,
+  border_gp = gpar(lwd = 0.1),
+  column_names_gp = gpar(fontsize = 10),
+  row_dend_width = unit(3, "cm"),
+  width = unit(4.5, 'cm'),
+  height = unit(7,'cm'),
+  row_title = NULL,
+  left_annotation = clade_ha,
+  right_annotation = anc_ha,
+  row_split = forcats::fct_shift(
+    relevel(
+      as.factor(combined_data_pir_ht$subfam),
+      ref = 'ref')
+  ),
+  heatmap_legend_param = list(
+    labels_gp = gpar(fontsize = 10),
+    title_gp = gpar(fontsize = 10,
+                    fontface = 'bold'),
+    at = c(0,
+           maxcol/4,
+           maxcol/2,
+           3*maxcol/4,
+           maxcol)
+  )
+)
+
+clade_lgd <- Legend(
+  title = 'sub-family',
+  at = c(labels_subfam),
+  legend_gp = gpar(fill = annotation_col)
+)
+
+draw(ht, 
+     row_title_side = 'left',
+     column_title_side = 'bottom',
+     heatmap_legend_side = 'right',
+     annotation_legend_side = 'right',
+     annotation_legend_list = list(clade_lgd),
+     merge_legend = TRUE)
 ```

### 2021-04-21 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Chapl/Aapl heatmap start

@@
             delim = '\t')
+
+#Add chapl_aapl numbers for easier heatmap drawing later.
+cir_info <- cir_info %>% 
+  group_by(chapl_aapl_name) %>% 
+  mutate(chapl_aapl_name_num = n_distinct(Geneid)) %>% 
+  mutate(chapl_aapl_name_num = paste0(
+                       chapl_aapl_name, 
+                       ' (n = ',
+                       chapl_aapl_name_num,
+                       ')'
+                     ))
 ```
@@
 
+chapl_aapl_name_col <- setNames(
+  c('#F0E442',
+    '#56B4E9', #L1
+    '#009E73',
+    '#0072B2',
+    '#E69F00', #S1
+    '#D55E00',
+    '#CC79A7',
+    '#999999'),
+  sort(
+    unique(cir_info$chapl_aapl_name)
+  )
+)
+
+chapl_aapl_name_col_num <- chapl_aapl_name_col
+names(chapl_aapl_name_col_num) <- sapply(
+  names(chapl_aapl_name_col), 
+  function(clade_name) {
+    paste0(
+      clade_name, 
+      ' (n = ',
+      sum(cir_info$chapl_aapl_name == clade_name, 
+          na.rm = TRUE),
+      ')'
+    )
+  })
 ```
+
+_pir_ genes selected by being expressed above 1 TPM in at least one sample.
+
+Adding in housekeeping genes for comparison:
+HSP70 putative  PCHAS_0721000
+MSP1 PCHAS_0831300
+AMA1 PCHAS_0931000
 
@@
                          'PCHAS_0931000')
-```
 
-```{r}
 # Get order of pir genes by highest expressed genes in each sample.
@@
   filter_all(
-      all_vars(. > 1)
+      any_vars(. > 1)
   )
@@
 
-# combined_data_pir_ht <- twentyfour_avgtpm$RMT$tpm %>% 
-#   column_to_rownames(var = 'Geneid')
-# 
-# #Get the times instead of the stage names.
-# colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
-#                                               pattern = '[[:digit:]]{2}h')
-# 
-# combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
-#                                              drop = FALSE]
-
 ht <- Heatmap(
@@
 
-## Figure 4 and supplementary figure 7
+# Supp figure 9 - Pir expression of AAPLs and ChAPLs.
 
-_pir_ genes selected by being expressed above 1 TPM in all samples.
+_pir_ genes selected by being expressed above 1 TPM in at least one sample.
 
-Adding in housekeeping genes for comparison:
-HSP70 putative  PCHAS_0721000
-MSP1 PCHAS_0831300
-AMA1 PCHAS_0931000
+```{r chapl_appl_heatmap_suppfig9}
 
-```{r pirs_side-by-side_figure4}
+# Get order of pir genes by highest expressed genes in each sample.
 
-housekeeping_genes <- data.frame('HSP70' = 'PCHAS_0721000',
-                                 'MSP1' = 'PCHAS_0831300',
-                                 'AMA1' = 'PCHAS_0931000')
+combined_data_pir <- twentyfour_avgtpm$RMT$tpm %>% 
+  filter(Geneid %in% cir_id) %>% 
+  mutate(chapl_aapl_name = cir_info$chapl_aapl_name_num[match(Geneid, cir_info$Geneid)]) %>% 
+  filter(!str_detect(chapl_aapl_name, pattern = 'NA')) %>% 
+  column_to_rownames(var = 'Geneid') %>% 
+  filter_all(
+      any_vars(. > 1)
+  )
 
-cir_housekeeping_id <- c(cir_id, 
-                         'PCHAS_0721000',
-                         'PCHAS_0831300',
-                         'PCHAS_0931000')
+#Ordering that a more understandable order
+
+combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
+                                                   order_vector)),
+                                       drop = FALSE]
+
+combined_data_ordered <- tibble(Geneid = rownames(combined_data_pir),
+                                means = apply(combined_data_pir, 1, mean),
+                                chapl_aapl_name = factor(
+                                  combined_data_pir$chapl_aapl_name
+                                )) %>% 
+  arrange(chapl_aapl_name, 
+          desc(means)) %>% 
+  .$Geneid
+
+#Order the heatmap data by the order defined above (chapl_aapl_name, then mean expression).
+combined_data_pir_ht <- combined_data_pir[
+  order(
+    match(rownames(combined_data_pir),
+          combined_data_ordered),
+    na.last = NA #Remove NAs
+  ),
+]
+
+#Change the colnames to the times
+colnames(combined_data_pir_ht) <- str_extract(colnames(combined_data_pir_ht),
+                                              pattern = '[[:digit:]]{2}h|chapl_aapl_name')
+
+#Order by time point
+combined_data_pir_ht <- combined_data_pir_ht[,order(colnames(combined_data_pir_ht)),
+                                             drop = FALSE]
+
+#Get the ref gene actual names instead of the Geneid
+rownames(combined_data_pir_ht) <- ifelse(
+  rownames(combined_data_pir_ht) %in% housekeeping_genes[1,],
+  colnames(housekeeping_genes)[ match(rownames(combined_data_pir_ht), housekeeping_genes[1,]) ],
+  rownames(combined_data_pir_ht)
+)
+
+#Get the colour range
+maxcol <- ceiling(
+  max(
+    log2(1+combined_data_pir_ht[!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name'])
+  ))
+
+# pir_highest_tpm_order <- rownames(combined_data_pir_ht)
+
+#Sub-family annotation
+chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
+annotation_col <- c(chapl_aapl_name_col_num, 'ref' = 'black')
+
+#chapl_aapl_name annotation
+
+chapl_aapl_name_ha <- rowAnnotation(
+  'sub-family' = anno_simple(
+    chapl_aapl_name_info,
+    col = annotation_col,
+  ),
+  annotation_name_gp = gpar(fontsize = 10)
+)
+
+#chapl_aapl_name name labels
+num_chapl_aapl_name <- cir_info %>% 
+  group_by(chapl_aapl_name) %>% 
+  summarise(num = n_distinct(Gene.ID))
+
+labels_chapl_aapl_name <- setNames(paste0(num_chapl_aapl_name$chapl_aapl_name, 
+                                          ' (n = ', num_chapl_aapl_name$num, ')'),
+                                   num_chapl_aapl_name$chapl_aapl_name)
+
+ht <- Heatmap(
+  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
+  cluster_rows = FALSE,
+  cluster_columns = FALSE,
+  show_row_names = FALSE,
+  name = 'log(TPM+1)', 
+  col = colorRamp2(c(0,
+                     maxcol/4,
+                     maxcol/2,
+                     3*maxcol/4,
+                     maxcol),
+                   viridis(5)),
+  na_col = 'white', 
+  border = TRUE,
+  border_gp = gpar(lwd = 0.1),
+  column_names_gp = gpar(fontsize = 10),
+  row_dend_width = unit(3, "cm"),
+  width = unit(4.5, 'cm'),
+  height = unit(7,'cm'),
+  row_title = NULL,
+  left_annotation = chapl_aapl_name_ha,
+  row_split = 
+    factor(combined_data_pir_ht$chapl_aapl_name,
+              levels = ordering(combined_data_pir_ht$chapl_aapl_name,
+                                order_vector = 'ChAPL')),
+  heatmap_legend_param = list(
+    labels_gp = gpar(fontsize = 10),
+    title_gp = gpar(fontsize = 10,
+                    fontface = 'bold'),
+    at = c(0,
+           maxcol/4,
+           maxcol/2,
+           3*maxcol/4,
+           maxcol)
+  )
+)
+
+chapl_aapl_name_lgd <- Legend(
+  title = 'sub-family',
+  at = factor(labels_chapl_aapl_name,
+              ordering(labels_chapl_aapl_name,
+                order_vector = 'ChAPL')),
+  legend_gp = gpar(fill = annotation_col)
+)
+
+draw(ht, 
+     row_title_side = 'left',
+     column_title_side = 'bottom',
+     heatmap_legend_side = 'right',
+     annotation_legend_side = 'right',
+     annotation_legend_list = list(chapl_aapl_name_lgd),
+     merge_legend = TRUE)
+```
+
+```{r Supp_fig_AapL-ChaAPl_expression}
 
 cir_plot <- twentyfour_rMT %>% 
-  filter(Geneid %in% cir_housekeeping_id) %>%
+  filter(Geneid %in% cir_id) %>%
   melt(id.vars = 'Geneid', 
@@
     time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)]
-  ) %>% 
-  mutate(subfam = ifelse(is.na(subfam), 'ref', subfam)) %>% 
-  filter(!(time %in% c('sch.culture', 
-                       'Unknown'))) %>% 
-  mutate(subfam = str_replace(subfam, 
-                              'ancestral',
-                              'a')) %>% 
-  mutate(subfam =
-           forcats::fct_shift(
-             relevel(
-               as.factor(subfam),
-               ref = 'ref')
-           )
-  )
-
-#Need to add the transmission name to the subfam so that they are ordered correctly.
-# Need to move 'ref' factor to the start of the subfam, then shift the factor levels to the 'left' (so now
-# ref is at the end) ensuring that the 'ref' genes are at the end of the graph for each transmission type.
-
-cir_plot$Geneid <- ifelse(
-  cir_plot$Geneid %in% housekeeping_genes[1,],
-  colnames(housekeeping_genes)[ match(cir_plot$Geneid, housekeeping_genes[1,]) ],
-  cir_plot$Geneid
-)
-
-limits_cir <- c(0, 
-                mround(max(cir_plot$tpm), 
-                       base = 100))
+    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
+    chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
+                                                       cir_info$Geneid)],
+                        levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
+                                                                            cir_info$Geneid)])))) %>% 
+  filter(!(time %in% c('sch', 
+                       'Unknown')),
+         !is.na(chapl_aapl) ) %>% 
+  group_by(Geneid) %>% 
+    arrange(desc(tpm))
 
@@
                                      levels = unique(cir_plot_order$Geneid)))
+  cir_plot_chapl_aapl_supp_info <- cir_plot
 
-transmissionSep_cir_plot <- function(time_vec = '14h',
-                                     log_it = TRUE,
-                                     limits = limits_cir,
-                                     point_size = 1,
-                                     text_size = 10,
-                                     threshold_min = 100,
-                                     threshold_max = Inf){
+  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))
 
-  cir_plot_filter <- group_by(cir_plot, 
-                              Geneid) %>% 
-    filter(
-      time == time_vec,
-      !Geneid %in% cir_id | any(tpm >= threshold_min),
-      !Geneid %in% cir_id | all(tpm <= threshold_max)
-    )
-  #Filter for pirs expressed above or below the thresholds, have the OR so that the ref genes aren't removed.
-  # e.g. Either 'Not a cir' OR 'Any TPM is above the minimum'.
+  breaks = seq(0, max_aapl_chapl, by = 1)
+    limits = c(0, max_aapl_chapl)
+    point_size = 0.25
 
-  # cir_plot_filter %>% group_by(Geneid) %>% summarise(n())
-  if(log_it){
-    cir_plot_filter$tpm <- log2(1+cir_plot_filter$tpm)
-    limits = log2(1+limits)
-    breaks <- round(seq.int(0, 
-                            max(limits), 
-                            length.out = round(max(limits)) + 1, 
-                            digits = 0), 
-                    digits = 0)
-  }else{
-    breaks <- seq.int(0, max(limits), length.out = 11)
-  }
+  lapply(
+    sort(unique(cir_plot$time)),
+    function(time_vec){
 
-  cir_plot_filter$Geneid <- factor(
-    str_remove(cir_plot_filter$Geneid, 
+    cir_plot <- filter(cir_plot, 
+                       time == time_vec)
+
+      cir_plot$tpm <- log2(1+cir_plot$tpm)
+
+    cir_plot$Geneid <- factor(
+      str_remove(cir_plot$Geneid, 
                  pattern = 'PCHAS_'),
-    levels = str_remove(levels(cir_plot_filter$Geneid), 
+      levels = str_remove(levels(cir_plot$Geneid), 
                           pattern = 'PCHAS_')
@@
     y <- ggplot(
-    data = cir_plot_filter,
+      data = cir_plot,
       mapping = aes(
@@
                    geom = "bar", 
-                 fill = 'lightblue',
                    width = 0.75, 
                    size = 0.25, 
-                 position = 'dodge') +
+                   position = 'dodge',
+                   fill = 'lightblue') +
       geom_point(stat = 'identity', 
@@
                          limits = limits) +
-    facet_grid(.~subfam, 
+      scale_fill_manual(values = c('skyblue','red')) +
+      ylab('log(TPM +1)') +
+      facet_grid(.~chapl_aapl, 
                  scales = "free_x",
-               space = "free_x",
-               labeller = labeller(subfam = function(string) {
-                 str_remove_all(string, 
-                                pattern = '(RMT|SBP)_')
-               })) +
+                 space = "free_x") +
       theme_classic() +
-    ylab('log(TPM + 1)')+
-    ggtitle(paste(unique(cir_plot_filter$time)))
-
-  if(grepl(time_vec, pattern = '20|23')) {
-    y <- y + theme(
-      axis.text.x = element_text(size = text_size,
-                                 angle = 90,
-                                 vjust = 0.5)
-    )
-  } else {
-    y <- y + theme(
-      axis.text.x = element_blank(),
-      axis.ticks.x = element_blank()
-    ) 
-  }
-
-           #This bit is to change the colour, and text colour, of the facet boxes
-
-           g <- ggplot_gtable(ggplot_build(y))
-
-           # g$widths[5] = 2*g$widths[5]
-
-           strip_t <- which(grepl('strip-t', g$layout$name))
-           subfam_vec <- sort(unique(cir_plot_filter$subfam)) %>% str_remove_all(pattern = '(RMT|SBP)_')
-           fills <- case_when(str_detect(subfam_vec, 
-                                         pattern = 'a') ~ clade_col['ancestral'],
-                              subfam_vec %in% names(clade_col) ~ clade_col[paste(subfam_vec)],
-                              str_detect(subfam_vec, 
-                                         pattern = 'ref') ~ 'white')
-           k <- 1
-           for (i in strip_t) {
-             j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-             g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-             k <- k+1
-           }
-
-           #subfam_vec also corresponds to which panel the different clades are, so for the small ones we can use this to
-           # manually make them a bit wider.
-
-           short_panels <- which(str_detect(subfam_vec, pattern = 'a|S1'))
-
-           for(i in short_panels){
-             num <- g$layout$l[grep(paste0('panel-1-', i, '$'), g$layout$name)]
-             g$widths[num] = 1.75*g$widths[num]
-             #1.75 times the width of the right panels.
-             # print(num)
-           }
+      theme(
+        axis.text.x = element_text(angle = 90,
+                                   size = 5)
+      ) +
+      ggtitle(paste(unique(cir_plot$time)))
 
@@
       ggsave2(
-               paste0('Pchabaudi_side-by-side_barchart_facetbySubfam_',
+        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
                time_vec,
                '.png'),
-               plot = g,
+        plot = y,
         dpi = 300,
-               width = 28,
-               height = 8,
+        width = 25,
+        height = 10,
         units = 'cm'
@@
 
-           grid.newpage()
-           grid.draw(g)
-
-}
-
-lapply(
-  sort(unique(cir_plot$time)),
-  transmissionSep_cir_plot,
-  threshold_min = 50,
-  threshold_max = Inf,
-  point_size = 1,
-  text_size = 7.5
-)
+    return(y)
+    })
 ```

### 2021-04-21 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Refined Chapl heatmap and added clustered ht

@@
 combined_data_ordered <- tibble(Geneid = rownames(combined_data_pir),
-                                means = apply(combined_data_pir, 1, mean),
+                                means = apply(combined_data_pir[,
+                                                                str_detect(colnames(combined_data_pir), 
+                                                                           'chapl_aapl',
+                                                                           negate = TRUE)],
+                                              1,
+                                              mean),
                                 chapl_aapl_name = factor(
@@
 chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
-annotation_col <- c(chapl_aapl_name_col_num, 'ref' = 'black')
+annotation_col <- chapl_aapl_name_col_num[ordering(labels_chapl_aapl_name,
+                                                    order_vector = 'ChAPL')]
 
@@
            maxcol)
-  )
+  ),
+  row_gap = unit(c(rep.int(1, times = 2),
+                   3,
+                   rep.int(1, times = 4)), "mm")
 )
@@
   title = 'sub-family',
-  at = factor(labels_chapl_aapl_name,
+  at = str_subset(
+    sort(
+      factor(labels_chapl_aapl_name,
              ordering(labels_chapl_aapl_name,
-                order_vector = 'ChAPL')),
-  legend_gp = gpar(fill = annotation_col)
+                      order_vector = 'ChAPL'))
+    ),
+    pattern = 'NA',
+    negate = TRUE
+  ),
+  legend_gp = gpar(fill = na.omit(annotation_col))
 )
@@
 
-```{r Supp_fig_AapL-ChaAPl_expression}
-
-cir_plot <- twentyfour_rMT %>% 
-  filter(Geneid %in% cir_id) %>%
-  melt(id.vars = 'Geneid', 
-       variable.name = 'sample', 
-       value.name = 'tpm') %>% 
-  mutate(
-    stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = ''),
-    diff_count = str_split_fixed(sample, pattern = '_', n = 4)[,2],
-    time = str_split_fixed(sample, pattern = '_', n = 3)[,1],
-    subfam = cir_clade_info$clade[match(Geneid, cir_clade_info$Geneid)],
-    chapl_aapl = factor(cir_info$chapl_aapl_name[match(Geneid, 
-                                                       cir_info$Geneid)],
-                        levels = sort(unique(cir_info$chapl_aapl_name[match(Geneid, 
-                                                                            cir_info$Geneid)])))) %>% 
-  filter(!(time %in% c('sch', 
-                       'Unknown')),
-         !is.na(chapl_aapl) ) %>% 
-  group_by(Geneid) %>% 
-    arrange(desc(tpm))
-
-  cir_plot_order <- cir_plot %>% 
-    filter(time == '14h') %>% 
-    arrange(desc(tpm))
-
-  cir_plot <- mutate(cir_plot, 
-                     Geneid = factor(Geneid, 
-                                     levels = unique(cir_plot_order$Geneid)))
-  cir_plot_chapl_aapl_supp_info <- cir_plot
-
-  max_aapl_chapl <- ceiling(max(log2(cir_plot$tpm+1)))
-
-  breaks = seq(0, max_aapl_chapl, by = 1)
-    limits = c(0, max_aapl_chapl)
-    point_size = 0.25
-
-  lapply(
-    sort(unique(cir_plot$time)),
-    function(time_vec){
-
-    cir_plot <- filter(cir_plot, 
-                       time == time_vec)
-
-      cir_plot$tpm <- log2(1+cir_plot$tpm)
-
-    cir_plot$Geneid <- factor(
-      str_remove(cir_plot$Geneid, 
-                 pattern = 'PCHAS_'),
-      levels = str_remove(levels(cir_plot$Geneid), 
-                          pattern = 'PCHAS_')
+```{r}
+ht <- Heatmap(
+  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'chapl_aapl_name']),
+  cluster_rows = TRUE,
+  cluster_columns = FALSE,
+  show_row_names = FALSE,
+  name = 'log(TPM+1)', 
+  col = colorRamp2(c(0,
+                     maxcol/4,
+                     maxcol/2,
+                     3*maxcol/4,
+                     maxcol),
+                   viridis(5)),
+  na_col = 'white', 
+  border = TRUE,
+  border_gp = gpar(lwd = 0.1),
+  column_names_gp = gpar(fontsize = 10),
+  row_dend_width = unit(3, "cm"),
+  width = unit(4.5, 'cm'),
+  height = unit(7,'cm'),
+  row_title = NULL,
+  left_annotation = chapl_aapl_name_ha,
+  # row_split = 
+  #   factor(combined_data_pir_ht$chapl_aapl_name,
+  #             levels = ordering(combined_data_pir_ht$chapl_aapl_name,
+  #                               order_vector = 'ChAPL')),
+  row_km = 9,
+  heatmap_legend_param = list(
+    labels_gp = gpar(fontsize = 10),
+    title_gp = gpar(fontsize = 10,
+                    fontface = 'bold'),
+    at = c(0,
+           maxcol/4,
+           maxcol/2,
+           3*maxcol/4,
+           maxcol)
+  )
 )
 
-    y <- ggplot(
-      data = cir_plot,
-      mapping = aes(
-        x = Geneid,
-        y = tpm
+chapl_aapl_name_lgd <- Legend(
+  title = 'sub-family',
+  at = str_subset(
+    sort(
+      factor(labels_chapl_aapl_name,
+             ordering(labels_chapl_aapl_name,
+                      order_vector = 'ChAPL'))
+    ),
+    pattern = 'NA',
+    negate = TRUE
+  ),
+  legend_gp = gpar(fill = na.omit(annotation_col))
 )
-    ) +
-      stat_summary(fun.y = median, 
-                   fun.ymin = median, 
-                   fun.ymax = median,
-                   geom = "bar", 
-                   width = 0.75, 
-                   size = 0.25, 
-                   position = 'dodge',
-                   fill = 'lightblue') +
-      geom_point(stat = 'identity', 
-                 position = position_dodge(width = 0.75),
-                 size = point_size) +
-      scale_y_continuous(breaks = breaks,
-                         limits = limits) +
-      scale_fill_manual(values = c('skyblue','red')) +
-      ylab('log(TPM +1)') +
-      facet_grid(.~chapl_aapl, 
-                 scales = "free_x",
-                 space = "free_x") +
-      theme_classic() +
-      theme(
-        axis.text.x = element_text(angle = 90,
-                                   size = 5)
-      ) +
-      ggtitle(paste(unique(cir_plot$time)))
 
-    if(global_save_var){
-      ggsave2(
-        paste0('Pchabaudi_sibe-by-side_chapl_aapl_',
-               time_vec,
-               '.png'),
-        plot = y,
-        dpi = 300,
-        width = 25,
-        height = 10,
-        units = 'cm'
-      )
-    }
-
-    return(y)
-    })
+draw(ht, 
+     row_title_side = 'left',
+     column_title_side = 'bottom',
+     heatmap_legend_side = 'right',
+     annotation_legend_side = 'right',
+     annotation_legend_list = list(chapl_aapl_name_lgd),
+     merge_legend = TRUE)
 ```

### 2021-04-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Save heatmaps and use labels

@@
 
-#Add chapl_aapl numbers for easier heatmap drawing later.
+#Add chapl_aapl/subfam numbers for easier heatmap drawing later.
 cir_info <- cir_info %>% 
@@
   mutate(chapl_aapl_name_num = n_distinct(Geneid)) %>% 
-  mutate(chapl_aapl_name_num = paste0(
+  mutate(
+    chapl_aapl_name_num = paste0(
       chapl_aapl_name, 
@@
       ')'
-                     ))
+    )) %>% 
+  group_by(subfam) %>% 
+  mutate(subfam_num = n_distinct(Geneid)) %>% 
+  mutate(
+    subfam_num = paste0(
+      subfam, 
+      ' (n = ',
+      subfam_num,
+      ')'
+    )
+  )
 ```
@@
   filter_all(
-      any_vars(. > 1)
+      any_vars(. > 10)
   )
@@
   height = unit(7,'cm'),
-  row_title = NULL,
-  left_annotation = clade_ha,
-  right_annotation = anc_ha,
+  # row_title = NULL,
+  # left_annotation = clade_ha,
+  # right_annotation = anc_ha,
+  row_title_rot = 0,
+  row_title_gp = gpar(fontsize = 10),
   row_split = forcats::fct_shift(
@@
      column_title_side = 'bottom',
+     # heatmap_legend_side = 'right',
+     # annotation_legend_side = 'right',
+     # annotation_legend_list = list(clade_lgd),
+     merge_legend = TRUE)
+
+if(global_save_var){
+  pdf(file = 'Manuscript_Fig_subfam_heatmap.pdf',
+      height = 5,
+      width = 5)
+draw(ht, 
+     row_title_side = 'left',
+     column_title_side = 'bottom',
      heatmap_legend_side = 'right',
-     annotation_legend_side = 'right',
-     annotation_legend_list = list(clade_lgd),
+     # annotation_legend_side = 'right',
+     # annotation_legend_list = list(clade_lgd),
      merge_legend = TRUE)
+  dev.off()
+}
 ```
@@
 
-chapl_aapl_name_ha <- rowAnnotation(
-  'sub-family' = anno_simple(
-    chapl_aapl_name_info,
-    col = annotation_col,
-  ),
-  annotation_name_gp = gpar(fontsize = 10)
-)
+# chapl_aapl_name_ha <- rowAnnotation(
+#   'CHAPL/AAPL' = anno_simple(
+#     chapl_aapl_name_info,
+#     col = annotation_col,
+#   ),
+#   annotation_name_gp = gpar(fontsize = 10)
+# )
 
@@
   height = unit(7,'cm'),
-  row_title = NULL,
-  left_annotation = chapl_aapl_name_ha,
+  # row_title = NULL,
+  row_title_rot = 0,
+  row_title_gp = gpar(fontsize = 10),
   row_split = 
@@
 
+# chapl_aapl_name_lgd <- Legend(
+#   title = 'CHAPL/AAPL',
+#   at = str_subset(
+#     sort(
+#       factor(labels_chapl_aapl_name,
+#              ordering(labels_chapl_aapl_name,
+#                       order_vector = 'ChAPL'))
+#     ),
+#     pattern = 'NA',
+#     negate = TRUE
+#   ),
+#   legend_gp = gpar(fill = na.omit(annotation_col))
+# )
+
+if(global_save_var){
+  pdf(file = 'Manuscript_SuppFig_ChAPL-AAPL_heatmap.pdf',
+      height = 5,
+      width = 5)
+  draw(ht, 
+       row_title_side = 'left',
+       column_title_side = 'bottom',
+       # heatmap_legend_side = 'right',
+       # annotation_legend_side = 'right',
+       # annotation_legend_list = list(chapl_aapl_name_lgd),
+       # merge_legend = TRUE
+  )
+  dev.off()
+}
+```
+
+```{r}
+chapl_aapl_name_ha <- rowAnnotation(
+  'CHAPL/AAPL' = anno_simple(
+    chapl_aapl_name_info,
+    col = annotation_col,
+  ),
+  annotation_name_gp = gpar(fontsize = 10)
+)
+
 chapl_aapl_name_lgd <- Legend(
-  title = 'sub-family',
+  title = 'CHAPL/AAPL',
   at = str_subset(
@@
 
-draw(ht, 
-     row_title_side = 'left',
-     column_title_side = 'bottom',
-     heatmap_legend_side = 'right',
-     annotation_legend_side = 'right',
-     annotation_legend_list = list(chapl_aapl_name_lgd),
-     merge_legend = TRUE)
-```
-
-```{r}
 ht <- Heatmap(
@@
 
-chapl_aapl_name_lgd <- Legend(
-  title = 'sub-family',
-  at = str_subset(
-    sort(
-      factor(labels_chapl_aapl_name,
-             ordering(labels_chapl_aapl_name,
-                      order_vector = 'ChAPL'))
-    ),
-    pattern = 'NA',
-    negate = TRUE
-  ),
-  legend_gp = gpar(fill = na.omit(annotation_col))
-)
-
 draw(ht, 
@@
      merge_legend = TRUE)
+
+if(global_save_var){
+  pdf(file = 'Interest_chapl-aapl_clustered_heatmap.pdf',
+      height = 5,
+      width = 5)
+draw(ht, 
+     row_title_side = 'left',
+     column_title_side = 'bottom',
+     heatmap_legend_side = 'right',
+     annotation_legend_side = 'right',
+     annotation_legend_list = list(chapl_aapl_name_lgd),
+     merge_legend = TRUE)
+  dev.off()
+}
 ```

### 2021-04-22 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
02h time point was missinggit status

@@
 
-combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
-                                                   order_vector)),
-                                       drop = FALSE]
-
 combined_data_ordered <- tibble(Geneid = names(apply(combined_data_pir, 1, mean)),
@@
 combined_data_pir_ht$subfam <- ifelse(rownames(combined_data_pir_ht) %in% cir_id,
-                                      cir_info$subfam[match(rownames(combined_data_pir_ht), 
+                                      cir_info$subfam_num[match(rownames(combined_data_pir_ht), 
                                                             cir_info$Geneid)],
@@
 
-anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
-                                        labels = 'ancestral',
-                                        labels_gp = gpar(fontsize = 7),
-                                        link_width = unit(3, "mm")))
+# anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
+#                                         labels = 'ancestral',
+#                                         labels_gp = gpar(fontsize = 7),
+#                                         link_width = unit(3, "mm")))
 
@@
   )
-
-#Ordering that a more understandable order
-
-combined_data_pir <- combined_data_pir[,c(ordering(colnames(combined_data_pir)[-1], 
-                                                   order_vector)),
-                                       drop = FALSE]
 
@@
   row_title_rot = 0,
-  row_title_gp = gpar(fontsize = 10),
+  row_title_gp = gpar(fontsize = 7),
   row_split = 

### 2021-04-26 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir.Rmd
New chab heatmaps file
Heatmap separated by sub-family
Added ancestral bar chart code to see if it should be included

@@
 
-```{r highest_expressed_stage_heatmap_fig3b}
+```{r highest_expressed_stage_heatmap_fig3c}
 
@@
 
+# Bar chart of ancestral _pir_ expression.
+
+```{r}
+  cir_plot <- twentyfour_rMT %>% 
+    filter(Geneid %in% 'PCHAS_0101200') %>% 
+    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
+    # mutate(stage = gsub(sample, pattern = '_[[:digit:]]$', replacement = '')) %>%
+    mutate(time = str_extract(sample, pattern = '[[:digit:]]{2}h')) %>% 
+    group_by(sample, time) %>% 
+    dplyr::summarise(sum_tpm = sum(tpm))
+
+y_scale <- seq(0, 
+               signif(ceiling(max(cir_plot$sum_tpm)), 
+                      digits = 2), 
+               length.out = 5)
+
+(
+  p <- ggplot(data = cir_plot, 
+              aes(x = time, 
+                  y = sum_tpm)) +
+    stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
+                 geom = "bar", width = 0.75, size = 0.25, position = 'dodge',
+                 fill = 'lightblue') +
+    scale_y_continuous(breaks = y_scale) +
+    geom_point(stat = 'identity', position = position_dodge(width = 0.75)) +
+    ylab(as.expression(bquote('Total'~italic(.('PCHAS_0101200'))~'TPM'))) +
+    xlab('Asexual blood stage development cycle') +
+    theme_classic() +
+    theme(axis.text.x=element_text(angle = 0, 
+                                   vjust = 0.5,
+                                   size = 13))
+)
+
+if(global_save_var){
+  ggsave2(filename = 'Manuscript_ancestralPir_BarChart.pdf',
+          plot = p,
+          dpi = 300,
+          width = 15,
+          height = 10,
+          units = 'cm')
+}
+
+```
+
 # Calculate and cluster with z-scores {.tabset}

### 2021-04-26 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Added ancestral bar chart code to see if it should be included

@@
 
+Add chapl/aapl annotation to above
+
+```{r}
+
+chapl_aapl_name_vec <- ifelse(grepl(cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), 
+                                                                                     cir_info$Geneid)], 
+                                                  pattern = 'NA'),
+                                            NA,
+                                            cir_info$chapl_aapl_name_num[match(rownames(combined_data_pir_ht), cir_info$Geneid)])
+
+#chapl_aapl_name annotation
+
+annotation_col <- chapl_aapl_name_col_num[ordering(chapl_aapl_name_vec,
+                                                   order_vector = 'ChAPL')]
+
+chapl_aapl_name_ha <- rowAnnotation(
+  'CHAPL/AAPL' = anno_simple(
+    chapl_aapl_name_vec,
+    col = annotation_col,
+    na_col = 'white'
+  ),
+  annotation_name_gp = gpar(fontsize = 10)
+)
+
+chapl_aapl_name_lgd <- Legend(
+  title = 'CHAPL/AAPL',
+  at = names(annotation_col)
+    # str_subset(
+    # sort(
+    #   factor(chapl_aapl_name_vec,
+    #          ordering(chapl_aapl_name_vec,
+    #                   order_vector = 'ChAPL'))
+    # )
+    # ,
+    # pattern = 'NA',
+    # negate = TRUE
+  # )
+,
+  legend_gp = gpar(fill = na.omit(annotation_col))
+)
+
+ht <- Heatmap(
+  log2(1+combined_data_pir_ht[,!colnames(combined_data_pir_ht) %in% 'subfam']),
+  cluster_rows = FALSE,
+  cluster_columns = FALSE,
+  show_row_names = FALSE,
+  name = 'log(TPM+1)', 
+  col = colorRamp2(c(0,
+                     maxcol/4,
+                     maxcol/2,
+                     3*maxcol/4,
+                     maxcol),
+                   viridis(5)),
+  na_col = 'white', 
+  border = TRUE,
+  border_gp = gpar(lwd = 0.1),
+  column_names_gp = gpar(fontsize = 10),
+  row_dend_width = unit(3, "cm"),
+  width = unit(4.5, 'cm'),
+  height = unit(7,'cm'),
+  # row_title = NULL,
+  left_annotation = chapl_aapl_name_ha,
+  row_title_rot = 0,
+  row_title_gp = gpar(fontsize = 10),
+  row_split = forcats::fct_shift(
+    relevel(
+      as.factor(combined_data_pir_ht$subfam),
+      ref = 'ref')
+  ),
+  heatmap_legend_param = list(
+    labels_gp = gpar(fontsize = 10),
+    title_gp = gpar(fontsize = 10,
+                    fontface = 'bold'),
+    at = c(0,
+           maxcol/4,
+           maxcol/2,
+           3*maxcol/4,
+           maxcol)
+  )
+)
+
+draw(ht, 
+     row_title_side = 'left',
+     column_title_side = 'bottom',
+     heatmap_legend_side = 'right',
+     annotation_legend_side = 'right',
+     annotation_legend_list = list(chapl_aapl_name_lgd),
+     merge_legend = TRUE)
+
+if(global_save_var){
+  pdf(file = 'Interest_SubfamSplit_ChAPL-AAPLAnnotate_heatmap.pdf',
+      height = 5,
+      width = 5)
+draw(ht, 
+     row_title_side = 'left',
+     column_title_side = 'bottom',
+     heatmap_legend_side = 'right',
+     annotation_legend_side = 'right',
+     annotation_legend_list = list(chapl_aapl_name_lgd),
+     merge_legend = TRUE)
+  dev.off()
+}
+
+```
+
 # Supp figure 9 - Pir expression of AAPLs and ChAPLs.

### 2021-04-29 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_berghei_pir_DESEQ.Rmd
Saving supp info for deseq stats berghei

@@
 library('readxl')
+library('writexl')
 library('data.table')
@@
 ) %>% 
-  mutate(contrast = coef)
+  mutate(contrast = coef) %>% 
+  rownames_to_column(var = 'Geneid')
 ```
@@
                                                                                               'PBANKA_0600031'),]) %>%  
-  mutate(contrast = coef)
+  mutate(contrast = coef) %>% 
+  rownames_to_column(var = 'Geneid')
 ```
@@
 ) %>% 
-  mutate(contrast = coef)
+  mutate(contrast = coef) %>% 
+    rownames_to_column(var = 'Geneid')
 ```
@@
   ) %>%  
-  mutate(contrast = coef))
+  mutate(contrast = coef) %>% 
+    rownames_to_column(var = 'Geneid'))
 ```
@@
 ```{r}
-bind_rows(livMeroz_v_gamMale_df, 
+(deseq_stats <- bind_rows(livMeroz_v_gamMale_df, 
           as.data.frame(gamMale_v_asexMixed_df),
@@
             padj, 
-            pvalue))
+            pvalue)))
 ```
+
+```{r save_stats_df}
+write_xlsx(deseq_stats, 
+           'SupplementaryInformation_Stats_DESEQ_berghei.xlsx')
+```

### 2021-04-29 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
Save Qusage results

@@
 library('RColorBrewer')
+library('writexl')
 ```
@@
 pir_geneset <- list(pirs = pir_info$Geneid)
-
-chrom_list <- lapply(unique(pir_info$chrom), 
-                     function(chrom_nam) filter(pir_info, 
-                                                chrom == chrom_nam)$Geneid)
-
-names(chrom_list) <- paste0('chr', 
-                            unique(pir_info$chrom))
-
-#Remove groups with only three members or fewer.
-chrom_list <- chrom_list[unlist(lapply(chrom_list, function(x) length(x) > 3))]
 
@@
     subfam_list, 
-    ShortLong_list, 
-    chrom_list), 
+    ShortLong_list), 
   function(gene_list){
@@
 )
-names(maleGam_v_Asex.Mix.results) <- c('all', 'subfam', 'SorL', 'chrom')
+names(maleGam_v_Asex.Mix.results) <- c('all', 'subfam', 'SorL')
 
@@
 mutate(maleGam_v_Asex.Mix.results$subfam$qstable , adjust = p.adjust(p.Value, method = 'fdr'))
-plotGeneSetDistributions(maleGam_v_Asex.Mix.results$chrom$results, path.index = 'chr14')
 ```
@@
 )
-names(Asex.Mix_v_Liv.Meroz.results) <- c('all', 'subfam', 'SorL', 'chrom')
+names(Asex.Mix_v_Liv.Meroz.results) <- c('all', 'subfam', 'SorL')
 
@@
 
-```{r}
-berghei_comparison_stats_df
+```{r save_xlsx}
+ bind_rows(
+  lapply(c(maleGam_v_Asex.Mix.results, 
+           maleGam_v_Liv.Meroz.results, 
+           Asex.Mix_v_Liv.Meroz.results,
+           maleVfemale, 
+           post48livvearlyliv, 
+           livmero_livpost48.results),
+         getElement,
+         'qstable')
+) %>% 
+  select(contrast, everything()) %>% 
+  mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
+  arrange(contrast, pathway.name) %>% 
+  write_xlsx(path = 'SupplementaryInformation_Stats_Qusage_berghei.xlsx')
 ```

### 2021-05-13 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Saving png files as PDFs
Fixing figure 2A bar chart order of stacks

@@
 
+pdf(file = 'Manuscript_suppfig1A_corrplot_all_experiments_rnaseq.pdf',
+    height = 5,
+    width = 5)
+corrplot(cor(big_top_recast), method = 'color', 
+         order = 'hclust', hclust.method = 'ward.D2',
+         cl.lim = c(-0.036,1), tl.cex = 0.50,
+         col = c(viridis(100), magma(115)), addrect = 6)
+dev.off()
+
 corrplot(cor(big_top_recast), method = 'color', 
@@
               fill = SorL,
-              group = stage_num
+              # group = stage_num
             )) +
@@
     )
-  )+
+  ) +
   scale_color_manual('black') + 
-  labs(fill = stack_label) +
+  # labs(fill = stack_label) +
   scale_y_continuous(breaks = max_tpm) +
@@
               fill = SorL,
-              group = stack_group_order
+              # group = stage_num
             )) +
@@
   scale_fill_manual(
-    breaks = unique(tpm_data_SorL_stackgenes_low$SorL),
-    values = SorL_col_num
+    breaks = unique(tpm_data_SorL_stackgenes$SorL),
+    values = c(
+      SorL_col_num,
+      fills
+    )
   ) +

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Updated .gitignore and berghei

@@
     } else {
-      print('Warning:Count data and length gene ids DO NOT MATCH')
+      stop('Warning:Count data and length gene ids DO NOT MATCH')
     }
   } else {
-    print('Count data and/or length table supplied do not have an id column, make sure they are ordered correctly')
+    stop('Count data and/or length table supplied do not have an id column, make sure they are ordered correctly')
   }
@@
 names(expt_tpms) <- expt_nams
-
+```
+
+```{r remove_E06-6h}
 #Remove the Caldelari (E06) 6h samples, as they did in their paper.
@@
 expt_tpms$E03$tpm$`Asex-Schiz_3` <- NULL
+```
+
+## Saving the TPM
+
+```{r save_tpm}
+
 ```

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
r functions moved to new folder

@@
 
-#Want the defaults for arrange, mutate and summarise to be from dplyr
+#Want the defaults for arrange, select, mutate and summarise to be from dplyr
 arrange <- dplyr::arrange
@@
 
+```{r sourcing_functions}
+source('../rfunctions/uniquify_fun.R')
+source('../rfunctions/ordering_fun.R')
+source('../rfunctions/tpm_fun.R')
+```
+
 ```{r function_mround}
@@
 my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) 
-```
-
-```{r function_uniquify}
-#Make unique identifiers from a vector of names by appending '_X' where X is the number of the 'replicate'.
-uniquify <- function(vector_names, first_one = TRUE){
-  new_vector <- c()
-  df_num <- data.frame(names = unique(vector_names), rep = 1)
-  if(first_one){
-    duplicated_names <- unique(vector_names[duplicated(vector_names)])
-  }else{
-    duplicated_names <- c()
-  }
-  for(i in 1:length(vector_names)){
-    if(vector_names[i] %in% duplicated_names|!first_one){
-    name <- paste(vector_names[i],
-                  df_num$rep[match(vector_names[i], df_num$names)],
-                  sep = '_')
-    #Increase rep counter
-    df_num$rep[match(vector_names[i], df_num$names)] <- df_num$rep[match(vector_names[i], df_num$names)] + 1
-    new_vector[i] <- name
-    }else{
-      new_vector[i] <- vector_names[i]
-    }
-  }
-  new_vector
-}
 ```
@@
 
-```{r function_ordering}
-#Use an order_vector of partial string matches to order.
-ordering = function(to_order, order_vector){
-  order = unlist(sapply(order_vector, 
-                        function(x){unique(str_subset(to_order, 
-                                                      pattern = paste(x)))}))
-  order_missing = unique(to_order[!to_order %in% order])
-  unique(c(order, order_missing), fromLast = TRUE)
-}
-
+```{r ordering_vectors}
 order_vector <- c('Sporo', 'Liv-pre4h', 'Liv-pre47h', 'Liv-post48h',
@@
                  'Mero','Asex','Schiz','Gam','-Pre3h','BldMl-24h','Ook')
+
 r_order_vector <- str_replace_all(order_vector, 
@@
                                   replacement = '\\.')
-```
-
-```{r tpm, results='hide'}
-
-# We need to write a function to calculate the TPM from the count data we have generated.
-This function uses the _lengths.transcript_ object and should generate errors if it detects discrepancies between the Geneids so that the wrong transcript length is not used.
-
-tpm = function(count_data,lengths) {
-
-  #Print an error if the id columns do not match
-  if(any(grepl(colnames(lengths), 
-               pattern = 'id', 
-               ignore.case = T)) == T & any(grepl(colnames(count_data), 
-                                                  pattern = 'id', 
-                                                  ignore.case = T)) == T){
-    if(all(count_data[grep(colnames(count_data), 
-                           pattern = 'id', ignore.case = T)] == lengths[grep(colnames(lengths), 
-                                                                             pattern = 'id', ignore.case = T)])) {
-      print('Count data and length gene ids both match')
-    } else {
-      stop('Warning:Count data and length gene ids DO NOT MATCH')
-    }
-  } else {
-    stop('Count data and/or length table supplied do not have an id column, make sure they are ordered correctly')
-  }
-
-  #Remove the id column from supplied count_data
-  if(any(grepl(colnames(count_data), pattern = 'id',ignore.case = T)) == T){
-    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'id', ignore.case = T)]
-  } else{
-    counts_tpm <- count_data}
-
-  #Remove the id column from supplied lengths
-  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
-    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
-  }else{
-    lengths.table <- lengths}
-
-  #Now for the actual calculation
-  rpm <- counts_tpm
-  for (i in 1:dim(counts_tpm)[2]) {
-    for (j in 1:dim(counts_tpm)[1]) {
-      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j,]/1000))
-    }
-    scale.factor <- sum(rpm[,i])/(1E6)
-    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
-  }
-  if(any(grepl(colnames(count_data), pattern = 'id', ignore.case = T)) == T){
-    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'id', ignore.case = T), drop = FALSE], 
-                           counts_tpm)
-  } 
-  return(counts_tpm)
-}
 ```

### 2021-08-31 Manuscript_pirAnalysis_Github/rfunctions/tpm_fun.R

@@ -1,0 +1,49 @@
+
+# We need to write a function to calculate the TPM from the count data we have generated. This function uses the _lengths.transcript_ object and should generate errors if it detects discrepancies between the Geneids so that the wrong transcript length is not used.
+
+tpm = function(count_data,lengths) {
+  
+  #Print an error if the id columns do not match
+  if(any(grepl(colnames(lengths), 
+               pattern = 'id', 
+               ignore.case = T)) == T & any(grepl(colnames(count_data), 
+                                                  pattern = 'id', 
+                                                  ignore.case = T)) == T){
+    if(all(count_data[grep(colnames(count_data), 
+                           pattern = 'id', ignore.case = T)] == lengths[grep(colnames(lengths), 
+                                                                             pattern = 'id', ignore.case = T)])) {
+      print('Count data and length gene ids both match')
+    } else {
+      stop('Warning:Count data and length gene ids DO NOT MATCH')
+    }
+  } else {
+    stop('Count data and/or length table supplied do not have an id column, make sure they are ordered correctly')
+  }
+  
+  #Remove the id column from supplied count_data
+  if(any(grepl(colnames(count_data), pattern = 'id',ignore.case = T)) == T){
+    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'id', ignore.case = T)]
+  } else{
+    counts_tpm <- count_data}
+  
+  #Remove the id column from supplied lengths
+  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
+    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
+  }else{
+    lengths.table <- lengths}
+  
+  #Now for the actual calculation
+  rpm <- counts_tpm
+  for (i in 1:dim(counts_tpm)[2]) {
+    for (j in 1:dim(counts_tpm)[1]) {
+      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j,]/1000))
+    }
+    scale.factor <- sum(rpm[,i])/(1E6)
+    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
+  }
+  if(any(grepl(colnames(count_data), pattern = 'id', ignore.case = T)) == T){
+    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'id', ignore.case = T), drop = FALSE], 
+                           counts_tpm)
+  } 
+  return(counts_tpm)
+}

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Change from counts.xlsx to .csv

@@
 author: "Timothy Little"
-date: "12/03/2021"
+date: "31/08/2021"
 output: 
@@
 knitr::opts_chunk$set(echo = FALSE, results = TRUE, message = FALSE, warning = FALSE)
-# Set below to your own working directory.
-knitr::opts_knit$set(root.dir = 
-                       ".")
 options(scipen = 999) # Means that numbers are displayed normally not as 1e6 (1000000)
@@
 
-```{r read_files}
+```{r read_files_calc_TPM}
 #Reading in the counts from file
-berghei_counts <- read_xlsx('Manuscript_Frontiers_berghei_counts.xlsx')
+berghei_counts <- read.csv('tl20-05_berghei_counts.csv',
+                               row.names = 1)
 
@@
 apply(berghei_tpm[-1], 2, sum)
-
+```
+
+```{r save_tpm_oldNames_allSamples}
+#Save the 'raw' TPM before any samples have been removed or sample names combined.
+
+write.csv()
+
+```
+
+```{r expt_tpms}
 #Get a list of df for each experiment.

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Save TPM (all samples) file

@@
 
-write.csv()
+write.csv(berghei_tpm,
+          'berghei_tpm_oldNamesAllSamples.csv')
 
@@
 
-hclusts = function(tpm){
+lapply(expt_tpms, function(tpm){
   if(ncol(tpm$tpm) > 3){
@@
   }
-}
-
-lapply(expt_tpms, hclusts)
+})
 ```

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Do name changes and cluster analysis on big TPM df

@@
 
+```{r remove_E06-6h}
+#Remove the Caldelari (E06) 6h samples, as they did in their paper.
+berghei_tpm[,str_subset(colnames(berghei_tpm), 'E06_Liv.6h')] <- NULL
+```
+
+```{r hclusts, message = FALSE}
+experiment_id <- unique(na.omit(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}')))
+
+lapply(experiment_id, 
+       function(experiment_name){
+         tpm_data <- select(berghei_tpm, matches(experiment_name))
+         if(ncol(tpm_data) > 3){
+           dists = dist(t(tpm_data))
+           par(cex=0.7)
+           plot(hclust(dists),
+                main = experiment_name)
+         }
+       })
+
+```
+
+```{r combine_similar_samples_byNames}
+col_stage <- colnames(berghei_tpm)
+
+#Combine all Liver samples 4h or before into LiverPre4h
+col_stage <- str_replace_all(col_stage,
+                             pattern = 'Liv\\.(2|4)h_[[:digit:]]+',
+                             replacement = 'Liv.pre4h')
+#Combine all Liver samples 48h or onwards into LiverPost48h
+col_stage <- str_replace_all(col_stage,
+                             pattern = 'Liv\\.(4[8-9]|[5-9][0-9])h_[[:digit:]]+',
+                             replacement = 'Liv.post48h')
+#Combine all blood meal samples 3h or before into BldMlPre3h
+col_stage <- str_replace_all(col_stage,
+                             pattern = '(1h|3h)_[[:digit:]]+', 
+                             replacement = 'Pre3h')
+#Uniquify the name names
+col_stage[grep(col_stage, 
+               perl = TRUE,
+               invert = TRUE,
+               pattern = '_[[:digit:]]{1,}$')] <- uniquify(col_stage[grep(col_stage, 
+                                                                          perl = TRUE,
+                                                                          invert = TRUE,
+                                                                          pattern = '_[[:digit:]]{1,}$')]) 
+```
+
 ```{r expt_tpms}
@@
 
-```{r remove_E06-6h}
-#Remove the Caldelari (E06) 6h samples, as they did in their paper.
-
-expt_tpms$E06$tpm <- select_at(expt_tpms$E06$tpm,
-                                          vars(-matches('Liv-6h')))
-```
-
 Going to combine certain samples based on their distinctive profiles, e.g. <4h, 6-48h, >48h liver stages.
@@
 
-```{r hclusts, message = FALSE}
-
-lapply(expt_tpms, function(tpm){
-  if(ncol(tpm$tpm) > 3){
-    dists = dist(t(tpm$tpm[,-1]))
-    par(cex=0.7)
-    plot(hclust(dists),
-         main = tpm$experiment)
-  }
-})
-```
-
 ```{r remove_badly_clustering_samples}
@@
 
+```{r big_tpm}
+
+```
+
 ## Saving the TPM
@@
 
+write.csv()
 ```
@@
 ```{r all_samples_collated_heatmap}
-
-big_tpm <- bind_cols(
-  lapply(
-    expt_tpms,
-    function(x){
-      tpm_df <- x$tpm
-      colnames(tpm_df)[-1] <- paste0(x$experiment,
-                                     '_',
-                                     colnames(tpm_df)[-1])
-      tpm_df
-    }
-  )
-) %>% 
-  dplyr::rename(Geneid = Geneid...1) %>% 
-  select_at(vars(-matches('Geneid\\.\\.\\.\\d+$'))) %>% 
-  data.frame
-
-rownames(big_tpm) <- big_tpm$Geneid
-
-big_tpm <- big_tpm[,-1,drop=FALSE]
 

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_CalculateTPM.Rmd
Separate TPM calculation Rmd file

--- 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
@@
 ---
-title: "P.berghei pir analysis"
+title: "P.berghei pir calculate TPM"
 author: "Timothy Little"
@@
 
+Going to combine certain samples based on their similar profiles, e.g. <4h, >48h liver stages.
+
 ```{r combine_similar_samples_byNames}
@@
 
-```{r expt_tpms}
-#Get a list of df for each experiment.
-
-expt_nams <- unique(
-  na.omit(
-    str_extract(colnames(berghei_tpm), 
-                pattern = 'E[[:digit:]]{2}')
-  )
-)
-
-expt_tpms <- lapply(
-  expt_nams,
-  function(expt_nam){
-    expt_df <- select(
-      berghei_tpm,
-      matches(paste0('Geneid|', expt_nam))
-    ) %>% 
-      rename_with(~ str_remove(.x, 'E[[:digit:]]{2}_'),
-                  matches('E[[:digit:]]{2}'))
-    list(tpm = expt_df,
-         experiment = expt_nam)
-  }
-)
-names(expt_tpms) <- expt_nams
-```
-
-Going to combine certain samples based on their distinctive profiles, e.g. <4h, 6-48h, >48h liver stages.
-
-```{r combining_samples}
-names_expt_tpms <- names(expt_tpms)
-
-expt_tpms <- lapply(names(expt_tpms), function(expt){
-  col_stage <- colnames(
-    getElement(
-      getElement(
-        expt_tpms, 
-        paste(expt)
-      ),
-      'tpm'
-    )
-  )
-
-  liver_stage <- col_stage[
-    str_detect(
-      col_stage,
-      pattern = 'Liv-[[:digit:]]{1,2}h'
-    )
-    ]
-
-  new_liver_stage <- case_when(
-    str_detect(liver_stage, pattern = 'Liv-(2|4)h') ~ 'Liv-pre4h',
-    # str_detect(liver_stage, pattern = 'Liv-([5-9]|[1-3][0-9]|4[0-7])h') ~ 'Liv-pre47h',
-    str_detect(liver_stage, pattern = 'Liv-(4[8-9]|[5-9][0-9])h') ~ 'Liv-post48h',
-    TRUE ~ as.character(liver_stage),
-  )
-
-  if(length(new_liver_stage) > 0){
-    new_liver_stage <-uniquify(new_liver_stage)
-  }
-
-  col_stage[
-    str_detect(
-      col_stage,
-      pattern = 'Liv-[[:digit:]]{1,2}h'
-    )
-    ] <- new_liver_stage
-
-  bloodmeal_stage <- col_stage[
-    str_detect(
-      col_stage, 
-      pattern = 'BldMl-(1h|3h)_[[:digit:]]+'
-    )
-    ]
-
-  bloodmeal_stage <- str_replace(
-    bloodmeal_stage, 
-    pattern = '(1h|3h)_[[:digit:]]+', 
-    replacement = 'Pre3h'
-  )
-  if(length(bloodmeal_stage)>0){
-    bloodmeal_stage <- uniquify(bloodmeal_stage)
-  }
-
-  col_stage[
-    str_detect(
-      col_stage, 
-      pattern = 'BldMl-(1h|3h)_[[:digit:]]+'
-    )
-    ] <- bloodmeal_stage
-
-  tpm_element <- getElement(
-    getElement(
-      expt_tpms, 
-      paste(expt)
-    ),
-    'tpm'
-  )
-
-  colnames(tpm_element) <- col_stage
-
-  list(tpm = tpm_element,
-       experiment =   getElement(
-         getElement(
-           expt_tpms,
-           paste(expt)
-         ),
-         'experiment'
-       )
-  )
-
-})
-names(expt_tpms) <- names_expt_tpms
-
-#Re-ordering the list
-expt_tpms <- (expt_tpms[order(names(expt_tpms))])
-```
-
 ```{r remove_badly_clustering_samples}
-expt_tpms$E08$tpm$`Liv-pre4h_1` <- NULL
-expt_tpms$E03$tpm$`Asex-Schiz_1` <- NULL
-expt_tpms$E03$tpm$`Asex-Schiz_3` <- NULL
-```
-
-```{r big_tpm}
-
+berghei_tpm$`E08_Liv.pre4h_1` <- NULL
+berghei_tpm$`E03_Asex.Schiz_1` <- NULL
+berghei_tpm$`E03_Asex.Schiz_3` <- NULL
 ```
@@
 ```{r save_tpm}
-
-write.csv()
+write.csv(berghei_tpm,
+          'berghei_tpm.csv')
 ```
-
-## The highest expressed _pir_ genes across the samples
-
-```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
-top_n_genes <- function(tpm_input, 
-                       n = 5, 
-                       threshold = 0){
-  tpm = tpm_input$tpm %>% 
-    #Melt the data frame for using dplyr
-    melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-    mutate(stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1))) %>% 
-    group_by(Geneid, stage) %>%  
-    dplyr::summarise(median_tpm = median(tpm),
-                     mean_tpm = mean(tpm), 
-                     min_tpm = min(tpm), 
-                     max_tpm = max(tpm), 
-                     sd_tpm = sd(tpm)) %>% 
-    group_by(stage) %>% 
-    mutate(experiment = tpm_input$experiment)
-
-  #Calculating the index used to make the kable table
-
-  kable_index = rep(n, times = length(unique(tpm$stage)))
-  names(kable_index) = unique(tpm$stage)
-
-  all_genes = tpm %>% 
-    top_n(wt = mean_tpm, n = n) %>% 
-    dplyr::arrange(stage, desc(mean_tpm)) %>% 
-    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
-    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
-                                color = ifelse(mean_tpm > threshold, "red", "blue")))
-  all_genes$short_descr = allgene.info$handy_descriptions[match(all_genes$Geneid, 
-                                                                allgene.info$Geneid)]
-  all_genes$description = allgene.info$Description[match(all_genes$Geneid, 
-                                                         allgene.info$Geneid)]
-  all_genes = all_genes %>% 
-    kable(format = 'html', 
-          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed total genes'), 
-          escape = FALSE) %>%
-    kable_styling() %>% 
-    pack_rows(index = kable_index)
-
-  pir_genes = tpm %>% 
-    filter(Geneid %in% pir_id) %>% 
-    top_n(wt = mean_tpm, n = n) %>% 
-    dplyr::arrange(stage, desc(mean_tpm)) %>% 
-    mutate_if(is.numeric, function(x){round(x, digits = 2)}) %>% 
-    mutate(mean_tpm = cell_spec(mean_tpm, "html", 
-                                color = ifelse(mean_tpm > threshold, "red", "blue"))) %>% 
-    mutate(Geneid = as.character(Geneid)) %>% 
-    mutate(Description = allgene.info$handy_descriptions[match(Geneid, 
-                                                               allgene.info$Geneid)]) %>% 
-    mutate(Geneid = ifelse(grepl(Geneid,
-                                 pattern = 'PBANKA_0100500'),
-                           cell_spec(Geneid, bold = TRUE),
-                           Geneid)) %>%
-    kable(format = 'html', 
-          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed pir genes'), 
-          escape = FALSE) %>% 
-    pack_rows(index = kable_index) %>% 
-    kable_styling()
-
-  # print(paste0('Completed analysis of ', tpm_input$experiment))
-  list = list(tpm, all_genes, pir_genes, tpm_input$experiment)
-  names(list) = c('tpm', 'all_genes', 'pir_genes', 'experiment')
-  return(list)
-}
-
-### Generate and save top genes - core dataset all_top
-
-all_top <- lapply(expt_tpms, 
-                  top_n_genes)
-
-lapply(all_top, `[`, c('all_genes', 'pir_genes'))
-
-```
-
-##Supp figure 1A - Pearson correlation plot of collated samples.
-
-```{r all_samples_collated_heatmap}
-
-## Collated across the stages for a more easily understood heatmap
-
-big_top = bind_rows(lapply(all_top,function(x){
-  data = x$tpm
-  data$stage = paste(x$experiment, data$stage, sep = '_')
-  data
-}))
-
-big_top_recast = dcast(big_top, Geneid ~ stage, value.var = 'mean_tpm')
-big_top_recast <- column_to_rownames(big_top_recast, var = 'Geneid')
-# big_top_recast = big_top_recast[,-1,drop=FALSE]
-
-# apply(big_top_recast, 2, sum)
-
-pdf(file = 'Manuscript_suppfig1A_corrplot_all_experiments_rnaseq.pdf',
-    height = 5,
-    width = 5)
-corrplot(cor(big_top_recast), method = 'color', 
-         order = 'hclust', hclust.method = 'ward.D2',
-         cl.lim = c(-0.036,1), tl.cex = 0.50,
-         col = c(viridis(100), magma(115)), addrect = 6)
-dev.off()
-
-corrplot(cor(big_top_recast), method = 'color', 
-         order = 'hclust', hclust.method = 'ward.D2',
-         cl.lim = c(-0.036,1), tl.cex = 0.50,
-         col = c(viridis(100), magma(115)), addrect = 6)
-```
-
-## Supp figure 1B - PCA of all the samples.
-
-```{r supp_figure_PCA_allgenes}
-
-big_tpm_pca <-  prcomp(big_tpm, 
-                       center = TRUE, 
-                       scale. = TRUE)
-
-big_tpm_pca_var <- get_pca_var(big_tpm_pca)
-big_tpm_pca_coords <- data.frame(sample = rownames(big_tpm_pca_var$coord),
-                                 big_tpm_pca_var$coord)
-big_tpm_pca_coords$stage_experiment <- str_extract(
-  big_tpm_pca_coords$sample,
-  pattern = '.+(?=_[[:digit:]]+$)'
-)
-big_tpm_pca_coords$stage <- str_extract(big_tpm_pca_coords$sample, 
-                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
-big_tpm_pca_coords$experiment <- str_extract(big_tpm_pca_coords$sample, 
-                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')
-
-colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
-                               length.out = length(unique(big_tpm_pca_coords$stage))),
-                       unique(big_tpm_pca_coords$stage))
-
-pca_importance <- as.data.frame(summary(big_tpm_pca)$importance)
-
-plot <- ggplot(big_tpm_pca_coords,
-       aes(
-         x = Dim.1,
-         y = Dim.2,
-         col = stage,
-         shape = experiment
-       )) +
-  geom_point() +
-  scale_fill_manual(values = colornames) +
-  scale_shape_manual(
-    values = setNames(
-      c(1:10),
-      unique(big_tpm_pca_coords$experiment)
-    )
-  ) +
-  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-  guides(shape=guide_legend(ncol=2),
-         col = guide_legend(ncol=2)) +
-  theme_classic() +
-  theme(legend.key.size = unit(5, 'mm'),
-        legend.text = element_text(size = 6))
-
-plot
-
-```
-
-#Supp figure 1C - PCA using just the pir genes.
-
-```{r supp_figure_PCA_pir}
-big_tpm_pir <- big_tpm[rownames(big_tpm) %in% pir_id,,drop = FALSE]
-
-big_tpm_pir_pca <-  prcomp(big_tpm_pir, 
-                       center = TRUE, 
-                       scale. = TRUE)
-
-big_tpm_pir_pca_var <- get_pca_var(big_tpm_pir_pca)
-big_tpm_pir_pca_coords <- data.frame(sample = rownames(big_tpm_pir_pca_var$coord),
-                                 big_tpm_pir_pca_var$coord)
-big_tpm_pir_pca_coords$stage_experiment <- str_extract(
-  big_tpm_pir_pca_coords$sample,
-  pattern = '.+(?=_[[:digit:]]+$)'
-)
-big_tpm_pir_pca_coords$stage <- str_extract(big_tpm_pir_pca_coords$sample, 
-                                        pattern = '(?<=E[[:digit:]]{2}_).+(?=_[[:digit:]]+$)')
-big_tpm_pir_pca_coords$experiment <- str_extract(big_tpm_pir_pca_coords$sample, 
-                                             pattern = '^.+(?=_.+_[[:digit:]]+$)')
-
-colornames <- setNames(rep_len(c(cbPalette, 'black', 'red'), 
-                               length.out = length(unique(big_tpm_pir_pca_coords$stage))),
-                       unique(big_tpm_pir_pca_coords$stage))
-
-pca_importance <- as.data.frame(summary(big_tpm_pir_pca)$importance)
-
-plot_pir <- ggplot(big_tpm_pir_pca_coords,
-       aes(
-         x = Dim.1,
-         y = Dim.2,
-         col = stage,
-         shape = experiment
-       )) +
-  geom_point() +
-  scale_fill_manual(values = colornames) +
-  scale_shape_manual(
-    values = setNames(
-      c(1:10),
-      unique(big_tpm_pca_coords$experiment)
-    )
-  ) +
-  xlab(paste0('PC1 ', round(pca_importance$PC1[2]*100, digits = 1),'%')) +
-  ylab(paste0('PC2 ', round(pca_importance$PC2[2]*100, digits = 1),'%')) +
-  guides(shape=guide_legend(ncol=2),
-         col = guide_legend(ncol=2)) +
-  theme_classic() +
-  theme(legend.key.size = unit(5, 'mm'),
-        legend.text = element_text(size = 6))
-
-plot_pir
-```
-
-```{r combining_stages_data}
-combined_data <- big_tpm %>% 
-  rownames_to_column(var = 'Geneid') %>% 
-  melt(value.name = 'tpm') %>% 
-  mutate(stage = str_extract(variable, 
-                             pattern = '(?<=_).+(?=_)')) %>% 
-  group_by(Geneid, stage) %>% 
-  dplyr::summarise(tpm_mean = mean(tpm),
-                   tpm_med = median(tpm),
-                   tpm_sd = sd(tpm),
-                   tpm_max = max(tpm),
-                   tpm_min = min(tpm))
-
-combined_data <- dcast(combined_data, 
-                       Geneid ~ stage, 
-                       value.var = 'tpm_mean') %>% 
-  column_to_rownames(var = 'Geneid')
-
-#Ordering by life cycle progression
-combined_data <- combined_data[,
-                               ordering(colnames(combined_data), 
-                                        order_vector = gsub(order_vector,
-                                                            pattern = '-',
-                                                            replacement = '\\.')
-                               )]
-
-#Check that they still sum to 1e6
-
-apply(combined_data, 2, sum)
-
-```
-
-```{r colours_clades_phase}
-SorL_col <- setNames(
-  c(    '#F0E442',
-        '#56B4E9',
-        '#E69F00'),
-  sort(
-    unique(pir_clade %>% 
-             mutate(
-               SorL = str_extract(clade, 
-                                  pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent')
-             ) %>% 
-             .$SorL)
-  )
-)
-
-SorL_col_num <- SorL_col
-names(SorL_col_num) <- sapply(names(SorL_col), 
-                              function(clade_name) {
-                                paste0(
-                                  clade_name, 
-                                  ' (n = ',
-                                  sum(pir_clade$SorL == clade_name),
-                                  ')'
-                                )
-                              })
-
-clade_col <- setNames(
-  c('#F0E442',
-    '#56B4E9',
-    '#009E73',
-    '#0072B2',
-    '#E69F00',
-    '#D55E00',
-    '#CC79A7',
-    '#999999',
-    'black',
-    'red'),
-  sort(
-   unique(pir_clade$clade)
-  )
-)
-
-clade_col_num <- clade_col
-names(clade_col_num) <- sapply(names(clade_col), 
-                               function(clade_name) {
-                                 paste0(
-                                   clade_name, 
-                                   ' (n = ',
-                                   sum(pir_clade$clade == clade_name),
-                                   ')'
-                                 )
-                               })
-```
-
-# pir expression, S or L-form, stacked bar chart, stack for each gene
-
-```{r SorL_stacked_eachGene_fig2a}
-  tpm_data <- combined_data %>% 
-    rownames_to_column(var = 'Geneid') %>% 
-    filter(Geneid %in% pir_id) %>% 
-    melt(id.vars = 'Geneid', 
-         variable.name = 'stage', 
-         value.name = 'tpm') %>% 
-  mutate(clade = pir_clade$clade[match(Geneid, 
-                                       pir_clade$Geneid)]
-         ) %>% 
-  group_by(clade) %>% 
-  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
-    mutate(stack_group = Geneid) %>% 
-    mutate(tpm = ifelse(tpm > 1, tpm, 0)) %>% #Filtering step for low value pir genes.
-    group_by(stage, 
-             stack_group,
-             clade) %>% 
-    dplyr::summarise(total_tpm = sum(tpm)) %>%
-  arrange(
-          desc(clade),
-          total_tpm
-)
-
-tpm_data <- mutate(tpm_data,
-                   order = row_number())
-
-tpm_data_SorL_stackgenes <- mutate(tpm_data,
-                        SorL = str_extract(clade, 
-                                           pattern = '(L|S(?=[[:digit:]]))|ancestral|pan-rodent'))
-tpm_data_SorL_stackgenes <- group_by(tpm_data_SorL_stackgenes,
-                                     stage, 
-                                     SorL,
-                                     stack_group) %>% 
-  dplyr::summarise(total_tpm = sum(total_tpm)) %>% 
-  #Include the number of genes in each fill category for the labels
-  mutate(SorL = paste0(SorL, 
-                       ' (n = ', 
-                       n_distinct(stack_group),
-                       ')')) %>% 
-#Order by SorL and then by total_tpm, then convert stack_group to a factor so ggplot orders correctly
-  arrange(SorL, 
-          total_tpm) %>% 
-  mutate(stage_num = as.integer(stage))
-tpm_data_SorL_stackgenes$stage %>% as.integer %>% unique
-tpm_data_SorL_stackgenes$stage %>% unique
-
-tpm_data_SorL_stackgenes$stack_group_order <- seq_len(length.out = nrow(tpm_data_SorL_stackgenes))
-```
-
-```{r stacked_bar_chart_figure2b}
-#Calculate the range for the y value to make it look nicer
-stacked_max <- tpm_data %>% 
-  ungroup %>% 
-  group_by(stage) %>% 
-  dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
-  .$max_tpm %>% 
-  max
-
-max_tpm <- c(
-  seq(
-    0,
-    mround(
-      stacked_max,
-      500),
-    length.out = 11)
-)
-
-stack_label <- 'sub-family'
-
-write_csv(x = tpm_data_SorL_stackgenes,
-          path = 'tl20-05_SorL-subfamily_pir.csv')
-
-stage_vec <- unique(
-  levels(tpm_data$stage)
-)
-
-stage_vec <- tpm_data_SorL_stackgenes$stage
-col_group <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'zzAsex' ,
-                   str_detect(stage_vec, pattern = 'Liv') ~ 'zzLiv',
-                   str_detect(stage_vec, pattern = 'Gam') ~ 'zzGam',
-                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'zzMos')
-tpm_data_SorL_stackgenes$stage_col <- col_group
-fills <- c('#F6B9AD','#96d3ff','#f0d4e3','#a7e9af') #Note that this order is influenced by sort(col_group)
-
-y <- ggplot(data = tpm_data_SorL_stackgenes, 
-            aes(
-              x = stage_num,
-              y = total_tpm, 
-              fill = SorL,
-              # group = stage_num
-            )) +
-  geom_rect(
-    aes(
-      xmin = stage_num - .7,
-      xmax = stage_num + .7,
-      ymin = -Inf,
-      ymax = Inf,
-      fill = stage_col
-    ),
-    alpha = .05
-  ) +
-  geom_bar(position = position_stack(),
-           stat = 'identity',
-           colour = 'black',
-           size = 0.1) + 
-  scale_fill_manual(
-    breaks = unique(tpm_data_SorL_stackgenes$SorL),
-    values = c(
-      SorL_col_num,
-      fills
-    )
-  ) +
-  scale_color_manual('black') + 
-  # labs(fill = stack_label) +
-  scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  # scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) +
-  scale_x_continuous(
-    breaks = tpm_data_SorL_stackgenes$stage_num %>% unique(),
-    labels = tpm_data_SorL_stackgenes$stage %>% unique(),
-    expand = c(0, 0)
-  ) +
-    theme(
-      axis.text.x = element_text(angle = 90, 
-                                 vjust = .5,
-                                 size = 10),
-      axis.text.y = element_text(size = 10),
-    axis.title.y = element_text(size = 10),
-    legend.title = element_text(size = 12),
-    legend.text = element_text(size = 10),
-    legend.key.size = unit(3,'mm'),
-    legend.key.width = unit(2,'mm'),
-    ) +
-    xlab('stage') + 
-    facet_grid(.~stage, 
-               scales = "free_x",
-               space = "free_x") +
-    scale_x_discrete(expand=c(0,0)) +
-    theme(
-      axis.title.x=element_blank(),
-      axis.text.x=element_blank(),
-      axis.ticks.x = element_blank(),
-      # axis.line.x = element_blank(),
-      axis.text.y = element_text(size = 10),
-      axis.title.y = element_text(size = 10),
-      legend.title = element_text(size = 12),
-      legend.text = element_text(size = 10),
-      legend.key.size = unit(3,'mm'),
-      legend.key.width = unit(2,'mm'),
-      strip.text.x = element_text(size=8, 
-                                  face = 'bold', 
-                                  angle = 90),
-      strip.background = element_rect(
-        color="black", 
-        size=0.1, 
-        linetype="solid"
-      ),
-      panel.spacing.x = unit(0, "null")
-    )
-
-g <- ggplot_gtable(ggplot_build(y))
-strip_t <- which(grepl('strip-t', g$layout$name))
-stage_vec <- unique(
-  levels(tpm_data$stage)
-)
-fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
-                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
-                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
-                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
-k <- 1
-for (i in strip_t) {
-  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-  k <- k+1
-}
-
-grid.newpage()
-grid.draw(g)
-
-```
-
-# figure2binset
-
-```{r stack_bar_graph_SorL_lowexpression-figure2binset}
-#Include all stages expressed lower than a certain amount
-
-low_express <- tpm_data_SorL_stackgenes %>% 
-  group_by(stage) %>% 
-  summarise(sum_tpm = sum(total_tpm)) %>% 
-  filter(sum_tpm < 350)
-
-#Calculate the range for the y value to make it look nicer
-
-stacked_max <- max(low_express$sum_tpm)
-
-max_tpm <- c(
-  seq(
-    0,
-    mround(
-      stacked_max,
-      100),
-    length.out = 11)
-)
-
-tpm_data_SorL_stackgenes_low <- filter(tpm_data_SorL_stackgenes,
-                                       stage %in% low_express$stage)
-
-y <- ggplot(data = tpm_data_SorL_stackgenes_low, 
-            aes(
-              x = stage,
-              y = total_tpm, 
-              fill = SorL,
-              # group = stage_num
-            )) +
-  geom_bar(position = position_stack(),
-           stat = 'identity',
-           colour = 'black',
-           size = 0.1) + 
-  scale_fill_manual(
-    breaks = unique(tpm_data_SorL_stackgenes$SorL),
-    values = c(
-      SorL_col_num,
-      fills
-    )
-  ) +
-  scale_color_manual('black') + 
-  labs(fill = stack_label) +
-  scale_y_continuous(breaks = max_tpm) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('total '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~stage, 
-             scales = "free_x",
-             space = "free_x") +
-  scale_x_discrete(expand=expand_scale(add=1)) +
-  theme(
-    axis.title.x=element_blank(),
-    axis.text.x=element_blank(),
-    axis.ticks.x = element_blank(),
-    # axis.line.x = element_blank(),
-    axis.text.y = element_text(size = 15),
-    axis.title.y = element_text(size = 8),
-    legend.title = element_text(size = 7),
-    legend.text = element_text(size = 5),
-    legend.key.size = unit(3,'mm'),
-    legend.key.width = unit(2,'mm'),
-    strip.text.x = element_text(size=6, 
-                                face = 'bold', 
-                                angle = 90),
-    strip.background = element_rect(
-      color="black", 
-      size=0.1, 
-      linetype="solid"
-    ),
-    panel.spacing.x = unit(0, "null")
-  )
-
-#This bit is to change the colour of the facet boxes
-
-g <- ggplot_gtable(ggplot_build(y))
-strip_t <- which(grepl('strip-t', g$layout$name))
-stage_vec <- unique(
-  levels(tpm_data_SorL_stackgenes_low$stage)[levels(tpm_data_SorL_stackgenes_low$stage) %in% as.character(tpm_data_SorL_stackgenes_low$stage)]
-)
-fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
-                   str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
-                   str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
-                   str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
-k <- 1
-for (i in strip_t) {
-  j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-  g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-  k <- k+1
-}
-grid.newpage()
-grid.draw(g)
-```
-
-```{r table_of_SorL_highest_supptable}
-#For a neat table
-SorL_kable_index <- rep.int(2, 
-                            times = length(unique(tpm_data_SorL_stackgenes$stage)))
-names(SorL_kable_index) <- levels(tpm_data_SorL_stackgenes$stage)
-
-tpm_data_SorL_stackgenes_top_table <- tpm_data_SorL_stackgenes %>% 
-  mutate(Geneid = stack_group) %>% 
-  filter(str_detect(SorL, 
-                    pattern = 'ancestral', 
-                    negate = TRUE)) %>% 
-  group_by(stage) %>% 
-  mutate(sum_tpm = sum(total_tpm)) %>% 
-  group_by(stage, 
-           SorL) %>% 
-  slice_max(order_by = total_tpm,
-            n = 1) %>% 
-  mutate(total_tpm = round(total_tpm, digits = 2),
-         prop_tpm = round(total_tpm/sum_tpm, digits = 2)) %>% 
-  select(stage, Geneid, SorL, total_tpm, prop_tpm)
-
-tpm_data_SorL_stackgenes_top_table %>% 
-  kable() %>% 
-  kable_styling() %>% 
-  pack_rows(index = SorL_kable_index) %>% 
-  scroll_box(width = "100%", height = "500px")
-```
-
-## Figure 2B - sub-family proportions
-
-```{r combined_tpm_stacked_df}
-
-  tpm_data_subfam <- combined_data %>% 
-    rownames_to_column(var = 'Geneid') %>% 
-    filter(Geneid %in% pir_id) %>% 
-    melt(id.vars = 'Geneid', 
-         variable.name = 'stage', 
-         value.name = 'tpm') %>% 
-  mutate(clade = pir_clade$clade[match(Geneid, 
-                                       pir_clade$Geneid)]
-         ) %>% 
-  group_by(clade) %>% 
-  mutate(clade = paste0(clade, ' (n = ', n_distinct(Geneid) ,')')) %>% 
-    mutate(stack_group = clade) %>% 
-    group_by(stage, 
-             stack_group,
-             clade) %>% 
-    dplyr::summarise(total_tpm = sum(tpm)) %>%
-  arrange(
-          desc(clade),
-          total_tpm
-)
-
-tpm_data_subfam <- mutate(tpm_data_subfam,
-                   order = row_number())
-
-  # #Apply ordering
-  # tpm_data_subfam$stage = factor(tpm_data_subfam$stage, 
-  #                         levels = ordering(tpm_data_subfam$stage,
-  #                                           order_vector))
-
-  #Calculate the range for the y value to make it look nicer
-  stacked_max <- tpm_data_subfam %>% 
-    ungroup %>% 
-    group_by(stage) %>% 
-    dplyr::summarise(max_tpm = sum(total_tpm)) %>% 
-    .$max_tpm %>% 
-    max
-
-  max_tpm <- c(
-    seq(
-      0,
-      mround(
-        stacked_max,
-        500),
-      length.out = 11)
-  )
-```
-
-```{r combined_tpm_stacked_plot}
-
-  stack_label <- 'sub-family'
-
-y <- ggplot(data = tpm_data_subfam, 
-            aes(
-              x = stage,
-              y = total_tpm, 
-              fill = clade,
-              group = order
-            )) +
-  geom_bar(
-    # position = position_stack(),
-    position = 'fill',
-    stat = 'identity',
-    colour = 'black',
-    size = 0.1) +
-  scale_fill_manual(
-    values = clade_col_num
-  ) +
-  scale_color_manual('black') + 
-  labs(fill = stack_label) +
-  scale_y_continuous(breaks = seq(0,1,0.1)) +
-  theme_classic() +
-  scale_size(range = c(0,3)) +
-  ylab(
-    bquote('Proportion of '~italic(.('pir'))~' TPM')
-  ) + 
-  facet_grid(.~stage, 
-             scales = "free_x",
-             space = "free_x") +
-  scale_x_discrete(expand=expand_scale(add=1)) +
-  theme(
-    axis.title.x=element_blank(),
-    axis.text.x=element_blank(),
-    axis.ticks.x = element_blank(),
-    # axis.line.x = element_blank(),
-    axis.text.y = element_text(size = 10),
-    axis.title.y = element_text(size = 10),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 8),
-    legend.key.size = unit(3,'mm'),
-    legend.key.width = unit(2,'mm'),
-    strip.text.x = element_text(size=8, 
-                                face = 'bold', 
-                                angle = 90),
-    strip.background = element_rect(
-      color="black", 
-      size=0.1, 
-      linetype="solid"
-    ),
-    panel.spacing.x = unit(0, "null")
-  )
-
-  #This bit is to change the colour of the facet boxes
-
-  g <- ggplot_gtable(ggplot_build(y))
-  strip_t <- which(grepl('strip-t', g$layout$name))
-  stage_vec <- unique(
-
-    levels(tpm_data_subfam$stage)
-  )
-  fills <- case_when(str_detect(stage_vec, pattern = 'Asex') ~ 'coral1' ,
-                     str_detect(stage_vec, pattern = 'Liv') ~ 'plum2',
-                     str_detect(stage_vec, pattern = 'Gam') ~ 'lightskyblue1',
-                     str_detect(stage_vec, pattern  = 'Bld|Sporo|Ook') ~ 'lightgreen')
-  k <- 1
-  for (i in strip_t) {
-    j <- which(grepl('rect', g$grobs[[i]]$grobs[[1]]$childrenOrder))
-    g$grobs[[i]]$grobs[[1]]$children[[j]]$gp$fill <- fills[k]
-    k <- k+1
-  }
-
-  grid.newpage()
-  grid.draw(g)
-```
-
-## Gametocyte marker gene expression - supp figure 2
-
-```{r}
-list_genes <- c(
-  list('P28' = 'PBANKA_0514900'),
-  list('nek4' = 'PBANKA_0616700'),
-  list('map2' = 'PBANKA_0933700'),
-  list('hap2' = 'PBANKA_1212600')
-)
-gam_markers <- unlist(list_genes)
-
-list_tpms_line <- lapply(expt_tpms, 
-                         function(input){
-                           tpm_data <- input$tpm %>% 
-                             filter(Geneid %in% gam_markers) %>% 
-                             mutate(Geneid = names(gam_markers)[match(Geneid, gam_markers)]) %>% 
-                             melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-                             mutate(stage = unlist(lapply(strsplit(as.character(sample), 
-                                                                   split = '_'),
-                                                          `[`,
-                                                          1))) %>% 
-                             group_by(stage, 
-                                      sample,
-                                      Geneid) %>% 
-                             dplyr::summarise(total_tpm = sum(tpm))
-
-                           #Make stages with no cirs in the sample give a zero 
-                           if(!all(colnames(input$tpm) %in% tpm_data$sample)){
-                             missing_stages = colnames(input$tpm)[-1][!colnames(input$tpm)[-1] %in% tpm_data$sample]
-                             missing_stages_data_zero = data.frame(lapply(missing_stages, function(missing_st){
-                               data.frame(stage = unlist(lapply(strsplit(as.character(missing_st), split = '_'),`[`,2)),
-                                          sample = missing_st,
-                                          total_tpm = 0)
-                             }))
-                             tpm_data <- bind_rows(tpm_data, missing_stages_data_zero)
-                           }
-
-                           #Add experiment column
-                           tpm_data$experiment = rep(input$experiment, times = nrow(tpm_data))
-                           max_tpm = if(max(tpm_data$total_tpm) < 10){
-                             c(seq(0,mround(max(tpm_data$total_tpm),1),1))
-                           }else{
-                             c(seq(0,mround(max(tpm_data$total_tpm),5),5))
-                           }
-
-                           #Get the average data for the line graph
-                           tpm_data_line <- tpm_data %>% 
-                             ungroup %>% 
-                             group_by(stage, 
-                                      experiment,
-                                      Geneid) %>% 
-                             dplyr::summarise(total_tpm_median = median(total_tpm),
-                                              Number_of_replicates = n(),
-                                              total_tpm_max = max(total_tpm),
-                                              total_tpm_min = min(total_tpm)) %>% 
-                             arrange(stage)
-
-                           return(tpm_data_line)
-                         })
-
-tpm_data_line <- bind_rows(list_tpms_line)
-
-tpm_data_line$stage <- factor(tpm_data_line$stage, 
-                             levels = ordering(tpm_data_line$stage,
-                                               order_vector))
-
-#Calculate the range for the y value to make it look nicer
-if(max(
-  tpm_data_line$total_tpm_max
-) >= 500){
-  max_tpm = c(
-    seq(
-      0,
-      mround(
-        max(
-          tpm_data_line$total_tpm_max
-        ),
-        500),
-      length.out = 11)
-  )
-}else{
-  max_tpm = c(
-    seq(
-      0,
-      mround(
-        max(
-          tpm_data_line$total_tpm_max
-        ),
-        10),
-      length.out = 11)
-  )
-}
-```
-
-super hacky way to change the scales in each facet from:
-https://fishandwhistle.net/post/2018/modifying-facet-scales-in-ggplot2/
-
-```{r}
-scale_override <- function(which, scale) {
-  if(!is.numeric(which) || (length(which) != 1) || (which %% 1 != 0)) {
-    stop("which must be an integer of length 1")
-  }
-
-  if(is.null(scale$aesthetics) || !any(c("x", "y") %in% scale$aesthetics)) {
-    stop("scale must be an x or y position scale")
-  }
-
-  structure(list(which = which, scale = scale), class = "scale_override")
-}
-
-CustomFacetWrap <- ggproto(
-  "CustomFacetWrap", FacetWrap,
-  init_scales = function(self, layout, x_scale = NULL, y_scale = NULL, params) {
-    # make the initial x, y scales list
-    scales <- ggproto_parent(FacetWrap, self)$init_scales(layout, x_scale, y_scale, params)
-
-    if(is.null(params$scale_overrides)) return(scales)
-
-    max_scale_x <- length(scales$x)
-    max_scale_y <- length(scales$y)
-
-    # ... do some modification of the scales$x and scales$y here based on params$scale_overrides
-    for(scale_override in params$scale_overrides) {
-      which <- scale_override$which
-      scale <- scale_override$scale
-
-      if("x" %in% scale$aesthetics) {
-        if(!is.null(scales$x)) {
-          if(which < 0 || which > max_scale_x) stop("Invalid index of x scale: ", which)
-          scales$x[[which]] <- scale$clone()
-        }
-      } else if("y" %in% scale$aesthetics) {
-        if(!is.null(scales$y)) {
-          if(which < 0 || which > max_scale_y) stop("Invalid index of y scale: ", which)
-          scales$y[[which]] <- scale$clone()
-        }
-      } else {
-        stop("Invalid scale")
-      }
-    }
-
-    # return scales
-    scales
-  }
-)
-
-facet_wrap_custom <- function(..., scale_overrides = NULL) {
-  # take advantage of the sanitizing that happens in facet_wrap
-  facet_super <- facet_wrap(...)
-
-  # sanitize scale overrides
-  if(inherits(scale_overrides, "scale_override")) {
-    scale_overrides <- list(scale_overrides)
-  } else if(!is.list(scale_overrides) || 
-            !all(vapply(scale_overrides, inherits, "scale_override", FUN.VALUE = logical(1)))) {
-    stop("scale_overrides must be a scale_override object or a list of scale_override objects")
-  }
-
-  facet_super$params$scale_overrides <- scale_overrides
-
-  ggproto(NULL, CustomFacetWrap,
-    shrink = facet_super$shrink,
-    params = facet_super$params
-  )
-}
-
-(
-  y <- ggplot(data = tpm_data_line, 
-              aes(x = stage, 
-                  y = total_tpm_median, 
-                  fill = experiment, 
-                  group = experiment)) +
-    geom_bar(position = position_dodge2(preserve = 'single'),
-             stat = 'identity') +
-    scale_fill_manual(values = c(cbPalette, 'red', 'purple')[1:(length(unique(tpm_data_line$experiment)))]) +
-    geom_errorbar(aes(x = stage,
-                      ymax = total_tpm_max, 
-                      ymin = total_tpm_min),
-                  position = position_dodge2(preserve = 'single'),
-                  size = 0.25) +
-    scale_y_continuous(breaks = max_tpm) +
-    theme_classic() +
-    theme(axis.title.x=element_blank(),
-          axis.text.x=element_text(angle = 90, 
-                                   size = 6, 
-                                   vjust = 0.5),
-          axis.text.y = element_text(size = 5),
-          axis.title.y = element_text(size = 8),
-          legend.title = element_text(size = 7),
-          legend.text = element_text(size = 5),
-          legend.key.size = unit(3,'mm'),
-          legend.key.width = unit(2,'mm')) +
-    scale_size(name = 'Number of genes detected over threshold', 
-               range = c(0,3)) +
-    ylab(
-      bquote('total '~italic(.('pir'))~' TPM')
-    ) +
-    facet_wrap_custom(~Geneid, 
-                      scales = 'free_y',
-                      scale_overrides = list(
-                        scale_override(3, scale_y_continuous(breaks = seq(0,650,length.out = 11))),
-                        scale_override(2, scale_y_continuous(breaks = seq(0,900,length.out = 11))),
-                        scale_override(1, scale_y_continuous(breaks = seq(0,310,length.out = 11)))
-                      )) +
-    annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)
-)
-#Could also use functions from a package called 'lemon' to add the lines to the top two facets.
-```
-
-## Lower expression of _pir_ genes versus the rest of the genome
-
-```{r low_pir_expression}
-set.seed(4908)
-
-gene_expression_df <- combined_data %>% 
-    rownames_to_column(var = 'Geneid') %>% 
-    melt(id.vars = 'Geneid', 
-         variable.name = 'stage', 
-         value.name = 'tpm') %>% 
-  mutate(pir_gene = ifelse(Geneid %in% pir_id,
-                           'pir',
-                           'non-pir')) %>% 
-  group_by(stage, pir_gene) %>% 
-  slice_sample(n = length(pir_id)) %>%
-  summarise(mean_tpm = mean(tpm),
-            med_tpm = round(median(tpm),
-                            digits = 6))
-
-g <- ggplot(gene_expression_df,
-       aes(x = pir_gene,
-           y = log2(1+med_tpm))) +
-  xlab('Gene group')+
-  ylab('l0g(Median TPM + 1)')+
-  # geom_violin()+
-  geom_dotplot(binaxis = 'y',
-               stackdir = 'center',
-               # binwidth = 1,
-               # dotsize = 2,
-               method = 'dotdensity') +
-  geom_path(aes(group = stage,
-                col = stage)) +
-  theme_classic() +
-  theme(
-    legend.text = element_text(size = 9),
-    legend.key.size = unit(4, 'mm')
-  )
-
-g
-```
-
-_pir_ expression on average ranges from minimum mean `r min(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$mean_tpm)` TPM and median `r min(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$med_tpm)` TPM to maximum mean `r max(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$mean_tpm)` TPM and median `r max(gene_expression_df[gene_expression_df$pir_gene == 'pir',]$med_tpm)` TPM.
-Non-\ _pir_ expression ranges from minimum mean `r min(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$mean_tpm)` TPM and median `r min(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$med_tpm)` TPM to maximum mean `r max(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$mean_tpm)` TPM and median `r max(gene_expression_df[gene_expression_df$pir_gene == 'non-pir',]$med_tpm)` TPM.

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Separate TPM calculation Rmd file

@@
 
+```{r remove_badly_clustering_samples}
+berghei_tpm$`E08_Liv.pre4h_1` <- NULL
+berghei_tpm$`E03_Asex.Schiz_1` <- NULL
+berghei_tpm$`E03_Asex.Schiz_3` <- NULL
+```
+
+## Saving the TPM
+
+```{r save_tpm}
+write.csv(berghei_tpm,
+          'berghei_tpm.csv')
+```
+
 ```{r expt_tpms}
@@
 names(expt_tpms) <- expt_nams
-```
-
-Going to combine certain samples based on their distinctive profiles, e.g. <4h, 6-48h, >48h liver stages.
-
-```{r combining_samples}
-names_expt_tpms <- names(expt_tpms)
-
-expt_tpms <- lapply(names(expt_tpms), function(expt){
-  col_stage <- colnames(
-    getElement(
-      getElement(
-        expt_tpms, 
-        paste(expt)
-      ),
-      'tpm'
-    )
-  )
-
-  liver_stage <- col_stage[
-    str_detect(
-      col_stage,
-      pattern = 'Liv-[[:digit:]]{1,2}h'
-    )
-    ]
-
-  new_liver_stage <- case_when(
-    str_detect(liver_stage, pattern = 'Liv-(2|4)h') ~ 'Liv-pre4h',
-    # str_detect(liver_stage, pattern = 'Liv-([5-9]|[1-3][0-9]|4[0-7])h') ~ 'Liv-pre47h',
-    str_detect(liver_stage, pattern = 'Liv-(4[8-9]|[5-9][0-9])h') ~ 'Liv-post48h',
-    TRUE ~ as.character(liver_stage),
-  )
-
-  if(length(new_liver_stage) > 0){
-    new_liver_stage <-uniquify(new_liver_stage)
-  }
-
-  col_stage[
-    str_detect(
-      col_stage,
-      pattern = 'Liv-[[:digit:]]{1,2}h'
-    )
-    ] <- new_liver_stage
-
-  bloodmeal_stage <- col_stage[
-    str_detect(
-      col_stage, 
-      pattern = 'BldMl-(1h|3h)_[[:digit:]]+'
-    )
-    ]
-
-  bloodmeal_stage <- str_replace(
-    bloodmeal_stage, 
-    pattern = '(1h|3h)_[[:digit:]]+', 
-    replacement = 'Pre3h'
-  )
-  if(length(bloodmeal_stage)>0){
-    bloodmeal_stage <- uniquify(bloodmeal_stage)
-  }
-
-  col_stage[
-    str_detect(
-      col_stage, 
-      pattern = 'BldMl-(1h|3h)_[[:digit:]]+'
-    )
-    ] <- bloodmeal_stage
-
-  tpm_element <- getElement(
-    getElement(
-      expt_tpms, 
-      paste(expt)
-    ),
-    'tpm'
-  )
-
-  colnames(tpm_element) <- col_stage
-
-  list(tpm = tpm_element,
-       experiment =   getElement(
-         getElement(
-           expt_tpms,
-           paste(expt)
-         ),
-         'experiment'
-       )
-  )
-
-})
-names(expt_tpms) <- names_expt_tpms
-
-#Re-ordering the list
-expt_tpms <- (expt_tpms[order(names(expt_tpms))])
-```
-
-```{r remove_badly_clustering_samples}
-expt_tpms$E08$tpm$`Liv-pre4h_1` <- NULL
-expt_tpms$E03$tpm$`Asex-Schiz_1` <- NULL
-expt_tpms$E03$tpm$`Asex-Schiz_3` <- NULL
-```
-
-```{r big_tpm}
-
-```
-
-## Saving the TPM
-
-```{r save_tpm}
-
-write.csv()
 ```

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
added more functions to separate files

@@
 source('../rfunctions/tpm_fun.R')
-```
-
-```{r function_mround}
-#Make nicer numbers rounded up, useful for making scales for a plot.
-mround <- function(x,base){ 
-  base*ceiling(x/base) 
-} 
-```
-
-```{r function_my_max}
-#max robust to NAs
-my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) 
-```
-
-```{r function_firstup}
-#Capitalize first letter
-firstup <- function(x) {
-  substr(x, 1, 1) <- toupper(substr(x, 1, 1))
-  x
-}
+source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
 ```
@@
 
-```{r transcript_redundancy, message = FALSE, results='hide'}
-
-#Note that there is some redundancy
-two.transcript <- allgene.info %>% 
-  dplyr::count(Geneid) %>% 
-  filter(n > 1)
-
-#Nine genes have two transcripts - need to give them one average of the transcripts since featurecounts has only
-# calculated one count for the locus
-two.transcript.info <- allgene.info[allgene.info$Geneid %in% two.transcript$Geneid,]
-two.transcript.mean = aggregate(transcript_length~Geneid, data=two.transcript.info, 
-                                FUN=function(x) c(mean=mean(x)))
-
-#Remove the first match
-allgene.info <- allgene.info[-(match(two.transcript.mean$Geneid, 
-                                     allgene.info$Geneid)),]
-
-#For the second match replace the transcript value with the mean calculated with aggregate
-allgene.info$transcript_length <- as.double(allgene.info$transcript_length)
-allgene.info[match(two.transcript.mean$Geneid, 
-                   allgene.info$Geneid),]$transcript_length <- two.transcript.mean$transcript_length
-lengths.transcript <- subset(allgene.info, select = c(Geneid, transcript_length), drop = F)
-```
-
-## Read in counts and calculate TPM
-
-```{r read_files_calc_TPM}
-#Reading in the counts from file
-berghei_counts <- read.csv('tl20-05_berghei_counts.csv',
+## Saving the TPM
+
+```{r save_tpm}
+berghei_tpm <- read.csv('berghei_tpm.csv', 
                         row.names = 1)
-
-berghei_tpm <- tpm(berghei_counts,
-                   lengths = lengths.transcript)
-
-apply(berghei_tpm[-1], 2, sum)
-```
-
-```{r save_tpm_oldNames_allSamples}
-#Save the 'raw' TPM before any samples have been removed or sample names combined.
-
-write.csv(berghei_tpm,
-          'berghei_tpm_oldNamesAllSamples.csv')
-
-```
-
-```{r remove_E06-6h}
-#Remove the Caldelari (E06) 6h samples, as they did in their paper.
-berghei_tpm[,str_subset(colnames(berghei_tpm), 'E06_Liv.6h')] <- NULL
-```
-
-```{r hclusts, message = FALSE}
-experiment_id <- unique(na.omit(str_extract(colnames(berghei_tpm), pattern = 'E[[:digit:]]{2}')))
-
-lapply(experiment_id, 
-       function(experiment_name){
-         tpm_data <- select(berghei_tpm, matches(experiment_name))
-         if(ncol(tpm_data) > 3){
-           dists = dist(t(tpm_data))
-           par(cex=0.7)
-           plot(hclust(dists),
-                main = experiment_name)
-         }
-       })
-
-```
-
-Going to combine certain samples based on their similar profiles, e.g. <4h, >48h liver stages.
-
-```{r combine_similar_samples_byNames}
-col_stage <- colnames(berghei_tpm)
-
-#Combine all Liver samples 4h or before into LiverPre4h
-col_stage <- str_replace_all(col_stage,
-                             pattern = 'Liv\\.(2|4)h_[[:digit:]]+',
-                             replacement = 'Liv.pre4h')
-#Combine all Liver samples 48h or onwards into LiverPost48h
-col_stage <- str_replace_all(col_stage,
-                             pattern = 'Liv\\.(4[8-9]|[5-9][0-9])h_[[:digit:]]+',
-                             replacement = 'Liv.post48h')
-#Combine all blood meal samples 3h or before into BldMlPre3h
-col_stage <- str_replace_all(col_stage,
-                             pattern = '(1h|3h)_[[:digit:]]+', 
-                             replacement = 'Pre3h')
-#Uniquify the name names
-col_stage[grep(col_stage, 
-               perl = TRUE,
-               invert = TRUE,
-               pattern = '_[[:digit:]]{1,}$')] <- uniquify(col_stage[grep(col_stage, 
-                                                                          perl = TRUE,
-                                                                          invert = TRUE,
-                                                                          pattern = '_[[:digit:]]{1,}$')]) 
-```
-
-```{r remove_badly_clustering_samples}
-berghei_tpm$`E08_Liv.pre4h_1` <- NULL
-berghei_tpm$`E03_Asex.Schiz_1` <- NULL
-berghei_tpm$`E03_Asex.Schiz_3` <- NULL
-```
-
-## Saving the TPM
-
-```{r save_tpm}
-write.csv(berghei_tpm,
-          'berghei_tpm.csv')
 ```
@@
 
-lapply(all_top, `[`, c('all_genes', 'pir_genes'))
+lapply(all_top[[1]], `[`, c('all_genes', 'pir_genes'))
 

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Doing the berghei df melting and averging b4 top_n fun

@@
 
-## The highest expressed _pir_ genes across the samples
-
-```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
-top_n_genes <- function(tpm_input, 
-                        n = 5, 
-                        threshold = 0){
-  tpm = tpm_input$tpm %>% 
+```{r melt_df}
+berghei_tpm_melt <- berghei_tpm %>% 
     #Melt the data frame for using dplyr
     melt(id.vars = 'Geneid', variable.name = 'sample', value.name = 'tpm') %>% 
-    mutate(stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1))) %>% 
-    group_by(Geneid, stage) %>%  
+    mutate(experiment = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,1)),
+           stage = unlist(lapply(strsplit(as.character(sample), split = '_'),`[`,2)))
+
+berghei_tpm_avg_byexpt <- berghei_tpm_melt %>% 
+  group_by(Geneid, experiment, stage) %>%  
   dplyr::summarise(median_tpm = median(tpm),
@@
                    max_tpm = max(tpm), 
-                     sd_tpm = sd(tpm)) %>% 
-    group_by(stage) %>% 
-    mutate(experiment = tpm_input$experiment)
-
+                   sd_tpm = sd(tpm)) 
+```
+
+## The highest expressed _pir_ genes across the samples
+
+```{r top_genes, echo = FALSE, message = FALSE, results = "asis"}
+top_n_genes <- function(tpm_avg, 
+                        n = 5, 
+                        threshold = 0){
   #Calculating the index used to make the kable table
 
-  kable_index = rep(n, times = length(unique(tpm$stage)))
-  names(kable_index) = unique(tpm$stage)
-
-  all_genes = tpm %>% 
+  kable_index = rep(n, times = length(unique(tpm_avg$stage)))
+  names(kable_index) = unique(tpm_avg$stage)
+
+  all_genes = tpm_avg %>% 
     top_n(wt = mean_tpm, n = n) %>% 
@@
     kable(format = 'html', 
-          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed total genes'), 
+          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed total genes'), 
           escape = FALSE) %>%
@@
 
-  pir_genes = tpm %>% 
+  pir_genes = tpm_avg %>% 
     filter(Geneid %in% pir_id) %>% 
@@
     kable(format = 'html', 
-          caption = paste0(paste0(paste0(tpm_input$experiment, ' - Top '), paste(n)), ' expressed pir genes'), 
+          caption = paste0(tpm_avg$experiment[1], ' - Top ', n, ' expressed pir genes'), 
           escape = FALSE) %>% 
@@
   # print(paste0('Completed analysis of ', tpm_input$experiment))
-  list = list(tpm, all_genes, pir_genes, tpm_input$experiment)
-  names(list) = c('tpm', 'all_genes', 'pir_genes', 'experiment')
+  list = list(all_genes, pir_genes, tpm_input$experiment)
+  names(list) = c('all_genes', 'pir_genes', 'experiment')
   return(list)
@@
 
-all_top <- lapply(expt_tpms, 
-                  top_n_genes)
+all_top <- lapply(unique(berghei_tpm_avg_byexpt$experiment), 
+                  function(expt) top_n_genes(berghei_tpm_avg_byexpt[berghei_tpm_avg_byexpt$experiment == expt,]))
 
@@
 }))
+big_top <- berghei_tpm_melt
 

### 2021-08-31 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_Github_Pberghei_pir.Rmd
Fixed grouping issue for new top_n_gene fun

@@
 
+  tpm_avg <- tpm_avg %>% 
+    ungroup %>% 
+    group_by(stage)
+
   all_genes = tpm_avg %>% 
-    top_n(wt = mean_tpm, n = n) %>% 
+    slice_max(order_by = mean_tpm, n = n) %>% 
     dplyr::arrange(stage, desc(mean_tpm)) %>% 
@@
     filter(Geneid %in% pir_id) %>% 
-    top_n(wt = mean_tpm, n = n) %>% 
+    slice_max(order_by = mean_tpm, n = n) %>% 
     dplyr::arrange(stage, desc(mean_tpm)) %>% 

### 2021-09-01 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_calculateTPM.Rmd
Chabaudi use csv count data, and no SBP at all

@@
 
-```{r tpm, results = 'hide', message = FALSE}
-# # Calculating the transcript-per-million 
-# 
-# We need to write a function to calculate the TPM from the count data we have generated.
-This function uses the 'chabaudi_transcript_lengths' object created in the chunk above and should generate errors if it detects discrepencies between the Geneids so that the wrong transcript length is not used.
-
-tpm = function(count_data, lengths = chabaudi_transcript_lengths) {
-
-  #Print an error if the id columns do not match
-  # first if statement to test that both have a GeneID column
-  if(any(grepl(colnames(lengths), 
-               pattern = 'id', 
-               ignore.case = T)) == T & any(grepl(colnames(count_data), 
-                                                  pattern = 'Geneid', 
-                                                  #This needs to be Geneid when some of the samples have 'id' in them
-                                                  ignore.case = T)) == T){
-    if(all(count_data[grep(colnames(count_data), 
-                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
-                                                                             pattern = 'id', ignore.case = T)])) {
-      print('Count data and length gene ids both match')
-    } else {
-      stop('Error: Count data and length gene ids DO NOT MATCH, please make sure they are ordered identically.')
-    }
-  } else {
-    stop('Error: Count data and/or length table supplied do not have an id column.')
-  }
-
-  #Remove the id column from supplied count_data
-  if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = T)) == T){
-    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = T)]
-  } else{
-    counts_tpm <- count_data}
-
-  #Remove the id column from supplied lengths
-  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
-    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
-  }else{
-    lengths.table <- lengths}
-  lengths.table = matrix(lengths.table)
-  #This may make lengths.table a vector which can bring up some dimension problems in the following for loop
-
-  #Now for the actual calculation
-  rpm <- counts_tpm
-  for (i in 1:dim(counts_tpm)[2]) {
-    for (j in 1:dim(counts_tpm)[1]) {
-      rpm[j,i] <- (counts_tpm[j,i]/(lengths.table[j]/1000))
-    }
-    scale.factor <- sum(rpm[,i])/(1E6)
-    counts_tpm[,i] <- apply(rpm[,i,drop=F],1,function(x){x/scale.factor})
-  }
-  if(any(grepl(colnames(count_data), pattern = 'Geneid', ignore.case = T)) == T){
-    counts_tpm = bind_cols(count_data[,grep(colnames(count_data), pattern = 'Geneid', ignore.case = T), drop = FALSE], 
-                           counts_tpm)
-  } 
-  return(counts_tpm)
-}
+```{r sourcing_functions}
+source('../rfunctions/uniquify_fun.R')
+source('../rfunctions/ordering_fun.R')
+source('../rfunctions/tpm_fun.R')
+source('../rfunctions/usefulFunctions_mround_my-max_firstup_fun.R')
 ```
 
-```{r function_mround}
-mround <- function(x,base){ 
-  base*ceiling(x/base) 
-} 
-```
-
-```{r function_my_max}
-my_max <- function(x) ifelse( !all(is.na(x)), max(x, na.rm=T), NA) #max robust to NAs
-```
-
-```{r function_ordering}
-ordering <- function(to_order, order_vector){
-  order = unlist(sapply(order_vector, 
-                        function(x){unique(str_subset(to_order, 
-                                                      pattern = paste(x)))}))
-  order_missing = unique(to_order[!to_order %in% order])
-  unique(c(order, order_missing), fromLast = TRUE)
-}
+```{r order_vec}
 
@@
 
-```{r read_tpm_rescaled_data}
+```{r calc_tpm}
 
-chab_count_table <- read_xlsx('Manuscript_Frontiers_chabaudi_24h_counts.xlsx')
+chab_count_table <- read.csv('Manuscript_Suppinfo_chabaudi_24h_counts.csv',
+                             row.names = 1)
 
@@
                           lengths = chabaudi_transcript_lengths)
+```
+
+```{r tpm_rescale}
 
@@
 twentyfour_ct_comb <- twentyfour_tpm_rescale_nopseudo
+```
+
+```{r}
 
@@
 
-twentyfour_rMT_avg <- melt(twentyfour_rMT, 
+twentyfour_rMT_avg <- melt(twentyfour_ct_comb, 
                            variable.name = 'sample',
@@
 
-# twentyfour_SBP <- select(twentyfour_ct_comb, 
-#                          matches('Geneid|SBP')) %>% 
-#   rename_with(~ str_extract(.x,pattern = '(?<=SBP_).+'),
-#               matches('SBP'))
-# 
-# twentyfour_SBP_avg <- melt(twentyfour_SBP, 
-#                            variable.name = 'sample',
-#                            value.name = 'tpm') %>% 
-#   mutate(stage = str_extract(sample, pattern = '.+(?=_[[:digit:]])')) %>% 
-#   group_by(Geneid, stage) %>% 
-#   summarise(mean_tpm = mean(tpm)) %>% 
-#   dcast(Geneid ~ stage)
-
 twentyfour_avgtpm <- list(
@@
     experiment = 'rMT'
-  # ),
-  # 'SBP' = list(
-  #   tpm = twentyfour_SBP_avg,
-  #   experiment = 'SBP'
-  )
 )

### 2021-09-01 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_calculateTPM.Rmd
Changing the tpm_fun to be more strict with inputs

@@
 chabaudi_transcript_lengths <- select(chabaudi_gene_info, c(Gene.ID, Transcript.Length)) %>% 
-  arrange(Gene.ID)
+  arrange(Gene.ID) %>% 
+  mutate(Geneid = Gene.ID) %>% 
+  select(Geneid, Transcript.Length)
 
@@
 
-twentyfour_ct_comb <- tpm(chab_count_table, 
+twentyfour_ct_comb <- tpm(count_data = chab_count_table, 
                           lengths = chabaudi_transcript_lengths)
@@
   )
+)
 
-write_csv(twentyfour_rMT,
-          file = 'Pchabaudi_TPM_allBioReps.csv')
+# write_csv(twentyfour_rMT,
+#           file = 'Pchabaudi_TPM_allBioReps.csv')
+
+#The current file is slightly different to the new one I wrote, this worries me a lot.
+
 write_csv(twentyfour_rMT_avg,
           file = 'Pchabaudi_TPM_average.csv')
-
 ```

### 2021-09-01 Manuscript_pirAnalysis_Github/rfunctions/tpm_fun.R

@@ -5,45 +5,48 @@ tpm = function(count_data,lengths) {
   
   #Print an error if the id columns do not match
   if(any(grepl(colnames(lengths), 
-               pattern = 'id', 
-               ignore.case = T)) == T & any(grepl(colnames(count_data), 
-                                                  pattern = 'id', 
-                                                  ignore.case = T)) == T){
+               pattern = 'Geneid', 
+               ignore.case = T)) & any(grepl(colnames(count_data), 
+                                                  pattern = 'Geneid', 
+                                                  ignore.case = T))){
     if(all(count_data[grep(colnames(count_data), 
-                           pattern = 'id', ignore.case = T)] == lengths[grep(colnames(lengths), 
-                                                                             pattern = 'id', ignore.case = T)])) {
+                           pattern = 'Geneid', ignore.case = T)] == lengths[grep(colnames(lengths), 
+                                                                             pattern = 'Geneid', ignore.case = T)])) {
       print('Count data and length gene ids both match')
     } else {
       stop('Warning:Count data and length gene ids DO NOT MATCH')
     }
   } else {
-    stop('Count data and/or length table supplied do not have an id column, make sure they are ordered correctly')
+    stop('Count data and/or length table supplied do not have a Geneid column, make sure they are ordered correctly')
   }
   
   #Remove the id column from supplied count_data
-  if(any(grepl(colnames(count_data), pattern = 'id',ignore.case = T)) == T){
-    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'id', ignore.case = T)]
-  } else{
-    counts_tpm <- count_data}
+  # if(any(grepl(colnames(count_data), pattern = 'Geneid',ignore.case = T))){
+    counts_tpm = count_data[,-grep(colnames(count_data), pattern = 'Geneid', ignore.case = T)]
+  # } else{
+  #   counts_tpm <- count_data}
   
   #Remove the id column from supplied lengths
-  if(any(grepl(colnames(lengths), pattern = 'id', ignore.case = T)) == T){
-    lengths.table = lengths[,-grep(colnames(lengths), pattern = 'id', ignore.case = T)]
-  }else{
-    lengths.table <- lengths}
[…]

### 2021-09-02 Manuscript_pirAnalysis_Github/Pchabaudi24h/Manuscript_Github_Pchabaudi_pir_TPMHeatmaps.Rmd
Fixing chabTPMheatmap file error
Removed commented out code, and draw a heatmap that b4 was only saved

@@
 #Global save var - set to TRUE if you want to save figures.
-global_save_var <- TRUE
+global_save_var <- FALSE
 ```
@@
 
-# pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
 #Sub-family annotation
@@
 
-# anc_ha <- rowAnnotation(foo = anno_mark(at =  na.omit(match(clade_info, 'ancestral')) , 
-#                                         labels = 'ancestral',
-#                                         labels_gp = gpar(fontsize = 7),
-#                                         link_width = unit(3, "mm")))
-
 ht <- Heatmap(
@@
   height = unit(7,'cm'),
-  # row_title = NULL,
-  # left_annotation = clade_ha,
-  # right_annotation = anc_ha,
   row_title_rot = 0,
@@
 
-```{r}
+```{r chapl_aapl_ann}
 
@@
   at = names(annotation_col)
-    # str_subset(
-    # sort(
-    #   factor(chapl_aapl_name_vec,
-    #          ordering(chapl_aapl_name_vec,
-    #                   order_vector = 'ChAPL'))
-    # )
-    # ,
-    # pattern = 'NA',
-    # negate = TRUE
-  # )
 ,
@@
 
-# pir_highest_tpm_order <- rownames(combined_data_pir_ht)
-
-#Sub-family annotation
-chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
-annotation_col <- chapl_aapl_name_col_num[ordering(labels_chapl_aapl_name,
-                                                    order_vector = 'ChAPL')]
-
-#chapl_aapl_name annotation
-
-# chapl_aapl_name_ha <- rowAnnotation(
-#   'CHAPL/AAPL' = anno_simple(
-#     chapl_aapl_name_info,
-#     col = annotation_col,
-#   ),
-#   annotation_name_gp = gpar(fontsize = 10)
-# )
-
 #chapl_aapl_name name labels
@@
                                    num_chapl_aapl_name$chapl_aapl_name)
+
+#loci annotation
+chapl_aapl_name_info <- combined_data_pir_ht$chapl_aapl_name
+annotation_col <- chapl_aapl_name_col_num[ordering(labels_chapl_aapl_name,
+                                                    order_vector = 'ChAPL')]
 
@@
 
-# chapl_aapl_name_lgd <- Legend(
-#   title = 'CHAPL/AAPL',
-#   at = str_subset(
-#     sort(
-#       factor(labels_chapl_aapl_name,
-#              ordering(labels_chapl_aapl_name,
-#                       order_vector = 'ChAPL'))
-#     ),
-#     pattern = 'NA',
-#     negate = TRUE
-#   ),
-#   legend_gp = gpar(fill = na.omit(annotation_col))
-# )
-
 if(global_save_var){
@@
        row_title_side = 'left',
-       column_title_side = 'bottom',
-       # heatmap_legend_side = 'right',
-       # annotation_legend_side = 'right',
-       # annotation_legend_list = list(chapl_aapl_name_lgd),
-       # merge_legend = TRUE
+       column_title_side = 'bottom'
   )
@@
 }
+
+  draw(ht, 
+       row_title_side = 'left',
+       column_title_side = 'bottom'
+  )
 ```
@@
   left_annotation = chapl_aapl_name_ha,
-  # row_split = 
-  #   factor(combined_data_pir_ht$chapl_aapl_name,
-  #             levels = ordering(combined_data_pir_ht$chapl_aapl_name,
-  #                               order_vector = 'ChAPL')),
   row_km = 9,

### 2021-09-02 Manuscript_pirAnalysis_Github/rfunctions/tpm_fun.R
Fixed tpm_fun for berghei and added sanity check

@@ -31,7 +31,7 @@ tpm = function(count_data,lengths) {
     stop('Length table needs a column labelled \"Transcript.Length\" (with capitals)')
   }
     
-    lengths.table = lengths[,'Transcript.Length']
+    lengths.table = lengths[,'Transcript.Length', drop = TRUE]
     # }else{
   #   lengths.table <- lengths}
   

### 2021-09-06 Manuscript_pirAnalysis_Github/Pberghei/Manuscript_berghei_pir_DESEQ.Rmd
SuppInfoName DESEq
berg DESEQ \'apeglm\' package, renv update and stats .csv file save

@@
 library('DESeq2')
+library('apeglm')
 ```
@@
 write_xlsx(deseq_stats, 
-           'SupplementaryInformation_Stats_DESEQ_berghei.xlsx')
+           'SupplementaryInformation4_Stats_DESEQ_berghei.xlsx')
 ```
